### 2016-01-23 featureImagesTable.tex
Initial commit.

\begin{table}[!htb]
  \centering
  \begin{tabular*}{0.65\textwidth}{@{\extracolsep{\fill}} ll}
    \multicolumn{1}{c}{\textbf{Feature type}} & \multicolumn{1}{c}{\textbf{Image source}} \\
    \toprule
    \midrule
    \multicolumn{2}{c}{\textbf{Intensities}} \\
    \midrule
    normalized/preprocessed & FLAIR, T1, and T2 \\
    \midrule
    \multicolumn{2}{c}{\textbf{Symmetric template}} \\
    \midrule
    template difference & FLAIR, T1, and T2 \\
    contralateral difference & FLAIR, T1, and T2 \\
    location indices & FLAIR, T1, and T2 \\
    \midrule
    \multicolumn{2}{c}{\textbf{Segmentation probabilities}} \\
    \midrule
    $Pr$(cerebrospinal fluid) & T1 \\
    $Pr$(gray matter) & T1 \\
    $Pr$(white matter) & T1 \\
    $Pr$(deep bray matter) & T1 \\
    $Pr$(brain stem) & T1 \\
    $Pr$(cerebellum) & T1 \\
    \midrule
    \multicolumn{2}{c}{\textbf{Distance maps}} \\
    \midrule
    cerebrospinal fluid & T1 brain segmentation \\
    gray matter & T1 brain segmentation \\
    deep gray matter & T1 brain segmentation \\
    whole brain & T1 brain segmentation \\
    \midrule
    \multicolumn{2}{c}{\textbf{Neighborhood statistics}} \\
    \midrule
    mean & FLAIR, T1, and T2 \\
    standard deviation & FLAIR, T1, and T2 \\
    skewness & FLAIR, T1, and T2 \\
    \midrule
    \bottomrule
  \end{tabular*}
 \label{table:indices}
 \caption{List of feature images used for Stage 1 of the proposed white matter
          hyperintensity segmentation framework.}
\end{table}

### 2016-01-23 format.Rmd
Initial commit.

---
output:
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.5}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
bibliography: references.bib
csl: national-science-foundation-grant-proposals.csl
fontsize: 11pt
mainfont: Georgia
geometry: margin=1.0in
---

<!--
   - \setlength{\parskip}{3pt}
   - \setlength{\topsep}{0pt}
   - \setlength{\partopsep}{0pt}
   - \setlength{\itemsep}{0pt}
   - \setlength{\floatsep}{0pt}
   - \setlength{\intextsep}{2pt}
   - \setlength{\abovecaptionskip}{2pt}
   - \setlength{\belowcaptionskip}{0pt}
-->

<style type="text/css">
body,
code.bash{
  font-size: 8px;
}
pre {
  font-size: 8px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set( cache=TRUE )
```

\pagenumbering{gobble}

### 2016-01-23 methods.Rmd
Initial commit.

# Materials and Methods

Feature generation for the proposed framework is critical. The performance of the selected
machine learning strategy is limited by the input used for training. The image features
described below were selected based on previous work in the literature and our own observations
regarding tumor characteristics and what could possibly be encoded algorithmically.
Observations regarding tumor intensity differences from normal tissue has been encoded in
previously proposed research by calculating normalized neighborhood intensity statistics
and intensity modeling (e.g., Bauer et al. 2011, 2012; Geremia et al. 2012; Zikic et al.
2012). In these approaches both the features and classification algorithm are limited to
voxelwise considerations.

Our success in the MICCAI 2013 challenge was largely due to the inclusion of an extremely
important set of features based on the common observation of bilateral asymmetrization
caused by tumor presence. Additionally, several features were designed to encode spatial
coherence even though the RF classification is performed on a voxelwise basis. For example,
connected component geometry of the intensity modeling labels were used to discriminate
observations of compact, isotropic objects (vs. highly anisotropic) typical of tumor
presence and growth (Greenspan 1972). Table 1 provides a listing of image features used
and the motivation for their inclusion.

\input{featureImagesTable.tex}

### 2016-01-23 README.R

@@ -1,0 +1,25 @@
+library( rmarkdown )
+library( ggplot2 )
+library( RISmed )
+
+stitchedFile <- "stitched.Rmd";
+
+formatFile <- "format.Rmd"
+
+rmdFiles <- c( formatFile, "methods.Rmd"
+   )
+
+for( i in 1:length( rmdFiles ) )
+  {
+  if( i == 1 )
+    {
+    cmd <- paste( "cat", rmdFiles[i], ">", stitchedFile )
+    } else {
+    cmd <- paste( "cat", rmdFiles[i], ">>", stitchedFile )
+    }
+  system( cmd )
+  }
+
+cat( '\n Pandoc rendering', stitchedFile, '\n' )
+render( stitchedFile, output_format = "all" )
+

### 2016-01-24 featureImagesTable.tex
ENH:  more stuff.

@@
     $Pr$(white matter) & T1 \\
-    $Pr$(deep bray matter) & T1 \\
+    $Pr$(deep gray matter) & T1 \\
     $Pr$(brain stem) & T1 \\

### 2016-01-24 methods.Rmd
ENH:  more stuff.

@@
 
-Feature generation for the proposed framework is critical. The performance of the selected
-machine learning strategy is limited by the input used for training. The image features
-described below were selected based on previous work in the literature and our own observations
-regarding tumor characteristics and what could possibly be encoded algorithmically.
-Observations regarding tumor intensity differences from normal tissue has been encoded in
-previously proposed research by calculating normalized neighborhood intensity statistics
-and intensity modeling (e.g., Bauer et al. 2011, 2012; Geremia et al. 2012; Zikic et al.
-2012). In these approaches both the features and classification algorithm are limited to
-voxelwise considerations.
+## Imaging
 
-Our success in the MICCAI 2013 challenge was largely due to the inclusion of an extremely
-important set of features based on the common observation of bilateral asymmetrization
-caused by tumor presence. Additionally, several features were designed to encode spatial
-coherence even though the RF classification is performed on a voxelwise basis. For example,
-connected component geometry of the intensity modeling labels were used to discriminate
-observations of compact, isotropic objects (vs. highly anisotropic) typical of tumor
-presence and growth (Greenspan 1972). Table 1 provides a listing of image features used
-and the motivation for their inclusion.
+## Quantitative analysis
+
+Machine learning and pattern recognition techniques have proven extremely useful
+for various medical image analysis workflows (see, for example, the annual
+Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
+Image Computing and Computer-Aided Intervention international meeting [@mlmi2015]).
+Popular techniques such as support vector machines and neural networks have been applied
+successfully to clinically relevant imaging tasks such as segmentation
+(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
+
+Another versatile technique that has acheieved much success
+
+The random forest machine learning framework provides an excelllent fromework for medical
+image segmentation
+
+Supervised methodologies are uniquely characterized, in part, by the feature images that
+are used to identify the regions of interest.
+
+Intensity information alone is insufficient for removing false positives.  For example,
+as pointed out in [@Neema:2009aa], higher intensities can be found at the periventricular
+caps in normal subjects which often confounds automated lesion detection algorithms.
 
 \input{featureImagesTable.tex}
+
+![Representation of Stage 1 feature images for subject 01C1019.  Input for each
+subject consists of FLAIR, T1-, and T2-weighted images which are rigidly pre-aligned
+[@Avants:2014aa] to the
+space of the T1 image.  The three images are then preprocessed (N4 bias correction
+[@Tustison:2010ac] and denoising [@Manjon:2010aa]) followed by application of standard ANTs
+brain extraction and $n$-tissue segmentation protocols using the MMRR
+symmetric template corresponding and priors [@Tustison:2014ab] to the T1 image.  The
+feature images are then generated for voxelwise input to the RF model which results in
+the voting maps illustrated on the right which gives a probabilistic classification of
+tissue type.
+](Figures/featureImages.png)
+
+\clearpage
+
+# References

### 2016-01-25 format.Rmd
ENH:  More.

@@
 output:
+  word_document:
+    fig_caption: true
   pdf_document:

### 2016-01-25 methods.Rmd
ENH:  More.

@@
 ## Quantitative analysis
+
+<!--
+
+* Importance of machine learning in medical image analysis
+    * give examples
+    * (tools) support vector machines
+    * (tools) neural networks
+    * (apps)
+    * Random forests
+        * MS lesion detection (Geremeria 2011)
+        * Brain tumor segmentation (BRATS 2012 and BRATS 2013)
+        * Tustison 2015 (importance of feature images)
+        * Pustina 2016
+
+* Overview of protocol
+    * Preprocess all images
+        * Denoise
+        * N4 bias correction
+    * Create Stage 1 feature images
+    * Use those as input to the first stage RF model to get RF voting maps
+        * CSF
+        * GM
+        * WM
+        * Deep GM
+        * Brain stem
+        * Cerebellum
+        * WMHs
+    * Use RF voting maps as priors for antsAtroposN4.sh to perform Bayesian-based
+      segmentation with an additional Markov random field prior to promote spatial
+      smoothness of the labels.
+    * Use Stage 1 feature images + RF voting maps + antsAtroposN4.sh posteriors as
+      Stage 2 feature images.
+    * Remove isolated voxels as final refinement.
+
+* Creation of feature images (all contained within ANTs)
+    * Benefits of open source
+    * best registration (for template building)
+    * Preprocessing
+        * Denoising
+        * N4 bias correction
+    * Types of feature images and motivation for their use (Stage 1)
+        * Normalized intensities and neighborhood statistics image
+        * Symmetric template --- the brain exhibits a shape and intensity bilateral
+          symmetry under normal conditions.  WMHs can "break" this symmetry.
+        * Probabilistic segmentations, distance maps and symmetric template indices ---
+          help remove false positives
+          caused by partial voluming at the gray/white matter interface and the location
+          indices since, for example, higher intensities can be found at the periventricular
+          caps in normal subjects [@Neema:2009aa] which often confounds automated lesion
+          detection algorithms.
+    * Types of feature images and motivation for their use (Stage 2)
+        * All previously images (for the reasons stated above)
+        * ``antsAtroposN4.sh`` with a large prior weight (w=0.5) on all three modalities
+          to classify voxels based on intensity (use MRF to smooth out remaining noise
+          in labeling).
+
+* Public avail
+
+* Leave one out evaluation strategy overview
+    * Create stage 1 feature images for all subjects
+    * For each of the 24 subjects, take out a subject and create a Stage 1 random forest
+      model from the remaining feature images + manually labeled truth data.
+    * For each of the remaining 23 subjects, send them through the Stage 1 RF model.
+    * The above step permits the creation of the Stage 2 feature images
+    * Create the Stage 2 RF model from the 23 sets of Stage 2 feature images.
+    * Send the leave-one-out subject through the two RF models (as described above).
+    * Compare with ground truth.
+
+* (Save for Discusion) Difficulties of WMH evaluation (outlined in greater detail here [@MIAreview].
+    *
+-->
 
@@
 
-Another versatile technique that has acheieved much success
+Another popular machine learning technique that has acheieved much success is the
+random forest
 
@@
 
+![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
+forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
+are useful in providing a quantitative assessment of the predictive importance of each feature.
+The error bars provide the 95^{th} percentile (i.e., $1.96 \times sigma$) and illustrate the
+stability of the feature importance across models.](Figures/averageLeaveOneOut.png)
+
+To calculate this quantity for a single feature from a single random forest model, the
+decrease in prediction accuracy produced by omitting the specified feature is calculated
+during the out-of-bag phase of model creation.
+
+During the out-of-bag error calculation stage of the random forest model creation, the
+decrease in prediction accuracy with the omission of a single feature or variable is
+tracked and averaged. Those features which have the greatest decrease in mean accuracy are
+considered to be the most discriminative. In this work, we do not use these measurements for
+feature pruning.
+
 \clearpage

### 2016-01-26 methods.Rmd
ENH:  More work.

@@
 * Importance of machine learning in medical image analysis
-    * give examples
-    * (tools) support vector machines
-    * (tools) neural networks
-    * (apps)
-    * Random forests
-        * MS lesion detection (Geremeria 2011)
-        * Brain tumor segmentation (BRATS 2012 and BRATS 2013)
-        * Tustison 2015 (importance of feature images)
-        * Pustina 2016
+    x give examples
+    x (tools) support vector machines
+    x (tools) neural networks
+    x (apps)
+    x Random forests algorithms
+        x MS lesion detection (Geremeria 2011)
+        x Brain tumor segmentation (BRATS 2012 and BRATS 2013)
+        x Tustison 2015 (importance of feature images)
+        x Pustina 2016
+    x Random forests overview
 
@@
 
-Machine learning and pattern recognition techniques have proven extremely useful
+### Random forests for machine learning
+
+Machine learning and pattern recognition techniques have seen increased application
 for various medical image analysis workflows (see, for example, the annual
@@
 Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as segmentation
+successfully to clinically relevant imaging tasks such as supervised image segmentation
 (e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
+Facilitating the current employment of such techniques are the number of available imaging
+data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
+SciPy [@scipy] and the R project for statistical computing [@R] and their associated add-on
+toolkits.
 
-Another popular machine learning technique that has acheieved much success is the
-random forest
+Random forests [@breiman2001] is a popular machine learning technique that has
+demonstrated significant utitility for supervised segmentation tasks (e.g.,
+normal human brain segmentation) and other computer vision applications
+(e.g., human gait detection [@viola2005]).
+In the context of neuropathology, random forest-based paradigms have been employed in the
+delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
+and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2014aa].  Of note, these
+latter random forest approaches for brain tumor segmentation have performed well in recent
+international competitions.  In response to the lack of objective comparisons between
+segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
+2012 [@Menze:2015aa] and has continued every year since under the auspices of the International Conference
+of Medical Image Computing and Computer Assisted Interventions (MICCAI).
 
-The random forest machine learning framework provides an excelllent fromework for medical
-image segmentation
+Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
+that are built from training data.  Once constructed, data to be classified is "pushed"
+through each decision tree resulting in a single classification "vote" per tree.  These
+votes are then be used for regression or classification of the data.  Although decision
+trees had been extensively studied previously, the success of employing collections of
+such weak learners for boosting machine learning performance
+(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
+of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
+Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
+data (i.e., "bagging") resulting in the current random forest framework.
 
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.
+Crucial to these supervised segmentation approaches are the creation and selection of
+"features" as input in conjunction with the ground-truth for model construction.  For the
+targeted application in this work (i.e., white matter hyperintensities),
+regression/classification are performed at the voxelwise level.  In other words, each voxel
+within the region of interest is sent through the ensemble of decision trees and receives
+a set of classification votes from each tree permitting a regression or classification
+solution.  Since this procedure is performed at the voxelwise level, intensity information
+alone is insufficient for good segmentation performance since it lacks spatial context.
+For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
+periventricular caps in normal subjects which often confounds automated lesion detection
+algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
+even though machine learning and pattern recognition techniques are extremely powerful and
+have significant potential, just as crucial to outcome is the creative construction and
+deployment of salient feature images which we detail below.
 
-Intensity information alone is insufficient for removing false positives.  For example,
-as pointed out in [@Neema:2009aa], higher intensities can be found at the periventricular
-caps in normal subjects which often confounds automated lesion detection algorithms.
+### Feature images for WMH segmentation
 
@@
 
-![Representation of Stage 1 feature images for subject 01C1019.  Input for each
-subject consists of FLAIR, T1-, and T2-weighted images which are rigidly pre-aligned
-[@Avants:2014aa] to the
-space of the T1 image.  The three images are then preprocessed (N4 bias correction
-[@Tustison:2010ac] and denoising [@Manjon:2010aa]) followed by application of standard ANTs
-brain extraction and $n$-tissue segmentation protocols using the MMRR
-symmetric template corresponding and priors [@Tustison:2014ab] to the T1 image.  The
-feature images are then generated for voxelwise input to the RF model which results in
+Supervised methodologies are uniquely characterized, in part, by the feature images that
+are used to identify the regions of interest.  In Table 1, we provide a list and basic
+categorization of the feature images used for the initial (i.e., Stage 1---more on the use
+of multiple random forest stages below) segmentation of the white matter hyperintensities.
+In addition Figure 1 provides a representation of a set of feature images for a single
+subject analyzed in this work.
+
+![Representation of Stage 1 feature images for subject 01C1019.  The
+FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
+[@Avants:2014aa] to the space of the T1 image.  The three images are then preprocessed
+(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
+application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
+the MMRR symmetric template corresponding and priors [@Tustison:2014ab] to the T1 image.
+The feature images are then generated for voxelwise input to the RF model which results in
 the voting maps illustrated on the right which gives a probabilistic classification of
-tissue type.
-](Figures/featureImages.png)
+tissue type.  Not shown are the probability and voting images for the brain stem and
+cerebellum.](Figures/featureImages.png)
+
+As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
+acquisitions.  The FLAIR and T2 images are rigidly registered to the T1 image using
+the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
+are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
+N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
+range $[0,1]$.  Although we could have used the intensity standardization algorithm of
+[@nyul2000], we found that a simple linear rescaling produced better results.
+
+The T1 image is then processed via the ANTs brain
+extraction and tissue segmentation protocols described in [@Tustison:2014ab] in order to
+produce a mask for the brain parenchyma and provide probabilistic estimates of the
+cerebrospinal fluid (csf), gray matter, white matter, deep gray matter, brain stem, and
+cerebellum.  Segmentation is performed using the ANTs Atropos tool [@Avants:2011aa] and
+multi-model optimal symmetric shape/intensity templates [@Tustison:2014aa] created from the public
+MMRR data set  [@landman2011].
+
+To model the intensity information the first set of images simply includes the
+preprocessed and normalized intensity FLAIR, T1, and T2 images.  We also calculate a set
+of neighborhood statistics (mean, standard deviation, and skewness) feature images using
+a radius of one voxel. For each of the normalized
+images, we calculate the difference in intensities with the corresponding warped template
+component.  Previous success in the international brain tumor segmentation
+competition [@Menze:2015aa] was based on an important set of intensity features that were created
+from multi-modal templates mentioned previously [@Tustison:2014aa].  We employ the
+same strategy here.  For example, the template
+difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
+
+$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$
+
+where $\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$ is the transform
+which maps from the individual subject space to the
+template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
+to take advantage of the bilateral symmetry of the normal brain (in terms of both shape
+and intensity), and the fact that the presence of WMH violates that assumption,
+we use the symmetric templates to compute the contralateral intensity
+differences as an additional intensity feature.  For the FLAIR component, this contralateral
+difference image is calculated from
+
+$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$
+
+where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
+the symmetric template.
+
+The segmentation probability images described above are used as feature images to provide a
+spatial context for the random forest model prediction step.  Additional spatial contextual
+feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
+deep gray matter images.  These latter images are intended to help distinguish white matter
+hyperintensities from false positives induced by the partial voluming at the gray/white
+matter interface.  A third set of images are based on the voxel location within the
+space of the template.  The T1 image of the subject is registered to the T1 template
+component using a B-spline variant [@Tustison:2013ac] of the well-known ANTs Symmetric
+Normalization (SyN) algorithm [@Avants:2011ab].  Since the inverse transform is also
+derived as part of the registration process, we then register the voxel index locations
+back to the space of the individual subject.  These index images are then warped back to
+the space of the individual subject.  Note that this is similar in motivation to the work
+of [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
+coordinate system provided by the template to dramatically improve spatial specificity
+across all subjects.
+
+### Stacked/cascaded/concatenated random forests for improved segmentation performance
+
+It has been previously observed that
+
+### Code and data availability
+
+### Evaluation protocol
+
+# Results
 
@@
 are useful in providing a quantitative assessment of the predictive importance of each feature.
-The error bars provide the 95^{th} percentile (i.e., $1.96 \times sigma$) and illustrate the
-stability of the feature importance across models.](Figures/averageLeaveOneOut.png)
+The error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
+stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
 

### 2016-01-26 methods.Rmd
ENH:  More work.

@@
 Normalization (SyN) algorithm [@Avants:2011ab].  Since the inverse transform is also
-derived as part of the registration process, we then register the voxel index locations
-back to the space of the individual subject.  These index images are then warped back to
-the space of the individual subject.  Note that this is similar in motivation to the work
+derived as part of the registration process, we can warp the voxel index locations
+back to the space of the individual subject.    Note that this is similar in motivation to the work
 of [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard

### 2016-01-27 intro.Rmd
ENH: More.

--- 2016-01-25 methods.Rmd
@@
-# Materials and Methods
+# Introduction
 
-## Imaging
+## White matter hyperintensities in TBI
 
-## Quantitative analysis
+## Random forests for WMH segmentation
 
-<!--
-
-* Importance of machine learning in medical image analysis
-    * give examples
-    * (tools) support vector machines
-    * (tools) neural networks
-    * (apps)
-    * Random forests
-        * MS lesion detection (Geremeria 2011)
-        * Brain tumor segmentation (BRATS 2012 and BRATS 2013)
-        * Tustison 2015 (importance of feature images)
-        * Pustina 2016
-
-* Overview of protocol
-    * Preprocess all images
-        * Denoise
-        * N4 bias correction
-    * Create Stage 1 feature images
-    * Use those as input to the first stage RF model to get RF voting maps
-        * CSF
-        * GM
-        * WM
-        * Deep GM
-        * Brain stem
-        * Cerebellum
-        * WMHs
-    * Use RF voting maps as priors for antsAtroposN4.sh to perform Bayesian-based
-      segmentation with an additional Markov random field prior to promote spatial
-      smoothness of the labels.
-    * Use Stage 1 feature images + RF voting maps + antsAtroposN4.sh posteriors as
-      Stage 2 feature images.
-    * Remove isolated voxels as final refinement.
-
-* Creation of feature images (all contained within ANTs)
-    * Benefits of open source
-    * best registration (for template building)
-    * Preprocessing
-        * Denoising
-        * N4 bias correction
-    * Types of feature images and motivation for their use (Stage 1)
-        * Normalized intensities and neighborhood statistics image
-        * Symmetric template --- the brain exhibits a shape and intensity bilateral
-          symmetry under normal conditions.  WMHs can "break" this symmetry.
-        * Probabilistic segmentations, distance maps and symmetric template indices ---
-          help remove false positives
-          caused by partial voluming at the gray/white matter interface and the location
-          indices since, for example, higher intensities can be found at the periventricular
-          caps in normal subjects [@Neema:2009aa] which often confounds automated lesion
-          detection algorithms.
-    * Types of feature images and motivation for their use (Stage 2)
-        * All previously images (for the reasons stated above)
-        * ``antsAtroposN4.sh`` with a large prior weight (w=0.5) on all three modalities
-          to classify voxels based on intensity (use MRF to smooth out remaining noise
-          in labeling).
-
-* Public avail
-
-* Leave one out evaluation strategy overview
-    * Create stage 1 feature images for all subjects
-    * For each of the 24 subjects, take out a subject and create a Stage 1 random forest
-      model from the remaining feature images + manually labeled truth data.
-    * For each of the remaining 23 subjects, send them through the Stage 1 RF model.
-    * The above step permits the creation of the Stage 2 feature images
-    * Create the Stage 2 RF model from the 23 sets of Stage 2 feature images.
-    * Send the leave-one-out subject through the two RF models (as described above).
-    * Compare with ground truth.
-
-* (Save for Discusion) Difficulties of WMH evaluation (outlined in greater detail here [@MIAreview].
-    *
--->
-
-Machine learning and pattern recognition techniques have proven extremely useful
+Machine learning and pattern recognition techniques have seen increased application
 for various medical image analysis workflows (see, for example, the annual
@@
 Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as segmentation
+successfully to clinically relevant imaging tasks such as supervised image segmentation
 (e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
+Facilitating the current employment of such techniques are the number of available imaging
+data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
+SciPy [@scipy] and the R project for statistical computing [@R] and their associated add-on
+toolkits.
 
-Another popular machine learning technique that has acheieved much success is the
-random forest
+Random forests [@breiman2001] is a popular machine learning technique that has
+demonstrated significant utitility for supervised segmentation tasks (e.g.,
+normal human brain segmentation) and other computer vision applications
+(e.g., human gait detection [@viola2005]).
+In the context of neuropathology, random forest-based paradigms have been employed in the
+delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
+and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2014aa].  Of note, these
+latter random forest approaches for brain tumor segmentation have performed well in recent
+international competitions.  In response to the lack of objective comparisons between
+segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
+2012 [@Menze:2015aa] and has continued every year since under the auspices of the International Conference
+of Medical Image Computing and Computer Assisted Interventions (MICCAI).
 
-The random forest machine learning framework provides an excelllent fromework for medical
-image segmentation
+Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
+that are built from training data.  Once constructed, data to be classified is "pushed"
+through each decision tree resulting in a single classification "vote" per tree.  These
+votes are then be used for regression or classification of the data.  Although decision
+trees had been extensively studied previously, the success of employing collections of
+such weak learners for boosting machine learning performance
+(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
+of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
+Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
+data (i.e., "bagging") resulting in the current random forest framework.
 
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.
+In this work, we develop a concatenated random forest framework with a tailored feature image set
+for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
+the entire framework is provided as open source through the well-known open-source
+ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the investigation of this work is
+the additional public availability of the TBI cohorts thus permitting full reproducibility
+of the results reported and discussed.
 
-Intensity information alone is insufficient for removing false positives.  For example,
-as pointed out in [@Neema:2009aa], higher intensities can be found at the periventricular
-caps in normal subjects which often confounds automated lesion detection algorithms.
-
-\input{featureImagesTable.tex}
-
-![Representation of Stage 1 feature images for subject 01C1019.  Input for each
-subject consists of FLAIR, T1-, and T2-weighted images which are rigidly pre-aligned
-[@Avants:2014aa] to the
-space of the T1 image.  The three images are then preprocessed (N4 bias correction
-[@Tustison:2010ac] and denoising [@Manjon:2010aa]) followed by application of standard ANTs
-brain extraction and $n$-tissue segmentation protocols using the MMRR
-symmetric template corresponding and priors [@Tustison:2014ab] to the T1 image.  The
-feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right which gives a probabilistic classification of
-tissue type.
-](Figures/featureImages.png)
-
-![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
-forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-The error bars provide the 95^{th} percentile (i.e., $1.96 \times sigma$) and illustrate the
-stability of the feature importance across models.](Figures/averageLeaveOneOut.png)
-
-To calculate this quantity for a single feature from a single random forest model, the
-decrease in prediction accuracy produced by omitting the specified feature is calculated
-during the out-of-bag phase of model creation.
-
-During the out-of-bag error calculation stage of the random forest model creation, the
-decrease in prediction accuracy with the omission of a single feature or variable is
-tracked and averaged. Those features which have the greatest decrease in mean accuracy are
-considered to be the most discriminative. In this work, we do not use these measurements for
-feature pruning.
-
-\clearpage
-
-# References
+[^i1]: https://github.com/stnava/ANTs
+[^i2]: https://github.com/stnava/ANTsR

### 2016-01-27 methods.Rmd
ENH: More.

@@
 
-### Random forests for machine learning
-
-Machine learning and pattern recognition techniques have seen increased application
-for various medical image analysis workflows (see, for example, the annual
-Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention international meeting [@mlmi2015]).
-Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as supervised image segmentation
-(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
-Facilitating the current employment of such techniques are the number of available imaging
-data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
-SciPy [@scipy] and the R project for statistical computing [@R] and their associated add-on
-toolkits.
-
-Random forests [@breiman2001] is a popular machine learning technique that has
-demonstrated significant utitility for supervised segmentation tasks (e.g.,
-normal human brain segmentation) and other computer vision applications
-(e.g., human gait detection [@viola2005]).
-In the context of neuropathology, random forest-based paradigms have been employed in the
-delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
-and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2014aa].  Of note, these
-latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions.  In response to the lack of objective comparisons between
-segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
-2012 [@Menze:2015aa] and has continued every year since under the auspices of the International Conference
-of Medical Image Computing and Computer Assisted Interventions (MICCAI).
-
-Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
-that are built from training data.  Once constructed, data to be classified is "pushed"
-through each decision tree resulting in a single classification "vote" per tree.  These
-votes are then be used for regression or classification of the data.  Although decision
-trees had been extensively studied previously, the success of employing collections of
-such weak learners for boosting machine learning performance
-(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
-of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
-Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
-data (i.e., "bagging") resulting in the current random forest framework.
-
 Crucial to these supervised segmentation approaches are the creation and selection of
@@
 In addition Figure 1 provides a representation of a set of feature images for a single
-subject analyzed in this work.
+subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
+with seven labels:
+* cerebrospinal fluid (label 1),
+* gray matter (label 2),
+* white matter (label 3),
+* deep gray matter (label 4),
+* brain stem (label 5),
+* cerebellum (label 6), and
+* white matter hyperintensities (label 7).
 
@@
 cerebrospinal fluid (csf), gray matter, white matter, deep gray matter, brain stem, and
-cerebellum.  Segmentation is performed using the ANTs Atropos tool [@Avants:2011aa] and
+cerebellum.  These provide the ground-truth labels for the first six tissue labels
+given above.  The white matter hyperintensities were manually identified by one of the authors
+(J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
+the ANTs Atropos tool [@Avants:2011aa] and
 multi-model optimal symmetric shape/intensity templates [@Tustison:2014aa] created from the public
@@
 
-It has been previously observed that
+In previous brain tumor segmentation work [@Tustison:2014aa], it was demonstrated that a
+concatenated supervised approach, whereby the prediction output from the first random forest
+model serves as partial input for a second random forest model, can significantly improve
+segmentation performance.  We do the same thing for the work described here.  The Stage 1
+feature images of the training data (as described previously) are used to construct the
+Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
+voting maps via the Stage 1 model.  All the Stage 1 features plus the Stage 1 voting maps
+are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
+tissue priors for a second application of the Atropos maximum aposteriori algorithm with an
+additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
+the second stage we use all three aligned preprocessed images for a multivariate
+analysis.  The resulting seven posterior probability images constitute a third additional
+feature image set for Stage 2.
 
@@
 
+As pointed out in a recent comprehensive multiple sclerosis lesion segmentation review
+[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
+is quite extensive, there were only four publicly available segmentation algorithms
+at the time of writing of which none are based on supervised learning.  As we did for
+our brain tumor segmentation algorithm [@Tustison:2014aa], all of the code described in
+this work is publicly available through the open-source ANTs/ANTsR toolkits.  In addition,
+similar to our previous offering [^1], we plan on creating
+a self-encapsulated example to showcase the proposed methodology.  The fact that this
+the data will also be made available \textcolor{red}{through the FITBIR repository} along
+with the ground truth data.
+
+[^1]: https://github.com/ntustison/ANTsAndArboles
+
 ### Evaluation protocol
-
-# Results
-
-![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
-forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-The error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
-stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
-
-To calculate this quantity for a single feature from a single random forest model, the
-decrease in prediction accuracy produced by omitting the specified feature is calculated
-during the out-of-bag phase of model creation.
-
-During the out-of-bag error calculation stage of the random forest model creation, the
-decrease in prediction accuracy with the omission of a single feature or variable is
-tracked and averaged. Those features which have the greatest decrease in mean accuracy are
-considered to be the most discriminative. In this work, we do not use these measurements for
-feature pruning.
-
-\clearpage
-
-# References

### 2016-01-27 results.Rmd
ENH: More.

# Results

![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
The error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)

To calculate this quantity for a single feature from a single random forest model, the
decrease in prediction accuracy produced by omitting the specified feature is calculated
during the out-of-bag phase of model creation.

During the out-of-bag error calculation stage of the random forest model creation, the
decrease in prediction accuracy with the omission of a single feature or variable is
tracked and averaged. Those features which have the greatest decrease in mean accuracy are
considered to be the most discriminative. In this work, we do not use these measurements for
feature pruning.

\clearpage

# References

### 2016-01-27 README.R

@@ -6,7 +6,10 @@ stitchedFile <- "stitched.Rmd";
 
 formatFile <- "format.Rmd"
 
-rmdFiles <- c( formatFile, "methods.Rmd"
+rmdFiles <- c( formatFile,
+               "intro.Rmd",
+               "methods.Rmd",
+               "results.Rmd"
    )
 
 for( i in 1:length( rmdFiles ) )

### 2016-01-27 methods.Rmd
ENH:  More.

@@
 our brain tumor segmentation algorithm [@Tustison:2014aa], all of the code described in
-this work is publicly available through the open-source ANTs/ANTsR toolkits.  In addition,
-similar to our previous offering [^1], we plan on creating
+this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
+ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
+use the \texttt{randomForest} package [@liaw2002] using the default settings with
+2000 trees per model and 500 randomly selected samples per label per image.
+
+In addition, similar to our previous offering [^1], we plan on creating
 a self-encapsulated example to showcase the proposed methodology.  The fact that this
@@
 
-### Evaluation protocol
+### Evaluation protocol overview
+
+In order to evaluate the protocol described, we perform a leave-one-out evaluation using
+the data acquired from the 24 subjects described above.  The Stage 1 feature images
+were created for all subjects.  The initial brain segmentation of each T1 image and the
+manual white matter hyperintensity tracings were combined to provide the truth labels
+for the training data.  The truth labels are the seven anatomical regions given above.
+
+The leave-one-out procedure is as follows:
+
+* Create Stage 1 feature images for all 24 subjects.
+* For each of the 24 subjects:
+    + sequester the current subject and corresponding feature images.
+    + construct the Stage 1 random forest model from the remaining 23 subjects.
+    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
+    + the previous step produces the Stage 1 voting maps for all seven labels.
+    + for each of the 23 subjects, perform a Bayesian-based segmentation with an MRF spatial
+      prior using the seven voting maps are used as additional tissue priors.
+    + construct the Stage 2 random forest model from all the Stage 1 feature images,
+      seven voting maps, and seven posterior probability maps from the previous step.
+    + send the sequestered subject through the random forest models for both stages.
+    + compare the final results with the manually-defined white matter hyperintensity
+      regions.

### 2016-01-28 results.Rmd
ENH:  more.

@@
+<!--
+
+* Evaluation
+    * Proper context
+        * We only have manual labelings from a single radiologist (
+        * TBI white matter hyperintensities are more difficult to segment than MS lesions
+            * individual lesions are smaller
+            * lesion load is typically smaller
+            * not prone to enhancement
+            * less distinct from surrounding tissue ?
+    * Despite the doubts regarding the efficacy of a typical segmentation validation,
+      our contribution is useful because
+         * the total segmentation framework is publicly available within the ANTs/ANTsR
+           framework
+         * although only a single site is analyzed, the feature images are site-agnostic
+              * we plan to apply the current RF models directly to the other site data
+              * we can then build the training database further by enlisting other experts
+         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
+         * The data (including manual tracings) will be made available as part of the FITBIR
+           effort.
+
+-->
+
 # Results
+
+As mentioned previously, segmentation of white matter hyperintensities in traumatic brain
+injury cohorts is a much more difficult problem than
 

### 2016-01-29 featureImagesTable.tex
ENH:  More.

@@
   \centering
-  \begin{tabular*}{0.65\textwidth}{@{\extracolsep{\fill}} ll}
-    \multicolumn{1}{c}{\textbf{Feature type}} & \multicolumn{1}{c}{\textbf{Image source}} \\
+  \begin{tabular*}{0.85\textwidth}{@{\extracolsep{\fill}} lll}
+    \multicolumn{1}{c}{\textbf{Feature type}} & \multicolumn{1}{c}{\textbf{Image source}} & \multicolumn{1}{c}{\textbf{Name}} \\
     \toprule
     \midrule
-    \multicolumn{2}{c}{\textbf{Intensities}} \\
+    \multicolumn{3}{c}{\textbf{Intensities}} \\
     \midrule
-    normalized/preprocessed & FLAIR, T1, and T2 \\
+    normalized/preprocessed & FLAIR, T1, and T2 & \texttt{Preprocessed} \\
     \midrule
-    \multicolumn{2}{c}{\textbf{Symmetric template}} \\
+    \multicolumn{3}{c}{\textbf{Symmetric template}} \\
     \midrule
-    template difference & FLAIR, T1, and T2 \\
-    contralateral difference & FLAIR, T1, and T2 \\
-    location indices & FLAIR, T1, and T2 \\
+    template difference & FLAIR, T1, and T2 & \texttt{SymmetricTemplateDifference}\\
+    contralateral difference & FLAIR, T1, and T2 & \texttt{ContralateralDifference}\\
+    template location indices & FLAIR, T1, and T2 & \texttt{TemplateIndicesWarped}\\
     \midrule
-    \multicolumn{2}{c}{\textbf{Segmentation probabilities}} \\
+    \multicolumn{3}{c}{\textbf{Segmentation probabilities}} \\
     \midrule
-    $Pr$(cerebrospinal fluid) & T1 \\
-    $Pr$(gray matter) & T1 \\
-    $Pr$(white matter) & T1 \\
-    $Pr$(deep gray matter) & T1 \\
-    $Pr$(brain stem) & T1 \\
-    $Pr$(cerebellum) & T1 \\
+    $Pr$(cerebrospinal fluid) & T1 & \texttt{BrainSegmentationPosteriors1}\\
+    $Pr$(gray matter) & T1 & \texttt{BrainSegmentationPosteriors2}\\
+    $Pr$(white matter) & T1 & \texttt{BrainSegmentationPosteriors3}\\
+    $Pr$(deep gray matter) & T1 & \texttt{BrainSegmentationPosteriors4}\\
+    $Pr$(brain stem) & T1 & \texttt{BrainSegmentationPosteriors5}\\
+    $Pr$(cerebellum) & T1 & \texttt{BrainSegmentationPosteriors6}\\
     \midrule
-    \multicolumn{2}{c}{\textbf{Distance maps}} \\
+    \multicolumn{3}{c}{\textbf{Distance maps}} \\
     \midrule
-    cerebrospinal fluid & T1 brain segmentation \\
-    gray matter & T1 brain segmentation \\
-    deep gray matter & T1 brain segmentation \\
-    whole brain & T1 brain segmentation \\
+    cerebrospinal fluid & T1 brain segmentation & \texttt{NormalizedCsfDistance}\\
+    gray matter & T1 brain segmentation & \texttt{NormalizedGmDistance}\\
+    deep gray matter & T1 brain segmentation & \texttt{NormalizedDeepGmDistance}\\
+    whole brain & T1 brain segmentation & \texttt{NormalizedMaskDistance}\\
     \midrule
-    \multicolumn{2}{c}{\textbf{Neighborhood statistics}} \\
+    \multicolumn{3}{c}{\textbf{Neighborhood statistics}} \\
     \midrule
-    mean & FLAIR, T1, and T2 \\
-    standard deviation & FLAIR, T1, and T2 \\
-    skewness & FLAIR, T1, and T2 \\
+    mean & FLAIR, T1, and T2 & \texttt{NeighborhoodMean}\\
+    standard deviation & FLAIR, T1, and T2  & \texttt{NeighborhoodSigma}\\
+    skewness & FLAIR, T1, and T2  & \texttt{NeighborhoodSkewness}\\
     \midrule
@@
  \caption{List of feature images used for Stage 1 of the proposed white matter
-          hyperintensity segmentation framework.}
+          hyperintensity segmentation framework.  These feature images encompass
+          salient location and intensity information used for voxelwise identification
+          of white matter hyperintensities in TBI.  The column on the right provides
+          the names used in the software developed for this methodology and
+          facilitates discussion of the results (cf Figure 2).
+          }
 \end{table}

### 2016-01-29 results.Rmd
ENH:  More.

@@
 
-As mentioned previously, segmentation of white matter hyperintensities in traumatic brain
-injury cohorts is a much more difficult problem than
+After performing the leave-one-out evaluation described at the end of the previous section,
+we calculated the \texttt{MeanDecreaseAccuracy} feature values for each of the 24 subjects $\times$
+2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
+during the out-of-bag phase of the random forest model construction and quantifies the decrease
+in prediction accuracy from omitting the specified feature.  In other words, this quantity
+helps determine the importance of a particular feature and, although we save such
+efforts for future work, this information provides us with guidance for future feature
+pruning and/or additions.
+
+The resulting rankings for both Stages are given in Figure 2 where the values for the
+separate stages are averaged over the entire corresponding model set.  In addition, we
+track the variance for each feature over all models to illustrate the stability of
+the chosen features during the evaluation.  This latter information is illustrated as
+horizontal errors bars providing the $95^{th}$ percentile (i.e., $1.96 \times \sigma$).
+
+One can also use these measurements as a type of sanity check.  For example, from the Stage
+1  plot, one can see that the \texttt{MeanDecreaseAccuracy} values for the location indices in the
+anterior-posterior direction (i.e., \texttt{TemplateIndicesWarped1}) are greater than
+those for either the inferior-superior (i.e., \texttt{TemplateIndicesWarped0}) or the
+left-right (i.e., \texttt{TemplateIndicesWarped0}) directions in the space of the symmetric
+template.  This is intuitive since, as discussed previously, manifestation of TBI white
+matter hyperintensities can often be confused with higher intensities at the
+periventricular caps in normal subjects [@Neema:2009aa] whereas there does not seem to
+be contralateral bias in manifestation of white matter hyperintensities in TBI [@reference].
+Additionally, it is interesting to see that the top two performing
+A couple of  other interesting
 
@@
 are useful in providing a quantitative assessment of the predictive importance of each feature.
-The error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
+The horizontal error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
 stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
 
-To calculate this quantity for a single feature from a single random forest model, the
-decrease in prediction accuracy produced by omitting the specified feature is calculated
-during the out-of-bag phase of model creation.
-
-During the out-of-bag error calculation stage of the random forest model creation, the
-decrease in prediction accuracy with the omission of a single feature or variable is
-tracked and averaged. Those features which have the greatest decrease in mean accuracy are
-considered to be the most discriminative. In this work, we do not use these measurements for
-feature pruning.
+![LLV](Figures/llvAndDice.png)
 

### 2016-01-29 intro.Rmd
ENH:  More.
ENH:  more.

@@
 delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
-and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2014aa].  Of note, these
+and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
 latter random forest approaches for brain tumor segmentation have performed well in recent

### 2016-01-29 methods.Rmd
ENH:  more.

@@
 with seven labels:
+
 * cerebrospinal fluid (label 1),
@@
 the ANTs Atropos tool [@Avants:2011aa] and
-multi-model optimal symmetric shape/intensity templates [@Tustison:2014aa] created from the public
+multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
 MMRR data set  [@landman2011].
@@
 competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2014aa].  We employ the
+from multi-modal templates mentioned previously [@Tustison:2015aa].  We employ the
 same strategy here.  For example, the template
@@
 
-### Stacked/cascaded/concatenated random forests for improved segmentation performance
+### Stacked (concatenated) random forests for improved segmentation performance
 
-In previous brain tumor segmentation work [@Tustison:2014aa], it was demonstrated that a
+In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
 concatenated supervised approach, whereby the prediction output from the first random forest
@@
 at the time of writing of which none are based on supervised learning.  As we did for
-our brain tumor segmentation algorithm [@Tustison:2014aa], all of the code described in
+our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
 this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through

### 2016-01-29 results.Rmd
ENH:  more.

@@
 
+## Ranking feature importance
+
 After performing the leave-one-out evaluation described at the end of the previous section,
@@
 
+![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
+forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
+are useful in providing a quantitative assessment of the predictive importance of each feature.
+The horizontal error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
+stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
+
 The resulting rankings for both Stages are given in Figure 2 where the values for the
@@
 horizontal errors bars providing the $95^{th}$ percentile (i.e., $1.96 \times \sigma$).
+Note that the reader can cross reference Table 1 for identifying corresponding feature types
+and names.
 
@@
 be contralateral bias in manifestation of white matter hyperintensities in TBI [@reference].
-Additionally, it is interesting to see that the top two performing
-A couple of  other interesting
 
-![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
-forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-The horizontal error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
-stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
+Additionally, it is interesting to note some of the other top performing features for Stage 1.
+The contralateral difference FLAIR image is  highly discriminative over the set of
+evaluation random forest models.  This accords with the known clinical relevance of
+FLAIR images for identifying white matter hyperintensities and the fact that such
+pathology does not manifest symmetrically in both hemispheres.  Interestingly, the
+posterior maps for the deep gray matter are extremely important for accurate white matter
+hyperintensity segmentation.  \textcolor{red}{Perhaps the spatial specification of deep
+gray matter aids in the removal of false positives}.  Inspection of the bottom of the
+plots demonstrates the lack of discriminating features associated with the
+T1 image which is also well-known in the clinical literature.
 
-![LLV](Figures/llvAndDice.png)
+As described earlier, for Stage 2, we used the output random forest voting maps from Stage
+1 as both features themselves and as priors for input to a Bayesian-based segmentation with
+an additional MRF spatial prior.  In Figure 2, the voting maps are labeled as
+"\texttt{RFStage1VotingMaps}" where the final numeral is associated with the brain
+parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
+feature probability maps are labeled as "\texttt{RFBrainSegmentationPosteriors}".
+The Stage 2 feature importance plot follows similar
+trends as that for Stage 1 with the T1 images not contributing much to the identification
+of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
+important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
+3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
+alone it is intuitive that such features would be important.
+
+## White matter hyperintensity segmentation evaluation
+
+In Figure 3 are the segmentation comparisons derived from manual segmentations of the same
+data.  Despite the large variability characteristic with manual labelings in related fields
+[@Grimaud:1996aa;styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
+current clinical practices and the methodology proposed herein is readily adapted to refinements
+in training data.  On the left of Figure 3 are the improvement in Dice values over all white
+matter hyperintensities when comparing the segmentations between the two stages.  Performing
+the second round of supervised learning improves Dice values.  One can also note from the
+right side of Figure 3 that the total lesion load volume illustrates a few subjects that are
+severe outliers in terms of the number of false positives.  The second round helps to
+correct this issue.
+
+![Comparison with manual delineation of white matter hyperintensities.  On the left are the
+calculate Dice values over all white matter hyperintensities.  Note the improvement
+in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
+(Right) Similar results can be seen by comparing the total lesion load volume between manual
+and automated detection strategies.  Although some outliers are found after the Stage 2
+processing in a couple subjects, the number of outliers caused by false positives is
+significantly with the second stage processing.](Figures/llvAndDice.png)
 

### 2016-01-29 methods.Rmd
ENH:  More.

@@
 multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
-MMRR data set  [@landman2011].
+MMRR data set  [@landman2011] (cf Figure X).
+
+![Canonical views of the mutlivariate, bilaterally symmetric template constructed
+from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
+the components relevant for this work).  Template construction is detailed in
+[@Tustison:2015aa].  These images are important for certain intensity-based
+features.](Figures/MMRR.png)
 

### 2016-01-29 intro.Rmd
ENH:  Read-through.

@@
 Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention international meeting [@mlmi2015]).
+Image Computing and Computer-Aided Intervention (MICCAI) international meeting [@mlmi2015]).
 Popular techniques such as support vector machines and neural networks have been applied
@@
 data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
-SciPy [@scipy] and the R project for statistical computing [@R] and their associated add-on
-toolkits.
+SciPy [@scipy] and the R project for statistical computing [@R] and their associated
+extensions.
 
-Random forests [@breiman2001] is a popular machine learning technique that has
+The random forests framework [@breiman2001] is a popular machine learning technique that has
 demonstrated significant utitility for supervised segmentation tasks (e.g.,
-normal human brain segmentation) and other computer vision applications
+normal human brain segmentation [@yi2009]) and other computer vision applications
 (e.g., human gait detection [@viola2005]).
@@
 segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
-2012 [@Menze:2015aa] and has continued every year since under the auspices of the International Conference
-of Medical Image Computing and Computer Assisted Interventions (MICCAI).
+2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI
+conference.
 
@@
 through each decision tree resulting in a single classification "vote" per tree.  These
-votes are then be used for regression or classification of the data.  Although decision
+votes are then used for regression or classification of the data.  Although decision
 trees had been extensively studied previously, the success of employing collections of
@@
 Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
-data (i.e., "bagging") resulting in the current random forest framework.
+data (i.e., "bagging") resulting in the current random forest technique.
 
-In this work, we develop a concatenated random forest framework with a tailored feature image set
+In this work, we develop a concatenated random forest framework with a tailored contextual
+feature image set (both spatial and intensity-based)
 for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
-the entire framework is provided as open source through the well-known open-source
-ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the investigation of this work is
-the additional public availability of the TBI cohorts thus permitting full reproducibility
+the entire framework is provided publicly through the well-known open-source
+ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the research of this work is
+the public availability of the imaging data thus permitting full reproducibility
 of the results reported and discussed.

### 2016-02-01 intro.Rmd
ENH:  James edits.

@@
 ## White matter hyperintensities in TBI
+
+White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
+These lesions are distinguished from prominent perivascular spaces seen on T2-weighted imaging by the lack of fluid suppression of signal on FLAIR sequences.
+WMHs in the periventricular and deep brain regions are associated with normal aging and neurological conditions including hypertension and stroke.
+WMHs are also a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult
+[@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa].
+Further, the regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa].
+Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa].
+Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa].
+Additionally, the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].
+
+Despite the above findings, outside of multiple sclerosis, WMHs are not routinely employed as a diagnostic measure in clinical practice.
+Their presence within asymptomatic patients or in association with a variety of conditions, such as stroke, dementia, neuroinflammatory conditions, and TBI challenge their utility in narrowing a radiological differential diagnosis.
+Further, performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
+Despite the limited inclusion of WMH observations in routine radiological reports, two large meta-analyses demonstrated an association between WMHs, cognitive function, increased risk of stroke, dementia, and death [@Debette:2010aa;@Kloppenborg:2014aa].
+As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state.
+Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
 

### 2016-02-01 methods.Rmd
ENH:  James edits.

@@
 "features" as input in conjunction with the ground-truth for model construction.  For the
-targeted application in this work (i.e., white matter hyperintensities),
+targeted application in this work (i.e., WMHs),
 regression/classification are performed at the voxelwise level.  In other words, each voxel
 within the region of interest is sent through the ensemble of decision trees and receives
-a set of classification votes from each tree permitting a regression or classification
+a set of classification votes from each tree thus permitting a regression or classification
 solution.  Since this procedure is performed at the voxelwise level, intensity information
-alone is insufficient for good segmentation performance since it lacks spatial context.
+alone is insufficient for good segmentation performance due to the lack of spatial context.
 For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
@@
 FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the T1 image.  The three images are then preprocessed
+[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
 (N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
 application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template corresponding and priors [@Tustison:2014ab] to the T1 image.
+the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
 The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right which gives a probabilistic classification of
+the voting maps illustrated on the right.  This gives a probabilistic classification of
 tissue type.  Not shown are the probability and voting images for the brain stem and
@@
 N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
-range $[0,1]$.  Although we could have used the intensity standardization algorithm of
-[@nyul2000], we found that a simple linear rescaling produced better results.
+range $[0,1]$.  Although we could have used an alternative intensity standardization
+algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results.
+
+![Sample FLAIR acquisition image slices showing both manual (top) and random forest
+(RDF) segmentations (bottom) obtained during the leave-one-out evaluation.  Manual
+segmentations were performed by one of the authors and provided the ground
+truth WMH labels for training the random forest models.](Figures/montage.png)
 
 The T1 image is then processed via the ANTs brain
-extraction and tissue segmentation protocols described in [@Tustison:2014ab] in order to
+extraction and normal tissue segmentation (i.e., the pathology is discounted which is
+a fairly safe assumption for the T1 image) protocols described in [@Tustison:2014ab] in order to
 produce a mask for the brain parenchyma and provide probabilistic estimates of the
@@
 multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
-MMRR data set  [@landman2011] (cf Figure X).
+MMRR data set  [@landman2011] (cf Figure 3).
 
@@
 the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for certain intensity-based
+[@Tustison:2015aa].  These images are important for specific intensity-based
 features.](Figures/MMRR.png)
@@
 To model the intensity information the first set of images simply includes the
-preprocessed and normalized intensity FLAIR, T1, and T2 images.  We also calculate a set
+preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.  We also calculate a set
 of neighborhood statistics (mean, standard deviation, and skewness) feature images using
@@
 to take advantage of the bilateral symmetry of the normal brain (in terms of both shape
-and intensity), and the fact that the presence of WMH violates that assumption,
+and intensity), and the fact that the presence of WMHs violates that assumption,
 we use the symmetric templates to compute the contralateral intensity
@@
 the second stage we use all three aligned preprocessed images for a multivariate
-analysis.  The resulting seven posterior probability images constitute a third additional
+segmentation.  The resulting seven posterior probability images constitute a third additional
 feature image set for Stage 2.
@@
 In addition, similar to our previous offering [^1], we plan on creating
-a self-encapsulated example to showcase the proposed methodology.  The fact that this
-the data will also be made available \textcolor{red}{through the FITBIR repository} along
-with the ground truth data.
+a self-encapsulated example to showcase the proposed methodology.  The fact that
+the data will also be made available through the FITBIR repository along
+with the manual labelings will facilitate reproducibility on the part of the reader
+as well as any interest in extending the proposed framework to other data sets.
 
@@
 
-In order to evaluate the protocol described, we perform a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.  The Stage 1 feature images
-were created for all subjects.  The initial brain segmentation of each T1 image and the
+In order to evaluate the protocol described, we performed a leave-one-out evaluation using
+the data acquired from the 24 subjects described above.  Initial processing included the creation
+of all Stage 1 feature images for all subjects.  The initial brain segmentation of each T1 image and the
 manual white matter hyperintensity tracings were combined to provide the truth labels

### 2016-02-01 results.Rmd
ENH:  James edits.

@@
 
-The resulting rankings for both Stages are given in Figure 2 where the values for the
+The resulting rankings for both Stages are given in Figure 4 where the values for the
 separate stages are averaged over the entire corresponding model set.  In addition, we
@@
 periventricular caps in normal subjects [@Neema:2009aa] whereas there does not seem to
-be contralateral bias in manifestation of white matter hyperintensities in TBI [@reference].
+be contralateral bias in manifestation of white matter hyperintensities in TBI.
 
@@
 posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  \textcolor{red}{Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives}.  Inspection of the bottom of the
+hyperintensity segmentation.  Perhaps the spatial specification of deep
+gray matter aids in the removal of false positives.  Inspection of the bottom of the
 plots demonstrates the lack of discriminating features associated with the
@@
 1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 2, the voting maps are labeled as
+an additional MRF spatial prior.  In Figure 4, the voting maps are labeled as
 "\texttt{RFStage1VotingMaps}" where the final numeral is associated with the brain
@@
 
-In Figure 3 are the segmentation comparisons derived from manual segmentations of the same
+In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
 data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
+[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
 current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.  On the left of Figure 3 are the improvement in Dice values over all white
+in training data.  On the left of Figure 5 are the improvement in Dice values over all white
 matter hyperintensities when comparing the segmentations between the two stages.  Performing
 the second round of supervised learning improves Dice values.  One can also note from the
-right side of Figure 3 that the total lesion load volume illustrates a few subjects that are
+right side of Figure 5 that the total lesion load volume illustrates a few subjects that are
 severe outliers in terms of the number of false positives.  The second round helps to

### 2016-02-01 stitched.Rmd
ENH:  James edits.

--- 2016-02-01 methods.Rmd
@@
+---
+output:
+  word_document:
+    fig_caption: true
+  pdf_document:
+    fig_caption: true
+    latex_engine: xelatex
+    keep_tex: yes
+header-includes:
+   - \usepackage{booktabs}
+   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
+   - \linespread{1.5}
+   - \usepackage[compact]{titlesec}
+   - \usepackage{enumitem}
+   - \usepackage{tikz}
+   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
+   - \setlist{nolistsep}
+   - \titlespacing{\section}{2pt}{*0}{*0}
+   - \titlespacing{\subsection}{2pt}{*0}{*0}
+   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
+   - \setlength{\parskip}{3pt}
+bibliography: references.bib
+csl: national-science-foundation-grant-proposals.csl
+fontsize: 11pt
+mainfont: Georgia
+geometry: margin=1.0in
+---
+
+<!--
+   - \setlength{\parskip}{3pt}
+   - \setlength{\topsep}{0pt}
+   - \setlength{\partopsep}{0pt}
+   - \setlength{\itemsep}{0pt}
+   - \setlength{\floatsep}{0pt}
+   - \setlength{\intextsep}{2pt}
+   - \setlength{\abovecaptionskip}{2pt}
+   - \setlength{\belowcaptionskip}{0pt}
+-->
+
+<style type="text/css">
+body,
+code.bash{
+  font-size: 8px;
+}
+pre {
+  font-size: 8px
+}
+</style>
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set( cache=TRUE )
+```
+
+\pagenumbering{gobble}
+
+# Introduction
+
+## White matter hyperintensities in TBI
+
+White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
+These lesions are distinguished from prominent perivascular spaces seen on T2-weighted imaging by the lack of fluid suppression of signal on FLAIR sequences.
+WMHs in the periventricular and deep brain regions are associated with normal aging and neurological conditions including hypertension and stroke.
+WMHs are also a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult
+[@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa].
+Further, the regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa].
+Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa].
+Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa].
+Additionally, the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].
+
+Despite the above findings, outside of multiple sclerosis, WMHs are not routinely employed as a diagnostic measure in clinical practice.
+Their presence within asymptomatic patients or in association with a variety of conditions, such as stroke, dementia, neuroinflammatory conditions, and TBI challenge their utility in narrowing a radiological differential diagnosis.
+Further, performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
+Despite the limited inclusion of WMH observations in routine radiological reports, two large meta-analyses demonstrated an association between WMHs, cognitive function, increased risk of stroke, dementia, and death [@Debette:2010aa;@Kloppenborg:2014aa].
+As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state.
+Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
+
+## Random forests for WMH segmentation
+
+Machine learning and pattern recognition techniques have seen increased application
+for various medical image analysis workflows (see, for example, the annual
+Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
+Image Computing and Computer-Aided Intervention (MICCAI) international meeting [@mlmi2015]).
+Popular techniques such as support vector machines and neural networks have been applied
+successfully to clinically relevant imaging tasks such as supervised image segmentation
+(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
+Facilitating the current employment of such techniques are the number of available imaging
+data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
+SciPy [@scipy] and the R project for statistical computing [@R] and their associated
+extensions.
+
+The random forests framework [@breiman2001] is a popular machine learning technique that has
+demonstrated significant utitility for supervised segmentation tasks (e.g.,
+normal human brain segmentation [@yi2009]) and other computer vision applications
+(e.g., human gait detection [@viola2005]).
+In the context of neuropathology, random forest-based paradigms have been employed in the
+delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
+and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
+latter random forest approaches for brain tumor segmentation have performed well in recent
+international competitions.  In response to the lack of objective comparisons between
+segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
+2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI
+conference.
+
+Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
+that are built from training data.  Once constructed, data to be classified is "pushed"
+through each decision tree resulting in a single classification "vote" per tree.  These
+votes are then used for regression or classification of the data.  Although decision
+trees had been extensively studied previously, the success of employing collections of
+such weak learners for boosting machine learning performance
+(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
+of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
+Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
+data (i.e., "bagging") resulting in the current random forest technique.
+
+In this work, we develop a concatenated random forest framework with a tailored contextual
+feature image set (both spatial and intensity-based)
+for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
+the entire framework is provided publicly through the well-known open-source
+ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the research of this work is
+the public availability of the imaging data thus permitting full reproducibility
+of the results reported and discussed.
+
+[^i1]: https://github.com/stnava/ANTs
+[^i2]: https://github.com/stnava/ANTsR
+
 # Materials and Methods
@@
       regions.
+
+<!--
+
+* Evaluation
+    * Proper context
+        * We only have manual labelings from a single radiologist (
+        * TBI white matter hyperintensities are more difficult to segment than MS lesions
+            * individual lesions are smaller
+            * lesion load is typically smaller
+            * not prone to enhancement
+            * less distinct from surrounding tissue ?
+    * Despite the doubts regarding the efficacy of a typical segmentation validation,
+      our contribution is useful because
+         * the total segmentation framework is publicly available within the ANTs/ANTsR
+           framework
+         * although only a single site is analyzed, the feature images are site-agnostic
+              * we plan to apply the current RF models directly to the other site data
+              * we can then build the training database further by enlisting other experts
+         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
+         * The data (including manual tracings) will be made available as part of the FITBIR
+           effort.
+
+-->
+
+# Results
+
+## Ranking feature importance
+
+After performing the leave-one-out evaluation described at the end of the previous section,
+we calculated the \texttt{MeanDecreaseAccuracy} feature values for each of the 24 subjects $\times$
+2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
+during the out-of-bag phase of the random forest model construction and quantifies the decrease
+in prediction accuracy from omitting the specified feature.  In other words, this quantity
+helps determine the importance of a particular feature and, although we save such
+efforts for future work, this information provides us with guidance for future feature
+pruning and/or additions.
+
+![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
+forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
+are useful in providing a quantitative assessment of the predictive importance of each feature.
+The horizontal error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
+stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
+
+The resulting rankings for both Stages are given in Figure 4 where the values for the
+separate stages are averaged over the entire corresponding model set.  In addition, we
+track the variance for each feature over all models to illustrate the stability of
+the chosen features during the evaluation.  This latter information is illustrated as
+horizontal errors bars providing the $95^{th}$ percentile (i.e., $1.96 \times \sigma$).
+Note that the reader can cross reference Table 1 for identifying corresponding feature types
+and names.
+
+One can also use these measurements as a type of sanity check.  For example, from the Stage
+1  plot, one can see that the \texttt{MeanDecreaseAccuracy} values for the location indices in the
+anterior-posterior direction (i.e., \texttt{TemplateIndicesWarped1}) are greater than
+those for either the inferior-superior (i.e., \texttt{TemplateIndicesWarped0}) or the
+left-right (i.e., \texttt{TemplateIndicesWarped0}) directions in the space of the symmetric
+template.  This is intuitive since, as discussed previously, manifestation of TBI white
+matter hyperintensities can often be confused with higher intensities at the
+periventricular caps in normal subjects [@Neema:2009aa] whereas there does not seem to
+be contralateral bias in manifestation of white matter hyperintensities in TBI.
+
+Additionally, it is interesting to note some of the other top performing features for Stage 1.
+The contralateral difference FLAIR image is  highly discriminative over the set of
+evaluation random forest models.  This accords with the known clinical relevance of
+FLAIR images for identifying white matter hyperintensities and the fact that such
+pathology does not manifest symmetrically in both hemispheres.  Interestingly, the
+posterior maps for the deep gray matter are extremely important for accurate white matter
+hyperintensity segmentation.  Perhaps the spatial specification of deep
+gray matter aids in the removal of false positives.  Inspection of the bottom of the
+plots demonstrates the lack of discriminating features associated with the
+T1 image which is also well-known in the clinical literature.
+
+As described earlier, for Stage 2, we used the output random forest voting maps from Stage
+1 as both features themselves and as priors for input to a Bayesian-based segmentation with
+an additional MRF spatial prior.  In Figure 4, the voting maps are labeled as
+"\texttt{RFStage1VotingMaps}" where the final numeral is associated with the brain
+parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
+feature probability maps are labeled as "\texttt{RFBrainSegmentationPosteriors}".
+The Stage 2 feature importance plot follows similar
+trends as that for Stage 1 with the T1 images not contributing much to the identification
+of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
+important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
+3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
+alone it is intuitive that such features would be important.
+
+## White matter hyperintensity segmentation evaluation
+
+In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
+data.  Despite the large variability characteristic with manual labelings in related fields
+[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
+current clinical practices and the methodology proposed herein is readily adapted to refinements
+in training data.  On the left of Figure 5 are the improvement in Dice values over all white
+matter hyperintensities when comparing the segmentations between the two stages.  Performing
+the second round of supervised learning improves Dice values.  One can also note from the
+right side of Figure 5 that the total lesion load volume illustrates a few subjects that are
+severe outliers in terms of the number of false positives.  The second round helps to
+correct this issue.
+
+![Comparison with manual delineation of white matter hyperintensities.  On the left are the
+calculate Dice values over all white matter hyperintensities.  Note the improvement
+in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
+(Right) Similar results can be seen by comparing the total lesion load volume between manual
+and automated detection strategies.  Although some outliers are found after the Stage 2
+processing in a couple subjects, the number of outliers caused by false positives is
+significantly with the second stage processing.](Figures/llvAndDice.png)
+
+\clearpage
+
+# References

### 2016-02-01 stitched.tex
ENH:  James edits.

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
    \setmainfont{Georgia}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{booktabs}
\usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
\linespread{1.5}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\setlist{nolistsep}
\titlespacing{\section}{2pt}{*0}{*0}
\titlespacing{\subsection}{2pt}{*0}{*0}
\titlespacing{\subsubsection}{2pt}{*0}{*0}
\setlength{\parskip}{3pt}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle

\pagenumbering{gobble}

\section{Introduction}\label{introduction}

\subsection{White matter hyperintensities in
TBI}\label{white-matter-hyperintensities-in-tbi}

White matter hyperintensities (WMHs) are foci of abnormally increased
signal intensity seen within white matter regions within the cerebrum
and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic
resonance imaging (MRI) sequences. These lesions are distinguished from
prominent perivascular spaces seen on T2-weighted imaging by the lack of
fluid suppression of signal on FLAIR sequences. WMHs in the
periventricular and deep brain regions are associated with normal aging
and neurological conditions including hypertension and stroke. WMHs are
also a frequent finding following traumatic brain injury (TBI) and have
been correlated with functional outcome and injury severity in both
pediatric {[}1, 2{]} and adult {[}3--6{]}. Further, the regional
distribution and volume of WMHs have been shown to possess prognostic
value in the TBI patient {[}2, 6--8{]}. Specifically, lesion volume in
corpus callosum correlates with functional scores in the acute phase
following injury, while lesion volume in frontal lobes correlates with
scores at 1 year following injury {[}6{]}. Further, volume of FLAIR
lesions within the corpus callosum, brainstem, and thalamus in patients
with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores
{[}4{]}. Additionally, the regional distribution of FLAIR lesions within
the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal,
occipital lobes, and insula along with the observation of grasping or
chewing behavior are associated with poor outcome {[}7{]}.

Despite the above findings, outside of multiple sclerosis, WMHs are not
routinely employed as a diagnostic measure in clinical practice. Their
presence within asymptomatic patients or in association with a variety
of conditions, such as stroke, dementia, neuroinflammatory conditions,
and TBI challenge their utility in narrowing a radiological differential
diagnosis. Further, performing a comprehensive manual counting of number
and distribution of lesions in the clinical setting is simply not
practical. Despite the limited inclusion of WMH observations in routine
radiological reports, two large meta-analyses demonstrated an
association between WMHs, cognitive function, increased risk of stroke,
dementia, and death {[}9, 10{]}. As such, the development of automated
methods for the rapid identification and quantification of WMHs within
individual patients may allow for identification of correlative patterns
between WMH number, volume, distribution, and disease state. Further,
the development of such lesion quantification approaches may allow for
the practical inclusion of this type of information within routine
radiological practice.

\subsection{Random forests for WMH
segmentation}\label{random-forests-for-wmh-segmentation}

Machine learning and pattern recognition techniques have seen increased
application for various medical image analysis workflows (see, for
example, the annual Workshop on Machine Learning in Medical Imaging held
in conjunction with the Medical Image Computing and Computer-Aided
Intervention (MICCAI) international meeting {[}11{]}). Popular
techniques such as support vector machines and neural networks have been
applied successfully to clinically relevant imaging tasks such as
supervised image segmentation (e.g., {[}12{]}) and diagnostic prediction
(e.g., {[}13, 14{]}). Facilitating the current employment of such
techniques are the number of available imaging data sets {[}15{]} and
the public availability of data science packages such as SciPy {[}16{]}
and the R project for statistical computing {[}17{]} and their
associated extensions.

The random forests framework {[}18{]} is a popular machine learning
technique that has demonstrated significant utitility for supervised
segmentation tasks (e.g., normal human brain segmentation {[}19{]}) and
other computer vision applications (e.g., human gait detection
{[}20{]}). In the context of neuropathology, random forest-based
paradigms have been employed in the delineation of multiple sclerosis
lesions {[}21{]}, stroke lesions {[}22{]}, and brain tumors
{[}23--26{]}. Of note, these latter random forest approaches for brain
tumor segmentation have performed well in recent international
competitions. In response to the lack of objective comparisons between
segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS)
challenge was initiated in 2012 {[}27{]} and has continued every year
since under the auspices of the MICCAI conference.

Random forests are conceptually simple {[}18{]}. They consist of
ensembles of decision trees that are built from training data. Once
constructed, data to be classified is ``pushed'' through each decision
tree resulting in a single classification ``vote'' per tree. These votes
are then used for regression or classification of the data. Although
decision trees had been extensively studied previously, the success of
employing collections of such weak learners for boosting machine
learning performance (e.g., AdaBoost {[}28, 29{]}) influenced the
similarly sytled conglomeration of decision trees into ``forests'' with
randomized node optimization {[}30, 31{]}. Finally, Breiman {[}18{]}
improved accuracy by random sampling of the training data (i.e.,
``bagging'') resulting in the current random forest technique.

In this work, we develop a concatenated random forest framework with a
tailored contextual feature image set (both spatial and intensity-based)
for segmenting white matter hyperintensities in traumatic brain injury
cohorts. Additionally, the entire framework is provided publicly through
the well-known open-source ANTs\footnote{\url{https://github.com/stnava/ANTs}}
and ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
Further motivating the research of this work is the public availability
of the imaging data thus permitting full reproducibility of the results
reported and discussed.

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Imaging}\label{imaging}

\subsection{Quantitative analysis}\label{quantitative-analysis}

Crucial to these supervised segmentation approaches are the creation and
selection of ``features'' as input in conjunction with the ground-truth
for model construction. For the targeted application in this work (i.e.,
WMHs), regression/classification are performed at the voxelwise level.
In other words, each voxel within the region of interest is sent through
the ensemble of decision trees and receives a set of classification
votes from each tree thus permitting a regression or classification
solution. Since this procedure is performed at the voxelwise level,
intensity information alone is insufficient for good segmentation
performance due to the lack of spatial context. For example, as pointed
out in {[}32{]}, higher intensities can be found at the periventricular
caps in normal subjects which often confounds automated lesion detection
algorithms. Other potential confounds include MR signal inhomogeneity
and noise. Therefore, even though machine learning and pattern
recognition techniques are extremely powerful and have significant
potential, just as crucial to outcome is the creative construction and
deployment of salient feature images which we detail below.

\subsubsection{Feature images for WMH
segmentation}\label{feature-images-for-wmh-segmentation}

\input{featureImagesTable.tex}

Supervised methodologies are uniquely characterized, in part, by the
feature images that are used to identify the regions of interest. In
Table 1, we provide a list and basic categorization of the feature
images used for the initial (i.e., Stage 1---more on the use of multiple
random forest stages below) segmentation of the white matter
hyperintensities. In addition Figure 1 provides a representation of a
set of feature images for a single subject analyzed in this work. Note
that in this work we categorize the brain parenchyma with seven labels:

\begin{itemize}
\tightlist
\item
  cerebrospinal fluid (label 1),
\item
  gray matter (label 2),
\item
  white matter (label 3),
\item
  deep gray matter (label 4),
\item
  brain stem (label 5),
\item
  cerebellum (label 6), and
\item
  white matter hyperintensities (label 7).
\end{itemize}

\begin{figure}[htbp]
\centering
\includegraphics{Figures/featureImages.png}
\caption{Representation of Stage 1 feature images for subject 01C1019.
The FLAIR, T1-, and T2-weighted images are rigidly pre-aligned {[}33{]}
to the space of the T1 image. The three modality images are then
preprocessed (N4 bias correction {[}34{]} and adaptive denoising
{[}35{]}) followed by application of standard ANTs brain extraction and
\(n\)-tissue segmentation protocols using the MMRR symmetric template
and corresponding priors {[}36{]} applied to the T1 image. The feature
images are then generated for voxelwise input to the RF model which
results in the voting maps illustrated on the right. This gives a
probabilistic classification of tissue type. Not shown are the
probability and voting images for the brain stem and cerebellum.}
\end{figure}

As mentioned previously, input for each subject comprises FLAIR, T1-,
and T2-weighted acquisitions. The FLAIR and T2 images are rigidly
registered to the T1 image using the open-source Advanced Normalization
Tools (ANTs) {[}33{]}. The aligned images are then preprocessed using
the denoising algorithm of {[}35{]} followed by N4 bias correction
{[}34{]} which are then normalized to the intensity range \([0,1]\).
Although we could have used an alternative intensity standardization
algorithm (e.g., {[}37{]}), we found that a simple linear rescaling
produced better results.

\begin{figure}[htbp]
\centering
\includegraphics{Figures/montage.png}
\caption{Sample FLAIR acquisition image slices showing both manual (top)
and random forest (RDF) segmentations (bottom) obtained during the
leave-one-out evaluation. Manual segmentations were performed by one of
the authors and provided the ground truth WMH labels for training the
random forest models.}
\end{figure}

The T1 image is then processed via the ANTs brain extraction and normal
tissue segmentation (i.e., the pathology is discounted which is a fairly
safe assumption for the T1 image) protocols described in {[}36{]} in
order to produce a mask for the brain parenchyma and provide
probabilistic estimates of the cerebrospinal fluid (csf), gray matter,
white matter, deep gray matter, brain stem, and cerebellum. These
provide the ground-truth labels for the first six tissue labels given
above. The white matter hyperintensities were manually identified by one
of the authors (J. R. S.) using the ITK-SNAP tool {[}38{]}. Segmentation
is performed using the ANTs Atropos tool {[}39{]} and multi-model
optimal symmetric shape/intensity templates {[}26{]} created from the
public MMRR data set {[}40{]} (cf Figure 3).

\begin{figure}[htbp]
\centering
\includegraphics{Figures/MMRR.png}
\caption{Canonical views of the mutlivariate, bilaterally symmetric
template constructed from the MMRR data set {[}40{]} (only shown are the
FLAIR, T1, and T2 modalities--- the components relevant for this work).
Template construction is detailed in {[}26{]}. These images are
important for specific intensity-based features.}
\end{figure}

To model the intensity information the first set of images simply
includes the preprocessed and normalized intensity FLAIR, T1, and T2
image voxel values. We also calculate a set of neighborhood statistics
(mean, standard deviation, and skewness) feature images using a radius
of one voxel. For each of the normalized images, we calculate the
difference in intensities with the corresponding warped template
component. Previous success in the international brain tumor
segmentation competition {[}27{]} was based on an important set of
intensity features that were created from multi-modal templates
mentioned previously {[}26{]}. We employ the same strategy here. For
example, the template difference feature image for the FLAIR image,
\(S_{FLAIR}\) is calculated as:

\[S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)\]

where \(\phi_b: S \leftrightarrow \underset{b}{\leftrightsquigarrow} T\)
is the transform which maps from the individual subject space to the
template space and \(T_{FLAIR}\) is the FLAIR template component. Also,
to take advantage of the bilateral symmetry of the normal brain (in
terms of both shape and intensity), and the fact that the presence of
WMHs violates that assumption, we use the symmetric templates to compute
the contralateral intensity differences as an additional intensity
feature. For the FLAIR component, this contralateral difference image is
calculated from

\[S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)\]

where \(\phi_R\) denotes a horizontal reflection perpendicular to the
mid-sagittal plane of the symmetric template.

The segmentation probability images described above are used as feature
images to provide a spatial context for the random forest model
prediction step. Additional spatial contextual feature images include
the distance maps {[}41{]} based on the csf, gray matter, and deep gray
matter images. These latter images are intended to help distinguish
white matter hyperintensities from false positives induced by the
partial voluming at the gray/white matter interface. A third set of
images are based on the voxel location within the space of the template.
The T1 image of the subject is registered to the T1 template component
using a B-spline variant {[}42{]} of the well-known ANTs Symmetric
Normalization (SyN) algorithm {[}43{]}. Since the inverse transform is
also derived as part of the registration process, we can warp the voxel
index locations back to the space of the individual subject. Note that
this is similar in motivation to the work of {[}44{]}. However, this
previous work lacks the normalization to the standard coordinate system
provided by the template to dramatically improve spatial specificity
across all subjects.

\subsubsection{Stacked (concatenated) random forests for improved
segmentation
performance}\label{stacked-concatenated-random-forests-for-improved-segmentation-performance}

In previous brain tumor segmentation work {[}26{]}, it was demonstrated
that a concatenated supervised approach, whereby the prediction output
from the first random forest model serves as partial input for a second
random forest model, can significantly improve segmentation performance.
We do the same thing for the work described here. The Stage 1 feature
images of the training data (as described previously) are used to
construct the Stage 1 model. The training data Stage 1 features are then
used to produce the voxelwise voting maps via the Stage 1 model. All the
Stage 1 features plus the Stage 1 voting maps are used as input to the
Stage 2 model. In addition, we use the Stage 1 voting maps as tissue
priors for a second application of the Atropos maximum aposteriori
algorithm with an additional Markov Random Field spatial prior (MAP-MRF)
{[}39{]}. However, for the second stage we use all three aligned
preprocessed images for a multivariate segmentation. The resulting seven
posterior probability images constitute a third additional feature image
set for Stage 2.

\subsubsection{Code and data
availability}\label{code-and-data-availability}

As pointed out in a recent comprehensive multiple sclerosis lesion
segmentation review {[}45{]}, although the number of algorithms reported
in the literature is quite extensive, there were only four publicly
available segmentation algorithms at the time of writing of which none
are based on supervised learning. As we did for our brain tumor
segmentation algorithm {[}26{]}, all of the code described in this work
is publicly available through the open-source ANTs/ANTsR toolkits.
Through ANTsR (an add-on toolkit which, in part, bridges ANTs and the R
statistical project) we use the \texttt{randomForest} package {[}46{]}
using the default settings with 2000 trees per model and 500 randomly
selected samples per label per image.

In addition, similar to our previous offering,\footnote{\url{https://github.com/ntustison/ANTsAndArboles}}
we plan on creating a self-encapsulated example to showcase the proposed
methodology. The fact that the data will also be made available through
the FITBIR repository along with the manual labelings will facilitate
reproducibility on the part of the reader as well as any interest in
extending the proposed framework to other data sets.

\subsubsection{Evaluation protocol
overview}\label{evaluation-protocol-overview}

In order to evaluate the protocol described, we performed a
leave-one-out evaluation using the data acquired from the 24 subjects
described above. Initial processing included the creation of all Stage 1
feature images for all subjects. The initial brain segmentation of each
T1 image and the manual white matter hyperintensity tracings were
combined to provide the truth labels for the training data. The truth
labels are the seven anatomical regions given above.

The leave-one-out procedure is as follows:

\begin{itemize}
\tightlist
\item
  Create Stage 1 feature images for all 24 subjects.
\item
  For each of the 24 subjects:

  \begin{itemize}
  \tightlist
  \item
    sequester the current subject and corresponding feature images.
  \item
    construct the Stage 1 random forest model from the remaining 23
    subjects.
  \item
    apply the Stage 1 random forest model to the feature images of the
    23 training subjects.
  \item
    the previous step produces the Stage 1 voting maps for all seven
    labels.
  \item
    for each of the 23 subjects, perform a Bayesian-based segmentation
    with an MRF spatial prior using the seven voting maps are used as
    additional tissue priors.
  \item
    construct the Stage 2 random forest model from all the Stage 1
    feature images, seven voting maps, and seven posterior probability
    maps from the previous step.
  \item
    send the sequestered subject through the random forest models for
    both stages.
  \item
    compare the final results with the manually-defined white matter
    hyperintensity regions.
  \end{itemize}
\end{itemize}

\section{Results}\label{results}

\subsection{Ranking feature
importance}\label{ranking-feature-importance}

After performing the leave-one-out evaluation described at the end of
the previous section, we calculated the \texttt{MeanDecreaseAccuracy}
feature values for each of the 24 subjects \(\times\) 2 models per
subject \(=48\) total models. This measure (per feature, per model) is
calculated during the out-of-bag phase of the random forest model
construction and quantifies the decrease in prediction accuracy from
omitting the specified feature. In other words, this quantity helps
determine the importance of a particular feature and, although we save
such efforts for future work, this information provides us with guidance
for future feature pruning and/or additions.

\begin{figure}[htbp]
\centering
\includegraphics{Figures/averageLeaveOneOut.png}
\caption{Average \texttt{MeanDecreaseAccuracy} plots generated from the
creation of all 24 random forest models for both Stage 1 and Stage 2
during the leave-one-out evaluation. These plots are useful in providing
a quantitative assessment of the predictive importance of each feature.
The horizontal error bars provide the \(95^{th}\) percentile (i.e.,
\(1.96 \times \sigma\)) and illustrate the stability of the feature
importance across the leave-one-out models.}
\end{figure}

The resulting rankings for both Stages are given in Figure 4 where the
values for the separate stages are averaged over the entire
corresponding model set. In addition, we track the variance for each
feature over all models to illustrate the stability of the chosen
features during the evaluation. This latter information is illustrated
as horizontal errors bars providing the \(95^{th}\) percentile (i.e.,
\(1.96 \times \sigma\)). Note that the reader can cross reference Table
1 for identifying corresponding feature types and names.

One can also use these measurements as a type of sanity check. For
example, from the Stage 1 plot, one can see that the
\texttt{MeanDecreaseAccuracy} values for the location indices in the
anterior-posterior direction (i.e., \texttt{TemplateIndicesWarped1}) are
greater than those for either the inferior-superior (i.e.,
\texttt{TemplateIndicesWarped0}) or the left-right (i.e.,
\texttt{TemplateIndicesWarped0}) directions in the space of the
symmetric template. This is intuitive since, as discussed previously,
manifestation of TBI white matter hyperintensities can often be confused
with higher intensities at the periventricular caps in normal subjects
{[}32{]} whereas there does not seem to be contralateral bias in
manifestation of white matter hyperintensities in TBI.

Additionally, it is interesting to note some of the other top performing
features for Stage 1. The contralateral difference FLAIR image is highly
discriminative over the set of evaluation random forest models. This
accords with the known clinical relevance of FLAIR images for
identifying white matter hyperintensities and the fact that such
pathology does not manifest symmetrically in both hemispheres.
Interestingly, the posterior maps for the deep gray matter are extremely
important for accurate white matter hyperintensity segmentation. Perhaps
the spatial specification of deep gray matter aids in the removal of
false positives. Inspection of the bottom of the plots demonstrates the
lack of discriminating features associated with the T1 image which is
also well-known in the clinical literature.

As described earlier, for Stage 2, we used the output random forest
voting maps from Stage 1 as both features themselves and as priors for
input to a Bayesian-based segmentation with an additional MRF spatial
prior. In Figure 4, the voting maps are labeled as
``\texttt{RFStage1VotingMaps}'' where the final numeral is associated
with the brain parenchymal labeling given previously. Similarly, the
additional RF prior segmentation feature probability maps are labeled as
``\texttt{RFBrainSegmentationPosteriors}''. The Stage 2 feature
importance plot follows similar trends as that for Stage 1 with the T1
images not contributing much to the identification of white matter
hyperintensity voxels. The initial voting maps from Stage 1 are
extremely important with the top 3 being the estimated locations of the
1) gray matter, 2) white matter, and 3) white matter hyperintensities.
Since these tissue type can be conflated based on intensity alone it is
intuitive that such features would be important.

\subsection{White matter hyperintensity segmentation
evaluation}\label{white-matter-hyperintensity-segmentation-evaluation}

In Figure 5 are the segmentation comparisons derived from manual
segmentations of the same data. Despite the large variability
characteristic with manual labelings in related fields {[}45, 47, 48{]},
such labelings are characteristic of current clinical practices and the
methodology proposed herein is readily adapted to refinements in
training data. On the left of Figure 5 are the improvement in Dice
values over all white matter hyperintensities when comparing the
segmentations between the two stages. Performing the second round of
supervised learning improves Dice values. One can also note from the
right side of Figure 5 that the total lesion load volume illustrates a
few subjects that are severe outliers in terms of the number of false
positives. The second round helps to correct this issue.

\begin{figure}[htbp]
\centering
\includegraphics{Figures/llvAndDice.png}
\caption{Comparison with manual delineation of white matter
hyperintensities. On the left are the calculate Dice values over all
white matter hyperintensities. Note the improvement in the Dice metric
from the employment of the Stage 2 component of the processing pipeline.
(Right) Similar results can be seen by comparing the total lesion load
volume between manual and automated detection strategies. Although some
outliers are found after the Stage 2 processing in a couple subjects,
the number of outliers caused by false positives is significantly with
the second stage processing.}
\end{figure}

\clearpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Bigler:2013aa}{}
1. Bigler, E. D., Abildskov, T. J., Petrie, J., Farrer, T. J., Dennis,
M., Simic, N., Taylor, H. G., Rubin, K. H., Vannatta, K., Gerhardt, C.
A., Stancin, T., and Owen Yeates, K. ``\textbf{Heterogeneity of Brain
Lesions in Pediatric Traumatic Brain Injury}'' \emph{Neuropsychology}
27, no. 4 (2013): 438--51.
doi:\href{https://doi.org/10.1037/a0032837}{10.1037/a0032837}

\hypertarget{ref-Smitherman:2016aa}{}
2. Smitherman, E., Hernandez, A., Stavinoha, P. L., Huang, R., Kernie,
S. G., Diaz-Arrastia, R., and Miles, D. K. ``\textbf{Predicting Outcome
After Pediatric Traumatic Brain Injury by Early Magnetic Resonance
Imaging Lesion Location and Volume}'' \emph{J Neurotrauma} 33, no. 1
(2016): 35--48.
doi:\href{https://doi.org/10.1089/neu.2014.3801}{10.1089/neu.2014.3801}

\hypertarget{ref-Marquez-de-la-Plata:2007aa}{}
3. Marquez de la Plata, C., Ardelean, A., Koovakkattu, D., Srinivasan,
P., Miller, A., Phuong, V., Harper, C., Moore, C., Whittemore, A.,
Madden, C., Diaz-Arrastia, R., and Devous, M., Sr. ``\textbf{Magnetic
Resonance Imaging of Diffuse Axonal Injury: Quantitative Assessment of
White Matter Lesion Volume}'' \emph{J Neurotrauma} 24, no. 4 (2007):
591--8.
doi:\href{https://doi.org/10.1089/neu.2006.0214}{10.1089/neu.2006.0214}

\hypertarget{ref-Moen:2014aa}{}
4. Moen, K. G., Brezova, V., Skandsen, T., Håberg, A. K., Folvik, M.,
and Vik, A. ``\textbf{Traumatic Axonal Injury: The Prognostic Value of
Lesion Load in Corpus Callosum, Brain Stem, and Thalamus in Different
Magnetic Resonance Imaging Sequences}'' \emph{J Neurotrauma} 31, no. 17
(2014): 1486--96.
doi:\href{https://doi.org/10.1089/neu.2013.3258}{10.1089/neu.2013.3258}

\hypertarget{ref-Ding:2008aa}{}
5. Ding, K., Marquez de la Plata, C., Wang, J. Y., Mumphrey, M., Moore,
C., Harper, C., Madden, C. J., McColl, R., Whittemore, A., Devous, M.
D., and Diaz-Arrastia, R. ``\textbf{Cerebral Atrophy After Traumatic
White Matter Injury: Correlation with Acute Neuroimaging and Outcome}''
\emph{J Neurotrauma} 25, no. 12 (2008): 1433--40.
doi:\href{https://doi.org/10.1089/neu.2008.0683}{10.1089/neu.2008.0683}

\hypertarget{ref-Pierallini:2000aa}{}
6. Pierallini, A., Pantano, P., Fantozzi, L. M., Bonamini, M., Vichi,
R., Zylberman, R., Pisarri, F., Colonnese, C., and Bozzao, L.
``\textbf{Correlation Between MRI Findings and Long-Term Outcome in
Patients with Severe Brain Trauma}'' \emph{Neuroradiology} 42, no. 12
(2000): 860--7.

\hypertarget{ref-Weiss:2008aa}{}
7. Weiss, N., Galanaud, D., Carpentier, A., Tezenas de Montcel, S.,
Naccache, L., Coriat, P., and Puybasset, L. ``\textbf{A Combined
Clinical and MRI Approach for Outcome Assessment of Traumatic Head
Injured Comatose Patients}'' \emph{J Neurol} 255, no. 2 (2008): 217--23.
doi:\href{https://doi.org/10.1007/s00415-008-0658-4}{10.1007/s00415-008-0658-4}

\hypertarget{ref-Levin:1988aa}{}
8. Levin, H. S., Williams, D., Crofford, M. J., High, W. M., Jr,
Eisenberg, H. M., Amparo, E. G., Guinto, F. C., Jr, Kalisky, Z., Handel,
S. F., and Goldman, A. M. ``\textbf{Relationship of Depth of Brain
Lesions to Consciousness and Outcome After Closed Head Injury}'' \emph{J
Neurosurg} 69, no. 6 (1988): 861--6.
doi:\href{https://doi.org/10.3171/jns.1988.69.6.0861}{10.3171/jns.1988.69.6.0861}

\hypertarget{ref-Debette:2010aa}{}
9. Debette, S. and Markus, H. S. ``\textbf{The Clinical Importance of
White Matter Hyperintensities on Brain Magnetic Resonance Imaging:
Systematic Review and Meta-Analysis}'' \emph{BMJ} 341, (2010): c3666.

\hypertarget{ref-Kloppenborg:2014aa}{}
10. Kloppenborg, R. P., Nederkoorn, P. J., Geerlings, M. I., and Berg,
E. van den. ``\textbf{Presence and Progression of White Matter
Hyperintensities and Cognition: A Meta-Analysis}'' \emph{Neurology} 82,
no. 23 (2014): 2127--38.
doi:\href{https://doi.org/10.1212/WNL.0000000000000505}{10.1212/WNL.0000000000000505}

\hypertarget{ref-mlmi2015}{}
11. Available at \url{http://mlmi2015.web.unc.edu}

\hypertarget{ref-Bauer:2011aa}{}
12. Bauer, S., Nolte, L.-P., and Reyes, M. ``\textbf{Fully Automatic
Segmentation of Brain Tumor Images Using Support Vector Machine
Classification in Combination with Hierarchical Conditional Random Field
Regularization}'' \emph{Med Image Comput Comput Assist Interv} 14, no.
Pt 3 (2011): 354--61.

\hypertarget{ref-Tong:2014aa}{}
13. Tong, T., Wolz, R., Gao, Q., Guerrero, R., Hajnal, J. V., Rueckert,
D., and Alzheimer's Disease Neuroimaging Initiative. ``\textbf{Multiple
Instance Learning for Classification of Dementia in Brain MRI}''
\emph{Med Image Anal} 18, no. 5 (2014): 808--18.
doi:\href{https://doi.org/10.1016/j.media.2014.04.006}{10.1016/j.media.2014.04.006}

\hypertarget{ref-Liu:2013aa}{}
14. Liu, X., Tosun, D., Weiner, M. W., Schuff, N., and Alzheimer's
Disease Neuroimaging Initiative. ``\textbf{Locally Linear Embedding
(LLE) for MRI Based Alzheimer's Disease Classification}''
\emph{Neuroimage} 83, (2013): 148--57.
doi:\href{https://doi.org/10.1016/j.neuroimage.2013.06.033}{10.1016/j.neuroimage.2013.06.033}

\hypertarget{ref-Van-Horn:2014aa}{}
15. Van Horn, J. D. and Toga, A. W. ``\textbf{Human Neuroimaging as a
`Big Data' Science}'' \emph{Brain Imaging Behav} 8, no. 2 (2014):
323--31.
doi:\href{https://doi.org/10.1007/s11682-013-9255-y}{10.1007/s11682-013-9255-y}

\hypertarget{ref-scipy}{}
16. Jones, E., Oliphant, T., Peterson, P., and others. ``\textbf{SciPy:
Open Source Scientific Tools for Python}'' (2001--2001-\/-): Available
at \url{http://www.scipy.org/}

\hypertarget{ref-R}{}
17. R Core Team. ``\textbf{R: A Language and Environment for Statistical
Computing}'' (2016):

\hypertarget{ref-breiman2001}{}
18. Breiman, L. ``\textbf{Random Forests}'' \emph{Machine learning}
(2001): 5--32.

\hypertarget{ref-yi2009}{}
19. Yi, Z., Criminisi, A., Shotton, J., and Blake, A.
``\textbf{Discriminative, Semantic Segmentation of Brain Tissue in MR
Images}'' \emph{Med Image Comput Comput Assist Interv} 12, no. Pt 2
(2009): 558--65.

\hypertarget{ref-viola2005}{}
20. Viola, P., Jones, M., and Snow, D. ``\textbf{Detecting Pedestrians
Using Patterns of Motion and Appearance}'' \emph{International Journal
of Computer Vision} 63, (2005): 153--161.

\hypertarget{ref-geremia2011}{}
21. Geremia, E., Clatz, O., Menze, B. H., Konukoglu, E., Criminisi, A.,
and Ayache, N. ``\textbf{Spatial Decision Forests for MS Lesion
Segmentation in Multi-Channel Magnetic Resonance Images}''
\emph{Neuroimage} 57, no. 2 (2011): 378--90.
doi:\href{https://doi.org/10.1016/j.neuroimage.2011.03.080}{10.1016/j.neuroimage.2011.03.080}

\hypertarget{ref-Pustina:2016aa}{}
22. Pustina, D., Coslett, H. B., Turkeltaub, P. E., Tustison, N.,
Schwartz, M. F., and Avants, B. ``\textbf{Automated Segmentation of
Chronic Stroke Lesions Using LINDA: Lesion Identification with
Neighborhood Data Analysis}'' \emph{Hum Brain Mapp} (2016):
doi:\href{https://doi.org/10.1002/hbm.23110}{10.1002/hbm.23110}

\hypertarget{ref-geremia2012}{}
23. Geremia, E., Menze, B. H., and Ayache, N. ``\textbf{Spatial Decision
Forests for Glioma Segmentation in Multi-Channel MR Images}''
\emph{Proceedings of MICCAI-BRATS 2012} (2012):

\hypertarget{ref-bauer2012}{}
24. Bauer, S., Fejes, T., Slotboom, J., Wiest, R., Nolte, L.-P., and
Reyes, M. ``\textbf{Segmentation of Brain Tumor Images Based on
Integrated Hierarchical Classification and Regularization}''
\emph{Proceedings of MICCAI-BRATS 2012} (2012): 10--13.

\hypertarget{ref-zikic2012}{}
25. Zikic, D., Glocker, B., Konukoglu, E., Shotton, J., Criminisi, A.,
Ye, D. H., Demiralp, C., Thomas, O. M., Das, T., Jena, R., and Price, S.
J. ``\textbf{Context-Sensitive Classification Forests for Segmentation
of Brain Tumor Tissues}'' \emph{Proceedings of MICCAI-BRATS 2012}
(2012): 1--9.

\hypertarget{ref-Tustison:2015aa}{}
26. Tustison, N. J., Shrinidhi, K. L., Wintermark, M., Durst, C. R.,
Kandel, B. M., Gee, J. C., Grossman, M. C., and Avants, B. B.
``\textbf{Optimal Symmetric Multimodal Templates and Concatenated Random
Forests for Supervised Brain Tumor Segmentation (Simplified) with
ANTsR}'' \emph{Neuroinformatics} 13, no. 2 (2015): 209--25.
doi:\href{https://doi.org/10.1007/s12021-014-9245-2}{10.1007/s12021-014-9245-2}

\hypertarget{ref-Menze:2015aa}{}
27. Menze, B. H., Jakab, A., Bauer, S., Kalpathy-Cramer, J., Farahani,
K., Kirby, J., Burren, Y., Porz, N., Slotboom, J., Wiest, R., Lanczi,
L., Gerstner, E., Weber, M.-A., Arbel, T., Avants, B. B., Ayache, N.,
Buendia, P., Collins, D. L., Cordier, N., Corso, J. J., Criminisi, A.,
Das, T., Delingette, H., Demiralp, Ç., Durst, C. R., Dojat, M., Doyle,
S., Festa, J., Forbes, F., Geremia, E., Glocker, B., Golland, P., Guo,
X., Hamamci, A., Iftekharuddin, K. M., Jena, R., John, N. M., Konukoglu,
E., Lashkari, D., Mariz, J. A., Meier, R., Pereira, S., Precup, D.,
Price, S. J., Raviv, T. R., Reza, S. M. S., Ryan, M., Sarikaya, D.,
Schwartz, L., Shin, H.-C., Shotton, J., Silva, C. A., Sousa, N.,
Subbanna, N. K., Szekely, G., Taylor, T. J., Thomas, O. M., Tustison, N.
J., Unal, G., Vasseur, F., Wintermark, M., Ye, D. H., Zhao, L., Zhao,
B., Zikic, D., Prastawa, M., Reyes, M., and Van Leemput, K.
``\textbf{The Multimodal Brain Tumor Image Segmentation Benchmark
(BRATS)}'' \emph{IEEE Trans Med Imaging} 34, no. 10 (2015): 1993--2024.
doi:\href{https://doi.org/10.1109/TMI.2014.2377694}{10.1109/TMI.2014.2377694}

\hypertarget{ref-schapire1990}{}
28. Schapire, R. ``\textbf{The Strength of Weak Learnability}''
\emph{Machine Learning} 5, (1990): 197--227.

\hypertarget{ref-freund1997}{}
29. Freund, Y. and Schapire, R. ``\textbf{A Decision-Theoretic
Generalization of on-Line Learning and an Application to Boosting}''
\emph{Journal of Computer and System Sciences} 55, (1997): 119--139.

\hypertarget{ref-ho1995}{}
30. Ho, T. K. ``\textbf{Random Decision Forests}'' \emph{Document
analysis and recognition, 1995., proceedings of the third international
conference on} 1, (1995): 278--282 vol.1.
doi:\href{https://doi.org/10.1109/ICDAR.1995.598994}{10.1109/ICDAR.1995.598994}

\hypertarget{ref-amit1997}{}
31. Amit, Y. and Geman, D. ``\textbf{Shape Quantization and Recognition
with Randomized Trees}'' \emph{Neural Computation} 9, (1997):
1545--1588.

\hypertarget{ref-Neema:2009aa}{}
32. Neema, M., Guss, Z. D., Stankiewicz, J. M., Arora, A., Healy, B. C.,
and Bakshi, R. ``\textbf{Normal Findings on Brain Fluid-Attenuated
Inversion Recovery MR Images at 3T}'' \emph{AJNR Am J Neuroradiol} 30,
no. 5 (2009): 911--6.
doi:\href{https://doi.org/10.3174/ajnr.A1514}{10.3174/ajnr.A1514}

\hypertarget{ref-Avants:2014aa}{}
33. Avants, B. B., Tustison, N. J., Stauffer, M., Song, G., Wu, B., and
Gee, J. C. ``\textbf{The Insight ToolKit Image Registration Framework}''
\emph{Front Neuroinform} 8, (2014): 44.
doi:\href{https://doi.org/10.3389/fninf.2014.00044}{10.3389/fninf.2014.00044}

\hypertarget{ref-Tustison:2010ac}{}
34. Tustison, N. J., Avants, B. B., Cook, P. A., Zheng, Y., Egan, A.,
Yushkevich, P. A., and Gee, J. C. ``\textbf{N4ITK: Improved N3 Bias
Correction}'' \emph{IEEE Trans Med Imaging} 29, no. 6 (2010): 1310--20.
doi:\href{https://doi.org/10.1109/TMI.2010.2046908}{10.1109/TMI.2010.2046908}

\hypertarget{ref-Manjon:2010aa}{}
35. Manjón, J. V., Coupé, P., Martí-Bonmatí, L., Collins, D. L., and
Robles, M. ``\textbf{Adaptive Non-Local Means Denoising of MR Images
with Spatially Varying Noise Levels}'' \emph{J Magn Reson Imaging} 31,
no. 1 (2010): 192--203.
doi:\href{https://doi.org/10.1002/jmri.22003}{10.1002/jmri.22003}

\hypertarget{ref-Tustison:2014ab}{}
36. Tustison, N. J., Cook, P. A., Klein, A., Song, G., Das, S. R., Duda,
J. T., Kandel, B. M., Strien, N. van, Stone, J. R., Gee, J. C., and
Avants, B. B. ``\textbf{Large-Scale Evaluation of ANTs and FreeSurfer
Cortical Thickness Measurements}'' \emph{Neuroimage} 99, (2014):
166--79.
doi:\href{https://doi.org/10.1016/j.neuroimage.2014.05.044}{10.1016/j.neuroimage.2014.05.044}

\hypertarget{ref-nyul2000}{}
37. Nyúl, L. G., Udupa, J. K., and Zhang, X. ``\textbf{New Variants of a
Method of MRI Scale Standardization}'' \emph{IEEE Trans Med Imaging} 19,
no. 2 (2000): 143--50.
doi:\href{https://doi.org/10.1109/42.836373}{10.1109/42.836373}

\hypertarget{ref-Yushkevich:2006aa}{}
38. Yushkevich, P. A., Piven, J., Hazlett, H. C., Smith, R. G., Ho, S.,
Gee, J. C., and Gerig, G. ``\textbf{User-Guided 3D Active Contour
Segmentation of Anatomical Structures: Significantly Improved Efficiency
and Reliability}'' \emph{Neuroimage} 31, no. 3 (2006): 1116--28.
doi:\href{https://doi.org/10.1016/j.neuroimage.2006.01.015}{10.1016/j.neuroimage.2006.01.015}

\hypertarget{ref-Avants:2011aa}{}
39. Avants, B. B., Tustison, N. J., Wu, J., Cook, P. A., and Gee, J. C.
``\textbf{An Open Source Multivariate Framework for \(n\)-Tissue
Segmentation with Evaluation on Public Data}'' \emph{Neuroinformatics}
9, no. 4 (2011): 381--400.
doi:\href{https://doi.org/10.1007/s12021-011-9109-y}{10.1007/s12021-011-9109-y}

\hypertarget{ref-landman2011}{}
40. Landman, B. A., Huang, A. J., Gifford, A., Vikram, D. S., Lim, I. A.
L., Farrell, J. A. D., Bogovic, J. A., Hua, J., Chen, M., Jarso, S.,
Smith, S. A., Joel, S., Mori, S., Pekar, J. J., Barker, P. B., Prince,
J. L., and Zijl, P. C. M. van. ``\textbf{Multi-Parametric Neuroimaging
Reproducibility: A 3-T Resource Study}'' \emph{Neuroimage} 54, no. 4
(2011): 2854--66.
doi:\href{https://doi.org/10.1016/j.neuroimage.2010.11.047}{10.1016/j.neuroimage.2010.11.047}

\hypertarget{ref-maurer2003}{}
41. Maurer, C. R., Rensheng, Q., and Raghavan, V. ``\textbf{A Linear
Time Algorithm for Computing Exact Euclidean Distance Transforms of
Binary Images in Arbitrary Dimensions}'' \emph{Pattern Analysis and
Machine Intelligence, IEEE Transactions on} 25, no. 2 (2003): 265--270.
doi:\href{https://doi.org/10.1109/TPAMI.2003.1177156}{10.1109/TPAMI.2003.1177156}

\hypertarget{ref-Tustison:2013ac}{}
42. Tustison, N. J. and Avants, B. B. ``\textbf{Explicit B-Spline
Regularization in Diffeomorphic Image Registration}'' \emph{Front
Neuroinform} 7, (2013): 39.
doi:\href{https://doi.org/10.3389/fninf.2013.00039}{10.3389/fninf.2013.00039}

\hypertarget{ref-Avants:2011ab}{}
43. Avants, B. B., Tustison, N. J., Song, G., Cook, P. A., Klein, A.,
and Gee, J. C. ``\textbf{A Reproducible Evaluation of ANTs Similarity
Metric Performance in Brain Image Registration}'' \emph{Neuroimage} 54,
no. 3 (2011): 2033--44.
doi:\href{https://doi.org/10.1016/j.neuroimage.2010.09.025}{10.1016/j.neuroimage.2010.09.025}

\hypertarget{ref-Anbeek:2004aa}{}
44. Anbeek, P., Vincken, K. L., Osch, M. J. P. van, Bisschops, R. H. C.,
and Grond, J. van der. ``\textbf{Probabilistic Segmentation of White
Matter Lesions in MR Imaging}'' \emph{Neuroimage} 21, no. 3 (2004):
1037--44.
doi:\href{https://doi.org/10.1016/j.neuroimage.2003.10.012}{10.1016/j.neuroimage.2003.10.012}

\hypertarget{ref-Garcia-Lorenzo:2013aa}{}
45. García-Lorenzo, D., Francis, S., Narayanan, S., Arnold, D. L., and
Collins, D. L. ``\textbf{Review of Automatic Segmentation Methods of
Multiple Sclerosis White Matter Lesions on Conventional Magnetic
Resonance Imaging}'' \emph{Med Image Anal} 17, no. 1 (2013): 1--18.
doi:\href{https://doi.org/10.1016/j.media.2012.09.004}{10.1016/j.media.2012.09.004}

\hypertarget{ref-liaw2002}{}
46. Liaw, A. and Wiener, M. ``\textbf{Classification and Regression by
randomForest}'' \emph{R News} 2/3, (2002): 18--22.

\hypertarget{ref-Grimaud:1996aa}{}
47. Grimaud, J., Lai, M., Thorpe, J., Adeleine, P., Wang, L., Barker, G.
J., Plummer, D. L., Tofts, P. S., McDonald, W. I., and Miller, D. H.
``\textbf{Quantification of MRI Lesion Load in Multiple Sclerosis: A
Comparison of Three Computer-Assisted Techniques}'' \emph{Magn Reson
Imaging} 14, no. 5 (1996): 495--505.

\hypertarget{ref-styner2008}{}
48. ``\textbf{Special Issue on 2008 MICCAI Workshop - MS Lesion
Segmentation}'' (2008):

\end{document}

### 2016-02-02 discussion.Rmd
ENH:  More James' edits.

# Discussion

<!--
Nick could you add some detail here on how this technique compares to other automated or semi automated approaches for WMH detection?
-->

<!--

* Evaluation
    * Proper context
        * We only have manual labelings from a single radiologist (
        * TBI white matter hyperintensities are more difficult to segment than MS lesions
            * individual lesions are smaller
            * lesion load is typically smaller
            * not prone to enhancement
            * less distinct from surrounding tissue ?
    * Despite the doubts regarding the efficacy of a typical segmentation validation,
      our contribution is useful because
         * the total segmentation framework is publicly available within the ANTs/ANTsR
           framework
         * although only a single site is analyzed, the feature images are site-agnostic
              * we plan to apply the current RF models directly to the other site data
              * we can then build the training database further by enlisting other experts
         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
         * The data (including manual tracings) will be made available as part of the FITBIR
           effort.

-->

The current communications describes a supervised statistical learning methodology for identifying WHMs within multimodal MR brain imaging. This effort utilized information acquired from the manual segmentation of WMHs from FLAIR images to help build two-stage ensembles of decision trees for the automated identification of these lesions. Although
only a single expert was used to produce the manual labelings, our intent is to further
refine the proposed paradigm by crowdsourcing with feedback from other experts who interact with both the data and methodology.  Also, we recognize that only a single site
was used for evaluating the proposed framework.  However, we are currently processing other site data with the models developed for this work and the results look promising
since the developed features are site-agnostic.

As far as we know, this is the first report utilizing a novel random forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs tend to be more difficult to segment than MS lesions as the former tend to be smaller with an overall smaller lesion load.  Also, enhancement protocols with the former tend to be less successful than with the latter.  As mentioned previously, the work in MS lesion segmentation is extensive
with a handful of techniques being publicly available.  Our framework is also available as open-source as part of well-known neuroimaging tools which easily allows for additions/extensions but is also, as far as we know, the first random forest-based technique available for such application.

Two major meta-analyses of WMHs have been published covering the periods of 1966-2009 and 2010-2015, respectively. The earlier meta-analysis covered 53 longitudinal studies that included samples of high-risk populations, i.e., patients selected for a specific disease or condition such as hypertension, whereas other studies recruited samples of the general population. Longitudinal studies of samples representative of the general population are more relevant to the focus of the present paper. Debette & Markus [@Debette:2010aa] found that the presence of WMHs was related to subsequent cognitive decline, a higher risk of developing dementia, stroke, and of mortality. Lesion volume at baseline was also predictive of cognitive decline. Limitations of this meta-analysis include heterogeneity in the method of measuring WMHs; some studies used automated volumetric measurement, whereas others used a visual rating scale. The studies analyzed by Debette & Markus were limited to the occurrence of one of the aforementioned conditions which they analyzed by hazard ratios.

A recent meta-analysis by Kloppenborg et al. [@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and concurrent neuropsychological results in patients with heterogeneous diagnoses but without previously diagnosed cognitive impairment, found that WMHs were associated with cognitive deficit (effect size of -0.10, 95% CI: -0.13 to -0.08) after controlling for age. These studies also differed in the metric used to measure the WMHs, including volume, % of total intracranial volume, and a visual rating score. The effect size for the association with cognitive deficit in these cross-sectional studies did not differ significantly across various cognitive domains or the method of measuring lesion volume.  Among eight longitudinal studies analyzed by Kloppenborg et al that included a follow-up MRI and also controlled for age, the effect size for the association of progression in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09). This association was stronger for attention and executive function than for memory and processing speed. Although baseline WMHs were predictive of cognitive deficit at follow-up in the seven studies which did not repeat MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in the longitudinal studies that calculated progression in WMHs. In summary, progression of WMHs seen on repeat MRI has a stronger relation to cognitive deficit than concurrent imaging findings. These meta-analyses support the rationale for repeating an MRI in patients younger than 50 years whose initial scan shows WMHs.
Despite the above-described associations between WMHs, cognitive decline, increased risk of developing dementia, and mortality, these lesions receive little attention in current clinical workflows. When reported in a standard neuroradiologist interpretation, they are typically handled as incidental findings and are assigned little clinical significance. This likely reflects the impracticality of performing a detailed assessment of number, volume, and distribution within a qualitative neuroradiologist interpretation as well as the lack of correlative information on how the presence and distribution of these lesions may inform a diagnosis and prognosis in the appropriate clinical setting. To date, automated or semi-automated tools for the detection of WMHs have lacked the specificity and efficiency for the mining of large-scale datasets to generate highly granular data on whether these lesions possess any true diagnostic or prognostic value in the setting of a specific disease process. The present communication describes a supervised statistical learning tool that is appropriate for the application to such large-scale datasets.

The currently described tool is just one example of how “deep learning” algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state. It is an important demonstration of the potential power of these analytical approaches in the rapid but comprehensive mining of information from neuroimaging examinations. Deep learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features. Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data. Similarly, in the context of the neuroimaging assessments, deep learning approaches may allow for the rapid identification of information predictive of disease state in an individual patient. These approaches may be applied to the identification of macroscopically visible lesions as described in the current communication. Additionally, these approaches might be applied to the interrogation of imaging data in the individual patient with a primary quantitative output metrics to include sequences such as diffusion tensor imaging (DTI) and its variants, functional connectivity, perfusion weighted imaging, and cortical thickness assessments. At present, these sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient. Application of deep learning approaches in the context of data with primary quantitative outputs will require large scale normative and disease specific databases. Acquisition of these data is resource intensive and requires a multi-center approach with harmonized scanner between sites and correlative non-imaging clinical data. Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through the Federal Interagency TBI repository (FITBIR). In concert with any available high quality normative neuroimaging data, deep learning algorithms may be well positioned to help transform how neuroimaging is interpreted for the clinical management of patients with this disease process.

\clearpage

# References

### 2016-02-02 results.Rmd
ENH:  More James' edits.

@@
-<!--
-
-* Evaluation
-    * Proper context
-        * We only have manual labelings from a single radiologist (
-        * TBI white matter hyperintensities are more difficult to segment than MS lesions
-            * individual lesions are smaller
-            * lesion load is typically smaller
-            * not prone to enhancement
-            * less distinct from surrounding tissue ?
-    * Despite the doubts regarding the efficacy of a typical segmentation validation,
-      our contribution is useful because
-         * the total segmentation framework is publicly available within the ANTs/ANTsR
-           framework
-         * although only a single site is analyzed, the feature images are site-agnostic
-              * we plan to apply the current RF models directly to the other site data
-              * we can then build the training database further by enlisting other experts
-         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
-         * The data (including manual tracings) will be made available as part of the FITBIR
-           effort.
-
--->
-
 # Results
@@
 significantly with the second stage processing.](Figures/llvAndDice.png)
-
-\clearpage
-
-# References

### 2016-02-02 stitched.Rmd
ENH:  More James' edits.

@@
 
-<!--
-
-* Evaluation
-    * Proper context
-        * We only have manual labelings from a single radiologist (
-        * TBI white matter hyperintensities are more difficult to segment than MS lesions
-            * individual lesions are smaller
-            * lesion load is typically smaller
-            * not prone to enhancement
-            * less distinct from surrounding tissue ?
-    * Despite the doubts regarding the efficacy of a typical segmentation validation,
-      our contribution is useful because
-         * the total segmentation framework is publicly available within the ANTs/ANTsR
-           framework
-         * although only a single site is analyzed, the feature images are site-agnostic
-              * we plan to apply the current RF models directly to the other site data
-              * we can then build the training database further by enlisting other experts
-         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
-         * The data (including manual tracings) will be made available as part of the FITBIR
-           effort.
-
--->
-
 # Results
@@
 
+# Discussion
+
+<!--
+Nick could you add some detail here on how this technique compares to other automated or semi automated approaches for WMH detection?
+-->
+
+<!--
+
+* Evaluation
+    * Proper context
+        * We only have manual labelings from a single radiologist (
+        * TBI white matter hyperintensities are more difficult to segment than MS lesions
+            * individual lesions are smaller
+            * lesion load is typically smaller
+            * not prone to enhancement
+            * less distinct from surrounding tissue ?
+    * Despite the doubts regarding the efficacy of a typical segmentation validation,
+      our contribution is useful because
+         * the total segmentation framework is publicly available within the ANTs/ANTsR
+           framework
+         * although only a single site is analyzed, the feature images are site-agnostic
+              * we plan to apply the current RF models directly to the other site data
+              * we can then build the training database further by enlisting other experts
+         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
+         * The data (including manual tracings) will be made available as part of the FITBIR
+           effort.
+
+-->
+
+The current communications describes a supervised statistical learning methodology for identifying WHMs within multimodal MR brain imaging.
+This effort utilized information acquired from the manual segmentation of WMHs from FLAIR images to help build two-stage ensembles of decision trees for the automated identification of these lesions.
+Although
+only a single expert was used to produce the manual labelings, our intent is to further
+refine the proposed paradigm by crowdsourcing with feedback from other experts who interact with both the data and methodology.  Also, we recognize that only a single site
+was used for evaluating the proposed framework.  However, we are currently processing other site data with the models developed for this work and the results look promising
+since the developed features are site-agnostic.
+
+As far as we know, this is the first report utilizing a novel random forest approach to identify WMHs in a cohort of TBI patients.
+TBI WMHs tend to be more difficult to segment than MS lesions as the former tend to be smaller with an overall smaller lesion load.
+Also, enhancement protocols with the former tend to be less successful than with the latter.
+As mentioned previously, the work in MS lesion segmentation is extensive
+with a handful of techniques being publicly available.
+Our framework is also available as open-source as part of well-known neuroimaging tools which easily allows for additions/extensions but is also, as far as we know, the first random forest-based technique available for such application.
+
+Two major meta-analyses of WMHs have been published covering the periods of 1966-2009 and 2010-2015, respectively.
+The earlier meta-analysis covered 53 longitudinal studies that included samples of high-risk populations, i.e., patients selected for a specific disease or condition such as hypertension, whereas other studies recruited samples of the general population.
+Longitudinal studies of samples representative of the general population are more relevant to the focus of the present paper.
+Debette & Markus [@Debette:2010aa] found that the presence of WMHs was related to subsequent cognitive decline, a higher risk of developing dementia, stroke, and of mortality.
+Lesion volume at baseline was also predictive of cognitive decline.
+Limitations of this meta-analysis include heterogeneity in the method of measuring WMHs; some studies used automated volumetric measurement, whereas others used a visual rating scale.
+The studies analyzed by Debette & Markus were limited to the occurrence of one of the aforementioned conditions which they analyzed by hazard ratios.
+
+A recent meta-analysis by Kloppenborg et al.
+[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and concurrent neuropsychological results in patients with heterogeneous diagnoses but without previously diagnosed cognitive impairment, found that WMHs were associated with cognitive deficit (effect size of -0.10, 95% CI: -0.13 to -0.08) after controlling for age.
+These studies also differed in the metric used to measure the WMHs, including volume, % of total intracranial volume, and a visual rating score.
+The effect size for the association with cognitive deficit in these cross-sectional studies did not differ significantly across various cognitive domains or the method of measuring lesion volume.
+Among eight longitudinal studies analyzed by Kloppenborg et al that included a follow-up MRI and also controlled for age, the effect size for the association of progression in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09).
+This association was stronger for attention and executive function than for memory and processing speed.
+Although baseline WMHs were predictive of cognitive deficit at follow-up in the seven studies which did not repeat MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in the longitudinal studies that calculated progression in WMHs.
+In summary, progression of WMHs seen on repeat MRI has a stronger relation to cognitive deficit than concurrent imaging findings.
+These meta-analyses support the rationale for repeating an MRI in patients younger than 50 years whose initial scan shows WMHs.
+Despite the above-described associations between WMHs, cognitive decline, increased risk of developing dementia, and mortality, these lesions receive little attention in current clinical workflows.
+When reported in a standard neuroradiologist interpretation, they are typically handled as incidental findings and are assigned little clinical significance.
+This likely reflects the impracticality of performing a detailed assessment of number, volume, and distribution within a qualitative neuroradiologist interpretation as well as the lack of correlative information on how the presence and distribution of these lesions may inform a diagnosis and prognosis in the appropriate clinical setting.
+To date, automated or semi-automated tools for the detection of WMHs have lacked the specificity and efficiency for the mining of large-scale datasets to generate highly granular data on whether these lesions possess any true diagnostic or prognostic value in the setting of a specific disease process.
+The present communication describes a supervised statistical learning tool that is appropriate for the application to such large-scale datasets.
+
+The currently described tool is just one example of how “deep learning” algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state.
+It is an important demonstration of the potential power of these analytical approaches in the rapid but comprehensive mining of information from neuroimaging examinations.
+Deep learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features.
+Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data.
+Similarly, in the context of the neuroimaging assessments, deep learning approaches may allow for the rapid identification of information predictive of disease state in an individual patient.
+These approaches may be applied to the identification of macroscopically visible lesions as described in the current communication.
+Additionally, these approaches might be applied to the interrogation of imaging data in the individual patient with a primary quantitative output metrics to include sequences such as diffusion tensor imaging (DTI) and its variants, functional connectivity, perfusion weighted imaging, and cortical thickness assessments.
+At present, these sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient.
+Application of deep learning approaches in the context of data with primary quantitative outputs will require large scale normative and disease specific databases.
+Acquisition of these data is resource intensive and requires a multi-center approach with harmonized scanner between sites and correlative non-imaging clinical data.
+Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through the Federal Interagency TBI repository (FITBIR).
+In concert with any available high quality normative neuroimaging data, deep learning algorithms may be well positioned to help transform how neuroimaging is interpreted for the clinical management of patients with this disease process.
+
 \clearpage

### 2016-02-02 stitched.tex
ENH:  More James' edits.

@@
 
+\section{Discussion}\label{discussion}
+
+The current communications describes a supervised statistical learning
+methodology for identifying WHMs within multimodal MR brain imaging.
+This effort utilized information acquired from the manual segmentation
+of WMHs from FLAIR images to help build two-stage ensembles of decision
+trees for the automated identification of these lesions. Although only a
+single expert was used to produce the manual labelings, our intent is to
+further refine the proposed paradigm by crowdsourcing with feedback from
+other experts who interact with both the data and methodology. Also, we
+recognize that only a single site was used for evaluating the proposed
+framework. However, we are currently processing other site data with the
+models developed for this work and the results look promising since the
+developed features are site-agnostic.
+
+As far as we know, this is the first report utilizing a novel random
+forest approach to identify WMHs in a cohort of TBI patients. TBI WMHs
+tend to be more difficult to segment than MS lesions as the former tend
+to be smaller with an overall smaller lesion load. Also, enhancement
+protocols with the former tend to be less successful than with the
+latter. As mentioned previously, the work in MS lesion segmentation is
+extensive with a handful of techniques being publicly available. Our
+framework is also available as open-source as part of well-known
+neuroimaging tools which easily allows for additions/extensions but is
+also, as far as we know, the first random forest-based technique
+available for such application.
+
+Two major meta-analyses of WMHs have been published covering the periods
+of 1966-2009 and 2010-2015, respectively. The earlier meta-analysis
+covered 53 longitudinal studies that included samples of high-risk
+populations, i.e., patients selected for a specific disease or condition
+such as hypertension, whereas other studies recruited samples of the
+general population. Longitudinal studies of samples representative of
+the general population are more relevant to the focus of the present
+paper. Debette \& Markus {[}9{]} found that the presence of WMHs was
+related to subsequent cognitive decline, a higher risk of developing
+dementia, stroke, and of mortality. Lesion volume at baseline was also
+predictive of cognitive decline. Limitations of this meta-analysis
+include heterogeneity in the method of measuring WMHs; some studies used
+automated volumetric measurement, whereas others used a visual rating
+scale. The studies analyzed by Debette \& Markus were limited to the
+occurrence of one of the aforementioned conditions which they analyzed
+by hazard ratios.
+
+A recent meta-analysis by Kloppenborg et al. {[}10{]} of 23
+cross-sectional studies reporting MRI and concurrent neuropsychological
+results in patients with heterogeneous diagnoses but without previously
+diagnosed cognitive impairment, found that WMHs were associated with
+cognitive deficit (effect size of -0.10, 95\% CI: -0.13 to -0.08) after
+controlling for age. These studies also differed in the metric used to
+measure the WMHs, including volume, \% of total intracranial volume, and
+a visual rating score. The effect size for the association with
+cognitive deficit in these cross-sectional studies did not differ
+significantly across various cognitive domains or the method of
+measuring lesion volume. Among eight longitudinal studies analyzed by
+Kloppenborg et al that included a follow-up MRI and also controlled for
+age, the effect size for the association of progression in WMHs and
+cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
+association was stronger for attention and executive function than for
+memory and processing speed. Although baseline WMHs were predictive of
+cognitive deficit at follow-up in the seven studies which did not repeat
+MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
+in the longitudinal studies that calculated progression in WMHs. In
+summary, progression of WMHs seen on repeat MRI has a stronger relation
+to cognitive deficit than concurrent imaging findings. These
+meta-analyses support the rationale for repeating an MRI in patients
+younger than 50 years whose initial scan shows WMHs. Despite the
+above-described associations between WMHs, cognitive decline, increased
+risk of developing dementia, and mortality, these lesions receive little
+attention in current clinical workflows. When reported in a standard
+neuroradiologist interpretation, they are typically handled as
+incidental findings and are assigned little clinical significance. This
+likely reflects the impracticality of performing a detailed assessment
+of number, volume, and distribution within a qualitative
+neuroradiologist interpretation as well as the lack of correlative
+information on how the presence and distribution of these lesions may
+inform a diagnosis and prognosis in the appropriate clinical setting. To
+date, automated or semi-automated tools for the detection of WMHs have
+lacked the specificity and efficiency for the mining of large-scale
+datasets to generate highly granular data on whether these lesions
+possess any true diagnostic or prognostic value in the setting of a
+specific disease process. The present communication describes a
+supervised statistical learning tool that is appropriate for the
+application to such large-scale datasets.
+
+The currently described tool is just one example of how ``deep
+learning'' algorithms might be applied to aid in the diagnosis of TBI
+and other disease processes through the specific identification of
+features predictive of a given disease state. It is an important
+demonstration of the potential power of these analytical approaches in
+the rapid but comprehensive mining of information from neuroimaging
+examinations. Deep learning algorithms are presently employed across a
+wide variety of settings for the rapid identification of predictive
+imaging features. Automobile manufacturers utilize these types of
+approaches to equip self-driving vehicles to recognize and respond to
+unique external surroundings through the identification of visual
+information sufficiently similar to previously assimilated training
+data. Similarly, in the context of the neuroimaging assessments, deep
+learning approaches may allow for the rapid identification of
+information predictive of disease state in an individual patient. These
+approaches may be applied to the identification of macroscopically
+visible lesions as described in the current communication. Additionally,
+these approaches might be applied to the interrogation of imaging data
+in the individual patient with a primary quantitative output metrics to
+include sequences such as diffusion tensor imaging (DTI) and its
+variants, functional connectivity, perfusion weighted imaging, and
+cortical thickness assessments. At present, these sequences are confined
+to cohort-based research studies due to the lack of available analytical
+tools to assess the information in the setting of the individual
+patient. Application of deep learning approaches in the context of data
+with primary quantitative outputs will require large scale normative and
+disease specific databases. Acquisition of these data is resource
+intensive and requires a multi-center approach with harmonized scanner
+between sites and correlative non-imaging clinical data. Large scale TBI
+data is becoming increasingly available through activities such as the
+Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research
+and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European
+Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of
+Defense Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and
+other data being consolidated through the Federal Interagency TBI
+repository (FITBIR). In concert with any available high quality
+normative neuroimaging data, deep learning algorithms may be well
+positioned to help transform how neuroimaging is interpreted for the
+clinical management of patients with this disease process.
+
 \clearpage

### 2016-02-02 README.R

@@ -9,7 +9,8 @@ formatFile <- "format.Rmd"
 rmdFiles <- c( formatFile,
                "intro.Rmd",
                "methods.Rmd",
-               "results.Rmd"
+               "results.Rmd",
+               "discussion.Rmd"
    )
 
 for( i in 1:length( rmdFiles ) )

### 2016-02-11 format.Rmd
ENH:  Edits.

--- 2016-01-23 format.Rmd
@@
    - \usepackage{booktabs}
+   - \usepackage{changes}
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
    - \setlength{\parskip}{3pt}
+   - \setremarkmarkup{(#2)}
+   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
+   - \definechangesauthor[name={James Stone}, color=magenta]{js}
+   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
+   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
+   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
+   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
 bibliography: references.bib

### 2016-02-11 methods.Rmd
ENH:  Edits.

@@
 
+MR images utilized for this initial report were acquired from 24 individuals scanned at a single site involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (\added[id=nt,remark=NT: reference needed?]{see [Walker], this issue}).
+Briefly, participants were Operation Enduring Freedom/ Operation Iraqi Freedom /Operation New dawn (OEF/OIF/OND) era service members and veterans between the ages of 18--60 who have experienced combat situations and are between X and X (period of time) from deployment.
+\added[id=js,remark={JS: Data needed.}]{The feature images consisted of subjects aged $X.X \pm X$ years (range X – X years), XX percent male, with at least one or more potential concussive events (mean = X, SD = X, range X-X).}
+
+Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
+3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.
+3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9°, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 x 256 mm FOV.
+In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.
+
 ## Quantitative analysis
 
-<!--
-
-* Importance of machine learning in medical image analysis
-    x give examples
-    x (tools) support vector machines
-    x (tools) neural networks
-    x (apps)
-    x Random forests algorithms
-        x MS lesion detection (Geremeria 2011)
-        x Brain tumor segmentation (BRATS 2012 and BRATS 2013)
-        x Tustison 2015 (importance of feature images)
-        x Pustina 2016
-    x Random forests overview
-
-* Overview of protocol
-    * Preprocess all images
-        * Denoise
-        * N4 bias correction
-    * Create Stage 1 feature images
-    * Use those as input to the first stage RF model to get RF voting maps
-        * CSF
-        * GM
-        * WM
-        * Deep GM
-        * Brain stem
-        * Cerebellum
-        * WMHs
-    * Use RF voting maps as priors for antsAtroposN4.sh to perform Bayesian-based
-      segmentation with an additional Markov random field prior to promote spatial
-      smoothness of the labels.
-    * Use Stage 1 feature images + RF voting maps + antsAtroposN4.sh posteriors as
-      Stage 2 feature images.
-    * Remove isolated voxels as final refinement.
-
-* Creation of feature images (all contained within ANTs)
-    * Benefits of open source
-    * best registration (for template building)
-    * Preprocessing
-        * Denoising
-        * N4 bias correction
-    * Types of feature images and motivation for their use (Stage 1)
-        * Normalized intensities and neighborhood statistics image
-        * Symmetric template --- the brain exhibits a shape and intensity bilateral
-          symmetry under normal conditions.  WMHs can "break" this symmetry.
-        * Probabilistic segmentations, distance maps and symmetric template indices ---
-          help remove false positives
-          caused by partial voluming at the gray/white matter interface and the location
-          indices since, for example, higher intensities can be found at the periventricular
-          caps in normal subjects [@Neema:2009aa] which often confounds automated lesion
-          detection algorithms.
-    * Types of feature images and motivation for their use (Stage 2)
-        * All previously images (for the reasons stated above)
-        * ``antsAtroposN4.sh`` with a large prior weight (w=0.5) on all three modalities
-          to classify voxels based on intensity (use MRF to smooth out remaining noise
-          in labeling).
-
-* Public avail
-
-* Leave one out evaluation strategy overview
-    * Create stage 1 feature images for all subjects
-    * For each of the 24 subjects, take out a subject and create a Stage 1 random forest
-      model from the remaining feature images + manually labeled truth data.
-    * For each of the remaining 23 subjects, send them through the Stage 1 RF model.
-    * The above step permits the creation of the Stage 2 feature images
-    * Create the Stage 2 RF model from the 23 sets of Stage 2 feature images.
-    * Send the leave-one-out subject through the two RF models (as described above).
-    * Compare with ground truth.
-
-* (Save for Discusion) Difficulties of WMH evaluation (outlined in greater detail here [@MIAreview].
-    *
--->
-
 Crucial to these supervised segmentation approaches are the creation and selection of
-"features" as input in conjunction with the ground-truth for model construction.  For the
-targeted application in this work (i.e., WMHs),
+"features" as input in conjunction with
+\replaced[id=lw,remark=NT: I replaced ground-truth since it is problematic anyway]
+{expertly identified structures of interest}{the ground-truth} for model construction.  For the targeted application in this work (i.e., WMHs),
 regression/classification are performed at the voxelwise level.  In other words, each voxel
@@
 ![Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
+FLAIR, T1-, and T2-weighted images are \added[id=am,remark=NT: no need for 12 dof for
+intra-subject normalization]{rigidly} pre-aligned
 [@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
@@
 extraction and normal tissue segmentation (i.e., the pathology is discounted which is
-a fairly safe assumption for the T1 image) protocols described in [@Tustison:2014ab] in order to
-produce a mask for the brain parenchyma and provide probabilistic estimates of the
+a fairly safe assumption for the T1 image) protocols as part of the
+cortical thickness pipeline described in [@Tustison:2014ab].
+\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
+brain has been extracted, we apply a Bayesian-based segmentation algorithm
+with a template-based prior probability strategy to segment the
+parenchymal tissue types.  The result is} a mask delineating the brain parenchyma and probabilistic estimates of the
 cerebrospinal fluid (csf), gray matter, white matter, deep gray matter, brain stem, and
-cerebellum.  These provide the ground-truth labels for the first six tissue labels
+cerebellum.  These provide the expertly annotated labels for the first six tissue labels
 given above.  The white matter hyperintensities were manually identified by one of the authors
@@
 of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a radius of one voxel. For each of the normalized
+a \added[id=am]{Manhattan} radius of one voxel. For each of the normalized
 images, we calculate the difference in intensities with the corresponding warped template
@@
 template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
-to take advantage of the bilateral symmetry of the normal brain (in terms of both shape
-and intensity), and the fact that the presence of WMHs violates that assumption,
+to take advantage of the \added[id=hl,remark=NT:  we are using the "symmetry" of the brain in an analogous sense that if we know the left eye location and color then it is highly probabilistic that we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross} bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that \replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
 we use the symmetric templates to compute the contralateral intensity
@@
 model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here.  The Stage 1
+segmentation performance.  We do the same thing for the work described here \added[id=lw]{where we employ two stacked random forests (or two "stages")}.  The Stage 1
 feature images of the training data (as described previously) are used to construct the
 Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-voting maps via the Stage 1 model.  All the Stage 1 features plus the Stage 1 voting maps
+\added[id=lw]{"voting maps" (i.e., the classification count of each decision tree for each tissue label)} via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
 are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors for a second application of the Atropos maximum aposteriori algorithm with an
+tissue priors \added[id=lw]{(i.e., probabilistic estimates of the tissue spatial locations)} for a second application of the Atropos maximum aposteriori algorithm with an
 additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
 the second stage we use all three aligned preprocessed images for a multivariate
-segmentation.  The resulting seven posterior probability images constitute a third additional
-feature image set for Stage 2.
+segmentation.  The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
 
@@
 2000 trees per model and 500 randomly selected samples per label per image.
+\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}
 
@@
 manual white matter hyperintensity tracings were combined to provide the truth labels
-for the training data.  The truth labels are the seven anatomical regions given above.
+for the training data.  The "truth" labels are the seven anatomical regions given above.
 

### 2016-02-11 results.Rmd
ENH:  Edits.

@@
 ![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
-forest models for both Stage 1 and Stage 2 during the leave-one-out evaluation.  These plots
+forest models for Stage 1 during the leave-one-out evaluation.  These plots
 are useful in providing a quantitative assessment of the predictive importance of each feature.
-The horizontal error bars provide the $95^{th}$ percentile (i.e., $1.96 \times \sigma$) and illustrate the
-stability of the feature importance across the leave-one-out models.](Figures/averageLeaveOneOut.png)
+\added[id=nt]{Features are ranked in descending order of importance.}
+The horizontal error bars provide the $95^{th}$ percentile
+\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
+stability of the feature importance across the leave-one-out models.
+\added[id=nt]{At this initial stage only 31 features images are used.}](Figures/averageLeaveOneOutStage1.png)
 
-The resulting rankings for both Stages are given in Figure 4 where the values for the
+![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
+forest models for Stage 2 during the leave-one-out evaluation.  These plots
+are useful in providing a quantitative assessment of the predictive importance of each feature.
+\added[id=nt]{Features are ranked in descending order of importance.}
+The horizontal error bars provide the $95^{th}$ percentile
+\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
+stability of the feature importance across the leave-one-out models.
+\added[id=nt]{We augment the 31 feature images from the first stage by adding an additional
+7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
+segmentation for a total of 45 images for the second stage.}](Figures/averageLeaveOneOutStage2.png)
+
+The resulting rankings for both Stages are given in Figures 4 and 5 where the values for the
 separate stages are averaged over the entire corresponding model set.  In addition, we
@@
 the chosen features during the evaluation.  This latter information is illustrated as
-horizontal errors bars providing the $95^{th}$ percentile (i.e., $1.96 \times \sigma$).
+horizontal errors bars providing the $95^{th}$ percentile \deleted[id=bt]{(i.e., $1.96 \times \sigma$)}.
 Note that the reader can cross reference Table 1 for identifying corresponding feature types
@@
 left-right (i.e., \texttt{TemplateIndicesWarped0}) directions in the space of the symmetric
-template.  This is intuitive since, as discussed previously, manifestation of TBI white
+template.  \deleted[id=bt]{This is intuitive since, as discussed previously, manifestation of TBI white
 matter hyperintensities can often be confused with higher intensities at the
-periventricular caps in normal subjects [@Neema:2009aa] whereas there does not seem to
-be contralateral bias in manifestation of white matter hyperintensities in TBI.
+periventricular caps in normal subjects [@Neema:2009aa]} \deleted[id=am]{whereas there does not seem to
+be contralateral bias in manifestation of white matter hyperintensities in TBI.}
 
@@
 The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models.  This accords with the known clinical relevance of
+evaluation random forest models \added[id=bt]{(see Figure 6)}.  This accords with the known clinical relevance of
 FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not manifest symmetrically in both hemispheres.  Interestingly, the
+pathology does not \added[id=nt]{typically} manifest symmetrically in both hemispheres.  Interestingly, the
 posterior maps for the deep gray matter are extremely important for accurate white matter
@@
 
+![\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated
+in (b).  (c) Contralateral FLAIR difference image.}](Figures/FLAIRcontralaleteralWithLesions.png)
+
 As described earlier, for Stage 2, we used the output random forest voting maps from Stage
 1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 4, the voting maps are labeled as
+an additional MRF spatial prior.  In Figure 5, the voting maps are labeled as
 "\texttt{RFStage1VotingMaps}" where the final numeral is associated with the brain
@@
 
-In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
+In Figure 7 are the segmentation comparisons derived from manual segmentations of the same
 data.  Despite the large variability characteristic with manual labelings in related fields
@@
 current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.  On the left of Figure 5 are the improvement in Dice values over all white
-matter hyperintensities when comparing the segmentations between the two stages.  Performing
-the second round of supervised learning improves Dice values.  One can also note from the
-right side of Figure 5 that the total lesion load volume illustrates a few subjects that are
+in training data.  On the left of Figure 7 are the improvement in Dice values [@tustison2009],
+\added[id=nt]{i.e.,}
+
+$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$
+
+\added[id=nt]{over all white matter hyperintensities when comparing the segmentations between the two stages
+where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
+represents the intersection between the manual/automated lesion pair.}  Performing
+the second round of supervised learning improves these Dice values.  One can also note from the
+right side of Figure 6 that the total lesion load volume illustrates a few subjects that are
 severe outliers in terms of the number of false positives.  The second round helps to

### 2016-02-11 stitched.Rmd
ENH:  Edits.

@@
 
+\listofchanges
+
+\clearpage
+
 # References

### 2016-02-11 stitched.tex
ENH:  Edits.

@@
 \usepackage{booktabs}
+\usepackage{changes}
 \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
 \setlength{\parskip}{3pt}
+\setremarkmarkup{(#2)}
+\definechangesauthor[name={Nick Tustison}, color=red]{nt}
+\definechangesauthor[name={James Stone}, color=magenta]{js}
+\definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
+\definechangesauthor[name={Andy Mayer}, color=green]{am}
+\definechangesauthor[name={Harvey Levin}, color=blue]{hl}
+\definechangesauthor[name={Brian Taylor}, color=purple]{bt}
 
@@
 
+MR images utilized for this initial report were acquired from 24
+individuals scanned at a single site involved in the Chronic Effects of
+Neurotrauma Consortium's (CENC) observational study
+(\added[id=nt,remark=NT: reference needed?]{see [Walker], this issue}).
+Briefly, participants were Operation Enduring Freedom/ Operation Iraqi
+Freedom /Operation New dawn (OEF/OIF/OND) era service members and
+veterans between the ages of 18--60 who have experienced combat
+situations and are between X and X (period of time) from deployment.
+\added[id=js,remark={JS: Data needed.}]{The feature images consisted of subjects aged $X.X \pm X$ years (range X – X years), XX percent male, with at least one or more potential concussive events (mean = X, SD = X, range X-X).}
+
+Images were acquired on a Philips 3.0T Ingenia system with an 8-channel
+SENSE head coil (Philips Medical Systems, Best, Netherlands). 3D FLAIR
+sequences were acquired with a turbo spin echo inversion recovery
+sequence with the following parameters: repetition time (TR) = 4800 ms,
+echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal
+slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and
+256 x 256 mm FOV. 3D T1-weighted sequences were acquired with a fast
+field echo (FFE) sequence with the following parameters: TR = 6.8 ms, TE
+= 3.2 ms, echo train length (ETL) = 240; Flip angle = 9°, 170 sagittal
+slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and
+256 x 256 mm FOV. In addition, 3D T2-weighted images were acquired with
+a turbo spin echo sequence with the following parameters: TR = 2500 ms,
+TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice
+thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.
+
 \subsection{Quantitative analysis}\label{quantitative-analysis}
@@
 Crucial to these supervised segmentation approaches are the creation and
-selection of ``features'' as input in conjunction with the ground-truth
-for model construction. For the targeted application in this work (i.e.,
-WMHs), regression/classification are performed at the voxelwise level.
-In other words, each voxel within the region of interest is sent through
-the ensemble of decision trees and receives a set of classification
-votes from each tree thus permitting a regression or classification
-solution. Since this procedure is performed at the voxelwise level,
-intensity information alone is insufficient for good segmentation
-performance due to the lack of spatial context. For example, as pointed
-out in {[}32{]}, higher intensities can be found at the periventricular
-caps in normal subjects which often confounds automated lesion detection
-algorithms. Other potential confounds include MR signal inhomogeneity
-and noise. Therefore, even though machine learning and pattern
-recognition techniques are extremely powerful and have significant
-potential, just as crucial to outcome is the creative construction and
-deployment of salient feature images which we detail below.
+selection of ``features'' as input in conjunction with
+\replaced[id=lw,remark=NT: I replaced ground-truth since it is problematic anyway]
+{expertly identified structures of interest}{the ground-truth} for model
+construction. For the targeted application in this work (i.e., WMHs),
+regression/classification are performed at the voxelwise level. In other
+words, each voxel within the region of interest is sent through the
+ensemble of decision trees and receives a set of classification votes
+from each tree thus permitting a regression or classification solution.
+Since this procedure is performed at the voxelwise level, intensity
+information alone is insufficient for good segmentation performance due
+to the lack of spatial context. For example, as pointed out in {[}32{]},
+higher intensities can be found at the periventricular caps in normal
+subjects which often confounds automated lesion detection algorithms.
+Other potential confounds include MR signal inhomogeneity and noise.
+Therefore, even though machine learning and pattern recognition
+techniques are extremely powerful and have significant potential, just
+as crucial to outcome is the creative construction and deployment of
+salient feature images which we detail below.
 
@@
 \caption{Representation of Stage 1 feature images for subject 01C1019.
-The FLAIR, T1-, and T2-weighted images are rigidly pre-aligned {[}33{]}
-to the space of the T1 image. The three modality images are then
-preprocessed (N4 bias correction {[}34{]} and adaptive denoising
-{[}35{]}) followed by application of standard ANTs brain extraction and
-\(n\)-tissue segmentation protocols using the MMRR symmetric template
-and corresponding priors {[}36{]} applied to the T1 image. The feature
+The FLAIR, T1-, and T2-weighted images are
+\added[id=am,remark=NT: no need for 12 dof for
+intra-subject normalization]{rigidly} pre-aligned {[}33{]} to the space
+of the T1 image. The three modality images are then preprocessed (N4
+bias correction {[}34{]} and adaptive denoising {[}35{]}) followed by
+application of standard ANTs brain extraction and \(n\)-tissue
+segmentation protocols using the MMRR symmetric template and
+corresponding priors {[}36{]} applied to the T1 image. The feature
 images are then generated for voxelwise input to the RF model which
@@
 tissue segmentation (i.e., the pathology is discounted which is a fairly
-safe assumption for the T1 image) protocols described in {[}36{]} in
-order to produce a mask for the brain parenchyma and provide
-probabilistic estimates of the cerebrospinal fluid (csf), gray matter,
-white matter, deep gray matter, brain stem, and cerebellum. These
-provide the ground-truth labels for the first six tissue labels given
-above. The white matter hyperintensities were manually identified by one
-of the authors (J. R. S.) using the ITK-SNAP tool {[}38{]}. Segmentation
-is performed using the ANTs Atropos tool {[}39{]} and multi-model
-optimal symmetric shape/intensity templates {[}26{]} created from the
-public MMRR data set {[}40{]} (cf Figure 3).
+safe assumption for the T1 image) protocols as part of the cortical
+thickness pipeline described in {[}36{]}.
+\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
+brain has been extracted, we apply a Bayesian-based segmentation algorithm
+with a template-based prior probability strategy to segment the
+parenchymal tissue types.  The result is} a mask delineating the brain
+parenchyma and probabilistic estimates of the cerebrospinal fluid (csf),
+gray matter, white matter, deep gray matter, brain stem, and cerebellum.
+These provide the expertly annotated labels for the first six tissue
+labels given above. The white matter hyperintensities were manually
+identified by one of the authors (J. R. S.) using the ITK-SNAP tool
+{[}38{]}. Segmentation is performed using the ANTs Atropos tool {[}39{]}
+and multi-model optimal symmetric shape/intensity templates {[}26{]}
+created from the public MMRR data set {[}40{]} (cf Figure 3).
 
@@
 image voxel values. We also calculate a set of neighborhood statistics
-(mean, standard deviation, and skewness) feature images using a radius
-of one voxel. For each of the normalized images, we calculate the
-difference in intensities with the corresponding warped template
-component. Previous success in the international brain tumor
-segmentation competition {[}27{]} was based on an important set of
-intensity features that were created from multi-modal templates
-mentioned previously {[}26{]}. We employ the same strategy here. For
-example, the template difference feature image for the FLAIR image,
-\(S_{FLAIR}\) is calculated as:
+(mean, standard deviation, and skewness) feature images using a
+\added[id=am]{Manhattan} radius of one voxel. For each of the normalized
+images, we calculate the difference in intensities with the
+corresponding warped template component. Previous success in the
+international brain tumor segmentation competition {[}27{]} was based on
+an important set of intensity features that were created from
+multi-modal templates mentioned previously {[}26{]}. We employ the same
+strategy here. For example, the template difference feature image for
+the FLAIR image, \(S_{FLAIR}\) is calculated as:
 
@@
 template space and \(T_{FLAIR}\) is the FLAIR template component. Also,
-to take advantage of the bilateral symmetry of the normal brain (in
-terms of both shape and intensity), and the fact that the presence of
-WMHs violates that assumption, we use the symmetric templates to compute
-the contralateral intensity differences as an additional intensity
-feature. For the FLAIR component, this contralateral difference image is
-calculated from
+to take advantage of the
+\added[id=hl,remark=NT:  we are using the "symmetry" of the brain in an analogous sense that if we know the left eye location and color then it is highly probabilistic that we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross}
+bilateral symmetry of the normal brain (in terms of both shape and
+intensity), and the fact that
+\replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
+we use the symmetric templates to compute the contralateral intensity
+differences as an additional intensity feature. For the FLAIR component,
+this contralateral difference image is calculated from
 
@@
 random forest model, can significantly improve segmentation performance.
-We do the same thing for the work described here. The Stage 1 feature
-images of the training data (as described previously) are used to
-construct the Stage 1 model. The training data Stage 1 features are then
-used to produce the voxelwise voting maps via the Stage 1 model. All the
-Stage 1 features plus the Stage 1 voting maps are used as input to the
-Stage 2 model. In addition, we use the Stage 1 voting maps as tissue
-priors for a second application of the Atropos maximum aposteriori
-algorithm with an additional Markov Random Field spatial prior (MAP-MRF)
-{[}39{]}. However, for the second stage we use all three aligned
-preprocessed images for a multivariate segmentation. The resulting seven
-posterior probability images constitute a third additional feature image
-set for Stage 2.
+We do the same thing for the work described here
+\added[id=lw]{where we employ two stacked random forests (or two "stages")}.
+The Stage 1 feature images of the training data (as described
+previously) are used to construct the Stage 1 model. The training data
+Stage 1 features are then used to produce the voxelwise
+\added[id=lw]{"voting maps" (i.e., the classification count of each decision tree for each tissue label)}
+via the Stage 1 random forest model. All the Stage 1 features plus the
+Stage 1 voting maps are used as input to the Stage 2 model. In addition,
+we use the Stage 1 voting maps as tissue priors
+\added[id=lw]{(i.e., probabilistic estimates of the tissue spatial locations)}
+for a second application of the Atropos maximum aposteriori algorithm
+with an additional Markov Random Field spatial prior (MAP-MRF) {[}39{]}.
+However, for the second stage we use all three aligned preprocessed
+images for a multivariate segmentation. The resulting seven posterior
+probability images constitute a third additional feature image set for
+Stage 2.
 
@@
 selected samples per label per image.
+\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}
 
@@
 T1 image and the manual white matter hyperintensity tracings were
-combined to provide the truth labels for the training data. The truth
-labels are the seven anatomical regions given above.
+combined to provide the truth labels for the training data. The
+``truth'' labels are the seven anatomical regions given above.
 
@@
 \centering
-\includegraphics{Figures/averageLeaveOneOut.png}
+\includegraphics{Figures/averageLeaveOneOutStage1.png}
 \caption{Average \texttt{MeanDecreaseAccuracy} plots generated from the
-creation of all 24 random forest models for both Stage 1 and Stage 2
-during the leave-one-out evaluation. These plots are useful in providing
-a quantitative assessment of the predictive importance of each feature.
-The horizontal error bars provide the \(95^{th}\) percentile (i.e.,
-\(1.96 \times \sigma\)) and illustrate the stability of the feature
-importance across the leave-one-out models.}
+creation of all 24 random forest models for Stage 1 during the
+leave-one-out evaluation. These plots are useful in providing a
+quantitative assessment of the predictive importance of each feature.
+\added[id=nt]{Features are ranked in descending order of importance.}
+The horizontal error bars provide the \(95^{th}\) percentile
+\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
+stability of the feature importance across the leave-one-out models.
+\added[id=nt]{At this initial stage only 31 features images are used.}}
 \end{figure}
 
-The resulting rankings for both Stages are given in Figure 4 where the
-values for the separate stages are averaged over the entire
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/averageLeaveOneOutStage2.png}
+\caption{Average \texttt{MeanDecreaseAccuracy} plots generated from the
+creation of all 24 random forest models for Stage 2 during the
+leave-one-out evaluation. These plots are useful in providing a
+quantitative assessment of the predictive importance of each feature.
+\added[id=nt]{Features are ranked in descending order of importance.}
+The horizontal error bars provide the \(95^{th}\) percentile
+\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
+stability of the feature importance across the leave-one-out models.
+\added[id=nt]{We augment the 31 feature images from the first stage by adding an additional
+7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
+segmentation for a total of 45 images for the second stage.}}
+\end{figure}
+
+The resulting rankings for both Stages are given in Figures 4 and 5
+where the values for the separate stages are averaged over the entire
 corresponding model set. In addition, we track the variance for each
@@
 features during the evaluation. This latter information is illustrated
-as horizontal errors bars providing the \(95^{th}\) percentile (i.e.,
-\(1.96 \times \sigma\)). Note that the reader can cross reference Table
-1 for identifying corresponding feature types and names.
+as horizontal errors bars providing the \(95^{th}\) percentile
+\deleted[id=bt]{(i.e., $1.96 \times \sigma$)}. Note that the reader can
+cross reference Table 1 for identifying corresponding feature types and
+names.
 
@@
 \texttt{TemplateIndicesWarped0}) directions in the space of the
-symmetric template. This is intuitive since, as discussed previously,
-manifestation of TBI white matter hyperintensities can often be confused
-with higher intensities at the periventricular caps in normal subjects
-{[}32{]} whereas there does not seem to be contralateral bias in
-manifestation of white matter hyperintensities in TBI.
+symmetric template.
+\deleted[id=bt]{This is intuitive since, as discussed previously, manifestation of TBI white
+matter hyperintensities can often be confused with higher intensities at the
+periventricular caps in normal subjects [@Neema:2009aa]}
+\deleted[id=am]{whereas there does not seem to
+be contralateral bias in manifestation of white matter hyperintensities in TBI.}
 
@@
 features for Stage 1. The contralateral difference FLAIR image is highly
-discriminative over the set of evaluation random forest models. This
-accords with the known clinical relevance of FLAIR images for
-identifying white matter hyperintensities and the fact that such
-pathology does not manifest symmetrically in both hemispheres.
-Interestingly, the posterior maps for the deep gray matter are extremely
-important for accurate white matter hyperintensity segmentation. Perhaps
-the spatial specification of deep gray matter aids in the removal of
-false positives. Inspection of the bottom of the plots demonstrates the
-lack of discriminating features associated with the T1 image which is
-also well-known in the clinical literature.
+discriminative over the set of evaluation random forest models
+\added[id=bt]{(see Figure 6)}. This accords with the known clinical
+relevance of FLAIR images for identifying white matter hyperintensities
+and the fact that such pathology does not \added[id=nt]{typically}
+manifest symmetrically in both hemispheres. Interestingly, the posterior
+maps for the deep gray matter are extremely important for accurate white
+matter hyperintensity segmentation. Perhaps the spatial specification of
+deep gray matter aids in the removal of false positives. Inspection of
+the bottom of the plots demonstrates the lack of discriminating features
+associated with the T1 image which is also well-known in the clinical
+literature.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/FLAIRcontralaleteralWithLesions.png}
+\caption{\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated
+in (b).  (c) Contralateral FLAIR difference image.}}
+\end{figure}
 
@@
 input to a Bayesian-based segmentation with an additional MRF spatial
-prior. In Figure 4, the voting maps are labeled as
+prior. In Figure 5, the voting maps are labeled as
 ``\texttt{RFStage1VotingMaps}'' where the final numeral is associated
@@
 
-In Figure 5 are the segmentation comparisons derived from manual
+In Figure 7 are the segmentation comparisons derived from manual
 segmentations of the same data. Despite the large variability
@@
 methodology proposed herein is readily adapted to refinements in
-training data. On the left of Figure 5 are the improvement in Dice
-values over all white matter hyperintensities when comparing the
-segmentations between the two stages. Performing the second round of
-supervised learning improves Dice values. One can also note from the
-right side of Figure 5 that the total lesion load volume illustrates a
-few subjects that are severe outliers in terms of the number of false
-positives. The second round helps to correct this issue.
+training data. On the left of Figure 7 are the improvement in Dice
+values {[}49{]}, \added[id=nt]{i.e.,}
+
+\[ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } \]
+
+\added[id=nt]{over all white matter hyperintensities when comparing the segmentations between the two stages
+where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
+represents the intersection between the manual/automated lesion pair.}
+Performing the second round of supervised learning improves these Dice
+values. One can also note from the right side of Figure 6 that the total
+lesion load volume illustrates a few subjects that are severe outliers
+in terms of the number of false positives. The second round helps to
+correct this issue.
 
@@
 
+\listofchanges
+
+\clearpage
+
 \section*{References}\label{references}
@@
 
+\hypertarget{ref-tustison2009}{}
+49. Tustison, N. J. and Gee, J. C. ``\textbf{Introducing Dice, Jaccard,
+and Other Label Overlap Measures to ITK}'' \emph{Insight Journal}
+(2009):
+
 \end{document}

### 2016-02-11 discussion.Rmd
ENH:  Edits.
ENH:  More edits.

@@
 
-A recent meta-analysis by Kloppenborg et al.
+\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
 [@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and concurrent neuropsychological results in patients with heterogeneous diagnoses but without previously diagnosed cognitive impairment, found that WMHs were associated with cognitive deficit (effect size of -0.10, 95% CI: -0.13 to -0.08) after controlling for age.
@@
 
-The currently described tool is just one example of how “deep learning” algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state.
+The currently described tool is just one example of how "supervised learning" algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state.
 It is an important demonstration of the potential power of these analytical approaches in the rapid but comprehensive mining of information from neuroimaging examinations.
-Deep learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features.
-Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data.
+Supervised learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
+Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
 Similarly, in the context of the neuroimaging assessments, deep learning approaches may allow for the rapid identification of information predictive of disease state in an individual patient.
-These approaches may be applied to the identification of macroscopically visible lesions as described in the current communication.
+These approaches have been applied to the segmentation of macroscopically visible structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
 Additionally, these approaches might be applied to the interrogation of imaging data in the individual patient with a primary quantitative output metrics to include sequences such as diffusion tensor imaging (DTI) and its variants, functional connectivity, perfusion weighted imaging, and cortical thickness assessments.
-At present, these sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient.
+At present, these advanced neuroimaging sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient [@Mayer:2014aa].
 Application of deep learning approaches in the context of data with primary quantitative outputs will require large scale normative and disease specific databases.
-Acquisition of these data is resource intensive and requires a multi-center approach with harmonized scanner between sites and correlative non-imaging clinical data.
-Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through the Federal Interagency TBI repository (FITBIR).
+Building these large scale imaging libraries is resource intensive and requires a multi-center approach with harmonized scanners between sites and correlative non-imaging clinical data.
+Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
 In concert with any available high quality normative neuroimaging data, deep learning algorithms may be well positioned to help transform how neuroimaging is interpreted for the clinical management of patients with this disease process.
@@
 
+<!--
+\listofchanges
+
+\clearpage
+-->
+
 # References

### 2016-02-11 format.Rmd
ENH:  More edits.

--- 2016-01-23 format.Rmd
@@
    - \usepackage{booktabs}
+   - \usepackage[deletedmarkup=xout]{changes}
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
    - \setlength{\parskip}{3pt}
+   - \setremarkmarkup{(#2)}
+   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
+   - \definechangesauthor[name={James Stone}, color=magenta]{js}
+   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
+   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
+   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
+   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
+   - \definechangesauthor[name={David Tate}, color=orange]{dt}
 bibliography: references.bib

### 2016-02-11 intro.Rmd
ENH:  More edits.

@@
+<!--
+A couple notes in response to comments:
+
+[David Tate, Comment 15]:  Stupid question, but why not show the same subject and slice for both for a better comparison?
+
+NT:  I see what he is saying and probably agree.
+
+[David Tate, Comment 28]:  Maybe this could go in a Figure as a flow chart??
+
+NT:  Flow charts are great and I included a couple as Figure 5 in my tumor paper.
+However, they take a long time to make correctly and I do not want to commit to producing one unless it is absolutely necessary.
+
+[Andy Mayer, Comment 30]: It will be difficult to see this font for labeling on Y axis at current size/resolution.
+You may also consider alerting your reader directly that the magnitudes of the scales are very different for stage 1 versus stage 2 in the legend or scale them to be similar.
+
+NT:  Comparing these two plots in Figure 4 in terms of their absolute MeanDecreaseAccuracy
+is meaningless so coordinating the scales is not important.
+
+[Andy Mayer, Comment 34]:  A reviewer is likely to ask for some sort of reliability rating on the labeling of the lesions.
+
+NT:  I agree with this but unless somebody else is willing to manually segment the WMHs,
+then there is not really much we can do other than point to other papers which have looked
+at reliability.
+
+[David Tate, Comment 37]:  Seems like we are missing a picture.
+That is the football player in me, but it seems like we need a picture that shows the manual compared to the automated at each stage if possible.
+
+NT:  We could do this but I do not think that such images in publications are that meaningful as they typically result from heavy selection bias.  I did do something similar
+for my brain segmentation paper where I showed multiple slices (skipped every 10, or something like that) for each subject in the challenge cohort but tumors are obviously much bigger and can be displayed in this manner.
+
+[David Tate, Comment 39]:  I would would move these comments down to a limitation like paragraph, otherwise it minimizes the cool findings.
+
+NT:  I have no opinion on this.  Do what you think is best.
+
+[Andy Mayer, Comment 40]:  I do not know if you can include any of this data (even one subject).
+It would be great to demonstrate that your algorithm worked on data collected on another scanner since the intensities are likely to be very different.
+
+NT:  As I mentioned to you previously, I think this is a great idea and I would be happy to include it but how do we showcase this unless somebody is willing to annotate data from these other sites.
+We could include a picture of a single subject but how persuasive that is will be reviewer-dependent.
+Personally, I would not be persuaded as the showing the results of a single subject is very prone to selection bias.
+
+-->
+
 # Introduction
@@
 
-The random forests framework [@breiman2001] is a popular machine learning technique that has
-demonstrated significant utitility for supervised segmentation tasks (e.g.,
+The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant \replaced[id=dt]{utility}{utitility} for supervised segmentation tasks (e.g.,
 normal human brain segmentation [@yi2009]) and other computer vision applications
@@
 latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions.  In response to the lack of objective comparisons between
-segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
-2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI
-conference.
+international competitions \replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
+In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}[@Menze:2015aa].
 
@@
 votes are then used for regression or classification of the data.  Although decision
-trees had been extensively studied previously, the success of employing collections of
+trees had been extensively studied \deleted[id=dt]{previously}, the success of employing collections of
 such weak learners for boosting machine learning performance
@@
 of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
-Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
-data (i.e., "bagging") resulting in the current random forest technique.
+Finally, Breiman [@breiman2001] improved accuracy by random sampling of \deleted[id=dt]{the} training
+data (i.e., "bagging") resulting in the current random forest technique \added[id=dt]{applied here}.
 
@@
 feature image set (both spatial and intensity-based)
-for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
+for segmenting \replaced[id=dt]{WMHs}{white matter hyperintensities} in \replaced[id=dt]{a arge TBI cohort}{traumatic brain injury cohorts}.  Additionally,
 the entire framework is provided publicly through the well-known open-source
-ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the research of this work is
+\added[id=lw]{Advanced Normalization Tools} (ANTs)[^i1] and ANTsR[^i2] toolkits.
+\replaced[id=dt]{Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automate routine for WMH segmentation and quantification.}{Further motivating this research of this work is
 the public availability of the imaging data thus permitting full reproducibility
-of the results reported and discussed.
+of the results reported and discussed.}
 

### 2016-02-11 methods.Rmd
ENH:  More edits.

@@
 categorization of the feature images used for the initial (i.e., Stage 1---more on the use
-of multiple random forest stages below) segmentation of the white matter hyperintensities.
+of multiple random forest stages below) segmentation of the WMHs.
 In addition Figure 1 provides a representation of a set of feature images for a single
@@
 range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results.
+algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar \added[id=dt]{to previous work}[@Tustison:2015aa].
 
@@
 parenchymal tissue types.  The result is} a mask delineating the brain parenchyma and probabilistic estimates of the
-cerebrospinal fluid (csf), gray matter, white matter, deep gray matter, brain stem, and
+\replaced[id=dt]{CSF}{cerebrospinal fluid (csf)}, gray matter, white matter, deep gray matter, brain stem, and
 cerebellum.  These provide the expertly annotated labels for the first six tissue labels
-given above.  The white matter hyperintensities were manually identified by one of the authors
+given above.  The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually identified by one of the authors
 (J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
@@
 of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a \added[id=am]{Manhattan} radius of one voxel. For each of the normalized
+a \added[id=am]{Manhattan} radius of one voxel \added[id=nt]{given the typical size
+of individual WMHs}. For each of the normalized
 images, we calculate the difference in intensities with the corresponding warped template
@@
 competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa].  We employ the
+from multi-modal templates mentioned previously [@Tustison:2015aa] \added[id=nt]{and listed in Table 1}.  We employ the
 same strategy here.  For example, the template
@@
 template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
-to take advantage of the \added[id=hl,remark=NT:  we are using the "symmetry" of the brain in an analogous sense that if we know the left eye location and color then it is highly probabilistic that we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross} bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that \replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
+to take advantage of the \added[id=hl,remark=NT:  we are using the gross bilateral symmetry of the brain in an analogous sense that if we know the left eye location and color then we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross} bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that \replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
 we use the symmetric templates to compute the contralateral intensity
@@
 derived as part of the registration process, we can warp the voxel index locations
-back to the space of the individual subject.    Note that this is similar in motivation to the work
-of [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
+back to the space of the individual subject \replaced[id=nt]{which motivates similar work by others}{
+Note that this is similar in motivation to the work
+of} [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
 coordinate system provided by the template to dramatically improve spatial specificity
@@
 is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing of which none are based on supervised learning.  As we did for
+at the time of writing \added[id=am]{this article}.  \replaced[id=dt]{In contrast to the current work,}{of which} none are based on supervised learning.  As we did for
 our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
@@
 a self-encapsulated example to showcase the proposed methodology.  The fact that
-the data will also be made available through the FITBIR repository along
+the data will also be made available through the \replaced[id=lw]{Federal Interagency Traumatic Brain Injury Research (FITBIR)}{FITBIR} repository along
 with the manual labelings will facilitate reproducibility on the part of the reader

### 2016-02-11 results.Rmd
ENH:  More edits.

@@
 
-After performing the leave-one-out evaluation described at the end of the previous section,
+After performing the leave-one-out evaluation\deleted[id=dt]{described at the end of the previous section},
 we calculated the \texttt{MeanDecreaseAccuracy} feature values for each of the 24 subjects $\times$
@@
 
-![Comparison with manual delineation of white matter hyperintensities.  On the left are the
-calculate Dice values over all white matter hyperintensities.  Note the improvement
+![\added[id=am]{Voxelwise} comparison with manual delineation of white matter hyperintensities.  On the left are the
+calculate\added[id=am]{d} Dice values over all white matter hyperintensities.  Note the improvement
 in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
@@
 processing in a couple subjects, the number of outliers caused by false positives is
-significantly with the second stage processing.](Figures/llvAndDice.png)
+\added[id=lw]{decreased} significantly with the second stage processing.](Figures/llvAndDice.png)

### 2016-02-11 stitched.Rmd
ENH:  More edits.

@@
    - \usepackage{booktabs}
-   - \usepackage{changes}
+   - \usepackage[deletedmarkup=xout]{changes}
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
    - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
+   - \definechangesauthor[name={David Tate}, color=orange]{dt}
 bibliography: references.bib
@@
 \pagenumbering{gobble}
+<!--
+A couple notes in response to comments:
+
+[David Tate, Comment 15]:  Stupid question, but why not show the same subject and slice for both for a better comparison?
+
+NT:  I see what he is saying and probably agree.
+
+[David Tate, Comment 28]:  Maybe this could go in a Figure as a flow chart??
+
+NT:  Flow charts are great and I included a couple as Figure 5 in my tumor paper.
+However, they take a long time to make correctly and I do not want to commit to producing one unless it is absolutely necessary.
+
+[Andy Mayer, Comment 30]: It will be difficult to see this font for labeling on Y axis at current size/resolution.
+You may also consider alerting your reader directly that the magnitudes of the scales are very different for stage 1 versus stage 2 in the legend or scale them to be similar.
+
+NT:  Comparing these two plots in Figure 4 in terms of their absolute MeanDecreaseAccuracy
+is meaningless so coordinating the scales is not important.
+
+[Andy Mayer, Comment 34]:  A reviewer is likely to ask for some sort of reliability rating on the labeling of the lesions.
+
+NT:  I agree with this but unless somebody else is willing to manually segment the WMHs,
+then there is not really much we can do other than point to other papers which have looked
+at reliability.
+
+[David Tate, Comment 37]:  Seems like we are missing a picture.
+That is the football player in me, but it seems like we need a picture that shows the manual compared to the automated at each stage if possible.
+
+NT:  We could do this but I do not think that such images in publications are that meaningful as they typically result from heavy selection bias.  I did do something similar
+for my brain segmentation paper where I showed multiple slices (skipped every 10, or something like that) for each subject in the challenge cohort but tumors are obviously much bigger and can be displayed in this manner.
+
+[David Tate, Comment 39]:  I would would move these comments down to a limitation like paragraph, otherwise it minimizes the cool findings.
+
+NT:  I have no opinion on this.  Do what you think is best.
+
+[Andy Mayer, Comment 40]:  I do not know if you can include any of this data (even one subject).
+It would be great to demonstrate that your algorithm worked on data collected on another scanner since the intensities are likely to be very different.
+
+NT:  As I mentioned to you previously, I think this is a great idea and I would be happy to include it but how do we showcase this unless somebody is willing to annotate data from these other sites.
+We could include a picture of a single subject but how persuasive that is will be reviewer-dependent.
+Personally, I would not be persuaded as the showing the results of a single subject is very prone to selection bias.
+
+-->
 
@@
 processing in a couple subjects, the number of outliers caused by false positives is
-significantly with the second stage processing.](Figures/llvAndDice.png)
+\added[id=lw]{decreased} significantly with the second stage processing.](Figures/llvAndDice.png)
 
@@
 
+<!--
 \listofchanges
@@
 \clearpage
+-->
 

### 2016-02-11 stitched.tex
ENH:  More edits.

@@
 \usepackage{booktabs}
-\usepackage{changes}
+\usepackage[deletedmarkup=xout]{changes}
 \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
 \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
+\definechangesauthor[name={David Tate}, color=orange]{dt}
 
@@
 The random forests framework {[}18{]} is a popular machine learning
-technique that has demonstrated significant utitility for supervised
-segmentation tasks (e.g., normal human brain segmentation {[}19{]}) and
-other computer vision applications (e.g., human gait detection
-{[}20{]}). In the context of neuropathology, random forest-based
-paradigms have been employed in the delineation of multiple sclerosis
-lesions {[}21{]}, stroke lesions {[}22{]}, and brain tumors
-{[}23--26{]}. Of note, these latter random forest approaches for brain
-tumor segmentation have performed well in recent international
-competitions. In response to the lack of objective comparisons between
-segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS)
-challenge was initiated in 2012 {[}27{]} and has continued every year
-since under the auspices of the MICCAI conference.
+technique that has demonstrated significant
+\replaced[id=dt]{utility}{utitility} for supervised segmentation tasks
+(e.g., normal human brain segmentation {[}19{]}) and other computer
+vision applications (e.g., human gait detection {[}20{]}). In the
+context of neuropathology, random forest-based paradigms have been
+employed in the delineation of multiple sclerosis lesions {[}21{]},
+stroke lesions {[}22{]}, and brain tumors {[}23--26{]}. Of note, these
+latter random forest approaches for brain tumor segmentation have
+performed well in recent international competitions
+\replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
+In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}{[}27{]}.
 
@@
 are then used for regression or classification of the data. Although
-decision trees had been extensively studied previously, the success of
-employing collections of such weak learners for boosting machine
-learning performance (e.g., AdaBoost {[}28, 29{]}) influenced the
-similarly sytled conglomeration of decision trees into ``forests'' with
-randomized node optimization {[}30, 31{]}. Finally, Breiman {[}18{]}
-improved accuracy by random sampling of the training data (i.e.,
-``bagging'') resulting in the current random forest technique.
+decision trees had been extensively studied \deleted[id=dt]{previously},
+the success of employing collections of such weak learners for boosting
+machine learning performance (e.g., AdaBoost {[}28, 29{]}) influenced
+the similarly sytled conglomeration of decision trees into ``forests''
+with randomized node optimization {[}30, 31{]}. Finally, Breiman
+{[}18{]} improved accuracy by random sampling of \deleted[id=dt]{the}
+training data (i.e., ``bagging'') resulting in the current random forest
+technique \added[id=dt]{applied here}.
 
@@
 tailored contextual feature image set (both spatial and intensity-based)
-for segmenting white matter hyperintensities in traumatic brain injury
-cohorts. Additionally, the entire framework is provided publicly through
-the well-known open-source ANTs\footnote{\url{https://github.com/stnava/ANTs}}
-and ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
-Further motivating the research of this work is the public availability
-of the imaging data thus permitting full reproducibility of the results
-reported and discussed.
+for segmenting \replaced[id=dt]{WMHs}{white matter hyperintensities} in
+\replaced[id=dt]{a arge TBI cohort}{traumatic brain injury cohorts}.
+Additionally, the entire framework is provided publicly through the
+well-known open-source \added[id=lw]{Advanced Normalization Tools}
+(ANTs)\footnote{\url{https://github.com/stnava/ANTs}} and
+ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
+\replaced[id=dt]{Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automate routine for WMH segmentation and quantification.}{Further motivating this research of this work is
+the public availability of the imaging data thus permitting full reproducibility
+of the results reported and discussed.}
 
@@
 images used for the initial (i.e., Stage 1---more on the use of multiple
-random forest stages below) segmentation of the white matter
-hyperintensities. In addition Figure 1 provides a representation of a
-set of feature images for a single subject analyzed in this work. Note
-that in this work we categorize the brain parenchyma with seven labels:
+random forest stages below) segmentation of the WMHs. In addition Figure
+1 provides a representation of a set of feature images for a single
+subject analyzed in this work. Note that in this work we categorize the
+brain parenchyma with seven labels:
 
@@
 algorithm (e.g., {[}37{]}), we found that a simple linear rescaling
-produced better results.
+produced better results similar \added[id=dt]{to previous work}{[}26{]}.
 
@@
 parenchymal tissue types.  The result is} a mask delineating the brain
-parenchyma and probabilistic estimates of the cerebrospinal fluid (csf),
-gray matter, white matter, deep gray matter, brain stem, and cerebellum.
-These provide the expertly annotated labels for the first six tissue
-labels given above. The white matter hyperintensities were manually
+parenchyma and probabilistic estimates of the
+\replaced[id=dt]{CSF}{cerebrospinal fluid (csf)}, gray matter, white
+matter, deep gray matter, brain stem, and cerebellum. These provide the
+expertly annotated labels for the first six tissue labels given above.
+The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually
 identified by one of the authors (J. R. S.) using the ITK-SNAP tool
@@
 (mean, standard deviation, and skewness) feature images using a
-\added[id=am]{Manhattan} radius of one voxel. For each of the normalized
-images, we calculate the difference in intensities with the
-corresponding warped template component. Previous success in the
-international brain tumor segmentation competition {[}27{]} was based on
-an important set of intensity features that were created from
-multi-modal templates mentioned previously {[}26{]}. We employ the same
-strategy here. For example, the template difference feature image for
-the FLAIR image, \(S_{FLAIR}\) is calculated as:
+\added[id=am]{Manhattan} radius of one voxel
+\added[id=nt]{given the typical size
+of individual WMHs}. For each of the normalized images, we calculate the
+difference in intensities with the corresponding warped template
+component. Previous success in the international brain tumor
+segmentation competition {[}27{]} was based on an important set of
+intensity features that were created from multi-modal templates
+mentioned previously {[}26{]} \added[id=nt]{and listed in Table 1}. We
+employ the same strategy here. For example, the template difference
+feature image for the FLAIR image, \(S_{FLAIR}\) is calculated as:
 
@@
 to take advantage of the
-\added[id=hl,remark=NT:  we are using the "symmetry" of the brain in an analogous sense that if we know the left eye location and color then it is highly probabilistic that we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross}
+\added[id=hl,remark=NT:  we are using the gross bilateral symmetry of the brain in an analogous sense that if we know the left eye location and color then we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross}
 bilateral symmetry of the normal brain (in terms of both shape and
@@
 also derived as part of the registration process, we can warp the voxel
-index locations back to the space of the individual subject. Note that
-this is similar in motivation to the work of {[}44{]}. However, this
-previous work lacks the normalization to the standard coordinate system
-provided by the template to dramatically improve spatial specificity
-across all subjects.
+index locations back to the space of the individual subject
+\replaced[id=nt]{which motivates similar work by others}{
+Note that this is similar in motivation to the work
+of} {[}44{]}. However, this previous work lacks the normalization to the
+standard coordinate system provided by the template to dramatically
+improve spatial specificity across all subjects.
 
@@
 in the literature is quite extensive, there were only four publicly
-available segmentation algorithms at the time of writing of which none
-are based on supervised learning. As we did for our brain tumor
-segmentation algorithm {[}26{]}, all of the code described in this work
-is publicly available through the open-source ANTs/ANTsR toolkits.
-Through ANTsR (an add-on toolkit which, in part, bridges ANTs and the R
-statistical project) we use the \texttt{randomForest} package {[}46{]}
-using the default settings with 2000 trees per model and 500 randomly
-selected samples per label per image.
+available segmentation algorithms at the time of writing
+\added[id=am]{this article}.
+\replaced[id=dt]{In contrast to the current work,}{of which} none are
+based on supervised learning. As we did for our brain tumor segmentation
+algorithm {[}26{]}, all of the code described in this work is publicly
+available through the open-source ANTs/ANTsR toolkits. Through ANTsR (an
+add-on toolkit which, in part, bridges ANTs and the R statistical
+project) we use the \texttt{randomForest} package {[}46{]} using the
+default settings with 2000 trees per model and 500 randomly selected
+samples per label per image.
 \added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}
@@
 methodology. The fact that the data will also be made available through
-the FITBIR repository along with the manual labelings will facilitate
+the
+\replaced[id=lw]{Federal Interagency Traumatic Brain Injury Research (FITBIR)}{FITBIR}
+repository along with the manual labelings will facilitate
 reproducibility on the part of the reader as well as any interest in
@@
 
-After performing the leave-one-out evaluation described at the end of
-the previous section, we calculated the \texttt{MeanDecreaseAccuracy}
-feature values for each of the 24 subjects \(\times\) 2 models per
-subject \(=48\) total models. This measure (per feature, per model) is
-calculated during the out-of-bag phase of the random forest model
-construction and quantifies the decrease in prediction accuracy from
-omitting the specified feature. In other words, this quantity helps
-determine the importance of a particular feature and, although we save
-such efforts for future work, this information provides us with guidance
-for future feature pruning and/or additions.
+After performing the leave-one-out
+evaluation\deleted[id=dt]{described at the end of the previous section},
+we calculated the \texttt{MeanDecreaseAccuracy} feature values for each
+of the 24 subjects \(\times\) 2 models per subject \(=48\) total models.
+This measure (per feature, per model) is calculated during the
+out-of-bag phase of the random forest model construction and quantifies
+the decrease in prediction accuracy from omitting the specified feature.
+In other words, this quantity helps determine the importance of a
+particular feature and, although we save such efforts for future work,
+this information provides us with guidance for future feature pruning
+and/or additions.
 
@@
 \includegraphics{Figures/llvAndDice.png}
-\caption{Comparison with manual delineation of white matter
-hyperintensities. On the left are the calculate Dice values over all
-white matter hyperintensities. Note the improvement in the Dice metric
-from the employment of the Stage 2 component of the processing pipeline.
-(Right) Similar results can be seen by comparing the total lesion load
-volume between manual and automated detection strategies. Although some
+\caption{\added[id=am]{Voxelwise} comparison with manual delineation of
+white matter hyperintensities. On the left are the
+calculate\added[id=am]{d} Dice values over all white matter
+hyperintensities. Note the improvement in the Dice metric from the
+employment of the Stage 2 component of the processing pipeline. (Right)
+Similar results can be seen by comparing the total lesion load volume
+between manual and automated detection strategies. Although some
 outliers are found after the Stage 2 processing in a couple subjects,
-the number of outliers caused by false positives is significantly with
-the second stage processing.}
+the number of outliers caused by false positives is
+\added[id=lw]{decreased} significantly with the second stage
+processing.}
 \end{figure}
@@
 
-A recent meta-analysis by Kloppenborg et al. {[}10{]} of 23
-cross-sectional studies reporting MRI and concurrent neuropsychological
-results in patients with heterogeneous diagnoses but without previously
-diagnosed cognitive impairment, found that WMHs were associated with
-cognitive deficit (effect size of -0.10, 95\% CI: -0.13 to -0.08) after
-controlling for age. These studies also differed in the metric used to
-measure the WMHs, including volume, \% of total intracranial volume, and
-a visual rating score. The effect size for the association with
-cognitive deficit in these cross-sectional studies did not differ
-significantly across various cognitive domains or the method of
-measuring lesion volume. Among eight longitudinal studies analyzed by
+\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
+{[}10{]} of 23 cross-sectional studies reporting MRI and concurrent
+neuropsychological results in patients with heterogeneous diagnoses but
+without previously diagnosed cognitive impairment, found that WMHs were
+associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
+to -0.08) after controlling for age. These studies also differed in the
+metric used to measure the WMHs, including volume, \% of total
+intracranial volume, and a visual rating score. The effect size for the
+association with cognitive deficit in these cross-sectional studies did
+not differ significantly across various cognitive domains or the method
+of measuring lesion volume. Among eight longitudinal studies analyzed by
 Kloppenborg et al that included a follow-up MRI and also controlled for
@@
 
-The currently described tool is just one example of how ``deep
+The currently described tool is just one example of how ``supervised
 learning'' algorithms might be applied to aid in the diagnosis of TBI
@@
 the rapid but comprehensive mining of information from neuroimaging
-examinations. Deep learning algorithms are presently employed across a
-wide variety of settings for the rapid identification of predictive
-imaging features. Automobile manufacturers utilize these types of
-approaches to equip self-driving vehicles to recognize and respond to
-unique external surroundings through the identification of visual
-information sufficiently similar to previously assimilated training
-data. Similarly, in the context of the neuroimaging assessments, deep
-learning approaches may allow for the rapid identification of
-information predictive of disease state in an individual patient. These
-approaches may be applied to the identification of macroscopically
-visible lesions as described in the current communication. Additionally,
-these approaches might be applied to the interrogation of imaging data
-in the individual patient with a primary quantitative output metrics to
-include sequences such as diffusion tensor imaging (DTI) and its
-variants, functional connectivity, perfusion weighted imaging, and
-cortical thickness assessments. At present, these sequences are confined
-to cohort-based research studies due to the lack of available analytical
-tools to assess the information in the setting of the individual
-patient. Application of deep learning approaches in the context of data
-with primary quantitative outputs will require large scale normative and
-disease specific databases. Acquisition of these data is resource
-intensive and requires a multi-center approach with harmonized scanner
+examinations. Supervised learning algorithms are presently employed
+across a wide variety of settings for the rapid identification of
+predictive imaging features {[}50--53{]}. Automobile manufacturers
+utilize these types of approaches to equip self-driving vehicles to
+recognize and respond to unique external surroundings through the
+identification of visual information sufficiently similar to previously
+assimilated training data {[}54, 55{]}. Similarly, in the context of the
+neuroimaging assessments, deep learning approaches may allow for the
+rapid identification of information predictive of disease state in an
+individual patient. These approaches have been applied to the
+segmentation of macroscopically visible structures {[}50--53{]}.
+Additionally, these approaches might be applied to the interrogation of
+imaging data in the individual patient with a primary quantitative
+output metrics to include sequences such as diffusion tensor imaging
+(DTI) and its variants, functional connectivity, perfusion weighted
+imaging, and cortical thickness assessments. At present, these advanced
+neuroimaging sequences are confined to cohort-based research studies due
+to the lack of available analytical tools to assess the information in
+the setting of the individual patient {[}56{]}. Application of deep
+learning approaches in the context of data with primary quantitative
+outputs will require large scale normative and disease specific
+databases. Building these large scale imaging libraries is resource
+intensive and requires a multi-center approach with harmonized scanners
 between sites and correlative non-imaging clinical data. Large scale TBI
@@
 Defense Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and
-other data being consolidated through the Federal Interagency TBI
-repository (FITBIR). In concert with any available high quality
-normative neuroimaging data, deep learning algorithms may be well
-positioned to help transform how neuroimaging is interpreted for the
-clinical management of patients with this disease process.
-
-\clearpage
-
-\listofchanges
+other data being consolidated through FITBIR. In concert with any
+available high quality normative neuroimaging data, deep learning
+algorithms may be well positioned to help transform how neuroimaging is
+interpreted for the clinical management of patients with this disease
+process.
 
@@
 
+\hypertarget{ref-Plis:2014aa}{}
+50. Plis, S. M., Hjelm, D. R., Salakhutdinov, R., Allen, E. A.,
+Bockholt, H. J., Long, J. D., Johnson, H. J., Paulsen, J. S., Turner, J.
+A., and Calhoun, V. D. ``\textbf{Deep Learning for Neuroimaging: A
+Validation Study}'' \emph{Front Neurosci} 8, (2014): 229.
+doi:\href{https://doi.org/10.3389/fnins.2014.00229}{10.3389/fnins.2014.00229}
+
+\hypertarget{ref-Suk:2015aa}{}
+51. Suk, H.-I., Lee, S.-W., Shen, D., and Alzheimer's Disease
+Neuroimaging Initiative. ``\textbf{Deep Sparse Multi-Task Learning for
+Feature Selection in Alzheimer's Disease Diagnosis}'' \emph{Brain Struct
+Funct} (2015):
+doi:\href{https://doi.org/10.1007/s00429-015-1059-y}{10.1007/s00429-015-1059-y}
+
+\hypertarget{ref-Li:2014aa}{}
+52. Li, R., Zhang, W., Suk, H.-I., Wang, L., Li, J., Shen, D., and Ji,
+S. ``\textbf{Deep Learning Based Imaging Data Completion for Improved
+Brain Disease Diagnosis}'' \emph{Med Image Comput Comput Assist Interv}
+17, no. Pt 3 (2014): 305--12.
+
+\hypertarget{ref-Liu:2015aa}{}
+53. Liu, S., Liu, S., Cai, W., Che, H., Pujol, S., Kikinis, R., Feng,
+D., Fulham, M. J., and ADNI. ``\textbf{Multimodal Neuroimaging Feature
+Learning for Multiclass Diagnosis of Alzheimer's Disease}'' \emph{IEEE
+Trans Biomed Eng} 62, no. 4 (2015): 1132--40.
+doi:\href{https://doi.org/10.1109/TBME.2014.2372011}{10.1109/TBME.2014.2372011}
+
+\hypertarget{ref-HadsellSBESKML09}{}
+54. Hadsell, R., Sermanet, P., Ben, J., Erkan, A., Scoffier, M.,
+Kavukcuoglu, K., Muller, U., and LeCun, Y. ``\textbf{Learning Long-Range
+Vision for Autonomous Off-Road Driving}'' \emph{J. Field Robotics} 26,
+no. 2 (2009): 120--144.
+doi:\href{https://doi.org/10.1002/rob.20276}{10.1002/rob.20276},
+Available at \url{http://dx.doi.org/10.1002/rob.20276}
+
+\hypertarget{ref-FarabetCNL12}{}
+55. Farabet, C., Couprie, C., Najman, L., and LeCun, Y. ``\textbf{Scene
+Parsing with Multiscale Feature Learning, Purity Trees, and Optimal
+Covers}'' \emph{Proceedings of the 29th international conference on
+machine learning, ICML 2012, edinburgh, scotland, uK, june 26 - july 1,
+2012} (2012): Available at \url{http://icml.cc/discuss/2012/319.html}
+
+\hypertarget{ref-Mayer:2014aa}{}
+56. Mayer, A. R., Bedrick, E. J., Ling, J. M., Toulouse, T., and Dodd,
+A. ``\textbf{Methods for Identifying Subject-Specific Abnormalities in
+Neuroimaging Data}'' \emph{Hum Brain Mapp} 35, no. 11 (2014): 5457--70.
+doi:\href{https://doi.org/10.1002/hbm.22563}{10.1002/hbm.22563}
+
 \end{document}

### 2016-02-11 format.Rmd
ENH: final

--- 2016-01-25 format.Rmd
@@
    - \usepackage{booktabs}
+   - \usepackage[final]{changes}
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}

### 2016-02-11 stitched.Rmd
ENH: final

@@
    - \usepackage{booktabs}
-   - \usepackage[deletedmarkup=xout]{changes}
+   - \usepackage[final]{changes}
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}

### 2016-02-18 abstract.Rmd
ENH:  more.

# Abstract

White matter hyperintensities (WMHs) are foci of abnormal signal
intensity in white matter regions seen with magnetic resonance imaging
(MRI). WMHs are associated with normal aging and have shown prognostic
value in neurological conditions such as traumatic brain injury (TBI).
The impracticality of manually quantifying these lesions limits their
clinical utility and motivates the utilization of machine learning
techniques for automated segmentation workflows.  Herein, we develop a
concatenated random forest framework with tailored features for
segmenting WMHs in a TBI cohort. The framework is provided publicly
through the Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
(CENC) observational study were acquired. Manual annotations were
employed for both training and evaluation using a leave-one-out
strategy.  Lesion load and overlap evaluative comparisons are
complimented by feature rankings which showcase the utility of the
concatenated approach.  Our findings suggest supervised learning methods
may be applied to quantify WMHs on routine brain imaging. Paired with
correlative outcome data, supervised learning methods may allow for
identification of imaging features predictive of diagnosis and prognosis
in individual TBI patients.

\clearpage

### 2016-02-18 discussion.Rmd
ENH:  more.

@@
 # Discussion
-
-<!--
-Nick could you add some detail here on how this technique compares to other automated or semi automated approaches for WMH detection?
--->
+#
+<!-- Nick could you add some detail here on how this technique compares
+to other automated or semi automated approaches for WMH detection? -->
 
@@
 
-* Evaluation
-    * Proper context
-        * We only have manual labelings from a single radiologist (
-        * TBI white matter hyperintensities are more difficult to segment than MS lesions
-            * individual lesions are smaller
-            * lesion load is typically smaller
-            * not prone to enhancement
-            * less distinct from surrounding tissue ?
-    * Despite the doubts regarding the efficacy of a typical segmentation validation,
-      our contribution is useful because
-         * the total segmentation framework is publicly available within the ANTs/ANTsR
-           framework
-         * although only a single site is analyzed, the feature images are site-agnostic
-              * we plan to apply the current RF models directly to the other site data
-              * we can then build the training database further by enlisting other experts
-         * Is this the first work exploring segmentation of white matter hyperintensities in TBI?
-         * The data (including manual tracings) will be made available as part of the FITBIR
-           effort.
+* Evaluation * Proper context * We only have manual labelings from a
+single radiologist ( * TBI white matter hyperintensities are more
+difficult to segment than MS lesions * individual lesions are smaller *
+lesion load is typically smaller * not prone to enhancement * less
+distinct from surrounding tissue ? * Despite the doubts regarding the
+efficacy of a typical segmentation validation, our contribution is
+useful because * the total segmentation framework is publicly available
+within the ANTs/ANTsR framework * although only a single site is
+analyzed, the feature images are site-agnostic * we plan to apply the
+current RF models directly to the other site data * we can then build
+the training database further by enlisting other experts * Is this the
+first work exploring segmentation of white matter hyperintensities in
+TBI? * The data (including manual tracings) will be made available as
+part of the FITBIR effort.
 
@@
 
-The current communications describes a supervised statistical learning methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation of WMHs from FLAIR images to help build two-stage ensembles of decision trees for the automated identification of these lesions.
-Although
-only a single expert was used to produce the manual labelings, our intent is to further
-refine the proposed paradigm by crowdsourcing with feedback from other experts who interact with both the data and methodology.  Also, we recognize that only a single site
-was used for evaluating the proposed framework.  However, we are currently processing other site data with the models developed for this work and the results look promising
-since the developed features are site-agnostic.
+The current communications describes a supervised statistical learning
+methodology for identifying WHMs within multimodal MR brain imaging.
+This effort utilized information acquired from the manual segmentation
+of WMHs from FLAIR images to help build two-stage ensembles of decision
+trees for the automated identification of these lesions. Although only a
+single expert was used to produce the manual labelings, our intent is to
+further refine the proposed paradigm by crowdsourcing with feedback from
+other experts who interact with both the data and methodology.  Also, we
+recognize that only a single site was used for evaluating the proposed
+framework.  However, we are currently processing other site data with
+the models developed for this work and the results look promising since
+the developed features are site-agnostic.
 
-As far as we know, this is the first report utilizing a novel random forest approach to identify WMHs in a cohort of TBI patients.
-TBI WMHs tend to be more difficult to segment than MS lesions as the former tend to be smaller with an overall smaller lesion load.
-Also, enhancement protocols with the former tend to be less successful than with the latter.
-As mentioned previously, the work in MS lesion segmentation is extensive
-with a handful of techniques being publicly available.
-Our framework is also available as open-source as part of well-known neuroimaging tools which easily allows for additions/extensions but is also, as far as we know, the first random forest-based technique available for such application.
+As far as we know, this is the first report utilizing a novel random
+forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
+tend to be more difficult to segment than MS lesions as the former tend
+to be smaller with an overall smaller lesion load.  Also, enhancement
+protocols with the former tend to be less successful than with the
+latter.  As mentioned previously, the work in MS lesion segmentation is
+extensive with a handful of techniques being publicly available.  Our
+framework is also available as open-source as part of well-known
+neuroimaging tools which easily allows for additions/extensions but is
+also, as far as we know, the first random forest-based technique
+available for such application.
 
-Two major meta-analyses of WMHs have been published covering the periods of 1966-2009 and 2010-2015, respectively.
-The earlier meta-analysis covered 53 longitudinal studies that included samples of high-risk populations, i.e., patients selected for a specific disease or condition such as hypertension, whereas other studies recruited samples of the general population.
-Longitudinal studies of samples representative of the general population are more relevant to the focus of the present paper.
-Debette & Markus [@Debette:2010aa] found that the presence of WMHs was related to subsequent cognitive decline, a higher risk of developing dementia, stroke, and of mortality.
-Lesion volume at baseline was also predictive of cognitive decline.
-Limitations of this meta-analysis include heterogeneity in the method of measuring WMHs; some studies used automated volumetric measurement, whereas others used a visual rating scale.
-The studies analyzed by Debette & Markus were limited to the occurrence of one of the aforementioned conditions which they analyzed by hazard ratios.
+Two major meta-analyses of WMHs have been published covering the periods
+prior to 2010 [@Debette:2010aa] and after [@Kloppenborg:2014aa]. The earlier meta-analysis
+covered 53 longitudinal studies that included samples of high-risk
+populations, i.e., patients selected for a specific disease or condition
+such as hypertension, whereas other studies recruited samples of the
+general population. Longitudinal studies of samples representative of
+the general population are more relevant to the focus of the present
+paper. Debette & Markus [@Debette:2010aa] found that the presence of
+WMHs was related to subsequent cognitive decline, a higher risk of
+developing dementia, stroke, and of mortality. Lesion volume at baseline
+was also predictive of cognitive decline. Limitations of this
+meta-analysis include heterogeneity in the method of measuring WMHs;
+some studies used automated volumetric measurement, whereas others used
+a visual rating scale. The studies analyzed by Debette & Markus were
+limited to the occurrence of one of the aforementioned conditions which
+they analyzed by hazard ratios.
 
 \replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and concurrent neuropsychological results in patients with heterogeneous diagnoses but without previously diagnosed cognitive impairment, found that WMHs were associated with cognitive deficit (effect size of -0.10, 95% CI: -0.13 to -0.08) after controlling for age.
-These studies also differed in the metric used to measure the WMHs, including volume, % of total intracranial volume, and a visual rating score.
-The effect size for the association with cognitive deficit in these cross-sectional studies did not differ significantly across various cognitive domains or the method of measuring lesion volume.
-Among eight longitudinal studies analyzed by Kloppenborg et al that included a follow-up MRI and also controlled for age, the effect size for the association of progression in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09).
-This association was stronger for attention and executive function than for memory and processing speed.
-Although baseline WMHs were predictive of cognitive deficit at follow-up in the seven studies which did not repeat MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in the longitudinal studies that calculated progression in WMHs.
-In summary, progression of WMHs seen on repeat MRI has a stronger relation to cognitive deficit than concurrent imaging findings.
-These meta-analyses support the rationale for repeating an MRI in patients younger than 50 years whose initial scan shows WMHs.
-Despite the above-described associations between WMHs, cognitive decline, increased risk of developing dementia, and mortality, these lesions receive little attention in current clinical workflows.
-When reported in a standard neuroradiologist interpretation, they are typically handled as incidental findings and are assigned little clinical significance.
-This likely reflects the impracticality of performing a detailed assessment of number, volume, and distribution within a qualitative neuroradiologist interpretation as well as the lack of correlative information on how the presence and distribution of these lesions may inform a diagnosis and prognosis in the appropriate clinical setting.
-To date, automated or semi-automated tools for the detection of WMHs have lacked the specificity and efficiency for the mining of large-scale datasets to generate highly granular data on whether these lesions possess any true diagnostic or prognostic value in the setting of a specific disease process.
-The present communication describes a supervised statistical learning tool that is appropriate for the application to such large-scale datasets.
+[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
+concurrent neuropsychological results in patients with heterogeneous
+diagnoses but without previously diagnosed cognitive impairment, found
+that WMHs were associated with cognitive deficit (effect size of -0.10,
+95% CI: -0.13 to -0.08) after controlling for age. These studies also
+differed in the metric used to measure the WMHs, including volume, % of
+total intracranial volume, and a visual rating score. The effect size
+for the association with cognitive deficit in these cross-sectional
+studies did not differ significantly across various cognitive domains or
+the method of measuring lesion volume.  Among eight longitudinal studies
+analyzed by Kloppenborg et al that included a follow-up MRI and also
+controlled for age, the effect size for the association of progression
+in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09). This
+association was stronger for attention and executive function than for
+memory and processing speed. Although baseline WMHs were predictive of
+cognitive deficit at follow-up in the seven studies which did not repeat
+MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in
+the longitudinal studies that calculated progression in WMHs. In
+summary, progression of WMHs seen on repeat MRI has a stronger relation
+to cognitive deficit than concurrent imaging findings. These
+meta-analyses support the rationale for repeating an MRI in patients
+younger than 50 years whose initial scan shows WMHs.
 
-The currently described tool is just one example of how "supervised learning" algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state.
-It is an important demonstration of the potential power of these analytical approaches in the rapid but comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning approaches may allow for the rapid identification of information predictive of disease state in an individual patient.
-These approaches have been applied to the segmentation of macroscopically visible structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of imaging data in the individual patient with a primary quantitative output metrics to include sequences such as diffusion tensor imaging (DTI) and its variants, functional connectivity, perfusion weighted imaging, and cortical thickness assessments.
-At present, these advanced neuroimaging sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient [@Mayer:2014aa].
-Application of deep learning approaches in the context of data with primary quantitative outputs will require large scale normative and disease specific databases.
-Building these large scale imaging libraries is resource intensive and requires a multi-center approach with harmonized scanners between sites and correlative non-imaging clinical data.
-Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data, deep learning algorithms may be well positioned to help transform how neuroimaging is interpreted for the clinical management of patients with this disease process.
+Despite the
+above-described associations between WMHs, cognitive decline, increased
+risk of developing dementia, and mortality, these lesions receive little
+attention in current clinical workflows. When reported in a standard
+neuroradiologist interpretation, they are typically handled as
+incidental findings and are assigned little clinical significance. This
+likely reflects the impracticality of performing a detailed assessment
+of number, volume, and distribution within a qualitative
+neuroradiologist interpretation as well as the lack of correlative
+information on how the presence and distribution of these lesions may
+inform a diagnosis and prognosis in the appropriate clinical setting. To
+date, automated or semi-automated tools for the detection of WMHs have
+lacked the specificity and efficiency for the mining of large-scale
+datasets to generate highly granular data on whether these lesions
+possess any true diagnostic or prognostic value in the setting of a
+specific disease process. The present communication describes a
+supervised statistical learning tool that is appropriate for the
+application to such large-scale datasets.
+
+The currently described tool is just one example of how "supervised
+learning" algorithms might be applied to aid in the diagnosis of TBI and
+other disease processes through the specific identification of features
+predictive of a given disease state. It is an important demonstration of
+the potential power of these analytical approaches in the rapid but
+comprehensive mining of information from neuroimaging examinations.
+Supervised learning algorithms are presently employed across a wide
+variety of settings for the rapid identification of predictive imaging
+features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
+manufacturers utilize these types of approaches to equip self-driving
+vehicles to recognize and respond to unique external surroundings
+through the identification of visual information sufficiently similar to
+previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
+Similarly, in the context of the neuroimaging assessments, deep learning
+approaches may allow for the rapid identification of information
+predictive of disease state in an individual patient. These approaches
+have been applied to the segmentation of macroscopically visible
+structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
+Additionally, these approaches might be applied to the interrogation of
+imaging data in the individual patient with a primary quantitative
+output metrics to include sequences such as diffusion tensor imaging
+(DTI) and its variants, functional connectivity, perfusion weighted
+imaging, and cortical thickness assessments. At present, these advanced
+neuroimaging sequences are confined to cohort-based research studies due
+to the lack of available analytical tools to assess the information in
+the setting of the individual patient [@Mayer:2014aa]. Application of
+deep learning approaches in the context of data with primary
+quantitative outputs will require large scale normative and disease
+specific databases. Building these large scale imaging libraries is
+resource intensive and requires a multi-center approach with harmonized
+scanners between sites and correlative non-imaging clinical data. Large
+scale TBI data is becoming increasingly available through activities
+such as the Chronic Effects of Neurotrauma Consortium (CENC),
+Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
+Collaborative European Neurotrauma Effectiveness Research in TBI
+(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
+Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
+In concert with any available high quality normative neuroimaging data,
+deep learning algorithms may be well positioned to help transform how
+neuroimaging is interpreted for the clinical management of patients with
+this disease process.
 
@@
 
-<!--
-\listofchanges
+<!-- \listofchanges
 
-\clearpage
--->
+\clearpage -->
 

### 2016-02-18 format.Rmd
ENH:  more.

--- 2016-02-11 format.Rmd
@@
    - \usepackage{booktabs}
-   - \usepackage{changes}
+   - \usepackage[final]{changes}
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
    - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
+   - \definechangesauthor[name={David Tate}, color=orange]{dt}
 bibliography: references.bib
-csl: national-science-foundation-grant-proposals.csl
+csl: council-of-science-editors.csl
 fontsize: 11pt

### 2016-02-18 intro.Rmd
ENH:  more.

@@
 Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa].
-Additionally, the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].
+Additionally,
+in patients who are comatose following severe TBI
+the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].
 
@@
 delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
-and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
+and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa] for both
+single and multi-modal acquisition protocols.  Of note, these
 latter random forest approaches for brain tumor segmentation have performed well in recent
@@
 
-Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
+Random forests are conceptually straightforward [@breiman2001].  They consist of ensembles of decision trees
 that are built from training data.  Once constructed, data to be classified is "pushed"

### 2016-02-18 methods.Rmd
ENH:  more.

@@
 
-![Sample FLAIR acquisition image slices showing both manual (top) and random forest
-(RDF) segmentations (bottom) obtained during the leave-one-out evaluation.  Manual
+![Sample FLAIR acquisition image slices showing both manual and random forest
+segmentations for both stages obtained during the leave-one-out evaluation.  Manual
 segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.](Figures/montage.png)
+truth WMH labels for training the random forest models.](Figures/sampleResults.png)
 
@@
 2000 trees per model and 500 randomly selected samples per label per image.
-\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}
+\added[id=bt]{Note that we saw little variation in performance when these parameters were changed (i.e.
+up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.}
 

### 2016-02-18 results.Rmd
ENH:  more.

@@
 
-![\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated
-in (b).  (c) Contralateral FLAIR difference image.}](Figures/FLAIRcontralaleteralWithLesions.png)
+![\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated.
+The region around the WMHs is enlarged (b) in the original FLAIR and the
+(c) contralateral FLAIR difference image.}](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
 

### 2016-02-18 stitched.Rmd
ENH:  more.

@@
 bibliography: references.bib
-csl: national-science-foundation-grant-proposals.csl
+csl: council-of-science-editors.csl
 fontsize: 11pt
@@
 \pagenumbering{gobble}
+
+# Abstract
+
+White matter hyperintensities (WMHs) are foci of abnormal signal
+intensity in white matter regions seen with magnetic resonance imaging
+(MRI). WMHs are associated with normal aging and have shown prognostic
+value in neurological conditions such as traumatic brain injury (TBI).
+The impracticality of manually quantifying these lesions limits their
+clinical utility and motivates the utilization of machine learning
+techniques for automated segmentation workflows.  Herein, we develop a
+concatenated random forest framework with tailored features for
+segmenting WMHs in a TBI cohort. The framework is provided publicly
+through the Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
+(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
+veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
+(CENC) observational study were acquired. Manual annotations were
+employed for both training and evaluation using a leave-one-out
+strategy.  Lesion load and overlap evaluative comparisons are
+complimented by feature rankings which showcase the utility of the
+concatenated approach.  Our findings suggest supervised learning methods
+may be applied to quantify WMHs on routine brain imaging. Paired with
+correlative outcome data, supervised learning methods may allow for
+identification of imaging features predictive of diagnosis and prognosis
+in individual TBI patients.
+
+\clearpage
 <!--

### 2016-02-18 stitched.tex
ENH: final
ENH:  more.

@@
 \usepackage{booktabs}
-\usepackage[deletedmarkup=xout]{changes}
+\usepackage[final]{changes}
 \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
@@
 
+\section{Abstract}\label{abstract}
+
+White matter hyperintensities (WMHs) are foci of abnormal signal
+intensity in white matter regions seen with magnetic resonance imaging
+(MRI). WMHs are associated with normal aging and have shown prognostic
+value in neurological conditions such as traumatic brain injury (TBI).
+The impracticality of manually quantifying these lesions limits their
+clinical utility and motivates the utilization of machine learning
+techniques for automated segmentation workflows. Herein, we develop a
+concatenated random forest framework with tailored features for
+segmenting WMHs in a TBI cohort. The framework is provided publicly
+through the Advanced Normalization Tools (ANTs) and ANTsR toolkits. MR
+(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
+veterans scanned in the Chronic Effects of Neurotrauma Consortium's
+(CENC) observational study were acquired. Manual annotations were
+employed for both training and evaluation using a leave-one-out
+strategy. Lesion load and overlap evaluative comparisons are
+complimented by feature rankings which showcase the utility of the
+concatenated approach. Our findings suggest supervised learning methods
+may be applied to quantify WMHs on routine brain imaging. Paired with
+correlative outcome data, supervised learning methods may allow for
+identification of imaging features predictive of diagnosis and prognosis
+in individual TBI patients.
+
+\clearpage
+
 \section{Introduction}\label{introduction}
@@
 been correlated with functional outcome and injury severity in both
-pediatric {[}1, 2{]} and adult {[}3--6{]}. Further, the regional
-distribution and volume of WMHs have been shown to possess prognostic
-value in the TBI patient {[}2, 6--8{]}. Specifically, lesion volume in
-corpus callosum correlates with functional scores in the acute phase
-following injury, while lesion volume in frontal lobes correlates with
-scores at 1 year following injury {[}6{]}. Further, volume of FLAIR
-lesions within the corpus callosum, brainstem, and thalamus in patients
-with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores
-{[}4{]}. Additionally, the regional distribution of FLAIR lesions within
+pediatric\textsuperscript{1,2} and adult\textsuperscript{3--6}. Further,
+the regional distribution and volume of WMHs have been shown to possess
+prognostic value in the TBI patient\textsuperscript{2,6--8}.
+Specifically, lesion volume in corpus callosum correlates with
+functional scores in the acute phase following injury, while lesion
+volume in frontal lobes correlates with scores at 1 year following
+injury\textsuperscript{6}. Further, volume of FLAIR lesions within the
+corpus callosum, brainstem, and thalamus in patients with severe TBI
+correlates with Glasgow Outcome-Extended (GOS-E)
+scores\textsuperscript{4}. Additionally, in patients who are comatose
+following severe TBI the regional distribution of FLAIR lesions within
 the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal,
 occipital lobes, and insula along with the observation of grasping or
-chewing behavior are associated with poor outcome {[}7{]}.
+chewing behavior are associated with poor outcome\textsuperscript{7}.
 
@@
 association between WMHs, cognitive function, increased risk of stroke,
-dementia, and death {[}9, 10{]}. As such, the development of automated
-methods for the rapid identification and quantification of WMHs within
-individual patients may allow for identification of correlative patterns
-between WMH number, volume, distribution, and disease state. Further,
-the development of such lesion quantification approaches may allow for
-the practical inclusion of this type of information within routine
-radiological practice.
+dementia, and death\textsuperscript{9,10}. As such, the development of
+automated methods for the rapid identification and quantification of
+WMHs within individual patients may allow for identification of
+correlative patterns between WMH number, volume, distribution, and
+disease state. Further, the development of such lesion quantification
+approaches may allow for the practical inclusion of this type of
+information within routine radiological practice.
 
@@
 in conjunction with the Medical Image Computing and Computer-Aided
-Intervention (MICCAI) international meeting {[}11{]}). Popular
-techniques such as support vector machines and neural networks have been
-applied successfully to clinically relevant imaging tasks such as
-supervised image segmentation (e.g., {[}12{]}) and diagnostic prediction
-(e.g., {[}13, 14{]}). Facilitating the current employment of such
-techniques are the number of available imaging data sets {[}15{]} and
-the public availability of data science packages such as SciPy {[}16{]}
-and the R project for statistical computing {[}17{]} and their
+Intervention (MICCAI) international meeting\textsuperscript{11}).
+Popular techniques such as support vector machines and neural networks
+have been applied successfully to clinically relevant imaging tasks such
+as supervised image segmentation (e.g.,\textsuperscript{12}) and
+diagnostic prediction (e.g.,\textsuperscript{13,14}). Facilitating the
+current employment of such techniques are the number of available
+imaging data sets\textsuperscript{15} and the public availability of
+data science packages such as SciPy\textsuperscript{16} and the R
+project for statistical computing\textsuperscript{17} and their
 associated extensions.
 
-The random forests framework {[}18{]} is a popular machine learning
-technique that has demonstrated significant
+The random forests framework\textsuperscript{18} is a popular machine
+learning technique that has demonstrated significant
 \replaced[id=dt]{utility}{utitility} for supervised segmentation tasks
-(e.g., normal human brain segmentation {[}19{]}) and other computer
-vision applications (e.g., human gait detection {[}20{]}). In the
-context of neuropathology, random forest-based paradigms have been
-employed in the delineation of multiple sclerosis lesions {[}21{]},
-stroke lesions {[}22{]}, and brain tumors {[}23--26{]}. Of note, these
+(e.g., normal human brain segmentation\textsuperscript{19}) and other
+computer vision applications (e.g., human gait
+detection\textsuperscript{20}). In the context of neuropathology, random
+forest-based paradigms have been employed in the delineation of multiple
+sclerosis lesions\textsuperscript{21}, stroke
+lesions\textsuperscript{22}, and brain tumors\textsuperscript{23--26}
+for both single and multi-modal acquisition protocols. Of note, these
 latter random forest approaches for brain tumor segmentation have
@@
 \replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
-In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}{[}27{]}.
-
-Random forests are conceptually simple {[}18{]}. They consist of
-ensembles of decision trees that are built from training data. Once
-constructed, data to be classified is ``pushed'' through each decision
-tree resulting in a single classification ``vote'' per tree. These votes
-are then used for regression or classification of the data. Although
-decision trees had been extensively studied \deleted[id=dt]{previously},
-the success of employing collections of such weak learners for boosting
-machine learning performance (e.g., AdaBoost {[}28, 29{]}) influenced
-the similarly sytled conglomeration of decision trees into ``forests''
-with randomized node optimization {[}30, 31{]}. Finally, Breiman
-{[}18{]} improved accuracy by random sampling of \deleted[id=dt]{the}
-training data (i.e., ``bagging'') resulting in the current random forest
-technique \added[id=dt]{applied here}.
+In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}\textsuperscript{27}.
+
+Random forests are conceptually straightforward\textsuperscript{18}.
+They consist of ensembles of decision trees that are built from training
+data. Once constructed, data to be classified is ``pushed'' through each
+decision tree resulting in a single classification ``vote'' per tree.
+These votes are then used for regression or classification of the data.
+Although decision trees had been extensively studied
+\deleted[id=dt]{previously}, the success of employing collections of
+such weak learners for boosting machine learning performance (e.g.,
+AdaBoost\textsuperscript{28,29}) influenced the similarly sytled
+conglomeration of decision trees into ``forests'' with randomized node
+optimization\textsuperscript{30,31}. Finally,
+Breiman\textsuperscript{18} improved accuracy by random sampling of
+\deleted[id=dt]{the} training data (i.e., ``bagging'') resulting in the
+current random forest technique \added[id=dt]{applied here}.
 
@@
 information alone is insufficient for good segmentation performance due
-to the lack of spatial context. For example, as pointed out in {[}32{]},
-higher intensities can be found at the periventricular caps in normal
-subjects which often confounds automated lesion detection algorithms.
-Other potential confounds include MR signal inhomogeneity and noise.
-Therefore, even though machine learning and pattern recognition
-techniques are extremely powerful and have significant potential, just
-as crucial to outcome is the creative construction and deployment of
-salient feature images which we detail below.
+to the lack of spatial context. For example, as pointed out
+in\textsuperscript{32}, higher intensities can be found at the
+periventricular caps in normal subjects which often confounds automated
+lesion detection algorithms. Other potential confounds include MR signal
+inhomogeneity and noise. Therefore, even though machine learning and
+pattern recognition techniques are extremely powerful and have
+significant potential, just as crucial to outcome is the creative
+construction and deployment of salient feature images which we detail
+below.
 
@@
 \added[id=am,remark=NT: no need for 12 dof for
-intra-subject normalization]{rigidly} pre-aligned {[}33{]} to the space
-of the T1 image. The three modality images are then preprocessed (N4
-bias correction {[}34{]} and adaptive denoising {[}35{]}) followed by
-application of standard ANTs brain extraction and \(n\)-tissue
-segmentation protocols using the MMRR symmetric template and
-corresponding priors {[}36{]} applied to the T1 image. The feature
-images are then generated for voxelwise input to the RF model which
-results in the voting maps illustrated on the right. This gives a
-probabilistic classification of tissue type. Not shown are the
-probability and voting images for the brain stem and cerebellum.}
+intra-subject normalization]{rigidly} pre-aligned\textsuperscript{33} to
+the space of the T1 image. The three modality images are then
+preprocessed (N4 bias correction\textsuperscript{34} and adaptive
+denoising\textsuperscript{35}) followed by application of standard ANTs
+brain extraction and \(n\)-tissue segmentation protocols using the MMRR
+symmetric template and corresponding priors\textsuperscript{36} applied
+to the T1 image. The feature images are then generated for voxelwise
+input to the RF model which results in the voting maps illustrated on
+the right. This gives a probabilistic classification of tissue type. Not
+shown are the probability and voting images for the brain stem and
+cerebellum.}
 \end{figure}
@@
 registered to the T1 image using the open-source Advanced Normalization
-Tools (ANTs) {[}33{]}. The aligned images are then preprocessed using
-the denoising algorithm of {[}35{]} followed by N4 bias correction
-{[}34{]} which are then normalized to the intensity range \([0,1]\).
-Although we could have used an alternative intensity standardization
-algorithm (e.g., {[}37{]}), we found that a simple linear rescaling
-produced better results similar \added[id=dt]{to previous work}{[}26{]}.
+Tools (ANTs)\textsuperscript{33}. The aligned images are then
+preprocessed using the denoising algorithm of\textsuperscript{35}
+followed by N4 bias correction\textsuperscript{34} which are then
+normalized to the intensity range \([0,1]\). Although we could have used
+an alternative intensity standardization algorithm
+(e.g.,\textsuperscript{37}), we found that a simple linear rescaling
+produced better results similar
+\added[id=dt]{to previous work}\textsuperscript{26}.
 
@@
 \centering
-\includegraphics{Figures/montage.png}
-\caption{Sample FLAIR acquisition image slices showing both manual (top)
-and random forest (RDF) segmentations (bottom) obtained during the
+\includegraphics{Figures/sampleResults.png}
+\caption{Sample FLAIR acquisition image slices showing both manual and
+random forest segmentations for both stages obtained during the
 leave-one-out evaluation. Manual segmentations were performed by one of
@@
 safe assumption for the T1 image) protocols as part of the cortical
-thickness pipeline described in {[}36{]}.
+thickness pipeline described in\textsuperscript{36}.
 \added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
@@
 The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually
-identified by one of the authors (J. R. S.) using the ITK-SNAP tool
-{[}38{]}. Segmentation is performed using the ANTs Atropos tool {[}39{]}
-and multi-model optimal symmetric shape/intensity templates {[}26{]}
-created from the public MMRR data set {[}40{]} (cf Figure 3).
+identified by one of the authors (J. R. S.) using the ITK-SNAP
+tool\textsuperscript{38}. Segmentation is performed using the ANTs
+Atropos tool\textsuperscript{39} and multi-model optimal symmetric
+shape/intensity templates\textsuperscript{26} created from the public
+MMRR data set\textsuperscript{40} (cf Figure 3).
 
@@
 \caption{Canonical views of the mutlivariate, bilaterally symmetric
-template constructed from the MMRR data set {[}40{]} (only shown are the
-FLAIR, T1, and T2 modalities--- the components relevant for this work).
-Template construction is detailed in {[}26{]}. These images are
-important for specific intensity-based features.}
+template constructed from the MMRR data set\textsuperscript{40} (only
+shown are the FLAIR, T1, and T2 modalities--- the components relevant
+for this work). Template construction is detailed
+in\textsuperscript{26}. These images are important for specific
+intensity-based features.}
 \end{figure}
@@
 component. Previous success in the international brain tumor
-segmentation competition {[}27{]} was based on an important set of
-intensity features that were created from multi-modal templates
-mentioned previously {[}26{]} \added[id=nt]{and listed in Table 1}. We
-employ the same strategy here. For example, the template difference
-feature image for the FLAIR image, \(S_{FLAIR}\) is calculated as:
+segmentation competition\textsuperscript{27} was based on an important
+set of intensity features that were created from multi-modal templates
+mentioned previously\textsuperscript{26}
+\added[id=nt]{and listed in Table 1}. We employ the same strategy here.
+For example, the template difference feature image for the FLAIR image,
+\(S_{FLAIR}\) is calculated as:
 
@@
 prediction step. Additional spatial contextual feature images include
-the distance maps {[}41{]} based on the csf, gray matter, and deep gray
-matter images. These latter images are intended to help distinguish
-white matter hyperintensities from false positives induced by the
-partial voluming at the gray/white matter interface. A third set of
-images are based on the voxel location within the space of the template.
-The T1 image of the subject is registered to the T1 template component
-using a B-spline variant {[}42{]} of the well-known ANTs Symmetric
-Normalization (SyN) algorithm {[}43{]}. Since the inverse transform is
-also derived as part of the registration process, we can warp the voxel
-index locations back to the space of the individual subject
+the distance maps\textsuperscript{41} based on the csf, gray matter, and
+deep gray matter images. These latter images are intended to help
+distinguish white matter hyperintensities from false positives induced
+by the partial voluming at the gray/white matter interface. A third set
+of images are based on the voxel location within the space of the
+template. The T1 image of the subject is registered to the T1 template
+component using a B-spline variant\textsuperscript{42} of the well-known
+ANTs Symmetric Normalization (SyN) algorithm\textsuperscript{43}. Since
+the inverse transform is also derived as part of the registration
+process, we can warp the voxel index locations back to the space of the
+individual subject
 \replaced[id=nt]{which motivates similar work by others}{
 Note that this is similar in motivation to the work
-of} {[}44{]}. However, this previous work lacks the normalization to the
-standard coordinate system provided by the template to dramatically
-improve spatial specificity across all subjects.
+of}\textsuperscript{44}. However, this previous work lacks the
+normalization to the standard coordinate system provided by the template
+to dramatically improve spatial specificity across all subjects.
 
@@
 
-In previous brain tumor segmentation work {[}26{]}, it was demonstrated
-that a concatenated supervised approach, whereby the prediction output
-from the first random forest model serves as partial input for a second
-random forest model, can significantly improve segmentation performance.
-We do the same thing for the work described here
+In previous brain tumor segmentation work\textsuperscript{26}, it was
+demonstrated that a concatenated supervised approach, whereby the
+prediction output from the first random forest model serves as partial
+input for a second random forest model, can significantly improve
+segmentation performance. We do the same thing for the work described
+here
 \added[id=lw]{where we employ two stacked random forests (or two "stages")}.
@@
 for a second application of the Atropos maximum aposteriori algorithm
-with an additional Markov Random Field spatial prior (MAP-MRF) {[}39{]}.
-However, for the second stage we use all three aligned preprocessed
-images for a multivariate segmentation. The resulting seven posterior
-probability images constitute a third additional feature image set for
-Stage 2.
+with an additional Markov Random Field spatial prior
+(MAP-MRF)\textsuperscript{39}. However, for the second stage we use all
+three aligned preprocessed images for a multivariate segmentation. The
+resulting seven posterior probability images constitute a third
+additional feature image set for Stage 2.
 
@@
 As pointed out in a recent comprehensive multiple sclerosis lesion
-segmentation review {[}45{]}, although the number of algorithms reported
-in the literature is quite extensive, there were only four publicly
-available segmentation algorithms at the time of writing
-\added[id=am]{this article}.
+segmentation review\textsuperscript{45}, although the number of
+algorithms reported in the literature is quite extensive, there were
+only four publicly available segmentation algorithms at the time of
+writing \added[id=am]{this article}.
 \replaced[id=dt]{In contrast to the current work,}{of which} none are
 based on supervised learning. As we did for our brain tumor segmentation
-algorithm {[}26{]}, all of the code described in this work is publicly
-available through the open-source ANTs/ANTsR toolkits. Through ANTsR (an
-add-on toolkit which, in part, bridges ANTs and the R statistical
-project) we use the \texttt{randomForest} package {[}46{]} using the
-default settings with 2000 trees per model and 500 randomly selected
-samples per label per image.
-\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}
+algorithm\textsuperscript{26}, all of the code described in this work is
+publicly available through the open-source ANTs/ANTsR toolkits. Through
+ANTsR (an add-on toolkit which, in part, bridges ANTs and the R
+statistical project) we use the \texttt{randomForest}
+package\textsuperscript{46} using the default settings with 2000 trees
+per model and 500 randomly selected samples per label per image.
+\added[id=bt]{Note that we saw little variation in performance when these parameters were changed (i.e.
+up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.}
 
@@
 \centering
-\includegraphics{Figures/FLAIRcontralaleteralWithLesions.png}
-\caption{\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated
-in (b).  (c) Contralateral FLAIR difference image.}}
+\includegraphics{Figures/FLAIRcontralaleteralWithLesionsAlternate.png}
+\caption{\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated.
+The region around the WMHs is enlarged (b) in the original FLAIR and the
+(c) contralateral FLAIR difference image.}}
 \end{figure}
@@
 segmentations of the same data. Despite the large variability
-characteristic with manual labelings in related fields {[}45, 47, 48{]},
-such labelings are characteristic of current clinical practices and the
-methodology proposed herein is readily adapted to refinements in
-training data. On the left of Figure 7 are the improvement in Dice
-values {[}49{]}, \added[id=nt]{i.e.,}
+characteristic with manual labelings in related
+fields\textsuperscript{45,47,48}, such labelings are characteristic of
+current clinical practices and the methodology proposed herein is
+readily adapted to refinements in training data. On the left of Figure 7
+are the improvement in Dice values\textsuperscript{49},
+\added[id=nt]{i.e.,}
 
@@
 
+\section{}\label{section}
+
 The current communications describes a supervised statistical learning
@@
 Two major meta-analyses of WMHs have been published covering the periods
-of 1966-2009 and 2010-2015, respectively. The earlier meta-analysis
-covered 53 longitudinal studies that included samples of high-risk
-populations, i.e., patients selected for a specific disease or condition
-such as hypertension, whereas other studies recruited samples of the
-general population. Longitudinal studies of samples representative of
-the general population are more relevant to the focus of the present
-paper. Debette \& Markus {[}9{]} found that the presence of WMHs was
-related to subsequent cognitive decline, a higher risk of developing
-dementia, stroke, and of mortality. Lesion volume at baseline was also
-predictive of cognitive decline. Limitations of this meta-analysis
-include heterogeneity in the method of measuring WMHs; some studies used
-automated volumetric measurement, whereas others used a visual rating
-scale. The studies analyzed by Debette \& Markus were limited to the
-occurrence of one of the aforementioned conditions which they analyzed
-by hazard ratios.
-
-\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
-{[}10{]} of 23 cross-sectional studies reporting MRI and concurrent
-neuropsychological results in patients with heterogeneous diagnoses but
-without previously diagnosed cognitive impairment, found that WMHs were
-associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
-to -0.08) after controlling for age. These studies also differed in the
-metric used to measure the WMHs, including volume, \% of total
-intracranial volume, and a visual rating score. The effect size for the
-association with cognitive deficit in these cross-sectional studies did
-not differ significantly across various cognitive domains or the method
-of measuring lesion volume. Among eight longitudinal studies analyzed by
-Kloppenborg et al that included a follow-up MRI and also controlled for
-age, the effect size for the association of progression in WMHs and
-cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
-association was stronger for attention and executive function than for
-memory and processing speed. Although baseline WMHs were predictive of
-cognitive deficit at follow-up in the seven studies which did not repeat
-MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
-in the longitudinal studies that calculated progression in WMHs. In
-summary, progression of WMHs seen on repeat MRI has a stronger relation
-to cognitive deficit than concurrent imaging findings. These
+prior to 2010\textsuperscript{9} and after\textsuperscript{10}. The
+earlier meta-analysis covered 53 longitudinal studies that included
+samples of high-risk populations, i.e., patients selected for a specific
+disease or condition such as hypertension, whereas other studies
+recruited samples of the general population. Longitudinal studies of
+samples representative of the general population are more relevant to
+the focus of the present paper. Debette \& Markus\textsuperscript{9}
+found that the presence of WMHs was related to subsequent cognitive
+decline, a higher risk of developing dementia, stroke, and of mortality.
+Lesion volume at baseline was also predictive of cognitive decline.
+Limitations of this meta-analysis include heterogeneity in the method of
+measuring WMHs; some studies used automated volumetric measurement,
+whereas others used a visual rating scale. The studies analyzed by
+Debette \& Markus were limited to the occurrence of one of the
+aforementioned conditions which they analyzed by hazard ratios.
+
+\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et
+al.\textsuperscript{10} of 23 cross-sectional studies reporting MRI and
+concurrent neuropsychological results in patients with heterogeneous
+diagnoses but without previously diagnosed cognitive impairment, found
+that WMHs were associated with cognitive deficit (effect size of -0.10,
+95\% CI: -0.13 to -0.08) after controlling for age. These studies also
+differed in the metric used to measure the WMHs, including volume, \% of
+total intracranial volume, and a visual rating score. The effect size
+for the association with cognitive deficit in these cross-sectional
+studies did not differ significantly across various cognitive domains or
+the method of measuring lesion volume. Among eight longitudinal studies
+analyzed by Kloppenborg et al that included a follow-up MRI and also
+controlled for age, the effect size for the association of progression
+in WMHs and cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09).
+This association was stronger for attention and executive function than
+for memory and processing speed. Although baseline WMHs were predictive
+of cognitive deficit at follow-up in the seven studies which did not
+repeat MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to
+-0.05) than in the longitudinal studies that calculated progression in
+WMHs. In summary, progression of WMHs seen on repeat MRI has a stronger
+relation to cognitive deficit than concurrent imaging findings. These
 meta-analyses support the rationale for repeating an MRI in patients
-younger than 50 years whose initial scan shows WMHs. Despite the
-above-described associations between WMHs, cognitive decline, increased
-risk of developing dementia, and mortality, these lesions receive little
-attention in current clinical workflows. When reported in a standard
-neuroradiologist interpretation, they are typically handled as
-incidental findings and are assigned little clinical significance. This
-likely reflects the impracticality of performing a detailed assessment
-of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
+younger than 50 years whose initial scan shows WMHs.
+
+Despite the above-described associations between WMHs, cognitive
+decline, increased risk of developing dementia, and mortality, these
+lesions receive little attention in current clinical workflows. When
+reported in a standard neuroradiologist interpretation, they are
+typically handled as incidental findings and are assigned little
+clinical significance. This likely reflects the impracticality of
+performing a detailed assessment of number, volume, and distribution
+within a qualitative neuroradiologist interpretation as well as the lack
+of correlative information on how the presence and distribution of these
+lesions may inform a diagnosis and prognosis in the appropriate clinical
+setting. To date, automated or semi-automated tools for the detection of
+WMHs have lacked the specificity and efficiency for the mining of
+large-scale datasets to generate highly granular data on whether these
+lesions possess any true diagnostic or prognostic value in the setting
+of a specific disease process. The present communication describes a
 supervised statistical learning tool that is appropriate for the
@@
 across a wide variety of settings for the rapid identification of
-predictive imaging features {[}50--53{]}. Automobile manufacturers
-utilize these types of approaches to equip self-driving vehicles to
-recognize and respond to unique external surroundings through the
-identification of visual information sufficiently similar to previously
-assimilated training data {[}54, 55{]}. Similarly, in the context of the
-neuroimaging assessments, deep learning approaches may allow for the
-rapid identification of information predictive of disease state in an
-individual patient. These approaches have been applied to the
-segmentation of macroscopically visible structures {[}50--53{]}.
-Additionally, these approaches might be applied to the interrogation of
-imaging data in the individual patient with a primary quantitative
-output metrics to include sequences such as diffusion tensor imaging
-(DTI) and its variants, functional connectivity, perfusion weighted
-imaging, and cortical thickness assessments. At present, these advanced
-neuroimaging sequences are confined to cohort-based research studies due
-to the lack of available analytical tools to assess the information in
-the setting of the individual patient {[}56{]}. Application of deep
-learning approaches in the context of data with primary quantitative
-outputs will require large scale normative and disease specific
-databases. Building these large scale imaging libraries is resource
-intensive and requires a multi-center approach with harmonized scanners
-between sites and correlative non-imaging clinical data. Large scale TBI
-data is becoming increasingly available through activities such as the
-Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research
-and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European
-Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of
-Defense Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and
-other data being consolidated through FITBIR. In concert with any
-available high quality normative neuroimaging data, deep learning
-algorithms may be well positioned to help transform how neuroimaging is
-interpreted for the clinical management of patients with this disease
-process.
+predictive imaging features\textsuperscript{50--53}. Automobile
+manufacturers utilize these types of approaches to equip self-driving
+vehicles to recognize and respond to unique external surroundings
+through the identification of visual information sufficiently similar to
+previously assimilated training data\textsuperscript{54,55}. Similarly,
+in the context of the neuroimaging assessments, deep learning approaches
+may allow for the rapid identification of information predictive of
+disease state in an individual patient. These approaches have been
+applied to the segmentation of macroscopically visible
+structures\textsuperscript{50--53}. Additionally, these approaches might
+be applied to the interrogation of imaging data in the individual
+patient with a primary quantitative output metrics to include sequences
+such as diffusion tensor imaging (DTI) and its variants, functional
+connectivity, perfusion weighted imaging, and cortical thickness
+assessments. At present, these advanced neuroimaging sequences are
+confined to cohort-based research studies due to the lack of available
+analytical tools to assess the information in the setting of the
+individual patient\textsuperscript{56}. Application of deep learning
+approaches in the context of data with primary quantitative outputs will
+require large scale normative and disease specific databases. Building
+these large scale imaging libraries is resource intensive and requires a
+multi-center approach with harmonized scanners between sites and
+correlative non-imaging clinical data. Large scale TBI data is becoming
+increasingly available through activities such as the Chronic Effects of
+Neurotrauma Consortium (CENC), Transforming Research and Clinical
+Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma
+Effectiveness Research in TBI (CENTER-TBI), Department of Defense
+Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and other data
+being consolidated through FITBIR. In concert with any available high
+quality normative neuroimaging data, deep learning algorithms may be
+well positioned to help transform how neuroimaging is interpreted for
+the clinical management of patients with this disease process.
 
@@
 \hypertarget{ref-Bigler:2013aa}{}
-1. Bigler, E. D., Abildskov, T. J., Petrie, J., Farrer, T. J., Dennis,
-M., Simic, N., Taylor, H. G., Rubin, K. H., Vannatta, K., Gerhardt, C.
-A., Stancin, T., and Owen Yeates, K. ``\textbf{Heterogeneity of Brain
-Lesions in Pediatric Traumatic Brain Injury}'' \emph{Neuropsychology}
-27, no. 4 (2013): 438--51.
-doi:\href{https://doi.org/10.1037/a0032837}{10.1037/a0032837}
+1. Bigler ED, Abildskov TJ, Petrie J, Farrer TJ, Dennis M, Simic N,
+Taylor HG, Rubin KH, Vannatta K, Gerhardt CA, et al. Heterogeneity of
+brain lesions in pediatric traumatic brain injury. Neuropsychology.
+2013;27(4):438--51.
 
 \hypertarget{ref-Smitherman:2016aa}{}
-2. Smitherman, E., Hernandez, A., Stavinoha, P. L., Huang, R., Kernie,
-S. G., Diaz-Arrastia, R., and Miles, D. K. ``\textbf{Predicting Outcome
-After Pediatric Traumatic Brain Injury by Early Magnetic Resonance
-Imaging Lesion Location and Volume}'' \emph{J Neurotrauma} 33, no. 1
-(2016): 35--48.
-doi:\href{https://doi.org/10.1089/neu.2014.3801}{10.1089/neu.2014.3801}
+2. Smitherman E, Hernandez A, Stavinoha PL, Huang R, Kernie SG,
+Diaz-Arrastia R, Miles DK. Predicting outcome after pediatric traumatic
+brain injury by early magnetic resonance imaging lesion location and
+volume. J Neurotrauma. 2016;33(1):35--48.
 
 \hypertarget{ref-Marquez-de-la-Plata:2007aa}{}
-3. Marquez de la Plata, C., Ardelean, A., Koovakkattu, D., Srinivasan,
-P., Miller, A., Phuong, V., Harper, C., Moore, C., Whittemore, A.,
-Madden, C., Diaz-Arrastia, R., and Devous, M., Sr. ``\textbf{Magnetic
-Resonance Imaging of Diffuse Axonal Injury: Quantitative Assessment of
-White Matter Lesion Volume}'' \emph{J Neurotrauma} 24, no. 4 (2007):
-591--8.
-doi:\href{https://doi.org/10.1089/neu.2006.0214}{10.1089/neu.2006.0214}
+3. Marquez de la Plata C, Ardelean A, Koovakkattu D, Srinivasan P,
+Miller A, Phuong V, Harper C, Moore C, Whittemore A, Madden C, et al.
+Magnetic resonance imaging of diffuse axonal injury: Quantitative
+assessment of white matter lesion volume. J Neurotrauma.
+2007;24(4):591--8.
 
 \hypertarget{ref-Moen:2014aa}{}
-4. Moen, K. G., Brezova, V., Skandsen, T., Håberg, A. K., Folvik, M.,
-and Vik, A. ``\textbf{Traumatic Axonal Injury: The Prognostic Value of
-Lesion Load in Corpus Callosum, Brain Stem, and Thalamus in Different
-Magnetic Resonance Imaging Sequences}'' \emph{J Neurotrauma} 31, no. 17
-(2014): 1486--96.
-doi:\href{https://doi.org/10.1089/neu.2013.3258}{10.1089/neu.2013.3258}
+4. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A. Traumatic
+axonal injury: The prognostic value of lesion load in corpus callosum,
+brain stem, and thalamus in different magnetic resonance imaging
+sequences. J Neurotrauma. 2014;31(17):1486--96.
 
 \hypertarget{ref-Ding:2008aa}{}
-5. Ding, K., Marquez de la Plata, C., Wang, J. Y., Mumphrey, M., Moore,
-C., Harper, C., Madden, C. J., McColl, R., Whittemore, A., Devous, M.
-D., and Diaz-Arrastia, R. ``\textbf{Cerebral Atrophy After Traumatic
-White Matter Injury: Correlation with Acute Neuroimaging and Outcome}''
-\emph{J Neurotrauma} 25, no. 12 (2008): 1433--40.
-doi:\href{https://doi.org/10.1089/neu.2008.0683}{10.1089/neu.2008.0683}
+5. Ding K, Marquez de la Plata C, Wang JY, Mumphrey M, Moore C, Harper
+C, Madden CJ, McColl R, Whittemore A, Devous MD, et al. Cerebral atrophy
+after traumatic white matter injury: Correlation with acute neuroimaging
+and outcome. J Neurotrauma. 2008;25(12):1433--40.
 
 \hypertarget{ref-Pierallini:2000aa}{}
-6. Pierallini, A., Pantano, P., Fantozzi, L. M., Bonamini, M., Vichi,
-R., Zylberman, R., Pisarri, F., Colonnese, C., and Bozzao, L.
-``\textbf{Correlation Between MRI Findings and Long-Term Outcome in
-Patients with Severe Brain Trauma}'' \emph{Neuroradiology} 42, no. 12
-(2000): 860--7.
+6. Pierallini A, Pantano P, Fantozzi LM, Bonamini M, Vichi R, Zylberman
+R, Pisarri F, Colonnese C, Bozzao L. Correlation between mRI findings
+and long-term outcome in patients with severe brain trauma.
+Neuroradiology. 2000;42(12):860--7.
 
 \hypertarget{ref-Weiss:2008aa}{}
-7. Weiss, N., Galanaud, D., Carpentier, A., Tezenas de Montcel, S.,
-Naccache, L., Coriat, P., and Puybasset, L. ``\textbf{A Combined
-Clinical and MRI Approach for Outcome Assessment of Traumatic Head
-Injured Comatose Patients}'' \emph{J Neurol} 255, no. 2 (2008): 217--23.
-doi:\href{https://doi.org/10.1007/s00415-008-0658-4}{10.1007/s00415-008-0658-4}
+7. Weiss N, Galanaud D, Carpentier A, Tezenas de Montcel S, Naccache L,
+Coriat P, Puybasset L. A combined clinical and mRI approach for outcome
+assessment of traumatic head injured comatose patients. J Neurol.
+2008;255(2):217--23.
 
 \hypertarget{ref-Levin:1988aa}{}
-8. Levin, H. S., Williams, D., Crofford, M. J., High, W. M., Jr,
-Eisenberg, H. M., Amparo, E. G., Guinto, F. C., Jr, Kalisky, Z., Handel,
-S. F., and Goldman, A. M. ``\textbf{Relationship of Depth of Brain
-Lesions to Consciousness and Outcome After Closed Head Injury}'' \emph{J
-Neurosurg} 69, no. 6 (1988): 861--6.
-doi:\href{https://doi.org/10.3171/jns.1988.69.6.0861}{10.3171/jns.1988.69.6.0861}
+8. Levin HS, Williams D, Crofford MJ, High WM Jr, Eisenberg HM, Amparo
+EG, Guinto FC Jr, Kalisky Z, Handel SF, Goldman AM. Relationship of
+depth of brain lesions to consciousness and outcome after closed head
+injury. J Neurosurg. 1988;69(6):861--6.
 
 \hypertarget{ref-Debette:2010aa}{}
-9. Debette, S. and Markus, H. S. ``\textbf{The Clinical Importance of
-White Matter Hyperintensities on Brain Magnetic Resonance Imaging:
-Systematic Review and Meta-Analysis}'' \emph{BMJ} 341, (2010): c3666.
+9. Debette S, Markus HS. The clinical importance of white matter
+hyperintensities on brain magnetic resonance imaging: Systematic review
+and meta-analysis. BMJ. 2010;341:c3666.
 
 \hypertarget{ref-Kloppenborg:2014aa}{}
-10. Kloppenborg, R. P., Nederkoorn, P. J., Geerlings, M. I., and Berg,
-E. van den. ``\textbf{Presence and Progression of White Matter
-Hyperintensities and Cognition: A Meta-Analysis}'' \emph{Neurology} 82,
-no. 23 (2014): 2127--38.
-doi:\href{https://doi.org/10.1212/WNL.0000000000000505}{10.1212/WNL.0000000000000505}
+10. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
+Presence and progression of white matter hyperintensities and cognition:
+A meta-analysis. Neurology. 2014;82(23):2127--38.
 
 \hypertarget{ref-mlmi2015}{}
-11. Available at \url{http://mlmi2015.web.unc.edu}
+11.
 
 \hypertarget{ref-Bauer:2011aa}{}
-12. Bauer, S., Nolte, L.-P., and Reyes, M. ``\textbf{Fully Automatic
-Segmentation of Brain Tumor Images Using Support Vector Machine
-Classification in Combination with Hierarchical Conditional Random Field
-Regularization}'' \emph{Med Image Comput Comput Assist Interv} 14, no.
-Pt 3 (2011): 354--61.
+12. Bauer S, Nolte L-P, Reyes M. Fully automatic segmentation of brain
+tumor images using support vector machine classification in combination
+with hierarchical conditional random field regularization. Med Image
+Comput Comput Assist Interv. 2011;14(Pt 3):354--61.
 
 \hypertarget{ref-Tong:2014aa}{}
-13. Tong, T., Wolz, R., Gao, Q., Guerrero, R., Hajnal, J. V., Rueckert,
-D., and Alzheimer's Disease Neuroimaging Initiative. ``\textbf{Multiple
-Instance Learning for Classification of Dementia in Brain MRI}''
-\emph{Med Image Anal} 18, no. 5 (2014): 808--18.
-doi:\href{https://doi.org/10.1016/j.media.2014.04.006}{10.1016/j.media.2014.04.006}
+13. Tong T, Wolz R, Gao Q, Guerrero R, Hajnal JV, Rueckert D,
+Alzheimer's Disease Neuroimaging Initiative. Multiple instance learning
+for classification of dementia in brain mRI. Med Image Anal.
+2014;18(5):808--18.
 
 \hypertarget{ref-Liu:2013aa}{}
-14. Liu, X., Tosun, D., Weiner, M. W., Schuff, N., and Alzheimer's
-Disease Neuroimaging Initiative. ``\textbf{Locally Linear Embedding
-(LLE) for MRI Based Alzheimer's Disease Classification}''
-\emph{Neuroimage} 83, (2013): 148--57.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2013.06.033}{10.1016/j.neuroimage.2013.06.033}
+14. Liu X, Tosun D, Weiner MW, Schuff N, Alzheimer's Disease
+Neuroimaging Initiative. Locally linear embedding (lLE) for mRI based
+alzheimer's disease classification. Neuroimage. 2013;83:148--57.
 
 \hypertarget{ref-Van-Horn:2014aa}{}
-15. Van Horn, J. D. and Toga, A. W. ``\textbf{Human Neuroimaging as a
-`Big Data' Science}'' \emph{Brain Imaging Behav} 8, no. 2 (2014):
-323--31.
-doi:\href{https://doi.org/10.1007/s11682-013-9255-y}{10.1007/s11682-013-9255-y}
+15. Van Horn JD, Toga AW. Human neuroimaging as a ``big data'' science.
+Brain Imaging Behav. 2014;8(2):323--31.
 
 \hypertarget{ref-scipy}{}
-16. Jones, E., Oliphant, T., Peterson, P., and others. ``\textbf{SciPy:
-Open Source Scientific Tools for Python}'' (2001--2001-\/-): Available
-at \url{http://www.scipy.org/}
+16. Jones E, Oliphant T, Peterson P, others. SciPy: Open source
+scientific tools for Python. 2001--2001-\/-.
 
 \hypertarget{ref-R}{}
-17. R Core Team. ``\textbf{R: A Language and Environment for Statistical
-Computing}'' (2016):
+17. R Core Team. R: A Language and Environment for Statistical
+Computing. Vienna, Austria: R Foundation for Statistical Computing;
+2016.
 
 \hypertarget{ref-breiman2001}{}
-18. Breiman, L. ``\textbf{Random Forests}'' \emph{Machine learning}
-(2001): 5--32.
+18. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
 
 \hypertarget{ref-yi2009}{}
-19. Yi, Z., Criminisi, A., Shotton, J., and Blake, A.
-``\textbf{Discriminative, Semantic Segmentation of Brain Tissue in MR
-Images}'' \emph{Med Image Comput Comput Assist Interv} 12, no. Pt 2
-(2009): 558--65.
+19. Yi Z, Criminisi A, Shotton J, Blake A. Discriminative, semantic
+segmentation of brain tissue in MR images. Med Image Comput Comput
+Assist Interv. 2009;12(Pt 2):558--65.
 
 \hypertarget{ref-viola2005}{}
-20. Viola, P., Jones, M., and Snow, D. ``\textbf{Detecting Pedestrians
-Using Patterns of Motion and Appearance}'' \emph{International Journal
-of Computer Vision} 63, (2005): 153--161.
+20. Viola P, Jones M, Snow D. Detecting pedestrians using patterns of
+motion and appearance. International Journal of Computer Vision.
+2005;63:153--161.
 
 \hypertarget{ref-geremia2011}{}
-21. Geremia, E., Clatz, O., Menze, B. H., Konukoglu, E., Criminisi, A.,
-and Ayache, N. ``\textbf{Spatial Decision Forests for MS Lesion
-Segmentation in Multi-Channel Magnetic Resonance Images}''
-\emph{Neuroimage} 57, no. 2 (2011): 378--90.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2011.03.080}{10.1016/j.neuroimage.2011.03.080}
+21. Geremia E, Clatz O, Menze BH, Konukoglu E, Criminisi A, Ayache N.
+Spatial decision forests for MS lesion segmentation in multi-channel
+magnetic resonance images. Neuroimage. 2011;57(2):378--90.
 
 \hypertarget{ref-Pustina:2016aa}{}
-22. Pustina, D., Coslett, H. B., Turkeltaub, P. E., Tustison, N.,
-Schwartz, M. F., and Avants, B. ``\textbf{Automated Segmentation of
-Chronic Stroke Lesions Using LINDA: Lesion Identification with
-Neighborhood Data Analysis}'' \emph{Hum Brain Mapp} (2016):
-doi:\href{https://doi.org/10.1002/hbm.23110}{10.1002/hbm.23110}
+22. Pustina D, Coslett HB, Turkeltaub PE, Tustison N, Schwartz MF,
+Avants B. Automated segmentation of chronic stroke lesions using lINDA:
+Lesion identification with neighborhood data analysis. Hum Brain Mapp.
+2016 Jan.
 
 \hypertarget{ref-geremia2012}{}
-23. Geremia, E., Menze, B. H., and Ayache, N. ``\textbf{Spatial Decision
-Forests for Glioma Segmentation in Multi-Channel MR Images}''
-\emph{Proceedings of MICCAI-BRATS 2012} (2012):
+23. Geremia E, Menze BH, Ayache N. Spatial decision forests for glioma
+segmentation in multi-channel MR images. In: Proceedings of MICCAI-BRATS
+2012. 2012.
 
 \hypertarget{ref-bauer2012}{}
-24. Bauer, S., Fejes, T., Slotboom, J., Wiest, R., Nolte, L.-P., and
-Reyes, M. ``\textbf{Segmentation of Brain Tumor Images Based on
-Integrated Hierarchical Classification and Regularization}''
-\emph{Proceedings of MICCAI-BRATS 2012} (2012): 10--13.
+24. Bauer S, Fejes T, Slotboom J, Wiest R, Nolte L-P, Reyes M.
+Segmentation of brain tumor images based on integrated hierarchical
+classification and regularization. In: Proceedings of MICCAI-BRATS 2012.
+2012. pp. 10--13.
 
 \hypertarget{ref-zikic2012}{}
-25. Zikic, D., Glocker, B., Konukoglu, E., Shotton, J., Criminisi, A.,
-Ye, D. H., Demiralp, C., Thomas, O. M., Das, T., Jena, R., and Price, S.
-J. ``\textbf{Context-Sensitive Classification Forests for Segmentation
-of Brain Tumor Tissues}'' \emph{Proceedings of MICCAI-BRATS 2012}
-(2012): 1--9.
+25. Zikic D, Glocker B, Konukoglu E, Shotton J, Criminisi A, Ye DH,
+Demiralp C, Thomas OM, Das T, Jena R, et al. Context-sensitive
+classification forests for segmentation of brain tumor tissues. In:
+Proceedings of MICCAI-BRATS 2012. 2012. pp. 1--9.
 
 \hypertarget{ref-Tustison:2015aa}{}
-26. Tustison, N. J., Shrinidhi, K. L., Wintermark, M., Durst, C. R.,
-Kandel, B. M., Gee, J. C., Grossman, M. C., and Avants, B. B.
-``\textbf{Optimal Symmetric Multimodal Templates and Concatenated Random
-Forests for Supervised Brain Tumor Segmentation (Simplified) with
-ANTsR}'' \emph{Neuroinformatics} 13, no. 2 (2015): 209--25.
-doi:\href{https://doi.org/10.1007/s12021-014-9245-2}{10.1007/s12021-014-9245-2}
+26. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee
+JC, Grossman MC, Avants BB. Optimal symmetric multimodal templates and
+concatenated random forests for supervised brain tumor segmentation
+(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.
 
 \hypertarget{ref-Menze:2015aa}{}
-27. Menze, B. H., Jakab, A., Bauer, S., Kalpathy-Cramer, J., Farahani,
-K., Kirby, J., Burren, Y., Porz, N., Slotboom, J., Wiest, R., Lanczi,
-L., Gerstner, E., Weber, M.-A., Arbel, T., Avants, B. B., Ayache, N.,
-Buendia, P., Collins, D. L., Cordier, N., Corso, J. J., Criminisi, A.,
-Das, T., Delingette, H., Demiralp, Ç., Durst, C. R., Dojat, M., Doyle,
-S., Festa, J., Forbes, F., Geremia, E., Glocker, B., Golland, P., Guo,
-X., Hamamci, A., Iftekharuddin, K. M., Jena, R., John, N. M., Konukoglu,
-E., Lashkari, D., Mariz, J. A., Meier, R., Pereira, S., Precup, D.,
-Price, S. J., Raviv, T. R., Reza, S. M. S., Ryan, M., Sarikaya, D.,
-Schwartz, L., Shin, H.-C., Shotton, J., Silva, C. A., Sousa, N.,
-Subbanna, N. K., Szekely, G., Taylor, T. J., Thomas, O. M., Tustison, N.
-J., Unal, G., Vasseur, F., Wintermark, M., Ye, D. H., Zhao, L., Zhao,
-B., Zikic, D., Prastawa, M., Reyes, M., and Van Leemput, K.
-``\textbf{The Multimodal Brain Tumor Image Segmentation Benchmark
-(BRATS)}'' \emph{IEEE Trans Med Imaging} 34, no. 10 (2015): 1993--2024.
-doi:\href{https://doi.org/10.1109/TMI.2014.2377694}{10.1109/TMI.2014.2377694}
+27. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
+Burren Y, Porz N, Slotboom J, Wiest R, et al. The multimodal brain tumor
+image segmentation benchmark (bRATS). IEEE Trans Med Imaging.
+2015;34(10):1993--2024.
 
 \hypertarget{ref-schapire1990}{}
-28. Schapire, R. ``\textbf{The Strength of Weak Learnability}''
-\emph{Machine Learning} 5, (1990): 197--227.
+28. Schapire R. The strength of weak learnability. Machine Learning.
+1990;5:197--227.
 
 \hypertarget{ref-freund1997}{}
-29. Freund, Y. and Schapire, R. ``\textbf{A Decision-Theoretic
-Generalization of on-Line Learning and an Application to Boosting}''
-\emph{Journal of Computer and System Sciences} 55, (1997): 119--139.
+29. Freund Y, Schapire R. A decision-theoretic generalization of on-line
+learning and an application to boosting. Journal of Computer and System
+Sciences. 1997;55:119--139.
 
 \hypertarget{ref-ho1995}{}
-30. Ho, T. K. ``\textbf{Random Decision Forests}'' \emph{Document
-analysis and recognition, 1995., proceedings of the third international
-conference on} 1, (1995): 278--282 vol.1.
-doi:\href{https://doi.org/10.1109/ICDAR.1995.598994}{10.1109/ICDAR.1995.598994}
+30. Ho TK. Random decision forests. In: Document analysis and
+recognition, 1995., proceedings of the third international conference
+on. Vol. 1. 1995. pp. 278--282 vol.1.
 
 \hypertarget{ref-amit1997}{}
-31. Amit, Y. and Geman, D. ``\textbf{Shape Quantization and Recognition
-with Randomized Trees}'' \emph{Neural Computation} 9, (1997):
-1545--1588.
+31. Amit Y, Geman D. Shape quantization and recognition with randomized
+trees. Neural Computation. 1997;9:1545--1588.
 
 \hypertarget{ref-Neema:2009aa}{}
-32. Neema, M., Guss, Z. D., Stankiewicz, J. M., Arora, A., Healy, B. C.,
-and Bakshi, R. ``\textbf{Normal Findings on Brain Fluid-Attenuated
-Inversion Recovery MR Images at 3T}'' \emph{AJNR Am J Neuroradiol} 30,
-no. 5 (2009): 911--6.
-doi:\href{https://doi.org/10.3174/ajnr.A1514}{10.3174/ajnr.A1514}
+32. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
+Normal findings on brain fluid-attenuated inversion recovery mR images
+at 3T. AJNR Am J Neuroradiol. 2009;30(5):911--6.
 
 \hypertarget{ref-Avants:2014aa}{}
-33. Avants, B. B., Tustison, N. J., Stauffer, M., Song, G., Wu, B., and
-Gee, J. C. ``\textbf{The Insight ToolKit Image Registration Framework}''
-\emph{Front Neuroinform} 8, (2014): 44.
-doi:\href{https://doi.org/10.3389/fninf.2014.00044}{10.3389/fninf.2014.00044}
+33. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The
+Insight ToolKit image registration framework. Front Neuroinform.
+2014;8:44.
 
 \hypertarget{ref-Tustison:2010ac}{}
-34. Tustison, N. J., Avants, B. B., Cook, P. A., Zheng, Y., Egan, A.,
-Yushkevich, P. A., and Gee, J. C. ``\textbf{N4ITK: Improved N3 Bias
-Correction}'' \emph{IEEE Trans Med Imaging} 29, no. 6 (2010): 1310--20.
-doi:\href{https://doi.org/10.1109/TMI.2010.2046908}{10.1109/TMI.2010.2046908}
+34. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
+JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
+2010;29(6):1310--20.
 
 \hypertarget{ref-Manjon:2010aa}{}
-35. Manjón, J. V., Coupé, P., Martí-Bonmatí, L., Collins, D. L., and
-Robles, M. ``\textbf{Adaptive Non-Local Means Denoising of MR Images
-with Spatially Varying Noise Levels}'' \emph{J Magn Reson Imaging} 31,
-no. 1 (2010): 192--203.
-doi:\href{https://doi.org/10.1002/jmri.22003}{10.1002/jmri.22003}
+35. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
+non-local means denoising of mR images with spatially varying noise
+levels. J Magn Reson Imaging. 2010;31(1):192--203.
 
 \hypertarget{ref-Tustison:2014ab}{}
-36. Tustison, N. J., Cook, P. A., Klein, A., Song, G., Das, S. R., Duda,
-J. T., Kandel, B. M., Strien, N. van, Stone, J. R., Gee, J. C., and
-Avants, B. B. ``\textbf{Large-Scale Evaluation of ANTs and FreeSurfer
-Cortical Thickness Measurements}'' \emph{Neuroimage} 99, (2014):
-166--79.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2014.05.044}{10.1016/j.neuroimage.2014.05.044}
+36. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
+Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
+and freeSurfer cortical thickness measurements. Neuroimage.
+2014;99:166--79.
 
 \hypertarget{ref-nyul2000}{}
-37. Nyúl, L. G., Udupa, J. K., and Zhang, X. ``\textbf{New Variants of a
-Method of MRI Scale Standardization}'' \emph{IEEE Trans Med Imaging} 19,
-no. 2 (2000): 143--50.
-doi:\href{https://doi.org/10.1109/42.836373}{10.1109/42.836373}
+37. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
+standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
 
 \hypertarget{ref-Yushkevich:2006aa}{}
-38. Yushkevich, P. A., Piven, J., Hazlett, H. C., Smith, R. G., Ho, S.,
-Gee, J. C., and Gerig, G. ``\textbf{User-Guided 3D Active Contour
-Segmentation of Anatomical Structures: Significantly Improved Efficiency
-and Reliability}'' \emph{Neuroimage} 31, no. 3 (2006): 1116--28.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2006.01.015}{10.1016/j.neuroimage.2006.01.015}
+38. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
+User-guided 3D active contour segmentation of anatomical structures:
+Significantly improved efficiency and reliability. Neuroimage.
+2006;31(3):1116--28.
 
 \hypertarget{ref-Avants:2011aa}{}
-39. Avants, B. B., Tustison, N. J., Wu, J., Cook, P. A., and Gee, J. C.
-``\textbf{An Open Source Multivariate Framework for \(n\)-Tissue
-Segmentation with Evaluation on Public Data}'' \emph{Neuroinformatics}
-9, no. 4 (2011): 381--400.
-doi:\href{https://doi.org/10.1007/s12021-011-9109-y}{10.1007/s12021-011-9109-y}
+39. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+multivariate framework for \(n\)-tissue segmentation with evaluation on
+public data. Neuroinformatics. 2011;9(4):381--400.
 
 \hypertarget{ref-landman2011}{}
-40. Landman, B. A., Huang, A. J., Gifford, A., Vikram, D. S., Lim, I. A.
-L., Farrell, J. A. D., Bogovic, J. A., Hua, J., Chen, M., Jarso, S.,
-Smith, S. A., Joel, S., Mori, S., Pekar, J. J., Barker, P. B., Prince,
-J. L., and Zijl, P. C. M. van. ``\textbf{Multi-Parametric Neuroimaging
-Reproducibility: A 3-T Resource Study}'' \emph{Neuroimage} 54, no. 4
-(2011): 2854--66.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2010.11.047}{10.1016/j.neuroimage.2010.11.047}
+40. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
+reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
 
 \hypertarget{ref-maurer2003}{}
-41. Maurer, C. R., Rensheng, Q., and Raghavan, V. ``\textbf{A Linear
-Time Algorithm for Computing Exact Euclidean Distance Transforms of
-Binary Images in Arbitrary Dimensions}'' \emph{Pattern Analysis and
-Machine Intelligence, IEEE Transactions on} 25, no. 2 (2003): 265--270.
-doi:\href{https://doi.org/10.1109/TPAMI.2003.1177156}{10.1109/TPAMI.2003.1177156}
+41. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
+computing exact Euclidean distance transforms of binary images in
+arbitrary dimensions. Pattern Analysis and Machine Intelligence, IEEE
+Transactions on. 2003;25(2):265--270.
 
 \hypertarget{ref-Tustison:2013ac}{}
-42. Tustison, N. J. and Avants, B. B. ``\textbf{Explicit B-Spline
-Regularization in Diffeomorphic Image Registration}'' \emph{Front
-Neuroinform} 7, (2013): 39.
-doi:\href{https://doi.org/10.3389/fninf.2013.00039}{10.3389/fninf.2013.00039}
+42. Tustison NJ, Avants BB. Explicit B-spline regularization in
+diffeomorphic image registration. Front Neuroinform. 2013;7:39.
 
 \hypertarget{ref-Avants:2011ab}{}
-43. Avants, B. B., Tustison, N. J., Song, G., Cook, P. A., Klein, A.,
-and Gee, J. C. ``\textbf{A Reproducible Evaluation of ANTs Similarity
-Metric Performance in Brain Image Registration}'' \emph{Neuroimage} 54,
-no. 3 (2011): 2033--44.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2010.09.025}{10.1016/j.neuroimage.2010.09.025}
+43. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+reproducible evaluation of ANTs similarity metric performance in brain
+image registration. Neuroimage. 2011;54(3):2033--44.
 
 \hypertarget{ref-Anbeek:2004aa}{}
-44. Anbeek, P., Vincken, K. L., Osch, M. J. P. van, Bisschops, R. H. C.,
-and Grond, J. van der. ``\textbf{Probabilistic Segmentation of White
-Matter Lesions in MR Imaging}'' \emph{Neuroimage} 21, no. 3 (2004):
-1037--44.
-doi:\href{https://doi.org/10.1016/j.neuroimage.2003.10.012}{10.1016/j.neuroimage.2003.10.012}
+44. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+Probabilistic segmentation of white matter lesions in mR imaging.
+Neuroimage. 2004;21(3):1037--44.
 
 \hypertarget{ref-Garcia-Lorenzo:2013aa}{}
-45. García-Lorenzo, D., Francis, S., Narayanan, S., Arnold, D. L., and
-Collins, D. L. ``\textbf{Review of Automatic Segmentation Methods of
-Multiple Sclerosis White Matter Lesions on Conventional Magnetic
-Resonance Imaging}'' \emph{Med Image Anal} 17, no. 1 (2013): 1--18.
-doi:\href{https://doi.org/10.1016/j.media.2012.09.004}{10.1016/j.media.2012.09.004}
+45. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
+Review of automatic segmentation methods of multiple sclerosis white
+matter lesions on conventional magnetic resonance imaging. Med Image
+Anal. 2013;17(1):1--18.
 
 \hypertarget{ref-liaw2002}{}
-46. Liaw, A. and Wiener, M. ``\textbf{Classification and Regression by
-randomForest}'' \emph{R News} 2/3, (2002): 18--22.
+46. Liaw A, Wiener M. Classification and regression by randomForest. R
+News. 2002;2/3:18--22.
 
 \hypertarget{ref-Grimaud:1996aa}{}
-47. Grimaud, J., Lai, M., Thorpe, J., Adeleine, P., Wang, L., Barker, G.
-J., Plummer, D. L., Tofts, P. S., McDonald, W. I., and Miller, D. H.
-``\textbf{Quantification of MRI Lesion Load in Multiple Sclerosis: A
-Comparison of Three Computer-Assisted Techniques}'' \emph{Magn Reson
-Imaging} 14, no. 5 (1996): 495--505.
+47. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
+DL, Tofts PS, McDonald WI, Miller DH. Quantification of mRI lesion load
+in multiple sclerosis: A comparison of three computer-assisted
+techniques. Magn Reson Imaging. 1996;14(5):495--505.
 
 \hypertarget{ref-styner2008}{}
-48. ``\textbf{Special Issue on 2008 MICCAI Workshop - MS Lesion
-Segmentation}'' (2008):
+48. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
+S, Jewells V, Warfield S, editors. Special Issue on 2008 MICCAI Workshop
+- MS Lesion Segmentation. MIDAS J; 2008.
 
 \hypertarget{ref-tustison2009}{}
-49. Tustison, N. J. and Gee, J. C. ``\textbf{Introducing Dice, Jaccard,
-and Other Label Overlap Measures to ITK}'' \emph{Insight Journal}
-(2009):
+49. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
+overlap measures to ITK. Insight Journal. 2009.
 
 \hypertarget{ref-Plis:2014aa}{}
-50. Plis, S. M., Hjelm, D. R., Salakhutdinov, R., Allen, E. A.,
-Bockholt, H. J., Long, J. D., Johnson, H. J., Paulsen, J. S., Turner, J.
-A., and Calhoun, V. D. ``\textbf{Deep Learning for Neuroimaging: A
-Validation Study}'' \emph{Front Neurosci} 8, (2014): 229.
-doi:\href{https://doi.org/10.3389/fnins.2014.00229}{10.3389/fnins.2014.00229}
+50. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
+Johnson HJ, Paulsen JS, Turner JA, Calhoun VD. Deep learning for
+neuroimaging: A validation study. Front Neurosci. 2014;8:229.
 
 \hypertarget{ref-Suk:2015aa}{}
-51. Suk, H.-I., Lee, S.-W., Shen, D., and Alzheimer's Disease
-Neuroimaging Initiative. ``\textbf{Deep Sparse Multi-Task Learning for
-Feature Selection in Alzheimer's Disease Diagnosis}'' \emph{Brain Struct
-Funct} (2015):
-doi:\href{https://doi.org/10.1007/s00429-015-1059-y}{10.1007/s00429-015-1059-y}
+51. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
+Initiative. Deep sparse multi-task learning for feature selection in
+alzheimer's disease diagnosis. Brain Struct Funct. 2015 May.
 
 \hypertarget{ref-Li:2014aa}{}
-52. Li, R., Zhang, W., Suk, H.-I., Wang, L., Li, J., Shen, D., and Ji,
-S. ``\textbf{Deep Learning Based Imaging Data Completion for Improved
-Brain Disease Diagnosis}'' \emph{Med Image Comput Comput Assist Interv}
-17, no. Pt 3 (2014): 305--12.
+52. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
+based imaging data completion for improved brain disease diagnosis. Med
+Image Comput Comput Assist Interv. 2014;17(Pt 3):305--12.
 
 \hypertarget{ref-Liu:2015aa}{}
-53. Liu, S., Liu, S., Cai, W., Che, H., Pujol, S., Kikinis, R., Feng,
-D., Fulham, M. J., and ADNI. ``\textbf{Multimodal Neuroimaging Feature
-Learning for Multiclass Diagnosis of Alzheimer's Disease}'' \emph{IEEE
-Trans Biomed Eng} 62, no. 4 (2015): 1132--40.
-doi:\href{https://doi.org/10.1109/TBME.2014.2372011}{10.1109/TBME.2014.2372011}
+53. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
+ADNI. Multimodal neuroimaging feature learning for multiclass diagnosis
+of alzheimer's disease. IEEE Trans Biomed Eng. 2015;62(4):1132--40.
 
 \hypertarget{ref-HadsellSBESKML09}{}
-54. Hadsell, R., Sermanet, P., Ben, J., Erkan, A., Scoffier, M.,
-Kavukcuoglu, K., Muller, U., and LeCun, Y. ``\textbf{Learning Long-Range
-Vision for Autonomous Off-Road Driving}'' \emph{J. Field Robotics} 26,
-no. 2 (2009): 120--144.
-doi:\href{https://doi.org/10.1002/rob.20276}{10.1002/rob.20276},
-Available at \url{http://dx.doi.org/10.1002/rob.20276}
+54. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
+Muller U, LeCun Y. Learning long-range vision for autonomous off-road
+driving. J. Field Robotics. 2009;26(2):120--144.
 
 \hypertarget{ref-FarabetCNL12}{}
-55. Farabet, C., Couprie, C., Najman, L., and LeCun, Y. ``\textbf{Scene
-Parsing with Multiscale Feature Learning, Purity Trees, and Optimal
-Covers}'' \emph{Proceedings of the 29th international conference on
-machine learning, ICML 2012, edinburgh, scotland, uK, june 26 - july 1,
-2012} (2012): Available at \url{http://icml.cc/discuss/2012/319.html}
+55. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
+multiscale feature learning, purity trees, and optimal covers. In:
+Proceedings of the 29th international conference on machine learning,
+ICML 2012, edinburgh, scotland, uK, june 26 - july 1, 2012. icml.cc /
+Omnipress; 2012.
 
 \hypertarget{ref-Mayer:2014aa}{}
-56. Mayer, A. R., Bedrick, E. J., Ling, J. M., Toulouse, T., and Dodd,
-A. ``\textbf{Methods for Identifying Subject-Specific Abnormalities in
-Neuroimaging Data}'' \emph{Hum Brain Mapp} 35, no. 11 (2014): 5457--70.
-doi:\href{https://doi.org/10.1002/hbm.22563}{10.1002/hbm.22563}
+56. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
+identifying subject-specific abnormalities in neuroimaging data. Hum
+Brain Mapp. 2014;35(11):5457--70.
 

### 2016-02-18 README.R

@@ -7,6 +7,7 @@ stitchedFile <- "stitched.Rmd";
 formatFile <- "format.Rmd"
 
 rmdFiles <- c( formatFile,
+               "abstract.Rmd",
                "intro.Rmd",
                "methods.Rmd",
                "results.Rmd",

### 2016-02-25 acknowledgementsDisclaimers.Rmd
Final submission.

## Acknowledgements

The authors wish to acknowledge all other members of
the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
and Kevin Wilson. We would also like to thank the study participants and
their family members. This project was jointly supported by the
Department of Defense (W81XWH-13-2-0095), the U.S. Department of
Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
Grant HU 0001-08-0001.

## Declaration of Interest/Disclaimer

The authors report no financial
disclosures or conflicts of interest. The views expressed here are those
of the authors and do not necessarily reflect the official policy of
position of the Department of the Navy, Department of Defense, nor the
U.S. Government. This work was prepared as a part of official duties;
Title 17 USC §105 provides that Copyright protection under this title is
not available for any work of the U.S. Government. Title 17 USC §101
defines a US Government work as a work prepared by a military service
member of employee of the US Government as part of that person’s
official duties.

\clearpage

<!-- \listofchanges

\clearpage -->

# References

### 2016-02-25 discussion.Rmd
Final submission.

--- 2016-02-11 discussion.Rmd
@@
 
-<!--
-Nick could you add some detail here on how this technique compares to other automated or semi automated approaches for WMH detection?
--->
+<!-- Nick could you add some detail here on how this technique compares
+to other automated or semi automated approaches for WMH detection? -->
 
@@
 
-The current communications describes a supervised statistical learning methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation of WMHs from FLAIR images to help build two-stage ensembles of decision trees for the automated identification of these lesions.
-Although
-only a single expert was used to produce the manual labelings, our intent is to further
-refine the proposed paradigm by crowdsourcing with feedback from other experts who interact with both the data and methodology.  Also, we recognize that only a single site
-was used for evaluating the proposed framework.  However, we are currently processing other site data with the models developed for this work and the results look promising
-since the developed features are site-agnostic.
+The current communications describes a supervised statistical learning
+methodology for identifying WHMs within multimodal MR brain imaging.
+This effort utilized information acquired from the manual segmentation
+of WMHs from FLAIR images to help build two-stage ensembles of decision
+trees for the automated identification of these lesions. Although only a
+single expert was used to produce the manual labelings, our intent is to
+further refine the proposed paradigm by crowdsourcing with feedback from
+other experts who interact with both the data and methodology.  Also, we
+recognize that only a single site was used for evaluating the proposed
+framework.  However, we are currently processing other site data with
+the models developed for this work and the results look promising since
+the developed features are site-agnostic.
 
-As far as we know, this is the first report utilizing a novel random forest approach to identify WMHs in a cohort of TBI patients.
-TBI WMHs tend to be more difficult to segment than MS lesions as the former tend to be smaller with an overall smaller lesion load.
-Also, enhancement protocols with the former tend to be less successful than with the latter.
-As mentioned previously, the work in MS lesion segmentation is extensive
-with a handful of techniques being publicly available.
-Our framework is also available as open-source as part of well-known neuroimaging tools which easily allows for additions/extensions but is also, as far as we know, the first random forest-based technique available for such application.
+As far as we know, this is the first report utilizing a novel random
+forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
+tend to be more difficult to segment than MS lesions as the former tend
+to be smaller with an overall smaller lesion load.  Also, enhancement
+protocols with the former tend to be less successful than with the
+latter.  As mentioned previously, the work in MS lesion segmentation is
+extensive with a handful of techniques being publicly available.  Our
+framework is also available as open-source as part of well-known
+neuroimaging tools which easily allows for additions/extensions but is
+also, as far as we know, the first random forest-based technique
+available for such application.
 
-Two major meta-analyses of WMHs have been published covering the periods of 1966-2009 and 2010-2015, respectively.
-The earlier meta-analysis covered 53 longitudinal studies that included samples of high-risk populations, i.e., patients selected for a specific disease or condition such as hypertension, whereas other studies recruited samples of the general population.
-Longitudinal studies of samples representative of the general population are more relevant to the focus of the present paper.
-Debette & Markus [@Debette:2010aa] found that the presence of WMHs was related to subsequent cognitive decline, a higher risk of developing dementia, stroke, and of mortality.
-Lesion volume at baseline was also predictive of cognitive decline.
-Limitations of this meta-analysis include heterogeneity in the method of measuring WMHs; some studies used automated volumetric measurement, whereas others used a visual rating scale.
-The studies analyzed by Debette & Markus were limited to the occurrence of one of the aforementioned conditions which they analyzed by hazard ratios.
+Two major meta-analyses of WMHs have been published covering the periods
+prior to 2010 [@Debette:2010aa] and after [@Kloppenborg:2014aa]. The earlier meta-analysis
+covered 53 longitudinal studies that included samples of high-risk
+populations, i.e., patients selected for a specific disease or condition
+such as hypertension, whereas other studies recruited samples of the
+general population. Longitudinal studies of samples representative of
+the general population are more relevant to the focus of the present
+paper. Debette & Markus [@Debette:2010aa] found that the presence of
+WMHs was related to subsequent cognitive decline, a higher risk of
+developing dementia, stroke, and of mortality. Lesion volume at baseline
+was also predictive of cognitive decline. Limitations of this
+meta-analysis include heterogeneity in the method of measuring WMHs;
+some studies used automated volumetric measurement, whereas others used
+a visual rating scale. The studies analyzed by Debette & Markus were
+limited to the occurrence of one of the aforementioned conditions which
+they analyzed by hazard ratios.
 
-\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and concurrent neuropsychological results in patients with heterogeneous diagnoses but without previously diagnosed cognitive impairment, found that WMHs were associated with cognitive deficit (effect size of -0.10, 95% CI: -0.13 to -0.08) after controlling for age.
-These studies also differed in the metric used to measure the WMHs, including volume, % of total intracranial volume, and a visual rating score.
-The effect size for the association with cognitive deficit in these cross-sectional studies did not differ significantly across various cognitive domains or the method of measuring lesion volume.
-Among eight longitudinal studies analyzed by Kloppenborg et al that included a follow-up MRI and also controlled for age, the effect size for the association of progression in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09).
-This association was stronger for attention and executive function than for memory and processing speed.
-Although baseline WMHs were predictive of cognitive deficit at follow-up in the seven studies which did not repeat MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in the longitudinal studies that calculated progression in WMHs.
-In summary, progression of WMHs seen on repeat MRI has a stronger relation to cognitive deficit than concurrent imaging findings.
-These meta-analyses support the rationale for repeating an MRI in patients younger than 50 years whose initial scan shows WMHs.
-Despite the above-described associations between WMHs, cognitive decline, increased risk of developing dementia, and mortality, these lesions receive little attention in current clinical workflows.
-When reported in a standard neuroradiologist interpretation, they are typically handled as incidental findings and are assigned little clinical significance.
-This likely reflects the impracticality of performing a detailed assessment of number, volume, and distribution within a qualitative neuroradiologist interpretation as well as the lack of correlative information on how the presence and distribution of these lesions may inform a diagnosis and prognosis in the appropriate clinical setting.
-To date, automated or semi-automated tools for the detection of WMHs have lacked the specificity and efficiency for the mining of large-scale datasets to generate highly granular data on whether these lesions possess any true diagnostic or prognostic value in the setting of a specific disease process.
-The present communication describes a supervised statistical learning tool that is appropriate for the application to such large-scale datasets.
+The more recent meta-analysis by Kloppenborg et al.
+[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
+concurrent neuropsychological results in patients with heterogeneous
+diagnoses but without previously diagnosed cognitive impairment, found
+that WMHs were associated with cognitive deficit (effect size of -0.10,
+95% CI: -0.13 to -0.08) after controlling for age. These studies also
+differed in the metric used to measure the WMHs, including volume, % of
+total intracranial volume, and a visual rating score. The effect size
+for the association with cognitive deficit in these cross-sectional
+studies did not differ significantly across various cognitive domains or
+the method of measuring lesion volume.  Among eight longitudinal studies
+analyzed by Kloppenborg et al that included a follow-up MRI and also
+controlled for age, the effect size for the association of progression
+in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09). This
+association was stronger for attention and executive function than for
+memory and processing speed. Although baseline WMHs were predictive of
+cognitive deficit at follow-up in the seven studies which did not repeat
+MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in
+the longitudinal studies that calculated progression in WMHs. In
+summary, progression of WMHs seen on repeat MRI has a stronger relation
+to cognitive deficit than concurrent imaging findings. These
+meta-analyses support the rationale for repeating an MRI in patients
+younger than 50 years whose initial scan shows WMHs.
 
-The currently described tool is just one example of how "supervised learning" algorithms might be applied to aid in the diagnosis of TBI and other disease processes through the specific identification of features predictive of a given disease state.
-It is an important demonstration of the potential power of these analytical approaches in the rapid but comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide variety of settings for the rapid identification of predictive imaging features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Automobile manufacturers utilize these types of approaches to equip self-driving vehicles to recognize and respond to unique external surroundings through the identification of visual information sufficiently similar to previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning approaches may allow for the rapid identification of information predictive of disease state in an individual patient.
-These approaches have been applied to the segmentation of macroscopically visible structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of imaging data in the individual patient with a primary quantitative output metrics to include sequences such as diffusion tensor imaging (DTI) and its variants, functional connectivity, perfusion weighted imaging, and cortical thickness assessments.
-At present, these advanced neuroimaging sequences are confined to cohort-based research studies due to the lack of available analytical tools to assess the information in the setting of the individual patient [@Mayer:2014aa].
-Application of deep learning approaches in the context of data with primary quantitative outputs will require large scale normative and disease specific databases.
-Building these large scale imaging libraries is resource intensive and requires a multi-center approach with harmonized scanners between sites and correlative non-imaging clinical data.
-Large scale TBI data is becoming increasingly available through activities such as the Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data, deep learning algorithms may be well positioned to help transform how neuroimaging is interpreted for the clinical management of patients with this disease process.
+Despite the
+above-described associations between WMHs, cognitive decline, increased
+risk of developing dementia, and mortality, these lesions receive little
+attention in current clinical workflows. When reported in a standard
+neuroradiologist interpretation, they are typically handled as
+incidental findings and are assigned little clinical significance. This
+likely reflects the impracticality of performing a detailed assessment
+of number, volume, and distribution within a qualitative
+neuroradiologist interpretation as well as the lack of correlative
+information on how the presence and distribution of these lesions may
+inform a diagnosis and prognosis in the appropriate clinical setting. To
+date, automated or semi-automated tools for the detection of WMHs have
+lacked the specificity and efficiency for the mining of large-scale
+datasets to generate highly granular data on whether these lesions
+possess any true diagnostic or prognostic value in the setting of a
+specific disease process. The present communication describes a
+supervised statistical learning tool that is appropriate for the
+application to such large-scale datasets.
+
+The currently described tool is just one example of how "supervised
+learning" algorithms might be applied to aid in the diagnosis of TBI and
+other disease processes through the specific identification of features
+predictive of a given disease state. It is an important demonstration of
+the potential power of these analytical approaches in the rapid but
+comprehensive mining of information from neuroimaging examinations.
+Supervised learning algorithms are presently employed across a wide
+variety of settings for the rapid identification of predictive imaging
+features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
+manufacturers utilize these types of approaches to equip self-driving
+vehicles to recognize and respond to unique external surroundings
+through the identification of visual information sufficiently similar to
+previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
+Similarly, in the context of the neuroimaging assessments, deep learning
+approaches may allow for the rapid identification of information
+predictive of disease state in an individual patient. These approaches
+have been applied to the segmentation of macroscopically visible
+structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
+Additionally, these approaches might be applied to the interrogation of
+imaging data in the individual patient with a primary quantitative
+output metrics to include sequences such as diffusion tensor imaging
+(DTI) and its variants, functional connectivity, perfusion weighted
+imaging, and cortical thickness assessments. At present, these advanced
+neuroimaging sequences are confined to cohort-based research studies due
+to the lack of available analytical tools to assess the information in
+the setting of the individual patient [@Mayer:2014aa]. Application of
+deep learning approaches in the context of data with primary
+quantitative outputs will require large scale normative and disease
+specific databases. Building these large scale imaging libraries is
+resource intensive and requires a multi-center approach with harmonized
+scanners between sites and correlative non-imaging clinical data. Large
+scale TBI data is becoming increasingly available through activities
+such as the Chronic Effects of Neurotrauma Consortium (CENC),
+Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
+Collaborative European Neurotrauma Effectiveness Research in TBI
+(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
+Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
+In concert with any available high quality normative neuroimaging data,
+deep learning algorithms may be well positioned to help transform how
+neuroimaging is interpreted for the clinical management of patients with
+this disease process.
 
 \clearpage
-
-<!--
-\listofchanges
-
-\clearpage
--->
-
-# References

### 2016-02-25 figureCaptions.Rmd
Final submission.

# Figure Captions

__Figure 1:__  Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.

__Figure 2:__  Sample FLAIR acquisition image slices showing both manual and random forest
segmentations for both stages obtained during the leave-one-out evaluation.  Manual
segmentations were performed by one of the authors and provided the ground
truth WMH labels for training the random forest models.

__Figure 3:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  Template construction is detailed in
[@Tustison:2015aa].  These images are important for specific intensity-based
features.

__Figure 4:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 1 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
and illustrate the
stability of the feature importance across the leave-one-out models.
At this initial stage only 31 feature images are used.

__Figure 5:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
forest models for Stage 2 during the leave-one-out evaluation.  These plots
are useful in providing a quantitative assessment of the predictive importance of each feature.
Features are ranked in descending order of importance.
The horizontal error bars provide the $95^{th}$ percentile
 and illustrate the
stability of the feature importance across the leave-one-out models.
We augment the 31 feature images from the first stage by adding an additional
seven voting maps and 7 segmentation posteriors from application of the Bayesian-based
segmentation for a total of 45 images for the second stage.

__Figure 6:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
The region around the WMHs is enlarged (b) in the original FLAIR and the
(c) contralateral FLAIR difference image.

__Figure 7:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
On the left are the
calculated Dice values over all white matter hyperintensities.  Note the improvement
in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
(Right) Similar results can be seen by comparing the total lesion load volume between manual
and automated detection strategies.  Although some outliers are found after the Stage 2
processing in a couple subjects, the number of outliers caused by false positives is
decreased significantly with the second stage processing.

\clearpage

### 2016-02-25 format.Rmd
Final submission.

@@
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.5}
+   - \linespread{2}
    - \usepackage[compact]{titlesec}

### 2016-02-25 intro.Rmd
Final submission.

@@
 Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention (MICCAI) international meeting [@mlmi2015]).
+Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
 Popular techniques such as support vector machines and neural networks have been applied
@@
 
-The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant \replaced[id=dt]{utility}{utitility} for supervised segmentation tasks (e.g.,
+The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
 normal human brain segmentation [@yi2009]) and other computer vision applications
@@
 latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions \replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
-In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}[@Menze:2015aa].
+international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
 
@@
 votes are then used for regression or classification of the data.  Although decision
-trees had been extensively studied \deleted[id=dt]{previously}, the success of employing collections of
+trees had been extensively studied, the success of employing collections of
 such weak learners for boosting machine learning performance
@@
 of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
-Finally, Breiman [@breiman2001] improved accuracy by random sampling of \deleted[id=dt]{the} training
-data (i.e., "bagging") resulting in the current random forest technique \added[id=dt]{applied here}.
+Finally, Breiman [@breiman2001] improved accuracy by random sampling of training
+data (i.e., "bagging") resulting in the current random forest technique applied here.
 
@@
 feature image set (both spatial and intensity-based)
-for segmenting \replaced[id=dt]{WMHs}{white matter hyperintensities} in \replaced[id=dt]{a arge TBI cohort}{traumatic brain injury cohorts}.  Additionally,
+for segmenting WMHs in a large TBI cohort.  Additionally,
 the entire framework is provided publicly through the well-known open-source
-\added[id=lw]{Advanced Normalization Tools} (ANTs)[^i1] and ANTsR[^i2] toolkits.
-\replaced[id=dt]{Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automate routine for WMH segmentation and quantification.}{Further motivating this research of this work is
-the public availability of the imaging data thus permitting full reproducibility
-of the results reported and discussed.}
+Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
+Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.
 

### 2016-02-25 methods.Rmd
Final submission.

@@
 
-MR images utilized for this initial report were acquired from 24 individuals scanned at a single site involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (\added[id=nt,remark=NT: reference needed?]{see [Walker], this issue}).
-Briefly, participants were Operation Enduring Freedom/ Operation Iraqi Freedom /Operation New dawn (OEF/OIF/OND) era service members and veterans between the ages of 18--60 who have experienced combat situations and are between X and X (period of time) from deployment.
-\added[id=js,remark={JS: Data needed.}]{The feature images consisted of subjects aged $X.X \pm X$ years (range X – X years), XX percent male, with at least one or more potential concussive events (mean = X, SD = X, range X-X).}
+MR images utilized for this initial report were acquired from a single scanner involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (see Walker et al., this issue).
+Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
+The feature images consisted of 26 subjects aged 39.6 ± 8.1 years (range 28 – 58 years).
+Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
+Each of the participants that were selected from the larger cohort of participants in this study demonstrated at least one white matter hyperintensity (but as many as 20) on FLAIR imaging.
 
@@
 "features" as input in conjunction with
-\replaced[id=lw,remark=NT: I replaced ground-truth since it is problematic anyway]
-{expertly identified structures of interest}{the ground-truth} for model construction.  For the targeted application in this work (i.e., WMHs),
-regression/classification are performed at the voxelwise level.  In other words, each voxel
+expertly identified structures of interest for model construction.  For the targeted application in this work (i.e., WMHs),
+tissue classification is performed at the voxelwise level.  In other words, each voxel
 within the region of interest is sent through the ensemble of decision trees and receives
@@
 
+<!--
+
 \input{featureImagesTable.tex}
+
+-->
 
@@
 
+<!--
+
 ![Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are \added[id=am,remark=NT: no need for 12 dof for
-intra-subject normalization]{rigidly} pre-aligned
+FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
 [@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
@@
 
+-->
+
 As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
@@
 range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar \added[id=dt]{to previous work}[@Tustison:2015aa].
+algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].
+
+<!--
 
@@
 
+-->
+
 The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation (i.e., the pathology is discounted which is
-a fairly safe assumption for the T1 image) protocols as part of the
-cortical thickness pipeline described in [@Tustison:2014ab].
-\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
+extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
+Any effect of pathology is minimal because WMHs are not commonly seen in T1-weighted MRI.
+This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.
+Once the
 brain has been extracted, we apply a Bayesian-based segmentation algorithm
 with a template-based prior probability strategy to segment the
-parenchymal tissue types.  The result is} a mask delineating the brain parenchyma and probabilistic estimates of the
-\replaced[id=dt]{CSF}{cerebrospinal fluid (csf)}, gray matter, white matter, deep gray matter, brain stem, and
+parenchymal tissue types.  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
+CSF, gray matter, white matter, deep gray matter, brain stem, and
 cerebellum.  These provide the expertly annotated labels for the first six tissue labels
-given above.  The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually identified by one of the authors
+given above.  The WMHs were manually identified by one of the authors
 (J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
@@
 MMRR data set  [@landman2011] (cf Figure 3).
+
+<!--
 
@@
 
+-->
+
 To model the intensity information the first set of images simply includes the
@@
 of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a \added[id=am]{Manhattan} radius of one voxel \added[id=nt]{given the typical size
-of individual WMHs}. For each of the normalized
+a Manhattan radius of one voxel given the typical size
+of individual WMHs. For each of the normalized
 images, we calculate the difference in intensities with the corresponding warped template
@@
 competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] \added[id=nt]{and listed in Table 1}.  We employ the
+from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 1.  We employ the
 same strategy here.  For example, the template
@@
 
-where $\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$ is the transform
+where
+
+$$\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$$
+
+is the transform
 which maps from the individual subject space to the
 template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
-to take advantage of the \added[id=hl,remark=NT:  we are using the gross bilateral symmetry of the brain in an analogous sense that if we know the left eye location and color then we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross} bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that \replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
+to take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
 we use the symmetric templates to compute the contralateral intensity
@@
 derived as part of the registration process, we can warp the voxel index locations
-back to the space of the individual subject \replaced[id=nt]{which motivates similar work by others}{
-Note that this is similar in motivation to the work
-of} [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
+back to the space of the individual subject which motivates similar work by others [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
 coordinate system provided by the template to dramatically improve spatial specificity
@@
 model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here \added[id=lw]{where we employ two stacked random forests (or two "stages")}.  The Stage 1
+segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
 feature images of the training data (as described previously) are used to construct the
 Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-\added[id=lw]{"voting maps" (i.e., the classification count of each decision tree for each tissue label)} via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
+"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
 are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors \added[id=lw]{(i.e., probabilistic estimates of the tissue spatial locations)} for a second application of the Atropos maximum aposteriori algorithm with an
+tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the Atropos maximum aposteriori algorithm with an
 additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
@@
 is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing \added[id=am]{this article}.  \replaced[id=dt]{In contrast to the current work,}{of which} none are based on supervised learning.  As we did for
+at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
 our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
@@
 ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
-use the \texttt{randomForest} package [@liaw2002] using the default settings with
+use the _randomForest_ package [@liaw2002] using the default settings with
 2000 trees per model and 500 randomly selected samples per label per image.
-\added[id=bt]{Note that we saw little variation in performance when these parameters were changed (i.e.
-up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.}
+Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.
 
@@
 a self-encapsulated example to showcase the proposed methodology.  The fact that
-the data will also be made available through the \replaced[id=lw]{Federal Interagency Traumatic Brain Injury Research (FITBIR)}{FITBIR} repository along
+the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
 with the manual labelings will facilitate reproducibility on the part of the reader
@@
     + for each of the 23 subjects, perform a Bayesian-based segmentation with an MRF spatial
-      prior using the seven voting maps are used as additional tissue priors.
+      prior using the seven voting maps as additional tissue priors.
     + construct the Stage 2 random forest model from all the Stage 1 feature images,

### 2016-02-25 results.Rmd
Final submission.

--- 2016-02-11 results.Rmd
@@
 
-After performing the leave-one-out evaluation described at the end of the previous section,
-we calculated the \texttt{MeanDecreaseAccuracy} feature values for each of the 24 subjects $\times$
+After performing the leave-one-out evaluation,
+we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
 2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
@@
 
-![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
+<!--
+
+![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
 forest models for Stage 1 during the leave-one-out evaluation.  These plots
 are useful in providing a quantitative assessment of the predictive importance of each feature.
-\added[id=nt]{Features are ranked in descending order of importance.}
+Features are ranked in descending order of importance.
 The horizontal error bars provide the $95^{th}$ percentile
-\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
+ and illustrate the
 stability of the feature importance across the leave-one-out models.
-\added[id=nt]{At this initial stage only 31 features images are used.}](Figures/averageLeaveOneOutStage1.png)
+At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)
 
-![Average \texttt{MeanDecreaseAccuracy} plots generated from the creation of all 24 random
+![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
 forest models for Stage 2 during the leave-one-out evaluation.  These plots
 are useful in providing a quantitative assessment of the predictive importance of each feature.
-\added[id=nt]{Features are ranked in descending order of importance.}
-The horizontal error bars provide the $95^{th}$ percentile
-\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
+Features are ranked in descending order of importance.
+The horizontal error bars provide the $95^{th}$ percentile and illustrate the
 stability of the feature importance across the leave-one-out models.
-\added[id=nt]{We augment the 31 feature images from the first stage by adding an additional
+We augment the 31 feature images from the first stage by adding an additional
 7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.}](Figures/averageLeaveOneOutStage2.png)
+segmentation for a total of 45 images for the second stage.](Figures/averageLeaveOneOutStage2.png)
+
+-->
 
@@
 the chosen features during the evaluation.  This latter information is illustrated as
-horizontal errors bars providing the $95^{th}$ percentile \deleted[id=bt]{(i.e., $1.96 \times \sigma$)}.
+horizontal errors bars providing the $95^{th}$ percentile
 Note that the reader can cross reference Table 1 for identifying corresponding feature types
@@
 One can also use these measurements as a type of sanity check.  For example, from the Stage
-1  plot, one can see that the \texttt{MeanDecreaseAccuracy} values for the location indices in the
-anterior-posterior direction (i.e., \texttt{TemplateIndicesWarped1}) are greater than
-those for either the inferior-superior (i.e., \texttt{TemplateIndicesWarped0}) or the
-left-right (i.e., \texttt{TemplateIndicesWarped0}) directions in the space of the symmetric
-template.  \deleted[id=bt]{This is intuitive since, as discussed previously, manifestation of TBI white
-matter hyperintensities can often be confused with higher intensities at the
-periventricular caps in normal subjects [@Neema:2009aa]} \deleted[id=am]{whereas there does not seem to
-be contralateral bias in manifestation of white matter hyperintensities in TBI.}
+1  plot, one can see that the _MeanDecreaseAccuracy_ values for the location indices in the
+anterior-posterior direction (i.e., _TemplateIndicesWarped1_) are greater than
+those for either the inferior-superior (i.e., _TemplateIndicesWarped2_) or the
+left-right (i.e., _TemplateIndicesWarped0_) directions in the space of the symmetric
+template.
 
@@
 The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models \added[id=bt]{(see Figure 6)}.  This accords with the known clinical relevance of
+evaluation random forest models (see Figure 6).  This accords with the known clinical relevance of
 FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not \added[id=nt]{typically} manifest symmetrically in both hemispheres.  Interestingly, the
+pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
 posterior maps for the deep gray matter are extremely important for accurate white matter
@@
 
-![\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated
-in (b).  (c) Contralateral FLAIR difference image.}](Figures/FLAIRcontralaleteralWithLesions.png)
+<!--
+
+![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
+The region around the WMHs is enlarged (b) in the original FLAIR and the
+(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
+
+-->
 
@@
 an additional MRF spatial prior.  In Figure 5, the voting maps are labeled as
-"\texttt{RFStage1VotingMaps}" where the final numeral is associated with the brain
+"_RFStage1VotingMaps_" where the final numeral is associated with the brain
 parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "\texttt{RFBrainSegmentationPosteriors}".
+feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
 The Stage 2 feature importance plot follows similar
@@
 in training data.  On the left of Figure 7 are the improvement in Dice values [@tustison2009],
-\added[id=nt]{i.e.,}
+i.e.,
 
@@
 
-\added[id=nt]{over all white matter hyperintensities when comparing the segmentations between the two stages
+over all white matter hyperintensities when comparing the segmentations between the two stages
 where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.}  Performing
+represents the intersection between the manual/automated lesion pair.  Performing
 the second round of supervised learning improves these Dice values.  One can also note from the
@@
 
-![Comparison with manual delineation of white matter hyperintensities.  On the left are the
-calculate Dice values over all white matter hyperintensities.  Note the improvement
+<!--
+
+![Voxelwise comparison with manual delineation of white matter hyperintensities.  On the left are the
+calculated Dice values over all white matter hyperintensities.  Note the improvement
 in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
@@
 processing in a couple subjects, the number of outliers caused by false positives is
-significantly with the second stage processing.](Figures/llvAndDice.png)
+decreased significantly with the second stage processing.](Figures/llvAndDice.png)
+
+-->

### 2016-02-25 stitched.Rmd
Final submission.

@@
 
-After performing the leave-one-out evaluation\deleted[id=dt]{described at the end of the previous section},
-we calculated the \texttt{MeanDecreaseAccuracy} feature values for each of the 24 subjects $\times$
+After performing the leave-one-out evaluation,
+we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
 2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
@@
 
-![\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated.
+<!--
+
+![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
 The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.}](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
+(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
+
+-->
 
@@
 
-![\added[id=am]{Voxelwise} comparison with manual delineation of white matter hyperintensities.  On the left are the
-calculate\added[id=am]{d} Dice values over all white matter hyperintensities.  Note the improvement
+<!--
+
+![Voxelwise comparison with manual delineation of white matter hyperintensities.  On the left are the
+calculated Dice values over all white matter hyperintensities.  Note the improvement
 in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
@@
 processing in a couple subjects, the number of outliers caused by false positives is
-\added[id=lw]{decreased} significantly with the second stage processing.](Figures/llvAndDice.png)
+decreased significantly with the second stage processing.](Figures/llvAndDice.png)
+
+-->
 
 # Discussion
-#
+
 <!-- Nick could you add some detail here on how this technique compares
@@
 
-\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
+The more recent meta-analysis by Kloppenborg et al.
 [@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
@@
 
+## Acknowledgements
+
+The authors wish to acknowledge all other members of
+the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
+X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
+also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
+and Kevin Wilson. We would also like to thank the study participants and
+their family members. This project was jointly supported by the
+Department of Defense (W81XWH-13-2-0095), the U.S. Department of
+Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
+Grant HU 0001-08-0001.
+
+## Declaration of Interest/Disclaimer
+
+The authors report no financial
+disclosures or conflicts of interest. The views expressed here are those
+of the authors and do not necessarily reflect the official policy of
+position of the Department of the Navy, Department of Defense, nor the
+U.S. Government. This work was prepared as a part of official duties;
+Title 17 USC §105 provides that Copyright protection under this title is
+not available for any work of the U.S. Government. Title 17 USC §101
+defines a US Government work as a work prepared by a military service
+member of employee of the US Government as part of that person’s
+official duties.
+
+\clearpage
+
 <!-- \listofchanges
@@
 # References
+
+# Figure Captions
+
+__Figure 1:__  Representation of Stage 1 feature images for subject 01C1019.  The
+FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
+[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
+(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
+application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
+the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
+The feature images are then generated for voxelwise input to the RF model which results in
+the voting maps illustrated on the right.  This gives a probabilistic classification of
+tissue type.  Not shown are the probability and voting images for the brain stem and
+cerebellum.
+
+__Figure 2:__  Sample FLAIR acquisition image slices showing both manual and random forest
+segmentations for both stages obtained during the leave-one-out evaluation.  Manual
+segmentations were performed by one of the authors and provided the ground
+truth WMH labels for training the random forest models.
+
+__Figure 3:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
+from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
+the components relevant for this work).  Template construction is detailed in
+[@Tustison:2015aa].  These images are important for specific intensity-based
+features.
+
+__Figure 4:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
+forest models for Stage 1 during the leave-one-out evaluation.  These plots
+are useful in providing a quantitative assessment of the predictive importance of each feature.
+Features are ranked in descending order of importance.
+The horizontal error bars provide the $95^{th}$ percentile
+and illustrate the
+stability of the feature importance across the leave-one-out models.
+At this initial stage only 31 feature images are used.
+
+__Figure 5:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
+forest models for Stage 2 during the leave-one-out evaluation.  These plots
+are useful in providing a quantitative assessment of the predictive importance of each feature.
+Features are ranked in descending order of importance.
+The horizontal error bars provide the $95^{th}$ percentile
+ and illustrate the
+stability of the feature importance across the leave-one-out models.
+We augment the 31 feature images from the first stage by adding an additional
+seven voting maps and 7 segmentation posteriors from application of the Bayesian-based
+segmentation for a total of 45 images for the second stage.
+
+__Figure 6:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
+The region around the WMHs is enlarged (b) in the original FLAIR and the
+(c) contralateral FLAIR difference image.
+
+__Figure 7:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
+On the left are the
+calculated Dice values over all white matter hyperintensities.  Note the improvement
+in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
+(Right) Similar results can be seen by comparing the total lesion load volume between manual
+and automated detection strategies.  Although some outliers are found after the Stage 2
+processing in a couple subjects, the number of outliers caused by false positives is
+decreased significantly with the second stage processing.
+
+\clearpage

### 2016-02-25 stitched.tex
Final submission.

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \setlength{\parindent}{0pt}
@@
 \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-\linespread{1.5}
+\linespread{2}
 \usepackage[compact]{titlesec}
@@
 in conjunction with the Medical Image Computing and Computer-Aided
-Intervention (MICCAI) international meeting\textsuperscript{11}).
-Popular techniques such as support vector machines and neural networks
-have been applied successfully to clinically relevant imaging tasks such
-as supervised image segmentation (e.g.,\textsuperscript{12}) and
-diagnostic prediction (e.g.,\textsuperscript{13,14}). Facilitating the
-current employment of such techniques are the number of available
-imaging data sets\textsuperscript{15} and the public availability of
-data science packages such as SciPy\textsuperscript{16} and the R
-project for statistical computing\textsuperscript{17} and their
-associated extensions.
-
-The random forests framework\textsuperscript{18} is a popular machine
-learning technique that has demonstrated significant
-\replaced[id=dt]{utility}{utitility} for supervised segmentation tasks
-(e.g., normal human brain segmentation\textsuperscript{19}) and other
-computer vision applications (e.g., human gait
-detection\textsuperscript{20}). In the context of neuropathology, random
-forest-based paradigms have been employed in the delineation of multiple
-sclerosis lesions\textsuperscript{21}, stroke
-lesions\textsuperscript{22}, and brain tumors\textsuperscript{23--26}
+Intervention (MICCAI) international meeting). Popular techniques such as
+support vector machines and neural networks have been applied
+successfully to clinically relevant imaging tasks such as supervised
+image segmentation (e.g.,\textsuperscript{11}) and diagnostic prediction
+(e.g.,\textsuperscript{12,13}). Facilitating the current employment of
+such techniques are the number of available imaging data
+sets\textsuperscript{14} and the public availability of data science
+packages such as SciPy\textsuperscript{15} and the R project for
+statistical computing\textsuperscript{16} and their associated
+extensions.
+
+The random forests framework\textsuperscript{17} is a popular machine
+learning technique that has demonstrated significant utility for
+supervised segmentation tasks (e.g., normal human brain
+segmentation\textsuperscript{18}) and other computer vision applications
+(e.g., human gait detection\textsuperscript{19}). In the context of
+neuropathology, random forest-based paradigms have been employed in the
+delineation of multiple sclerosis lesions\textsuperscript{20}, stroke
+lesions\textsuperscript{21}, and brain tumors\textsuperscript{22--25}
 for both single and multi-modal acquisition protocols. Of note, these
 latter random forest approaches for brain tumor segmentation have
-performed well in recent international competitions
-\replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
-In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}\textsuperscript{27}.
-
-Random forests are conceptually straightforward\textsuperscript{18}.
+performed well in recent international competitions established in
+response to the lack of objective comparisons between segmentation
+algorithms (i.e., the Multimodal Brain Tumor Segmentation (BRATS)
+challenge was initiated in 2012\textsuperscript{26}.
+
+Random forests are conceptually straightforward\textsuperscript{17}.
 They consist of ensembles of decision trees that are built from training
@@
 These votes are then used for regression or classification of the data.
-Although decision trees had been extensively studied
-\deleted[id=dt]{previously}, the success of employing collections of
-such weak learners for boosting machine learning performance (e.g.,
-AdaBoost\textsuperscript{28,29}) influenced the similarly sytled
-conglomeration of decision trees into ``forests'' with randomized node
-optimization\textsuperscript{30,31}. Finally,
-Breiman\textsuperscript{18} improved accuracy by random sampling of
-\deleted[id=dt]{the} training data (i.e., ``bagging'') resulting in the
-current random forest technique \added[id=dt]{applied here}.
+Although decision trees had been extensively studied, the success of
+employing collections of such weak learners for boosting machine
+learning performance (e.g., AdaBoost\textsuperscript{27,28}) influenced
+the similarly sytled conglomeration of decision trees into ``forests''
+with randomized node optimization\textsuperscript{29,30}. Finally,
+Breiman\textsuperscript{17} improved accuracy by random sampling of
+training data (i.e., ``bagging'') resulting in the current random forest
+technique applied here.
 
@@
 tailored contextual feature image set (both spatial and intensity-based)
-for segmenting \replaced[id=dt]{WMHs}{white matter hyperintensities} in
-\replaced[id=dt]{a arge TBI cohort}{traumatic brain injury cohorts}.
-Additionally, the entire framework is provided publicly through the
-well-known open-source \added[id=lw]{Advanced Normalization Tools}
-(ANTs)\footnote{\url{https://github.com/stnava/ANTs}} and
-ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
-\replaced[id=dt]{Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automate routine for WMH segmentation and quantification.}{Further motivating this research of this work is
-the public availability of the imaging data thus permitting full reproducibility
-of the results reported and discussed.}
+for segmenting WMHs in a large TBI cohort. Additionally, the entire
+framework is provided publicly through the well-known open-source
+Advanced Normalization Tools (ANTs)\footnote{\url{https://github.com/stnava/ANTs}}
+and ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
+Further motivating this research is the availability of several large
+publicly available imaging data sets that permits testing
+reproducibility of this automated routine for WMH segmentation and
+quantification.
 
@@
 
-MR images utilized for this initial report were acquired from 24
-individuals scanned at a single site involved in the Chronic Effects of
-Neurotrauma Consortium's (CENC) observational study
-(\added[id=nt,remark=NT: reference needed?]{see [Walker], this issue}).
-Briefly, participants were Operation Enduring Freedom/ Operation Iraqi
-Freedom /Operation New dawn (OEF/OIF/OND) era service members and
-veterans between the ages of 18--60 who have experienced combat
-situations and are between X and X (period of time) from deployment.
-\added[id=js,remark={JS: Data needed.}]{The feature images consisted of subjects aged $X.X \pm X$ years (range X – X years), XX percent male, with at least one or more potential concussive events (mean = X, SD = X, range X-X).}
+MR images utilized for this initial report were acquired from a single
+scanner involved in the Chronic Effects of Neurotrauma Consortium's
+(CENC) observational study (see Walker et al., this issue). Briefly,
+participants were Operation Iraqi Freedom/Operation Enduring Freedom
+(OIF/OEF) era Service Members and Veterans between the ages of 18-60
+years with prior combat exposure and deployment(s). The feature images
+consisted of 26 subjects aged 39.6 ± 8.1 years (range 28 -- 58 years).
+Within this cohort, 24 (92\%) were considered positive for TBI based
+upon the potential concussive events (PCE) interview process described
+in detail in Walker et al., this issue). Each of the participants that
+were selected from the larger cohort of participants in this study
+demonstrated at least one white matter hyperintensity (but as many as
+20) on FLAIR imaging.
 
@@
 Crucial to these supervised segmentation approaches are the creation and
-selection of ``features'' as input in conjunction with
-\replaced[id=lw,remark=NT: I replaced ground-truth since it is problematic anyway]
-{expertly identified structures of interest}{the ground-truth} for model
-construction. For the targeted application in this work (i.e., WMHs),
-regression/classification are performed at the voxelwise level. In other
-words, each voxel within the region of interest is sent through the
-ensemble of decision trees and receives a set of classification votes
-from each tree thus permitting a regression or classification solution.
-Since this procedure is performed at the voxelwise level, intensity
-information alone is insufficient for good segmentation performance due
-to the lack of spatial context. For example, as pointed out
-in\textsuperscript{32}, higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated
-lesion detection algorithms. Other potential confounds include MR signal
-inhomogeneity and noise. Therefore, even though machine learning and
-pattern recognition techniques are extremely powerful and have
-significant potential, just as crucial to outcome is the creative
-construction and deployment of salient feature images which we detail
-below.
+selection of ``features'' as input in conjunction with expertly
+identified structures of interest for model construction. For the
+targeted application in this work (i.e., WMHs), tissue classification is
+performed at the voxelwise level. In other words, each voxel within the
+region of interest is sent through the ensemble of decision trees and
+receives a set of classification votes from each tree thus permitting a
+regression or classification solution. Since this procedure is performed
+at the voxelwise level, intensity information alone is insufficient for
+good segmentation performance due to the lack of spatial context. For
+example, as pointed out in\textsuperscript{31}, higher intensities can
+be found at the periventricular caps in normal subjects which often
+confounds automated lesion detection algorithms. Other potential
+confounds include MR signal inhomogeneity and noise. Therefore, even
+though machine learning and pattern recognition techniques are extremely
+powerful and have significant potential, just as crucial to outcome is
+the creative construction and deployment of salient feature images which
+we detail below.
 
@@
 segmentation}\label{feature-images-for-wmh-segmentation}
-
-\input{featureImagesTable.tex}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/featureImages.png}
-\caption{Representation of Stage 1 feature images for subject 01C1019.
-The FLAIR, T1-, and T2-weighted images are
-\added[id=am,remark=NT: no need for 12 dof for
-intra-subject normalization]{rigidly} pre-aligned\textsuperscript{33} to
-the space of the T1 image. The three modality images are then
-preprocessed (N4 bias correction\textsuperscript{34} and adaptive
-denoising\textsuperscript{35}) followed by application of standard ANTs
-brain extraction and \(n\)-tissue segmentation protocols using the MMRR
-symmetric template and corresponding priors\textsuperscript{36} applied
-to the T1 image. The feature images are then generated for voxelwise
-input to the RF model which results in the voting maps illustrated on
-the right. This gives a probabilistic classification of tissue type. Not
-shown are the probability and voting images for the brain stem and
-cerebellum.}
-\end{figure}
-
 As mentioned previously, input for each subject comprises FLAIR, T1-,
@@
 registered to the T1 image using the open-source Advanced Normalization
-Tools (ANTs)\textsuperscript{33}. The aligned images are then
-preprocessed using the denoising algorithm of\textsuperscript{35}
+Tools (ANTs)\textsuperscript{32}. The aligned images are then
+preprocessed using the denoising algorithm of\textsuperscript{33}
 followed by N4 bias correction\textsuperscript{34} which are then
@@
 an alternative intensity standardization algorithm
-(e.g.,\textsuperscript{37}), we found that a simple linear rescaling
-produced better results similar
-\added[id=dt]{to previous work}\textsuperscript{26}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/sampleResults.png}
-\caption{Sample FLAIR acquisition image slices showing both manual and
-random forest segmentations for both stages obtained during the
-leave-one-out evaluation. Manual segmentations were performed by one of
-the authors and provided the ground truth WMH labels for training the
-random forest models.}
-\end{figure}
+(e.g.,\textsuperscript{35}), we found that a simple linear rescaling
+produced better results similar to previous work\textsuperscript{25}.
 
 The T1 image is then processed via the ANTs brain extraction and normal
-tissue segmentation (i.e., the pathology is discounted which is a fairly
-safe assumption for the T1 image) protocols as part of the cortical
-thickness pipeline described in\textsuperscript{36}.
-\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
-brain has been extracted, we apply a Bayesian-based segmentation algorithm
-with a template-based prior probability strategy to segment the
-parenchymal tissue types.  The result is} a mask delineating the brain
-parenchyma and probabilistic estimates of the
-\replaced[id=dt]{CSF}{cerebrospinal fluid (csf)}, gray matter, white
+tissue segmentation pipelines\textsuperscript{36}. Any effect of
+pathology is minimal because WMHs are not commonly seen in T1-weighted
+MRI. This protocol involves preprocessing using N4 bias correction
+followed by a template-based strategy for brain extraction. Once the
+brain has been extracted, we apply a Bayesian-based segmentation
+algorithm with a template-based prior probability strategy to segment
+the parenchymal tissue types. The result is a mask delineating the brain
+parenchyma and probabilistic estimates of the CSF, gray matter, white
 matter, deep gray matter, brain stem, and cerebellum. These provide the
 expertly annotated labels for the first six tissue labels given above.
-The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually
-identified by one of the authors (J. R. S.) using the ITK-SNAP
-tool\textsuperscript{38}. Segmentation is performed using the ANTs
-Atropos tool\textsuperscript{39} and multi-model optimal symmetric
-shape/intensity templates\textsuperscript{26} created from the public
-MMRR data set\textsuperscript{40} (cf Figure 3).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/MMRR.png}
-\caption{Canonical views of the mutlivariate, bilaterally symmetric
-template constructed from the MMRR data set\textsuperscript{40} (only
-shown are the FLAIR, T1, and T2 modalities--- the components relevant
-for this work). Template construction is detailed
-in\textsuperscript{26}. These images are important for specific
-intensity-based features.}
-\end{figure}
+The WMHs were manually identified by one of the authors (J. R. S.) using
+the ITK-SNAP tool\textsuperscript{37}. Segmentation is performed using
+the ANTs Atropos tool\textsuperscript{38} and multi-model optimal
+symmetric shape/intensity templates\textsuperscript{25} created from the
+public MMRR data set\textsuperscript{39} (cf Figure 3).
 
@@
 (mean, standard deviation, and skewness) feature images using a
-\added[id=am]{Manhattan} radius of one voxel
-\added[id=nt]{given the typical size
-of individual WMHs}. For each of the normalized images, we calculate the
-difference in intensities with the corresponding warped template
-component. Previous success in the international brain tumor
-segmentation competition\textsuperscript{27} was based on an important
-set of intensity features that were created from multi-modal templates
-mentioned previously\textsuperscript{26}
-\added[id=nt]{and listed in Table 1}. We employ the same strategy here.
-For example, the template difference feature image for the FLAIR image,
-\(S_{FLAIR}\) is calculated as:
+Manhattan radius of one voxel given the typical size of individual WMHs.
+For each of the normalized images, we calculate the difference in
+intensities with the corresponding warped template component. Previous
+success in the international brain tumor segmentation
+competition\textsuperscript{26} was based on an important set of
+intensity features that were created from multi-modal templates
+mentioned previously\textsuperscript{25} and listed in Table 1. We
+employ the same strategy here. For example, the template difference
+feature image for the FLAIR image, \(S_{FLAIR}\) is calculated as:
 
@@
 
-where \(\phi_b: S \leftrightarrow \underset{b}{\leftrightsquigarrow} T\)
+where
+
+\[\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T\]
+
 is the transform which maps from the individual subject space to the
 template space and \(T_{FLAIR}\) is the FLAIR template component. Also,
-to take advantage of the
-\added[id=hl,remark=NT:  we are using the gross bilateral symmetry of the brain in an analogous sense that if we know the left eye location and color then we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross}
-bilateral symmetry of the normal brain (in terms of both shape and
-intensity), and the fact that
-\replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
-we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature. For the FLAIR component,
-this contralateral difference image is calculated from
+to take advantage of the gross bilateral symmetry of the normal brain
+(in terms of both shape and intensity), and the fact that WMHs do not
+generally manifest symmetrically across hemispheres, we use the
+symmetric templates to compute the contralateral intensity differences
+as an additional intensity feature. For the FLAIR component, this
+contralateral difference image is calculated from
 
@@
 prediction step. Additional spatial contextual feature images include
-the distance maps\textsuperscript{41} based on the csf, gray matter, and
+the distance maps\textsuperscript{40} based on the csf, gray matter, and
 deep gray matter images. These latter images are intended to help
@@
 template. The T1 image of the subject is registered to the T1 template
-component using a B-spline variant\textsuperscript{42} of the well-known
-ANTs Symmetric Normalization (SyN) algorithm\textsuperscript{43}. Since
+component using a B-spline variant\textsuperscript{41} of the well-known
+ANTs Symmetric Normalization (SyN) algorithm\textsuperscript{42}. Since
 the inverse transform is also derived as part of the registration
 process, we can warp the voxel index locations back to the space of the
-individual subject
-\replaced[id=nt]{which motivates similar work by others}{
-Note that this is similar in motivation to the work
-of}\textsuperscript{44}. However, this previous work lacks the
+individual subject which motivates similar work by
+others\textsuperscript{43}. However, this previous work lacks the
 normalization to the standard coordinate system provided by the template
@@
 
-In previous brain tumor segmentation work\textsuperscript{26}, it was
+In previous brain tumor segmentation work\textsuperscript{25}, it was
 demonstrated that a concatenated supervised approach, whereby the
@@
 segmentation performance. We do the same thing for the work described
-here
-\added[id=lw]{where we employ two stacked random forests (or two "stages")}.
-The Stage 1 feature images of the training data (as described
-previously) are used to construct the Stage 1 model. The training data
-Stage 1 features are then used to produce the voxelwise
-\added[id=lw]{"voting maps" (i.e., the classification count of each decision tree for each tissue label)}
+here where we employ two stacked random forests (or two ``stages''). The
+Stage 1 feature images of the training data (as described previously)
+are used to construct the Stage 1 model. The training data Stage 1
+features are then used to produce the voxelwise ``voting maps'' (i.e.,
+the classification count of each decision tree for each tissue label)
 via the Stage 1 random forest model. All the Stage 1 features plus the
 Stage 1 voting maps are used as input to the Stage 2 model. In addition,
-we use the Stage 1 voting maps as tissue priors
-\added[id=lw]{(i.e., probabilistic estimates of the tissue spatial locations)}
-for a second application of the Atropos maximum aposteriori algorithm
-with an additional Markov Random Field spatial prior
-(MAP-MRF)\textsuperscript{39}. However, for the second stage we use all
-three aligned preprocessed images for a multivariate segmentation. The
-resulting seven posterior probability images constitute a third
-additional feature image set for Stage 2.
+we use the Stage 1 voting maps as tissue priors (i.e., probabilistic
+estimates of the tissue spatial locations) for a second application of
+the Atropos maximum aposteriori algorithm with an additional Markov
+Random Field spatial prior (MAP-MRF)\textsuperscript{38}. However, for
+the second stage we use all three aligned preprocessed images for a
+multivariate segmentation. The resulting seven posterior probability
+images constitute a third additional feature image set for Stage 2.
 
@@
 As pointed out in a recent comprehensive multiple sclerosis lesion
-segmentation review\textsuperscript{45}, although the number of
+segmentation review\textsuperscript{44}, although the number of
 algorithms reported in the literature is quite extensive, there were
 only four publicly available segmentation algorithms at the time of
-writing \added[id=am]{this article}.
-\replaced[id=dt]{In contrast to the current work,}{of which} none are
-based on supervised learning. As we did for our brain tumor segmentation
-algorithm\textsuperscript{26}, all of the code described in this work is
+writing this article. In contrast to the current work, none are based on
+supervised learning. As we did for our brain tumor segmentation
+algorithm\textsuperscript{25}, all of the code described in this work is
 publicly available through the open-source ANTs/ANTsR toolkits. Through
 ANTsR (an add-on toolkit which, in part, bridges ANTs and the R
-statistical project) we use the \texttt{randomForest}
-package\textsuperscript{46} using the default settings with 2000 trees
-per model and 500 randomly selected samples per label per image.
-\added[id=bt]{Note that we saw little variation in performance when these parameters were changed (i.e.
-up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.}
+statistical project) we use the \emph{randomForest}
+package\textsuperscript{45} using the default settings with 2000 trees
+per model and 500 randomly selected samples per label per image. Note
+that we saw little variation in performance when these parameters were
+changed (i.e.~up to 1000 random samples and as little as 1000 trees)
+which is consistent with our previous experience.
 
@@
 methodology. The fact that the data will also be made available through
-the
-\replaced[id=lw]{Federal Interagency Traumatic Brain Injury Research (FITBIR)}{FITBIR}
+the Federal Interagency Traumatic Brain Injury Research (FITBIR)
 repository along with the manual labelings will facilitate
@@
     for each of the 23 subjects, perform a Bayesian-based segmentation
-    with an MRF spatial prior using the seven voting maps are used as
-    additional tissue priors.
+    with an MRF spatial prior using the seven voting maps as additional
+    tissue priors.
   \item
@@
 
-After performing the leave-one-out
-evaluation\deleted[id=dt]{described at the end of the previous section},
-we calculated the \texttt{MeanDecreaseAccuracy} feature values for each
-of the 24 subjects \(\times\) 2 models per subject \(=48\) total models.
-This measure (per feature, per model) is calculated during the
-out-of-bag phase of the random forest model construction and quantifies
-the decrease in prediction accuracy from omitting the specified feature.
-In other words, this quantity helps determine the importance of a
-particular feature and, although we save such efforts for future work,
-this information provides us with guidance for future feature pruning
-and/or additions.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/averageLeaveOneOutStage1.png}
-\caption{Average \texttt{MeanDecreaseAccuracy} plots generated from the
-creation of all 24 random forest models for Stage 1 during the
-leave-one-out evaluation. These plots are useful in providing a
-quantitative assessment of the predictive importance of each feature.
-\added[id=nt]{Features are ranked in descending order of importance.}
-The horizontal error bars provide the \(95^{th}\) percentile
-\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
-stability of the feature importance across the leave-one-out models.
-\added[id=nt]{At this initial stage only 31 features images are used.}}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/averageLeaveOneOutStage2.png}
-\caption{Average \texttt{MeanDecreaseAccuracy} plots generated from the
-creation of all 24 random forest models for Stage 2 during the
-leave-one-out evaluation. These plots are useful in providing a
-quantitative assessment of the predictive importance of each feature.
-\added[id=nt]{Features are ranked in descending order of importance.}
-The horizontal error bars provide the \(95^{th}\) percentile
-\deleted[id=bt]{(i.e., $1.96 \times \sigma$)} and illustrate the
-stability of the feature importance across the leave-one-out models.
-\added[id=nt]{We augment the 31 feature images from the first stage by adding an additional
-7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.}}
-\end{figure}
+After performing the leave-one-out evaluation, we calculated the
+\emph{MeanDecreaseAccuracy} feature values for each of the 24 subjects
+\(\times\) 2 models per subject \(=48\) total models. This measure (per
+feature, per model) is calculated during the out-of-bag phase of the
+random forest model construction and quantifies the decrease in
+prediction accuracy from omitting the specified feature. In other words,
+this quantity helps determine the importance of a particular feature
+and, although we save such efforts for future work, this information
+provides us with guidance for future feature pruning and/or additions.
 
@@
 features during the evaluation. This latter information is illustrated
-as horizontal errors bars providing the \(95^{th}\) percentile
-\deleted[id=bt]{(i.e., $1.96 \times \sigma$)}. Note that the reader can
-cross reference Table 1 for identifying corresponding feature types and
-names.
+as horizontal errors bars providing the \(95^{th}\) percentile Note that
+the reader can cross reference Table 1 for identifying corresponding
+feature types and names.
 
@@
 example, from the Stage 1 plot, one can see that the
-\texttt{MeanDecreaseAccuracy} values for the location indices in the
-anterior-posterior direction (i.e., \texttt{TemplateIndicesWarped1}) are
+\emph{MeanDecreaseAccuracy} values for the location indices in the
+anterior-posterior direction (i.e., \emph{TemplateIndicesWarped1}) are
 greater than those for either the inferior-superior (i.e.,
-\texttt{TemplateIndicesWarped0}) or the left-right (i.e.,
-\texttt{TemplateIndicesWarped0}) directions in the space of the
-symmetric template.
-\deleted[id=bt]{This is intuitive since, as discussed previously, manifestation of TBI white
-matter hyperintensities can often be confused with higher intensities at the
-periventricular caps in normal subjects [@Neema:2009aa]}
-\deleted[id=am]{whereas there does not seem to
-be contralateral bias in manifestation of white matter hyperintensities in TBI.}
+\emph{TemplateIndicesWarped2}) or the left-right (i.e.,
+\emph{TemplateIndicesWarped0}) directions in the space of the symmetric
+template.
 
@@
 features for Stage 1. The contralateral difference FLAIR image is highly
-discriminative over the set of evaluation random forest models
-\added[id=bt]{(see Figure 6)}. This accords with the known clinical
-relevance of FLAIR images for identifying white matter hyperintensities
-and the fact that such pathology does not \added[id=nt]{typically}
-manifest symmetrically in both hemispheres. Interestingly, the posterior
-maps for the deep gray matter are extremely important for accurate white
-matter hyperintensity segmentation. Perhaps the spatial specification of
-deep gray matter aids in the removal of false positives. Inspection of
-the bottom of the plots demonstrates the lack of discriminating features
-associated with the T1 image which is also well-known in the clinical
-literature.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/FLAIRcontralaleteralWithLesionsAlternate.png}
-\caption{\added[id=nt]{(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.}}
-\end{figure}
+discriminative over the set of evaluation random forest models (see
+Figure 6). This accords with the known clinical relevance of FLAIR
+images for identifying white matter hyperintensities and the fact that
+such pathology does not typically manifest symmetrically in both
+hemispheres. Interestingly, the posterior maps for the deep gray matter
+are extremely important for accurate white matter hyperintensity
+segmentation. Perhaps the spatial specification of deep gray matter aids
+in the removal of false positives. Inspection of the bottom of the plots
+demonstrates the lack of discriminating features associated with the T1
+image which is also well-known in the clinical literature.
 
@@
 prior. In Figure 5, the voting maps are labeled as
-``\texttt{RFStage1VotingMaps}'' where the final numeral is associated
-with the brain parenchymal labeling given previously. Similarly, the
+``\emph{RFStage1VotingMaps}'' where the final numeral is associated with
+the brain parenchymal labeling given previously. Similarly, the
 additional RF prior segmentation feature probability maps are labeled as
-``\texttt{RFBrainSegmentationPosteriors}''. The Stage 2 feature
-importance plot follows similar trends as that for Stage 1 with the T1
-images not contributing much to the identification of white matter
-hyperintensity voxels. The initial voting maps from Stage 1 are
-extremely important with the top 3 being the estimated locations of the
-1) gray matter, 2) white matter, and 3) white matter hyperintensities.
-Since these tissue type can be conflated based on intensity alone it is
-intuitive that such features would be important.
+``\emph{RFBrainSegmentationPosteriors}''. The Stage 2 feature importance
+plot follows similar trends as that for Stage 1 with the T1 images not
+contributing much to the identification of white matter hyperintensity
+voxels. The initial voting maps from Stage 1 are extremely important
+with the top 3 being the estimated locations of the 1) gray matter, 2)
+white matter, and 3) white matter hyperintensities. Since these tissue
+type can be conflated based on intensity alone it is intuitive that such
+features would be important.
 
@@
 characteristic with manual labelings in related
-fields\textsuperscript{45,47,48}, such labelings are characteristic of
+fields\textsuperscript{44,46,47}, such labelings are characteristic of
 current clinical practices and the methodology proposed herein is
 readily adapted to refinements in training data. On the left of Figure 7
-are the improvement in Dice values\textsuperscript{49},
-\added[id=nt]{i.e.,}
+are the improvement in Dice values\textsuperscript{48}, i.e.,
 
@@
 
-\added[id=nt]{over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.}
+over all white matter hyperintensities when comparing the segmentations
+between the two stages where the sum is taken over all individually
+labeled manual, \(T_r\), and automated, \(S_r\), lesions and \(\cap\)
+represents the intersection between the manual/automated lesion pair.
 Performing the second round of supervised learning improves these Dice
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/llvAndDice.png}
-\caption{\added[id=am]{Voxelwise} comparison with manual delineation of
-white matter hyperintensities. On the left are the
-calculate\added[id=am]{d} Dice values over all white matter
-hyperintensities. Note the improvement in the Dice metric from the
-employment of the Stage 2 component of the processing pipeline. (Right)
-Similar results can be seen by comparing the total lesion load volume
-between manual and automated detection strategies. Although some
-outliers are found after the Stage 2 processing in a couple subjects,
-the number of outliers caused by false positives is
-\added[id=lw]{decreased} significantly with the second stage
-processing.}
-\end{figure}
-
 \section{Discussion}\label{discussion}
-
-\section{}\label{section}
 
@@
 
-\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et
-al.\textsuperscript{10} of 23 cross-sectional studies reporting MRI and
-concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
-that WMHs were associated with cognitive deficit (effect size of -0.10,
-95\% CI: -0.13 to -0.08) after controlling for age. These studies also
-differed in the metric used to measure the WMHs, including volume, \% of
-total intracranial volume, and a visual rating score. The effect size
-for the association with cognitive deficit in these cross-sectional
-studies did not differ significantly across various cognitive domains or
-the method of measuring lesion volume. Among eight longitudinal studies
-analyzed by Kloppenborg et al that included a follow-up MRI and also
-controlled for age, the effect size for the association of progression
-in WMHs and cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09).
-This association was stronger for attention and executive function than
-for memory and processing speed. Although baseline WMHs were predictive
-of cognitive deficit at follow-up in the seven studies which did not
-repeat MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to
--0.05) than in the longitudinal studies that calculated progression in
-WMHs. In summary, progression of WMHs seen on repeat MRI has a stronger
-relation to cognitive deficit than concurrent imaging findings. These
+The more recent meta-analysis by Kloppenborg et al.\textsuperscript{10}
+of 23 cross-sectional studies reporting MRI and concurrent
+neuropsychological results in patients with heterogeneous diagnoses but
+without previously diagnosed cognitive impairment, found that WMHs were
+associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
+to -0.08) after controlling for age. These studies also differed in the
+metric used to measure the WMHs, including volume, \% of total
+intracranial volume, and a visual rating score. The effect size for the
+association with cognitive deficit in these cross-sectional studies did
+not differ significantly across various cognitive domains or the method
+of measuring lesion volume. Among eight longitudinal studies analyzed by
+Kloppenborg et al that included a follow-up MRI and also controlled for
+age, the effect size for the association of progression in WMHs and
+cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
+association was stronger for attention and executive function than for
+memory and processing speed. Although baseline WMHs were predictive of
+cognitive deficit at follow-up in the seven studies which did not repeat
+MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
+in the longitudinal studies that calculated progression in WMHs. In
+summary, progression of WMHs seen on repeat MRI has a stronger relation
+to cognitive deficit than concurrent imaging findings. These
 meta-analyses support the rationale for repeating an MRI in patients
@@
 across a wide variety of settings for the rapid identification of
-predictive imaging features\textsuperscript{50--53}. Automobile
+predictive imaging features\textsuperscript{49--52}. Automobile
 manufacturers utilize these types of approaches to equip self-driving
@@
 through the identification of visual information sufficiently similar to
-previously assimilated training data\textsuperscript{54,55}. Similarly,
+previously assimilated training data\textsuperscript{53,54}. Similarly,
 in the context of the neuroimaging assessments, deep learning approaches
@@
 applied to the segmentation of macroscopically visible
-structures\textsuperscript{50--53}. Additionally, these approaches might
+structures\textsuperscript{49--52}. Additionally, these approaches might
 be applied to the interrogation of imaging data in the individual
@@
 analytical tools to assess the information in the setting of the
-individual patient\textsuperscript{56}. Application of deep learning
+individual patient\textsuperscript{55}. Application of deep learning
 approaches in the context of data with primary quantitative outputs will
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
+\subsection{Acknowledgements}\label{acknowledgements}
+
+The authors wish to acknowledge all other members of the CENC
+Neuroimaging Steering Committee and CENC leadership (Drs. David X. Cifu,
+Ramon Diaz-Arrastia, and Rick Williams) for their support. We also
+gratefully acknowledge the assistance of Tracy Nolen, Chris Siege and
+Kevin Wilson. We would also like to thank the study participants and
+their family members. This project was jointly supported by the
+Department of Defense (W81XWH-13-2-0095), the U.S. Department of
+Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
+Grant HU 0001-08-0001.
+
+\subsection{Declaration of
+Interest/Disclaimer}\label{declaration-of-interestdisclaimer}
+
+The authors report no financial disclosures or conflicts of interest.
+The views expressed here are those of the authors and do not necessarily
+reflect the official policy of position of the Department of the Navy,
+Department of Defense, nor the U.S. Government. This work was prepared
+as a part of official duties; Title 17 USC §105 provides that Copyright
+protection under this title is not available for any work of the U.S.
+Government. Title 17 USC §101 defines a US Government work as a work
+prepared by a military service member of employee of the US Government
+as part of that person's official duties.
+
+\clearpage
+
+\section{References}\label{references}
+
+\section{Figure Captions}\label{figure-captions}
+
+\textbf{Figure 1:} Representation of Stage 1 feature images for subject
+01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
+pre-aligned\textsuperscript{32} to the space of the T1 image. The three
+modality images are then preprocessed (N4 bias
+correction\textsuperscript{34} and adaptive
+denoising\textsuperscript{33}) followed by application of standard ANTs
+brain extraction and \(n\)-tissue segmentation protocols using the MMRR
+symmetric template and corresponding priors\textsuperscript{36} applied
+to the T1 image. The feature images are then generated for voxelwise
+input to the RF model which results in the voting maps illustrated on
+the right. This gives a probabilistic classification of tissue type. Not
+shown are the probability and voting images for the brain stem and
+cerebellum.
+
+\textbf{Figure 2:} Sample FLAIR acquisition image slices showing both
+manual and random forest segmentations for both stages obtained during
+the leave-one-out evaluation. Manual segmentations were performed by one
+of the authors and provided the ground truth WMH labels for training the
+random forest models.
+
+\textbf{Figure 3:} Canonical views of the mutlivariate, bilaterally
+symmetric template constructed from the MMRR data
+set\textsuperscript{39} (only shown are the FLAIR, T1, and T2
+modalities--- the components relevant for this work). Template
+construction is detailed in\textsuperscript{25}. These images are
+important for specific intensity-based features.
+
+\textbf{Figure 4:} Average \emph{MeanDecreaseAccuracy} plots generated
+from the creation of all 24 random forest models for Stage 1 during the
+leave-one-out evaluation. These plots are useful in providing a
+quantitative assessment of the predictive importance of each feature.
+Features are ranked in descending order of importance. The horizontal
+error bars provide the \(95^{th}\) percentile and illustrate the
+stability of the feature importance across the leave-one-out models. At
+this initial stage only 31 feature images are used.
+
+\textbf{Figure 5:} Average \emph{MeanDecreaseAccuracy} plots generated
+from the creation of all 24 random forest models for Stage 2 during the
+leave-one-out evaluation. These plots are useful in providing a
+quantitative assessment of the predictive importance of each feature.
+Features are ranked in descending order of importance. The horizontal
+error bars provide the \(95^{th}\) percentile and illustrate the
+stability of the feature importance across the leave-one-out models. We
+augment the 31 feature images from the first stage by adding an
+additional seven voting maps and 7 segmentation posteriors from
+application of the Bayesian-based segmentation for a total of 45 images
+for the second stage.
+
+\textbf{Figure 6:} (a) FLAIR image slice illustrating WMHs which have
+been manually delineated. The region around the WMHs is enlarged (b) in
+the original FLAIR and the (c) contralateral FLAIR difference image.
+
+\textbf{Figure 7:} Voxelwise comparison with manual delineation of white
+matter hyperintensities. On the left are the calculated Dice values over
+all white matter hyperintensities. Note the improvement in the Dice
+metric from the employment of the Stage 2 component of the processing
+pipeline. (Right) Similar results can be seen by comparing the total
+lesion load volume between manual and automated detection strategies.
+Although some outliers are found after the Stage 2 processing in a
+couple subjects, the number of outliers caused by false positives is
+decreased significantly with the second stage processing.
+
+\clearpage
 
@@
 
-\hypertarget{ref-mlmi2015}{}
-11.
-
 \hypertarget{ref-Bauer:2011aa}{}
-12. Bauer S, Nolte L-P, Reyes M. Fully automatic segmentation of brain
+11. Bauer S, Nolte L-P, Reyes M. Fully automatic segmentation of brain
 tumor images using support vector machine classification in combination
@@
 \hypertarget{ref-Tong:2014aa}{}
-13. Tong T, Wolz R, Gao Q, Guerrero R, Hajnal JV, Rueckert D,
+12. Tong T, Wolz R, Gao Q, Guerrero R, Hajnal JV, Rueckert D,
 Alzheimer's Disease Neuroimaging Initiative. Multiple instance learning
@@
 \hypertarget{ref-Liu:2013aa}{}
-14. Liu X, Tosun D, Weiner MW, Schuff N, Alzheimer's Disease
+13. Liu X, Tosun D, Weiner MW, Schuff N, Alzheimer's Disease
 Neuroimaging Initiative. Locally linear embedding (lLE) for mRI based
@@
 \hypertarget{ref-Van-Horn:2014aa}{}
-15. Van Horn JD, Toga AW. Human neuroimaging as a ``big data'' science.
+14. Van Horn JD, Toga AW. Human neuroimaging as a ``big data'' science.
 Brain Imaging Behav. 2014;8(2):323--31.
@@
 \hypertarget{ref-scipy}{}
-16. Jones E, Oliphant T, Peterson P, others. SciPy: Open source
+15. Jones E, Oliphant T, Peterson P, others. SciPy: Open source
 scientific tools for Python. 2001--2001-\/-.
@@
 \hypertarget{ref-R}{}
-17. R Core Team. R: A Language and Environment for Statistical
+16. R Core Team. R: A Language and Environment for Statistical
 Computing. Vienna, Austria: R Foundation for Statistical Computing;
@@
 \hypertarget{ref-breiman2001}{}
-18. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
+17. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
 
 \hypertarget{ref-yi2009}{}
-19. Yi Z, Criminisi A, Shotton J, Blake A. Discriminative, semantic
+18. Yi Z, Criminisi A, Shotton J, Blake A. Discriminative, semantic
 segmentation of brain tissue in MR images. Med Image Comput Comput
@@
 \hypertarget{ref-viola2005}{}
-20. Viola P, Jones M, Snow D. Detecting pedestrians using patterns of
+19. Viola P, Jones M, Snow D. Detecting pedestrians using patterns of
 motion and appearance. International Journal of Computer Vision.
@@
 \hypertarget{ref-geremia2011}{}
-21. Geremia E, Clatz O, Menze BH, Konukoglu E, Criminisi A, Ayache N.
+20. Geremia E, Clatz O, Menze BH, Konukoglu E, Criminisi A, Ayache N.
 Spatial decision forests for MS lesion segmentation in multi-channel
@@
 \hypertarget{ref-Pustina:2016aa}{}
-22. Pustina D, Coslett HB, Turkeltaub PE, Tustison N, Schwartz MF,
+21. Pustina D, Coslett HB, Turkeltaub PE, Tustison N, Schwartz MF,
 Avants B. Automated segmentation of chronic stroke lesions using lINDA:
@@
 \hypertarget{ref-geremia2012}{}
-23. Geremia E, Menze BH, Ayache N. Spatial decision forests for glioma
+22. Geremia E, Menze BH, Ayache N. Spatial decision forests for glioma
 segmentation in multi-channel MR images. In: Proceedings of MICCAI-BRATS
@@
 \hypertarget{ref-bauer2012}{}
-24. Bauer S, Fejes T, Slotboom J, Wiest R, Nolte L-P, Reyes M.
+23. Bauer S, Fejes T, Slotboom J, Wiest R, Nolte L-P, Reyes M.
 Segmentation of brain tumor images based on integrated hierarchical
@@
 \hypertarget{ref-zikic2012}{}
-25. Zikic D, Glocker B, Konukoglu E, Shotton J, Criminisi A, Ye DH,
+24. Zikic D, Glocker B, Konukoglu E, Shotton J, Criminisi A, Ye DH,
 Demiralp C, Thomas OM, Das T, Jena R, et al. Context-sensitive
@@
 \hypertarget{ref-Tustison:2015aa}{}
-26. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee
+25. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee
 JC, Grossman MC, Avants BB. Optimal symmetric multimodal templates and
@@
 \hypertarget{ref-Menze:2015aa}{}
-27. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
+26. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
 Burren Y, Porz N, Slotboom J, Wiest R, et al. The multimodal brain tumor
@@
 \hypertarget{ref-schapire1990}{}
-28. Schapire R. The strength of weak learnability. Machine Learning.
+27. Schapire R. The strength of weak learnability. Machine Learning.
 1990;5:197--227.
@@
 \hypertarget{ref-freund1997}{}
-29. Freund Y, Schapire R. A decision-theoretic generalization of on-line
+28. Freund Y, Schapire R. A decision-theoretic generalization of on-line
 learning and an application to boosting. Journal of Computer and System
@@
 \hypertarget{ref-ho1995}{}
-30. Ho TK. Random decision forests. In: Document analysis and
+29. Ho TK. Random decision forests. In: Document analysis and
 recognition, 1995., proceedings of the third international conference
@@
 \hypertarget{ref-amit1997}{}
-31. Amit Y, Geman D. Shape quantization and recognition with randomized
+30. Amit Y, Geman D. Shape quantization and recognition with randomized
 trees. Neural Computation. 1997;9:1545--1588.
@@
 \hypertarget{ref-Neema:2009aa}{}
-32. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
+31. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
 Normal findings on brain fluid-attenuated inversion recovery mR images
@@
 \hypertarget{ref-Avants:2014aa}{}
-33. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The
+32. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The
 Insight ToolKit image registration framework. Front Neuroinform.
 2014;8:44.
+
+\hypertarget{ref-Manjon:2010aa}{}
+33. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
+non-local means denoising of mR images with spatially varying noise
+levels. J Magn Reson Imaging. 2010;31(1):192--203.
 
@@
 
-\hypertarget{ref-Manjon:2010aa}{}
-35. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
-non-local means denoising of mR images with spatially varying noise
-levels. J Magn Reson Imaging. 2010;31(1):192--203.
+\hypertarget{ref-nyul2000}{}
+35. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
+standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
 
@@
 
-\hypertarget{ref-nyul2000}{}
-37. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
-standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
-
 \hypertarget{ref-Yushkevich:2006aa}{}
-38. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
+37. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
 User-guided 3D active contour segmentation of anatomical structures:
@@
 \hypertarget{ref-Avants:2011aa}{}
-39. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+38. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
 multivariate framework for \(n\)-tissue segmentation with evaluation on
@@
 \hypertarget{ref-landman2011}{}
-40. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+39. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
 Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
@@
 \hypertarget{ref-maurer2003}{}
-41. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
+40. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
 computing exact Euclidean distance transforms of binary images in
@@
 \hypertarget{ref-Tustison:2013ac}{}
-42. Tustison NJ, Avants BB. Explicit B-spline regularization in
+41. Tustison NJ, Avants BB. Explicit B-spline regularization in
 diffeomorphic image registration. Front Neuroinform. 2013;7:39.
@@
 \hypertarget{ref-Avants:2011ab}{}
-43. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+42. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
 reproducible evaluation of ANTs similarity metric performance in brain
@@
 \hypertarget{ref-Anbeek:2004aa}{}
-44. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+43. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
 Probabilistic segmentation of white matter lesions in mR imaging.
@@
 \hypertarget{ref-Garcia-Lorenzo:2013aa}{}
-45. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
+44. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
 Review of automatic segmentation methods of multiple sclerosis white
@@
 \hypertarget{ref-liaw2002}{}
-46. Liaw A, Wiener M. Classification and regression by randomForest. R
+45. Liaw A, Wiener M. Classification and regression by randomForest. R
 News. 2002;2/3:18--22.
@@
 \hypertarget{ref-Grimaud:1996aa}{}
-47. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
+46. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
 DL, Tofts PS, McDonald WI, Miller DH. Quantification of mRI lesion load
@@
 \hypertarget{ref-styner2008}{}
-48. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
+47. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
 S, Jewells V, Warfield S, editors. Special Issue on 2008 MICCAI Workshop
@@
 \hypertarget{ref-tustison2009}{}
-49. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
+48. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
 overlap measures to ITK. Insight Journal. 2009.
@@
 \hypertarget{ref-Plis:2014aa}{}
-50. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
+49. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
 Johnson HJ, Paulsen JS, Turner JA, Calhoun VD. Deep learning for
@@
 \hypertarget{ref-Suk:2015aa}{}
-51. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
+50. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
 Initiative. Deep sparse multi-task learning for feature selection in
@@
 \hypertarget{ref-Li:2014aa}{}
-52. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
+51. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
 based imaging data completion for improved brain disease diagnosis. Med
@@
 \hypertarget{ref-Liu:2015aa}{}
-53. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
+52. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
 ADNI. Multimodal neuroimaging feature learning for multiclass diagnosis
@@
 \hypertarget{ref-HadsellSBESKML09}{}
-54. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
+53. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
 Muller U, LeCun Y. Learning long-range vision for autonomous off-road
@@
 \hypertarget{ref-FarabetCNL12}{}
-55. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
+54. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
 multiscale feature learning, purity trees, and optimal covers. In:
@@
 \hypertarget{ref-Mayer:2014aa}{}
-56. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
+55. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
 identifying subject-specific abnormalities in neuroimaging data. Hum

### 2016-02-25 README.R

@@ -11,7 +11,9 @@ rmdFiles <- c( formatFile,
                "intro.Rmd",
                "methods.Rmd",
                "results.Rmd",
-               "discussion.Rmd"
+               "discussion.Rmd",
+               "acknowledgementsDisclaimers.Rmd",
+               "figureCaptions.Rmd"
    )
 
 for( i in 1:length( rmdFiles ) )

### 2016-04-29 Presentation/wmh.Rmd
ENH:  Add presentation.

## WMH segmentation

![](./Figures/lesions.png)

## Feature images

$label \sim_{RF} feature_{1} + \ldots + feature_{n}$

<BR><BR>

![](../Figures/featureImages.png)

## Feature importance

![](../Figures/averageLeaveOneOut.png)

## Sample results

![](../Figures/sampleResults.png)

### 2016-04-29 Presentation/wmhBuild.Rmd
ENH:  Add presentation.

---
title: ANTs and WMHs
author: James Stone and Nick Tustison
bibliography:
logo:
output:
  beamer_presentation:
    colortheme: dolphin
    fonttheme:  structurebold
    highlight: tango
    incremental: false
    theme:  AnnArbor
    includes:
      in_header: mystyle.tex
    toc: no
  pdf_document:
    toc: true
    highlight: zenburn
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

##  ANTs beginnings:  image registration

<BR><BR>

![](./Figures/highresdiffeos.jpg)

##  ANTs tools

* MRI bias correction
* image denoising
* image and point set registration
* $n$-tissue segmentation
* template building
* multi-atlas label fusion
* brain extraction
* cortical thickness estimation
* ANTsR for statistics and visualization

publicly available:  https://github.com/stnava/ANTs

## Competitions

![](./Figures/evalhistory.png)

## Competition:  BRATS 2013

![](./Figures/bratsData.png)

*_only team out of 20 to make code publicly available._

## WMH segmentation

![](./Figures/lesions.png)

## Feature images

$label \sim_{RF} feature_{1} + \ldots + feature_{n}$

<BR><BR>

![](../Figures/featureImages.png)

## Feature importance

![](../Figures/averageLeaveOneOut.png)

## Sample results

![](../Figures/sampleResults.png)

### 2016-04-29 Presentation/README.R

@@ -1,0 +1,22 @@
+#!/usr/bin/env Rscript
+#
+# ants wip / future work
+#
+library(ggplot2)
+library(shiny)
+library(ggvis)
+library(rmarkdown)
+
+srcdir<-"./"
+buildrmd='wmhBuild.Rmd'
+
+rawrmds <- c( "format.Rmd",
+              "antsBackground.Rmd",
+              "wmh.Rmd"
+               )
+
+for ( x in 1:length(rawrmds) ) {
+  if ( x == 1 )  cmd<-paste( "cat ",rawrmds[x]," > ",buildrmd ) else cmd<-paste( "cat ",rawrmds[x]," >> ",buildrmd )
+  system(cmd)
+}
+render(buildrmd, clean=TRUE )

### 2016-06-22 responseToReviewers.Rmd
ENH:  Respnose to reviewers.

---
output:
  word_document:
    fig_caption: true
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.2}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
bibliography: references.bib
csl: council-of-science-editors.csl
fontsize: 12pt
mainfont: Georgia
---

<style type="text/css">
body,
code.bash{
  font-size: 8px;
}
pre {
  font-size: 8px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set( cache=TRUE )
```

# Response to reviewers

_\textcolor{blue}{We appreciate the time spent by the editors and reviewers in
assessing our manuscript.  Please see below for a point-by-point response to the
issues raised.}_

## Reviewer 1

Thank you for the opportunity to review the paper "Supervised learning
technique for the automated identification of white matter
hyperintensities in traumatic brain injury". In this study, the authors
present a machine learning algorithm for automated segmentation of white
matter hyperintensities, a form of white matter degeneration visible in
FLAIR-MRI that is observed in a wide range of clinical syndromes, and
even in healthy controls. The method proposed by the authors relies on
training random forests to learn the relationship between voxel
intensities in a set of feature images derived from multimodal MRI (T1,
T2, FLAIR) and lesions drawn manually by an expert. Since the method uses
examples to learn what constitutes a lesion, this is correctly considered
a supervised method. The method is applied on a set of 24 patients with
clinically confirmed traumatic brain injury, and results are obtained
with a leave-one-out cross validation.

Overall, the method proposed by the authors is innovative and thoughtful,
and has great potential for bridging the gap between research findings
and clinical needs. The method uses a series of cutting edge solutions to
the segmentation problem: (1) multimodal imaging inputs, which is known
to increase segmentation accuracy, (2) random forests, which are an
excellent tool for finding multivariate nonlinear relationships, (3)
voxel neighborhood information, which provide contextual information for
each voxel, and (4) an algorithm that improves the segmentation in
stages. However, the manuscript itself, to my opinion, is not well
focused and needs significant improvement. Below are my specific comments
in support of this.

> _\textcolor{blue}{We thank the reviewer for the general positive comments concerning our work.  We also
> appreciate the detailed critique below to help us improve specific aspects of the
> presentation.}_

Major comments:

1. The biggest issue I had while reading the manuscript is the lack of a
coherent theme of what this paper is about. The authors span their exposé
from multiple sclerosis to the automobile industry, without stating why
these links to other diseases (or industries) have any relevance. It
appears in some parts that the closest disease with similar white matter
lesions is multiple sclerosis, but this is not clearly stated, nor the
two types of lesions are compared. It is, therefore, unclear why counting
how many MS lesion segmentation methods are publicly available is of any
importance. When explaining methods, the authors mention the utility of
support vector machines, or the existence of a python package called
SciPy. Yet, it is unclear why these things have any importance, unless
the authors make connections to their topic/method.

> _\textcolor{blue}{Nick}_

2. Second, the manuscript has no results. All the results section is focused
on talking about feature ranking, which by the way is explained in the
Results section instead of the Methods section. The most important
results (the prediction accuracy of automated segmentation), are not
mentioned in the manuscript but are only plotted in Figure 7. My advice
is to put the focus of the results section on accuracy, i.e., give
precise dice values and perhaps add other measures (metric displacement,
Hausdorf distance, sensitivity, specificity, etc.). Also, the gain in
accuracy from stage 1 to stage 2 cannot be guessed by looking at a plot,
the authors must provide a statistical comparison. When investigating
accuracy, the error in lesion size should be reported, both in terms of
volume and in terms of individual lesions counts (i.e. out of 10 lesions
9 were found, 1 was missed, etc.). This information is crucial to
clinicians, if the authors foresee the use of the method in a clinical
setting, as it seems from the discussion. Beside adding results to the
manuscript, it is equally important to give some results in the abstract.
The abstract must be self-sufficient in reporting all the major points of
the paper.

> _\textcolor{blue}{Nick}_

3. Third, there is no information about the manual lesion tracing procedure.
Were lesion drawn on FLAIR or were other modalities used as well? If
FLAIR was used as reference it is not surprising that FLAIR emerges as
one of the top predictive modalities. Descriptive statistics on the
lesions are also necessary: how big were the lesions, how many lesions
were found in each patient (the range 1-20 is not sufficiently
informative), where were the lesions located? This information helps
understand why, for example, brain stem or cerebellum labels were not
that important for achieving good lesion predictions.

> _\textcolor{blue}{Nick}_

4. Fourth, although the authors claim the method is available, I couldn't
find any link for the public. Note, it would be equally important if the
publication of the method online is associated with proper documentation
on how a new (naive) user can apply it.

> _\textcolor{blue}{Nick}_

Minor comments:

5. In the abstract, the authors mention the creation of tailored features
for segmenting WMHs. My understanding is that the method proposed by the
authors has no tailoring (if we think of tailoring as custom cutting to
WMH needs). There was no feature selection process all the features were
included as predictors. A similar pitfall if found on pg. 6 ln 56, where
the manuscript reads ``Crucial to these supervised segmentation approaches
are the creation and selection of features as input in conjunction with
expertly identified structures.'' What are the structures identified by
the expert in this study?

6. The methods section is very confusing. The section "Feature images for
WMH segmentation" starts with no explanation of what is Stage 1. The
logical flow would require template registration explained first, and all
feature computation performed later, and finally RF stages can be
explained with the available features. The pipeline also seem to be
repetitive, the N4 correction is mentioned in ln. 26 and then again in
ln. 43, which sounds like it was performed twice. Tissue segmentation is
mentioned in ln. 39, then it is stated is a Bayesian method (ln 45), and
then it is referred to as Atropos in the next page. It is unclear what is
³the first set of images² in pg. 9, or if there is a second set. Also,
mathematical formulas seem superfluous, particularly since it takes more
space to explain the formula than to say in plain English that
transformation matrices were applied to bring subject¹s images in
template space.

7. Random forests are explained in a way that can be understood only by
readers who already know how they work; a new reader (especially a
clinician) would not be able to understand the method. For example, data
pushed onto a tree is does not mean much, unless the reader knows what is
an identification tree.

8. The relationship between WMH and GOS-E is mentioned in pg. 2 but the
reader is not informed what is GOS-E. A similar issue is met few lines
after, where the ``outcome'' is mentioned, but there is no information what
is referred to as outcome.

> \textcolor{blue}{James}

9. In pg. 3 the authors propose that automated methods ``may allow for
identification of correlative patterns between WMH number, volume,
distribution, and disease state''. Some of these correlations have been
already identified in the literature, as the authors mention few lines
before. Moreover, the authors do not report number, volume, or
distribution of the lesions, to back up the utility of their method in
this regard.

10. In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think this
is the right verb.

> _\textcolor{blue}{We agree that the sentence in question could be read this way.
We have, therefore, clarified the sentence in question to read
Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
through each decision tree resulting in a single classification ``vote'' per tree.}_

> A similar issue is in pg 5 ln 52 where it is stated that ``feature images
> consisted of 26 subjects.'' It is not ethical to consider participants as feature
> images.

> _\textcolor{blue}{It was a grammatical mistake, not a question of
ethics.  It has been fixed: ``The feature images
were derived from MR acquisitions of 26 subjects\ldots''}_

11. It is unclear to me the meaning of the sentence: ``Since the inverse
transform is also derived as part of the registration process, we can
warp the voxel index locations back to the space of the individual
subject which motivates similar work by others [43].''

12. Why was Stage 1 segmentation based only on T1, while Stage 2 segmentation
was based on multivariate segmentation using all three modalities?

13. Similar to the feature ranking method, which is explain in Results, Dice
is also explained in Results. Also, it is unclear why finding which
direction of space is informative for the segmentation process would be a
``sanity check.''

14. The discussion starts with an ``intent to further refine the proposed
paradigm,'' and proceeds by mentioning other work currently being
performed which looks promising. The opening of the discussion with such
statements truly make the manuscript sound like an incomplete effort that
is not ready for publication.

> _\textcolor{blue}{We appreciate the positive view of our current work.
The full sentence in question is ``Although only a
single expert was used to produce the manual labelings, our intent is to
further refine the proposed paradigm by crowdsourcing with feedback from
other experts who interact with both the data and methodology.''
In our opinion, simply mentioning specific details regarding future work
does not mean that the manuscript ``sounds like an incomplete effort''.
Rather, it is simply an admission, similar to much of what is reported
in the literature, that our research program is ongoing.}_

15. The discussion also spends several paragraphs on defining literature that
shows how WMHs are related to cognitive performance, or how WMHs can be
applied in diagnosing TBI. Yet, this study makes no effort to test these
efforts in these 24 patients. Consequently, those paragraphs sound
wishful and speculative. Given the methodological nature of this paper, a
more appropriate discussion would perhaps focus on the comparison with
other methods in terms of accuracy, similarities, differences, etc.

> _\textcolor{red}{Get rid of the paragraphs in question.  Need to redo evaluation
to discuss.}_

## Reviewer: 2

Interesting paper which offers those without a current background in
state-of-the-art brain analysis tools. A good introduction to their capabilities.

1. However the authors do little to establish clinical utility of the
algorithm.

> _\textcolor{red}{Similar concerns were voiced by Reviewer 1.  We appreciate
> the opportunity to clarify the precise nature of our contribution which is methodological
> (vs. clinical).  Given the saliency of WMHs as a potential biomarker for TBI
> (as established by previous research), we believe that the automatic identification
> of WMHs is of sufficient merit for publication.  However, our follow-up research plans
> concern the application of our work to a much larger and investigation into the clinical
> correlates of WMH load.}_

2. A graphical description of the pipeline would help most
readers to understand how this might get into the clinic.

> _\textcolor{blue}{Done.  We added the following figure as a graphical description of
> the pipeline to the manuscript:}_

![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
input MRI for a single subject, using the multi-modal symmetric template, results in
the generation of the feature images.  These feature images are used as input to the
Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
voting maps, the original feature images, and the Stage 2 RF model result in the
final voting maps which includes the WMH probability estimate.  Note that the RF models
are constructed once from a set of training data which are processed using the
same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)

### 2016-06-22 responseToReviewers.tex
ENH:  Respnose to reviewers.

--- 2016-06-22 responseToReviewers.Rmd
@@
----
-output:
-  word_document:
-    fig_caption: true
-  pdf_document:
-    fig_caption: true
-    latex_engine: xelatex
-    keep_tex: yes
-header-includes:
-   - \usepackage{booktabs}
-   - \usepackage[final]{changes}
-   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.2}
-   - \usepackage[compact]{titlesec}
-   - \usepackage{enumitem}
-   - \usepackage{tikz}
-   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-   - \setlist{nolistsep}
-   - \titlespacing{\section}{2pt}{*0}{*0}
-   - \titlespacing{\subsection}{2pt}{*0}{*0}
-   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
-   - \setlength{\parskip}{3pt}
-   - \setremarkmarkup{(#2)}
-bibliography: references.bib
-csl: council-of-science-editors.csl
-fontsize: 12pt
-mainfont: Georgia
----
+\documentclass[12pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+    \setmainfont{Georgia}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
 
-<style type="text/css">
-body,
-code.bash{
-  font-size: 8px;
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
     }
-pre {
-  font-size: 8px
 }
-</style>
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set( cache=TRUE )
-```
+\setlength{\droptitle}{-2em}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
 
-# Response to reviewers
+\usepackage{booktabs}
+\usepackage[final]{changes}
+\usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
+\linespread{1.2}
+\usepackage[compact]{titlesec}
+\usepackage{enumitem}
+\usepackage{tikz}
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
+\setlist{nolistsep}
+\titlespacing{\section}{2pt}{*0}{*0}
+\titlespacing{\subsection}{2pt}{*0}{*0}
+\titlespacing{\subsubsection}{2pt}{*0}{*0}
+\setlength{\parskip}{3pt}
+\setremarkmarkup{(#2)}
 
-_\textcolor{blue}{We appreciate the time spent by the editors and reviewers in
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\begin{document}
+\maketitle
+
+\section{Response to reviewers}\label{response-to-reviewers}
+
+\emph{\textcolor{blue}{We appreciate the time spent by the editors and reviewers in
 assessing our manuscript.  Please see below for a point-by-point response to the
-issues raised.}_
+issues raised.}}
 
-## Reviewer 1
+\subsection{Reviewer 1}\label{reviewer-1}
 
-Thank you for the opportunity to review the paper "Supervised learning
+Thank you for the opportunity to review the paper ``Supervised learning
 technique for the automated identification of white matter
-hyperintensities in traumatic brain injury". In this study, the authors
+hyperintensities in traumatic brain injury''. In this study, the authors
 present a machine learning algorithm for automated segmentation of white
@@
 intensities in a set of feature images derived from multimodal MRI (T1,
-T2, FLAIR) and lesions drawn manually by an expert. Since the method uses
-examples to learn what constitutes a lesion, this is correctly considered
-a supervised method. The method is applied on a set of 24 patients with
-clinically confirmed traumatic brain injury, and results are obtained
-with a leave-one-out cross validation.
+T2, FLAIR) and lesions drawn manually by an expert. Since the method
+uses examples to learn what constitutes a lesion, this is correctly
+considered a supervised method. The method is applied on a set of 24
+patients with clinically confirmed traumatic brain injury, and results
+are obtained with a leave-one-out cross validation.
 
-Overall, the method proposed by the authors is innovative and thoughtful,
-and has great potential for bridging the gap between research findings
-and clinical needs. The method uses a series of cutting edge solutions to
-the segmentation problem: (1) multimodal imaging inputs, which is known
-to increase segmentation accuracy, (2) random forests, which are an
-excellent tool for finding multivariate nonlinear relationships, (3)
-voxel neighborhood information, which provide contextual information for
-each voxel, and (4) an algorithm that improves the segmentation in
-stages. However, the manuscript itself, to my opinion, is not well
-focused and needs significant improvement. Below are my specific comments
-in support of this.
+Overall, the method proposed by the authors is innovative and
+thoughtful, and has great potential for bridging the gap between
+research findings and clinical needs. The method uses a series of
+cutting edge solutions to the segmentation problem: (1) multimodal
+imaging inputs, which is known to increase segmentation accuracy, (2)
+random forests, which are an excellent tool for finding multivariate
+nonlinear relationships, (3) voxel neighborhood information, which
+provide contextual information for each voxel, and (4) an algorithm that
+improves the segmentation in stages. However, the manuscript itself, to
+my opinion, is not well focused and needs significant improvement. Below
+are my specific comments in support of this.
 
-> _\textcolor{blue}{We thank the reviewer for the general positive comments concerning our work.  We also
-> appreciate the detailed critique below to help us improve specific aspects of the
-> presentation.}_
+\begin{quote}
+\emph{\textcolor{blue}{We thank the reviewer for the general positive comments concerning our work.  We also
+appreciate the detailed critique below to help us improve specific aspects of the
+presentation.}}
+\end{quote}
 
@@
 
-1. The biggest issue I had while reading the manuscript is the lack of a
-coherent theme of what this paper is about. The authors span their exposé
-from multiple sclerosis to the automobile industry, without stating why
-these links to other diseases (or industries) have any relevance. It
-appears in some parts that the closest disease with similar white matter
-lesions is multiple sclerosis, but this is not clearly stated, nor the
-two types of lesions are compared. It is, therefore, unclear why counting
-how many MS lesion segmentation methods are publicly available is of any
-importance. When explaining methods, the authors mention the utility of
-support vector machines, or the existence of a python package called
-SciPy. Yet, it is unclear why these things have any importance, unless
-the authors make connections to their topic/method.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The biggest issue I had while reading the manuscript is the lack of a
+  coherent theme of what this paper is about. The authors span their
+  exposé from multiple sclerosis to the automobile industry, without
+  stating why these links to other diseases (or industries) have any
+  relevance. It appears in some parts that the closest disease with
+  similar white matter lesions is multiple sclerosis, but this is not
+  clearly stated, nor the two types of lesions are compared. It is,
+  therefore, unclear why counting how many MS lesion segmentation
+  methods are publicly available is of any importance. When explaining
+  methods, the authors mention the utility of support vector machines,
+  or the existence of a python package called SciPy. Yet, it is unclear
+  why these things have any importance, unless the authors make
+  connections to their topic/method.
+\end{enumerate}
 
-> _\textcolor{blue}{Nick}_
+\begin{quote}
+\emph{\textcolor{blue}{Nick}}
+\end{quote}
 
-2. Second, the manuscript has no results. All the results section is focused
-on talking about feature ranking, which by the way is explained in the
-Results section instead of the Methods section. The most important
-results (the prediction accuracy of automated segmentation), are not
-mentioned in the manuscript but are only plotted in Figure 7. My advice
-is to put the focus of the results section on accuracy, i.e., give
-precise dice values and perhaps add other measures (metric displacement,
-Hausdorf distance, sensitivity, specificity, etc.). Also, the gain in
-accuracy from stage 1 to stage 2 cannot be guessed by looking at a plot,
-the authors must provide a statistical comparison. When investigating
-accuracy, the error in lesion size should be reported, both in terms of
-volume and in terms of individual lesions counts (i.e. out of 10 lesions
-9 were found, 1 was missed, etc.). This information is crucial to
-clinicians, if the authors foresee the use of the method in a clinical
-setting, as it seems from the discussion. Beside adding results to the
-manuscript, it is equally important to give some results in the abstract.
-The abstract must be self-sufficient in reporting all the major points of
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Second, the manuscript has no results. All the results section is
+  focused on talking about feature ranking, which by the way is
+  explained in the Results section instead of the Methods section. The
+  most important results (the prediction accuracy of automated
+  segmentation), are not mentioned in the manuscript but are only
+  plotted in Figure 7. My advice is to put the focus of the results
+  section on accuracy, i.e., give precise dice values and perhaps add
+  other measures (metric displacement, Hausdorf distance, sensitivity,
+  specificity, etc.). Also, the gain in accuracy from stage 1 to stage 2
+  cannot be guessed by looking at a plot, the authors must provide a
+  statistical comparison. When investigating accuracy, the error in
+  lesion size should be reported, both in terms of volume and in terms
+  of individual lesions counts (i.e.~out of 10 lesions 9 were found, 1
+  was missed, etc.). This information is crucial to clinicians, if the
+  authors foresee the use of the method in a clinical setting, as it
+  seems from the discussion. Beside adding results to the manuscript, it
+  is equally important to give some results in the abstract. The
+  abstract must be self-sufficient in reporting all the major points of
   the paper.
+\end{enumerate}
 
-> _\textcolor{blue}{Nick}_
+\begin{quote}
+\emph{\textcolor{blue}{Nick}}
+\end{quote}
 
-3. Third, there is no information about the manual lesion tracing procedure.
-Were lesion drawn on FLAIR or were other modalities used as well? If
-FLAIR was used as reference it is not surprising that FLAIR emerges as
-one of the top predictive modalities. Descriptive statistics on the
-lesions are also necessary: how big were the lesions, how many lesions
-were found in each patient (the range 1-20 is not sufficiently
-informative), where were the lesions located? This information helps
-understand why, for example, brain stem or cerebellum labels were not
-that important for achieving good lesion predictions.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Third, there is no information about the manual lesion tracing
+  procedure. Were lesion drawn on FLAIR or were other modalities used as
+  well? If FLAIR was used as reference it is not surprising that FLAIR
+  emerges as one of the top predictive modalities. Descriptive
+  statistics on the lesions are also necessary: how big were the
+  lesions, how many lesions were found in each patient (the range 1-20
+  is not sufficiently informative), where were the lesions located? This
+  information helps understand why, for example, brain stem or
+  cerebellum labels were not that important for achieving good lesion
+  predictions.
+\end{enumerate}
 
-> _\textcolor{blue}{Nick}_
+\begin{quote}
+\emph{\textcolor{blue}{Nick}}
+\end{quote}
 
-4. Fourth, although the authors claim the method is available, I couldn't
-find any link for the public. Note, it would be equally important if the
-publication of the method online is associated with proper documentation
-on how a new (naive) user can apply it.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Fourth, although the authors claim the method is available, I couldn't
+  find any link for the public. Note, it would be equally important if
+  the publication of the method online is associated with proper
+  documentation on how a new (naive) user can apply it.
+\end{enumerate}
 
-> _\textcolor{blue}{Nick}_
+\begin{quote}
+\emph{\textcolor{blue}{Nick}}
+\end{quote}
 
@@
 
-5. In the abstract, the authors mention the creation of tailored features
-for segmenting WMHs. My understanding is that the method proposed by the
-authors has no tailoring (if we think of tailoring as custom cutting to
-WMH needs). There was no feature selection process all the features were
-included as predictors. A similar pitfall if found on pg. 6 ln 56, where
-the manuscript reads ``Crucial to these supervised segmentation approaches
-are the creation and selection of features as input in conjunction with
-expertly identified structures.'' What are the structures identified by
-the expert in this study?
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\item
+  In the abstract, the authors mention the creation of tailored features
+  for segmenting WMHs. My understanding is that the method proposed by
+  the authors has no tailoring (if we think of tailoring as custom
+  cutting to WMH needs). There was no feature selection process all the
+  features were included as predictors. A similar pitfall if found on
+  pg. 6 ln 56, where the manuscript reads ``Crucial to these supervised
+  segmentation approaches are the creation and selection of features as
+  input in conjunction with expertly identified structures.'' What are
+  the structures identified by the expert in this study?
+\item
+  The methods section is very confusing. The section ``Feature images
+  for WMH segmentation'' starts with no explanation of what is Stage 1.
+  The logical flow would require template registration explained first,
+  and all feature computation performed later, and finally RF stages can
+  be explained with the available features. The pipeline also seem to be
+  repetitive, the N4 correction is mentioned in ln. 26 and then again in
+  ln. 43, which sounds like it was performed twice. Tissue segmentation
+  is mentioned in ln. 39, then it is stated is a Bayesian method (ln
+  45), and then it is referred to as Atropos in the next page. It is
+  unclear what is ³the first set of images² in pg. 9, or if there is a
+  second set. Also, mathematical formulas seem superfluous, particularly
+  since it takes more space to explain the formula than to say in plain
+  English that transformation matrices were applied to bring subject¹s
+  images in template space.
+\item
+  Random forests are explained in a way that can be understood only by
+  readers who already know how they work; a new reader (especially a
+  clinician) would not be able to understand the method. For example,
+  data pushed onto a tree is does not mean much, unless the reader knows
+  what is an identification tree.
+\item
+  The relationship between WMH and GOS-E is mentioned in pg. 2 but the
+  reader is not informed what is GOS-E. A similar issue is met few lines
+  after, where the ``outcome'' is mentioned, but there is no information
+  what is referred to as outcome.
+\end{enumerate}
 
-6. The methods section is very confusing. The section "Feature images for
-WMH segmentation" starts with no explanation of what is Stage 1. The
-logical flow would require template registration explained first, and all
-feature computation performed later, and finally RF stages can be
-explained with the available features. The pipeline also seem to be
-repetitive, the N4 correction is mentioned in ln. 26 and then again in
-ln. 43, which sounds like it was performed twice. Tissue segmentation is
-mentioned in ln. 39, then it is stated is a Bayesian method (ln 45), and
-then it is referred to as Atropos in the next page. It is unclear what is
-³the first set of images² in pg. 9, or if there is a second set. Also,
-mathematical formulas seem superfluous, particularly since it takes more
-space to explain the formula than to say in plain English that
-transformation matrices were applied to bring subject¹s images in
-template space.
+\begin{quote}
+\textcolor{blue}{James}
+\end{quote}
 
-7. Random forests are explained in a way that can be understood only by
-readers who already know how they work; a new reader (especially a
-clinician) would not be able to understand the method. For example, data
-pushed onto a tree is does not mean much, unless the reader knows what is
-an identification tree.
-
-8. The relationship between WMH and GOS-E is mentioned in pg. 2 but the
-reader is not informed what is GOS-E. A similar issue is met few lines
-after, where the ``outcome'' is mentioned, but there is no information what
-is referred to as outcome.
-
-> \textcolor{blue}{James}
-
-9. In pg. 3 the authors propose that automated methods ``may allow for
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{8}
+\item
+  In pg. 3 the authors propose that automated methods ``may allow for
   identification of correlative patterns between WMH number, volume,
-distribution, and disease state''. Some of these correlations have been
-already identified in the literature, as the authors mention few lines
-before. Moreover, the authors do not report number, volume, or
+  distribution, and disease state''. Some of these correlations have
+  been already identified in the literature, as the authors mention few
+  lines before. Moreover, the authors do not report number, volume, or
   distribution of the lesions, to back up the utility of their method in
   this regard.
+\item
+  In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think
+  this is the right verb.
+\end{enumerate}
 
-10. In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think this
-is the right verb.
-
-> _\textcolor{blue}{We agree that the sentence in question could be read this way.
+\begin{quote}
+\emph{\textcolor{blue}{We agree that the sentence in question could be read this way.
 We have, therefore, clarified the sentence in question to read
 Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
-through each decision tree resulting in a single classification ``vote'' per tree.}_
+through each decision tree resulting in a single classification ``vote'' per tree.}}
+\end{quote}
 
-> A similar issue is in pg 5 ln 52 where it is stated that ``feature images
-> consisted of 26 subjects.'' It is not ethical to consider participants as feature
-> images.
+\begin{quote}
+A similar issue is in pg 5 ln 52 where it is stated that ``feature
+images consisted of 26 subjects.'' It is not ethical to consider
+participants as feature images.
+\end{quote}
 
-> _\textcolor{blue}{It was a grammatical mistake, not a question of
+\begin{quote}
+\emph{\textcolor{blue}{It was a grammatical mistake, not a question of
 ethics.  It has been fixed: ``The feature images
-were derived from MR acquisitions of 26 subjects\ldots''}_
+were derived from MR acquisitions of 26 subjects\ldots''}}
+\end{quote}
 
-11. It is unclear to me the meaning of the sentence: ``Since the inverse
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{10}
+\item
+  It is unclear to me the meaning of the sentence: ``Since the inverse
   transform is also derived as part of the registration process, we can
   warp the voxel index locations back to the space of the individual
-subject which motivates similar work by others [43].''
+  subject which motivates similar work by others {[}43{]}.''
+\item
+  Why was Stage 1 segmentation based only on T1, while Stage 2
+  segmentation was based on multivariate segmentation using all three
+  modalities?
+\item
+  Similar to the feature ranking method, which is explain in Results,
+  Dice is also explained in Results. Also, it is unclear why finding
+  which direction of space is informative for the segmentation process
+  would be a ``sanity check.''
+\item
+  The discussion starts with an ``intent to further refine the proposed
+  paradigm,'' and proceeds by mentioning other work currently being
+  performed which looks promising. The opening of the discussion with
+  such statements truly make the manuscript sound like an incomplete
+  effort that is not ready for publication.
+\end{enumerate}
 
-12. Why was Stage 1 segmentation based only on T1, while Stage 2 segmentation
-was based on multivariate segmentation using all three modalities?
-
-13. Similar to the feature ranking method, which is explain in Results, Dice
-is also explained in Results. Also, it is unclear why finding which
-direction of space is informative for the segmentation process would be a
-``sanity check.''
-
-14. The discussion starts with an ``intent to further refine the proposed
-paradigm,'' and proceeds by mentioning other work currently being
-performed which looks promising. The opening of the discussion with such
-statements truly make the manuscript sound like an incomplete effort that
-is not ready for publication.
-
-> _\textcolor{blue}{We appreciate the positive view of our current work.
+\begin{quote}
+\emph{\textcolor{blue}{We appreciate the positive view of our current work.
 The full sentence in question is ``Although only a
@@
 Rather, it is simply an admission, similar to much of what is reported
-in the literature, that our research program is ongoing.}_
+in the literature, that our research program is ongoing.}}
+\end{quote}
 
-15. The discussion also spends several paragraphs on defining literature that
-shows how WMHs are related to cognitive performance, or how WMHs can be
-applied in diagnosing TBI. Yet, this study makes no effort to test these
-efforts in these 24 patients. Consequently, those paragraphs sound
-wishful and speculative. Given the methodological nature of this paper, a
-more appropriate discussion would perhaps focus on the comparison with
-other methods in terms of accuracy, similarities, differences, etc.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{14}
+\tightlist
+\item
+  The discussion also spends several paragraphs on defining literature
+  that shows how WMHs are related to cognitive performance, or how WMHs
+  can be applied in diagnosing TBI. Yet, this study makes no effort to
+  test these efforts in these 24 patients. Consequently, those
+  paragraphs sound wishful and speculative. Given the methodological
+  nature of this paper, a more appropriate discussion would perhaps
+  focus on the comparison with other methods in terms of accuracy,
+  similarities, differences, etc.
+\end{enumerate}
 
-> _\textcolor{red}{Get rid of the paragraphs in question.  Need to redo evaluation
-to discuss.}_
+\begin{quote}
+\emph{\textcolor{red}{Get rid of the paragraphs in question.  Need to redo evaluation
+to discuss.}}
+\end{quote}
 
-## Reviewer: 2
+\subsection{Reviewer: 2}\label{reviewer-2}
 
 Interesting paper which offers those without a current background in
-state-of-the-art brain analysis tools. A good introduction to their capabilities.
+state-of-the-art brain analysis tools. A good introduction to their
+capabilities.
 
-1. However the authors do little to establish clinical utility of the
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  However the authors do little to establish clinical utility of the
   algorithm.
+\end{enumerate}
 
-> _\textcolor{red}{Similar concerns were voiced by Reviewer 1.  We appreciate
-> the opportunity to clarify the precise nature of our contribution which is methodological
-> (vs. clinical).  Given the saliency of WMHs as a potential biomarker for TBI
-> (as established by previous research), we believe that the automatic identification
-> of WMHs is of sufficient merit for publication.  However, our follow-up research plans
-> concern the application of our work to a much larger and investigation into the clinical
-> correlates of WMH load.}_
+\begin{quote}
+\emph{\textcolor{red}{Similar concerns were voiced by Reviewer 1.  We appreciate
+the opportunity to clarify the precise nature of our contribution which is methodological
+(vs. clinical).  Given the saliency of WMHs as a potential biomarker for TBI
+(as established by previous research), we believe that the automatic identification
+of WMHs is of sufficient merit for publication.  However, our follow-up research plans
+concern the application of our work to a much larger and investigation into the clinical
+correlates of WMH load.}}
+\end{quote}
 
-2. A graphical description of the pipeline would help most
-readers to understand how this might get into the clinic.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  A graphical description of the pipeline would help most readers to
+  understand how this might get into the clinic.
+\end{enumerate}
 
-> _\textcolor{blue}{Done.  We added the following figure as a graphical description of
-> the pipeline to the manuscript:}_
+\begin{quote}
+\emph{\textcolor{blue}{Done.  We added the following figure as a graphical description of
+the pipeline to the manuscript:}}
+\end{quote}
 
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/wmhPipeline.png}
+\caption{\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
 input MRI for a single subject, using the multi-modal symmetric template, results in
@@
 are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
+same feature-construction pipeline as the single-subject input MRI.}}
+\end{figure}
+
+\hypertarget{refs}{}
+
+\end{document}

### 2016-06-22 figureCaptions.Rmd
ENH:  Respnose to reviewers.

@@
 # Figure Captions
+
+__Figure X:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+input MRI for a single subject, using the multi-modal symmetric template, results in
+the generation of the feature images.  These feature images are used as input to the
+Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
+voting maps, the original feature images, and the Stage 2 RF model result in the
+final voting maps which includes the WMH probability estimate.  Note that the RF models
+are constructed once from a set of training data which are processed using the
+same feature-construction pipeline as the single-subject input MRI.}
 

### 2016-06-22 intro.Rmd
ENH:  Respnose to reviewers.

@@
 Random forests are conceptually straightforward [@breiman2001].  They consist of ensembles of decision trees
-that are built from training data.  Once constructed, data to be classified is "pushed"
+that are built from training data.  Once \textcolor{blue}{the ensemble of decision trees is} constructed, data to be classified is "pushed"
 through each decision tree resulting in a single classification "vote" per tree.  These

### 2016-06-22 methods.Rmd
ENH:  Respnose to reviewers.

@@
 Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images consisted of 26 subjects aged 39.6 ± 8.1 years (range 28 – 58 years).
+The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
 Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
@@
 Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
-3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9°, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 x 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.
+3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
+3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
+In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
 
 ## Quantitative analysis
+
+\textcolor{blue}{Figure X provides a graphical overview of the proposed workflow.  The major components
+include offline generation of symmetric multimodal templates, the creation of
+feature images from the training data which are then used to model the statistical
+prediction using a concatenated random forest framework.  Once these offline steps
+are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+
+<!--
+
+![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+input MRI for a single subject, using the multi-modal symmetric template, results in
+the generation of the feature images.  These feature images are used as input to the
+Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
+voting maps, the original feature images, and the Stage 2 RF model result in the
+final voting maps which includes the WMH probability estimate.  Note that the RF models
+are constructed once from a set of training data which are processed using the
+same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
+
+-->
+
+### Feature images for WMH segmentation
 
@@
 deployment of salient feature images which we detail below.
-
-### Feature images for WMH segmentation
 

### 2016-06-22 stitched.Rmd
ENH:  Respnose to reviewers.

@@
 
+__Figure X:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+input MRI for a single subject, using the multi-modal symmetric template, results in
+the generation of the feature images.  These feature images are used as input to the
+Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
+voting maps, the original feature images, and the Stage 2 RF model result in the
+final voting maps which includes the WMH probability estimate.  Note that the RF models
+are constructed once from a set of training data which are processed using the
+same feature-construction pipeline as the single-subject input MRI.}
+
 __Figure 1:__  Representation of Stage 1 feature images for subject 01C1019.  The

### 2016-06-22 stitched.tex
ENH:  Respnose to reviewers.

@@
 They consist of ensembles of decision trees that are built from training
-data. Once constructed, data to be classified is ``pushed'' through each
-decision tree resulting in a single classification ``vote'' per tree.
-These votes are then used for regression or classification of the data.
-Although decision trees had been extensively studied, the success of
-employing collections of such weak learners for boosting machine
-learning performance (e.g., AdaBoost\textsuperscript{27,28}) influenced
-the similarly sytled conglomeration of decision trees into ``forests''
-with randomized node optimization\textsuperscript{29,30}. Finally,
+data. Once \textcolor{blue}{the ensemble of decision trees is}
+constructed, data to be classified is ``pushed'' through each decision
+tree resulting in a single classification ``vote'' per tree. These votes
+are then used for regression or classification of the data. Although
+decision trees had been extensively studied, the success of employing
+collections of such weak learners for boosting machine learning
+performance (e.g., AdaBoost\textsuperscript{27,28}) influenced the
+similarly sytled conglomeration of decision trees into ``forests'' with
+randomized node optimization\textsuperscript{29,30}. Finally,
 Breiman\textsuperscript{17} improved accuracy by random sampling of
@@
 years with prior combat exposure and deployment(s). The feature images
-consisted of 26 subjects aged 39.6 ± 8.1 years (range 28 -- 58 years).
-Within this cohort, 24 (92\%) were considered positive for TBI based
-upon the potential concussive events (PCE) interview process described
-in detail in Walker et al., this issue). Each of the participants that
-were selected from the larger cohort of participants in this study
-demonstrated at least one white matter hyperintensity (but as many as
-20) on FLAIR imaging.
+\textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged
+39.6 \(\pm\) 8.1 years (range 28--58 years). Within this cohort, 24
+(92\%) were considered positive for TBI based upon the potential
+concussive events (PCE) interview process described in detail in Walker
+et al., this issue). Each of the participants that were selected from
+the larger cohort of participants in this study demonstrated at least
+one white matter hyperintensity (but as many as 20) on FLAIR imaging.
 
@@
 echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal
-slices with a 1.2 mm slice thickness, 256 x 256 acquisition matrix, and
-256 x 256 mm FOV. 3D T1-weighted sequences were acquired with a fast
-field echo (FFE) sequence with the following parameters: TR = 6.8 ms, TE
-= 3.2 ms, echo train length (ETL) = 240; Flip angle = 9°, 170 sagittal
-slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and
-256 x 256 mm FOV. In addition, 3D T2-weighted images were acquired with
-a turbo spin echo sequence with the following parameters: TR = 2500 ms,
-TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice
-thickness, 256 x 256 acquisition matrix, and 256 x 256 mm FOV.
+slices with a 1.2 mm slice thickness, 256 \(\times\) 256 acquisition
+matrix, and 256 \(\times\) 256 mm FOV. 3D T1-weighted sequences were
+acquired with a fast field echo (FFE) sequence with the following
+parameters: TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240;
+Flip angle = 9\(^\circ\), 170 sagittal slices with a 1.2 mm slice
+thickness, 256x240 acquisition matrix, and 256 \(\times\) 256 mm FOV. In
+addition, 3D T2-weighted images were acquired with a turbo spin echo
+sequence with the following parameters: TR = 2500 ms, TE = 245 ms, ET: =
+133; 170 sagittal slices with a 1.2 mm slice thickness, 256 \(\times\)
+256 acquisition matrix, and 256 x 256 mm FOV.
 
 \subsection{Quantitative analysis}\label{quantitative-analysis}
+
+\textcolor{blue}{Figure X provides a graphical overview of the proposed workflow.  The major components
+include offline generation of symmetric multimodal templates, the creation of
+feature images from the training data which are then used to model the statistical
+prediction using a concatenated random forest framework.  Once these offline steps
+are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+
+\subsubsection{Feature images for WMH
+segmentation}\label{feature-images-for-wmh-segmentation}
 
@@
 we detail below.
-
-\subsubsection{Feature images for WMH
-segmentation}\label{feature-images-for-wmh-segmentation}
 
@@
 
+\textbf{Figure X:}
+\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+input MRI for a single subject, using the multi-modal symmetric template, results in
+the generation of the feature images.  These feature images are used as input to the
+Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
+voting maps, the original feature images, and the Stage 2 RF model result in the
+final voting maps which includes the WMH probability estimate.  Note that the RF models
+are constructed once from a set of training data which are processed using the
+same feature-construction pipeline as the single-subject input MRI.}
+
 \textbf{Figure 1:} Representation of Stage 1 feature images for subject

### 2016-06-22 README.R

@@ -30,3 +30,15 @@ for( i in 1:length( rmdFiles ) )
 cat( '\n Pandoc rendering', stitchedFile, '\n' )
 render( stitchedFile, output_format = "all" )
 
+
+singleRmdFiles <- c(
+               "responseToReviewers.Rmd"
+   )
+
+for( i in 1:length( singleRmdFiles ) )
+  {
+  cat( '\n Pandoc rendering', singleRmdFiles[i], '\n' )
+  render( singleRmdFiles[i], output_format = "all" )
+  }
+
+

### 2016-06-24 responseToReviewers.Rmd
ENH:  More responses.

@@
 
-> _\textcolor{blue}{We thank the reviewer for the general positive comments concerning our work.  We also
+> _\textcolor{blue}{We thank the reviewer for the generally positive assessment of our work.  We also
 > appreciate the detailed critique below to help us improve specific aspects of the
@@
 
-> _\textcolor{blue}{Nick}_
+> _\textcolor{blue}{We reduced the introduction to limit the clinical motivation to a
+TBI context and removed mention of multiple sclerosis to remove any assumption of
+disease similarity (although the segmentation approaches can be similar).  We also
+removed the paragraph dealing with publicly available software for neuroimaging analysis.
+}_
 
-2. Second, the manuscript has no results. All the results section is focused
+2. Second, the manuscript has no results. All the results section is [sic] focused
 on talking about feature ranking, which by the way is explained in the
@@
 
-> _\textcolor{blue}{Nick}_
+> _\textcolor{red}{Nick}_
 
@@
 FLAIR was used as reference it is not surprising that FLAIR emerges as
-one of the top predictive modalities. Descriptive statistics on the
-lesions are also necessary: how big were the lesions, how many lesions
-were found in each patient (the range 1-20 is not sufficiently
-informative), where were the lesions located? This information helps
-understand why, for example, brain stem or cerebellum labels were not
-that important for achieving good lesion predictions.
+one of the top predictive modalities.
 
-> _\textcolor{blue}{Nick}_
+> _\textcolor{blue}{We added the following to the Methods section:}_
+
+> \textcolor{blue}{The first author (J.
+R.
+S.) performed the manual WMH tracings for all 26  subjects.
+J.
+R.
+S.
+is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
+All multi-modal MR dicom image slices were converted to the nifti file format.
+Given the utility of the FLAIR sequence in detecting white matter lesions in TBI}  [@Marquez-de-la-Plata:2007aa]\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
+The normalized MRI volumes were then provided to J.
+R.
+S.
+who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
+> \textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
+
+> Descriptive statistics on the lesions are also necessary: how big were the
+> lesions, how many lesions were found in each patient (the range 1-20 is not sufficiently
+> informative), where were the lesions located? This information helps
+> understand why, for example, brain stem or cerebellum labels were not
+> that important for achieving good lesion predictions.
+
+> _\textcolor{red}{Nick---Add a table for each patient and get all this
+information before doing the evaluation.}_
 
@@
 
-> _\textcolor{blue}{Nick}_
+> _\textcolor{blue}{Our methodological publication history demonstrates that we are
+very supportive of open science and have released both code and data in conjunction
+with these publications in public forums, such as github.  However, permission to
+publicly release the data for this data has not yet been obtained and so we removed
+the mention of public availability.  However, when we do get permission, it will
+be posted to the github repository associated with this manuscript
+(https://github.com/ntustison/WatchMeHyperventilate) for future readers.
+}_
 
@@
 authors has no tailoring (if we think of tailoring as custom cutting to
-WMH needs). There was no feature selection process all the features were
-included as predictors. A similar pitfall if found on pg. 6 ln 56, where
-the manuscript reads ``Crucial to these supervised segmentation approaches
-are the creation and selection of features as input in conjunction with
-expertly identified structures.'' What are the structures identified by
-the expert in this study?
+WMH needs).
+
+> _\textcolor{blue}{We removed the word ``tailor'' in its various forms in the
+manuscript to avoid the misreading described by the reviewer.}_
+
+> There was no feature selection process all the features were
+> included as predictors.
+
+> _\textcolor{blue}{Although feature selection is mentioned in the manuscript, nowhere
+do we mention or imply any type of computational feature selection.  There was
+a feature selection process in that we selected ones from the myriad feature images that
+one can possibly produce based on our targeted application, i.e. WMHs.}_
+
+> A similar pitfall if found on pg. 6 ln 56, where
+> the manuscript reads ``Crucial to these supervised segmentation approaches
+> are the creation and selection of features as input in conjunction with
+> expertly identified structures.'' What are the structures identified by
+> the expert in this study?
+
+> _\textcolor{blue}{For this particular application, white matter hyperintensities
+are the structure of interest.  We have re-emphasized this point by writing:
+}_
+
+> \textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
+``features'' as input (i.e., feature images constructed from the training data)
+in conjunction with expertly identified structures of interest
+(i.e., WMHs) for model construction.}
 
@@
 feature computation performed later, and finally RF stages can be
-explained with the available features. The pipeline also seem to be
-repetitive, the N4 correction is mentioned in ln. 26 and then again in
-ln. 43, which sounds like it was performed twice. Tissue segmentation is
-mentioned in ln. 39, then it is stated is a Bayesian method (ln 45), and
-then it is referred to as Atropos in the next page. It is unclear what is
-³the first set of images² in pg. 9, or if there is a second set. Also,
-mathematical formulas seem superfluous, particularly since it takes more
-space to explain the formula than to say in plain English that
-transformation matrices were applied to bring subject¹s images in
-template space.
+explained with the available features.
+
+> _\textcolor{red}{Here.}_
+
+> The pipeline also seem to be
+> repetitive, the N4 correction is mentioned in ln. 26 and then again in
+> ln. 43, which sounds like it was performed twice. Tissue segmentation is
+> mentioned in ln. 39, then it is stated is a Bayesian method (ln 45), and
+> then it is referred to as Atropos in the next page.
+
+> _\textcolor{blue}{The two mentions of N4 are due to the fact that they are associated
+with two distinct pipelines:  1) preparation of the images for extracting feature values
+and 2) brain extration and $n$-tissue segmentation.  We rewrote the section in question
+to avoid any confusing redundancy:}_
+
+> _\textcolor{blue}{The T1 image is then processed via the ANTs brain
+extraction and normal tissue segmentation pipelines}_ [@Tustison:2014ab]
+> _\textcolor{blue}{.  The result is a mask delineating the brain
+parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
+deep gray matter, brain stem, and cerebellum}_ [@Avants:2011aa]
+> _\textcolor{blue}{.  These provide the expertly annotated labels for the
+first six tissue labels given above.  Tissue prior probability maps for segmentation
+are from multi-model optimal symmetric shape/intensity templates}_ [@Tustison:2015aa]
+> _\textcolor{blue}{created from the public MMRR data set}_  [@landman2011]
+> _\textcolor{blue}{(cf Figure 3).}_
+
+> It is unclear what is ``the first set of images'' in pg. 9, or if there is a second set.
+
+> _\textcolor{blue}{We clarified this confusion by changing the sentence in question from}_
+
+> \textcolor{blue}{To model the intensity information the first set of images simply includes the
+> preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.}
+
+> _\textcolor{blue}{to}_
+
+> \textcolor{blue}{Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.}
+
+> Also, mathematical formulas seem superfluous, particularly since it takes more
+> space to explain the formula than to say in plain English that
+> transformation matrices were applied to bring subject's images in
+> template space.
+
+> _\textcolor{blue}{Done.  We removed the mathematical formulas.}_
 
@@
 
+> \textcolor{red}{We confess that we are also ignorant concerning the
+definition of an ``identification tree.''}
+
 8. The relationship between WMH and GOS-E is mentioned in pg. 2 but the
-reader is not informed what is GOS-E. A similar issue is met few lines
+reader is not informed what is GOS-E.
+A similar issue is met few lines
 after, where the ``outcome'' is mentioned, but there is no information what
@@
 
-> \textcolor{blue}{James}
+> _\textcolor{blue}{The sentence in question reads:}_
+
+> \textcolor{blue}{Further, volume of FLAIR lesions within the corpus
+callosum, brainstem, and thalamus in patients with severe TBI correlates
+with Glasgow Outcome-Extended (GOS-E) scores.} [@Moen:2014aa]
+
+> _\textcolor{blue}{Although the full phrase corresponding to the acronym
+in question is provided prior to the acronym in the text (along with an
+explanatory reference), we have added the following clarification:}_
+
+> \textcolor{blue}{Further, volume of FLAIR lesions within the corpus
+callosum, brainstem, and thalamus in patients with severe TBI correlates
+with Glasgow Outcome-Extended (GOS-E) scores} [@Moen:2014aa]
+> \textcolor{blue}{---a numeric groupwise assessment used to
+classify ``outcome'' in TBI patients where ``outcome'' refers to the
+spectrum of possible prognoses from death to disability to recovery.}
 
@@
 
+> _\textcolor{red}{Here.}_
+
 10. In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think this
@@
 We have, therefore, clarified the sentence in question to read
-Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
-through each decision tree resulting in a single classification ``vote'' per tree.}_
+``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
+through each decision tree resulting in a single classification ``vote'' per tree.``}_
 
@@
 
-> _\textcolor{blue}{It was a grammatical mistake, not a question of
+> _\textcolor{blue}{This was a grammatical mistake, not a question of
 ethics.  It has been fixed: ``The feature images
@@
 
+> _\textcolor{blue}{The sentences were rewritten as follows:}_
+
+> \textcolor{blue}{Similar feature images were used in}
+[@Anbeek:2004aa]
+> \textcolor{blue}{although, unlike the proposed framework, this previous work lacks
+> normalization to the standard
+> coordinate system provided by the template to dramatically improve spatial specificity
+> across all subjects.  To generate these images, the T1 image of each subject is
+> registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
+> \textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
+> \textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
+
 12. Why was Stage 1 segmentation based only on T1, while Stage 2 segmentation
@@
 
+> _\textcolor{blue}{We added the following clarification:}_
+
+> \textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation
+process following the voxelwise RF classification of Stage 1, we use all three
+aligned preprocessed images for multivariate segmentation during the second stage.}
+
 13. Similar to the feature ranking method, which is explain in Results, Dice
-is also explained in Results. Also, it is unclear why finding which
-direction of space is informative for the segmentation process would be a
-``sanity check.''
+is also explained in Results.
+
+> _\textcolor{blue}{We removed the equation for the Dice metric and described the
+evaluation procedure in the Methods section.}_
+
+> Also, it is unclear why finding which direction of space is informative for the
+> segmentation process would be a ``sanity check.''
+
+> _\textcolor{blue}{We removed this section to avoid unnecessary confusion.}_
 
@@
 In our opinion, simply mentioning specific details regarding future work
-does not mean that the manuscript ``sounds like an incomplete effort''.
+does not mean that the manuscript ``sounds like an incomplete effort.''
 Rather, it is simply an admission, similar to much of what is reported
@@
 same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
+
+\clearpage
+
+## References

### 2016-06-24 responseToReviewers.tex
ENH:  More responses.

@@
 \begin{quote}
-\emph{\textcolor{blue}{We thank the reviewer for the general positive comments concerning our work.  We also
+\emph{\textcolor{blue}{We thank the reviewer for the generally positive assessment of our work.  We also
 appreciate the detailed critique below to help us improve specific aspects of the
@@
 \begin{quote}
-\emph{\textcolor{blue}{Nick}}
+\emph{\textcolor{blue}{We reduced the introduction to limit the clinical motivation to a
+TBI context and removed mention of multiple sclerosis to remove any assumption of
+disease similarity (although the segmentation approaches can be similar).  We also
+removed the paragraph dealing with publicly available software for neuroimaging analysis.
+}}
 \end{quote}
@@
   Second, the manuscript has no results. All the results section is
-  focused on talking about feature ranking, which by the way is
-  explained in the Results section instead of the Methods section. The
-  most important results (the prediction accuracy of automated
+  {[}sic{]} focused on talking about feature ranking, which by the way
+  is explained in the Results section instead of the Methods section.
+  The most important results (the prediction accuracy of automated
   segmentation), are not mentioned in the manuscript but are only
@@
 \begin{quote}
-\emph{\textcolor{blue}{Nick}}
+\emph{\textcolor{red}{Nick}}
 \end{quote}
@@
   well? If FLAIR was used as reference it is not surprising that FLAIR
-  emerges as one of the top predictive modalities. Descriptive
-  statistics on the lesions are also necessary: how big were the
-  lesions, how many lesions were found in each patient (the range 1-20
-  is not sufficiently informative), where were the lesions located? This
-  information helps understand why, for example, brain stem or
-  cerebellum labels were not that important for achieving good lesion
-  predictions.
+  emerges as one of the top predictive modalities.
 \end{enumerate}
@@
 \begin{quote}
-\emph{\textcolor{blue}{Nick}}
+\emph{\textcolor{blue}{We added the following to the Methods section:}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{The first author (J.
+R.
+S.) performed the manual WMH tracings for all 26  subjects.
+J.
+R.
+S.
+is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
+All multi-modal MR dicom image slices were converted to the nifti file format.
+Given the utility of the FLAIR sequence in detecting white matter lesions in TBI}\textsuperscript{1}\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software}\textsuperscript{2}
+\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{3}
+\textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
+\end{quote}
+
+\begin{quote}
+Descriptive statistics on the lesions are also necessary: how big were
+the lesions, how many lesions were found in each patient (the range 1-20
+is not sufficiently informative), where were the lesions located? This
+information helps understand why, for example, brain stem or cerebellum
+labels were not that important for achieving good lesion predictions.
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{red}{Nick---Add a table for each patient and get all this
+information before doing the evaluation.}}
 \end{quote}
@@
 \begin{quote}
-\emph{\textcolor{blue}{Nick}}
+\emph{\textcolor{blue}{Our methodological publication history demonstrates that we are
+very supportive of open science and have released both code and data in conjunction
+with these publications in public forums, such as github.  However, permission to
+publicly release the data for this data has not yet been obtained and so we removed
+the mention of public availability.  However, when we do get permission, it will
+be posted to the github repository associated with this manuscript
+(https://github.com/ntustison/WatchMeHyperventilate) for future readers.
+}}
 \end{quote}
@@
 \setcounter{enumi}{4}
+\tightlist
 \item
@@
   the authors has no tailoring (if we think of tailoring as custom
-  cutting to WMH needs). There was no feature selection process all the
-  features were included as predictors. A similar pitfall if found on
-  pg. 6 ln 56, where the manuscript reads ``Crucial to these supervised
-  segmentation approaches are the creation and selection of features as
-  input in conjunction with expertly identified structures.'' What are
-  the structures identified by the expert in this study?
+  cutting to WMH needs).
+\end{enumerate}
+
+\begin{quote}
+\emph{\textcolor{blue}{We removed the word ``tailor'' in its various forms in the
+manuscript to avoid the misreading described by the reviewer.}}
+\end{quote}
+
+\begin{quote}
+There was no feature selection process all the features were included as
+predictors.
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{Although feature selection is mentioned in the manuscript, nowhere
+do we mention or imply any type of computational feature selection.  There was
+a feature selection process in that we selected ones from the myriad feature images that
+one can possibly produce based on our targeted application, i.e. WMHs.}}
+\end{quote}
+
+\begin{quote}
+A similar pitfall if found on pg. 6 ln 56, where the manuscript reads
+``Crucial to these supervised segmentation approaches are the creation
+and selection of features as input in conjunction with expertly
+identified structures.'' What are the structures identified by the
+expert in this study?
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{For this particular application, white matter hyperintensities
+are the structure of interest.  We have re-emphasized this point by writing:
+}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
+``features'' as input (i.e., feature images constructed from the training data)
+in conjunction with expertly identified structures of interest
+(i.e., WMHs) for model construction.}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{5}
+\tightlist
 \item
@@
   and all feature computation performed later, and finally RF stages can
-  be explained with the available features. The pipeline also seem to be
-  repetitive, the N4 correction is mentioned in ln. 26 and then again in
-  ln. 43, which sounds like it was performed twice. Tissue segmentation
-  is mentioned in ln. 39, then it is stated is a Bayesian method (ln
-  45), and then it is referred to as Atropos in the next page. It is
-  unclear what is ³the first set of images² in pg. 9, or if there is a
-  second set. Also, mathematical formulas seem superfluous, particularly
-  since it takes more space to explain the formula than to say in plain
-  English that transformation matrices were applied to bring subject¹s
-  images in template space.
+  be explained with the available features.
+\end{enumerate}
+
+\begin{quote}
+\emph{\textcolor{red}{Here.}}
+\end{quote}
+
+\begin{quote}
+The pipeline also seem to be repetitive, the N4 correction is mentioned
+in ln. 26 and then again in ln. 43, which sounds like it was performed
+twice. Tissue segmentation is mentioned in ln. 39, then it is stated is
+a Bayesian method (ln 45), and then it is referred to as Atropos in the
+next page.
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{The two mentions of N4 are due to the fact that they are associated
+with two distinct pipelines:  1) preparation of the images for extracting feature values
+and 2) brain extration and $n$-tissue segmentation.  We rewrote the section in question
+to avoid any confusing redundancy:}}
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{The T1 image is then processed via the ANTs brain
+extraction and normal tissue segmentation pipelines}}\textsuperscript{4}
+\emph{\textcolor{blue}{.  The result is a mask delineating the brain
+parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
+deep gray matter, brain stem, and cerebellum}}\textsuperscript{5}
+\emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
+first six tissue labels given above.  Tissue prior probability maps for segmentation
+are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{6}
+\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{7}
+\emph{\textcolor{blue}{(cf Figure 3).}}
+\end{quote}
+
+\begin{quote}
+It is unclear what is ``the first set of images'' in pg. 9, or if there
+is a second set.
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{We clarified this confusion by changing the sentence in question from}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{To model the intensity information the first set of images simply includes the
+preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.}
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{to}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.}
+\end{quote}
+
+\begin{quote}
+Also, mathematical formulas seem superfluous, particularly since it
+takes more space to explain the formula than to say in plain English
+that transformation matrices were applied to bring subject's images in
+template space.
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{Done.  We removed the mathematical formulas.}}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{6}
+\tightlist
 \item
@@
   what is an identification tree.
+\end{enumerate}
+
+\begin{quote}
+\textcolor{red}{We confess that we are also ignorant concerning the
+definition of an ``identification tree.''}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{7}
+\tightlist
 \item
@@
 \begin{quote}
-\textcolor{blue}{James}
+\emph{\textcolor{blue}{The sentence in question reads:}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
+callosum, brainstem, and thalamus in patients with severe TBI correlates
+with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{8}
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{Although the full phrase corresponding to the acronym
+in question is provided prior to the acronym in the text (along with an
+explanatory reference), we have added the following clarification:}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
+callosum, brainstem, and thalamus in patients with severe TBI correlates
+with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{8}
+\textcolor{blue}{---a numeric groupwise assessment used to
+classify ``outcome'' in TBI patients where ``outcome'' refers to the
+spectrum of possible prognoses from death to disability to recovery.}
 \end{quote}
@@
 \setcounter{enumi}{8}
+\tightlist
 \item
@@
   this regard.
+\end{enumerate}
+
+\begin{quote}
+\emph{\textcolor{red}{Here.}}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{9}
+\tightlist
 \item
@@
 We have, therefore, clarified the sentence in question to read
-Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
-through each decision tree resulting in a single classification ``vote'' per tree.}}
+``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
+through each decision tree resulting in a single classification ``vote'' per tree.``}}
 \end{quote}
@@
 \begin{quote}
-\emph{\textcolor{blue}{It was a grammatical mistake, not a question of
+\emph{\textcolor{blue}{This was a grammatical mistake, not a question of
 ethics.  It has been fixed: ``The feature images
@@
 \setcounter{enumi}{10}
+\tightlist
 \item
@@
   subject which motivates similar work by others {[}43{]}.''
+\end{enumerate}
+
+\begin{quote}
+\emph{\textcolor{blue}{The sentences were rewritten as follows:}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{Similar feature images were used in}\textsuperscript{9}
+\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
+normalization to the standard
+coordinate system provided by the template to dramatically improve spatial specificity
+across all subjects.  To generate these images, the T1 image of each subject is
+registered to the T1 template component using a B-spline variant}\textsuperscript{10}
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{11}.
+\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{11}
+\tightlist
 \item
@@
   modalities?
+\end{enumerate}
+
+\begin{quote}
+\emph{\textcolor{blue}{We added the following clarification:}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation
+process following the voxelwise RF classification of Stage 1, we use all three
+aligned preprocessed images for multivariate segmentation during the second stage.}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{12}
+\tightlist
 \item
   Similar to the feature ranking method, which is explain in Results,
-  Dice is also explained in Results. Also, it is unclear why finding
-  which direction of space is informative for the segmentation process
-  would be a ``sanity check.''
+  Dice is also explained in Results.
+\end{enumerate}
+
+\begin{quote}
+\emph{\textcolor{blue}{We removed the equation for the Dice metric and described the
+evaluation procedure in the Methods section.}}
+\end{quote}
+
+\begin{quote}
+Also, it is unclear why finding which direction of space is informative
+for the segmentation process would be a ``sanity check.''
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{We removed this section to avoid unnecessary confusion.}}
+\end{quote}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{13}
+\tightlist
 \item
@@
 
+\clearpage
+
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
 \hypertarget{refs}{}
+\hypertarget{ref-Marquez-de-la-Plata:2007aa}{}
+1. Marquez de la Plata C, Ardelean A, Koovakkattu D, Srinivasan P,
+Miller A, Phuong V, Harper C, Moore C, Whittemore A, Madden C, et al.
+Magnetic resonance imaging of diffuse axonal injury: Quantitative
+assessment of white matter lesion volume. J Neurotrauma.
+2007;24(4):591--8.
+
+\hypertarget{ref-Avants:2014aa}{}
+2. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
+ToolKit image registration framework. Front Neuroinform. 2014;8:44.
+
+\hypertarget{ref-Yushkevich:2006aa}{}
+3. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
+User-guided 3D active contour segmentation of anatomical structures:
+Significantly improved efficiency and reliability. Neuroimage.
+2006;31(3):1116--28.
+
+\hypertarget{ref-Tustison:2014ab}{}
+4. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
+Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
+and freeSurfer cortical thickness measurements. Neuroimage.
+2014;99:166--79.
+
+\hypertarget{ref-Avants:2011aa}{}
+5. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+multivariate framework for \(n\)-tissue segmentation with evaluation on
+public data. Neuroinformatics. 2011;9(4):381--400.
+
+\hypertarget{ref-Tustison:2015aa}{}
+6. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
+Grossman MC, Avants BB. Optimal symmetric multimodal templates and
+concatenated random forests for supervised brain tumor segmentation
+(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.
+
+\hypertarget{ref-landman2011}{}
+7. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
+reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
+
+\hypertarget{ref-Moen:2014aa}{}
+8. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A. Traumatic
+axonal injury: The prognostic value of lesion load in corpus callosum,
+brain stem, and thalamus in different magnetic resonance imaging
+sequences. J Neurotrauma. 2014;31(17):1486--96.
+
+\hypertarget{ref-Anbeek:2004aa}{}
+9. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+Probabilistic segmentation of white matter lesions in mR imaging.
+Neuroimage. 2004;21(3):1037--44.
+
+\hypertarget{ref-Tustison:2013ac}{}
+10. Tustison NJ, Avants BB. Explicit B-spline regularization in
+diffeomorphic image registration. Front Neuroinform. 2013;7:39.
+
+\hypertarget{ref-Avants:2011ab}{}
+11. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+reproducible evaluation of ANTs similarity metric performance in brain
+image registration. Neuroimage. 2011;54(3):2033--44.
 

### 2016-06-24 discussion.Rmd
ENH:  More responses.

@@
 latter.  As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.  Our
+extensive with a handful of techniques being publicly available.
+
+<!--
+Our
 framework is also available as open-source as part of well-known
@@
 available for such application.
+-->
 

### 2016-06-24 intro.Rmd
ENH:  More responses.

@@
 
-White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-These lesions are distinguished from prominent perivascular spaces seen on T2-weighted imaging by the lack of fluid suppression of signal on FLAIR sequences.
-WMHs in the periventricular and deep brain regions are associated with normal aging and neurological conditions including hypertension and stroke.
-WMHs are also a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult
-[@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa].
-Further, the regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa].
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa].
-Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa].
-Additionally,
+\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
+WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
+
+\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
+Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
+Further, volume of FLAIR lesions within the corpus
+callosum, brainstem, and thalamus in patients with severe TBI correlates
+with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
+classify ``outcome'' in TBI patients where ``outcome'' refers to the
+spectrum of possible prognoses from death to disability to recovery.} [@Moen:2014aa]\textcolor{blue}{. Additionally,
 in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].
+the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].
 
-Despite the above findings, outside of multiple sclerosis, WMHs are not routinely employed as a diagnostic measure in clinical practice.
-Their presence within asymptomatic patients or in association with a variety of conditions, such as stroke, dementia, neuroinflammatory conditions, and TBI challenge their utility in narrowing a radiological differential diagnosis.
-Further, performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-Despite the limited inclusion of WMH observations in routine radiological reports, two large meta-analyses demonstrated an association between WMHs, cognitive function, increased risk of stroke, dementia, and death [@Debette:2010aa;@Kloppenborg:2014aa].
+\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
+Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
 As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state.
 Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
+In this work, we present an automated framework for quantification of WMHs in
+multi-modal MRI using the random forest machine learning technique.}
 
@@
 
+<!--
 Machine learning and pattern recognition techniques have seen increased application
@@
 extensions.
+-->
 
@@
 normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., human gait detection [@viola2005]).
-In the context of neuropathology, random forest-based paradigms have been employed in the
-delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
-and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa] for both
-single and multi-modal acquisition protocols.  Of note, these
+(e.g., [@viola2005]).
+\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
+delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
+\textcolor{blue}{for both single and multi-modal acquisition protocols.}
+
+<!--
+Of note, these
 latter random forest approaches for brain tumor segmentation have performed well in recent
 international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
+-->
 
@@
 
-In this work, we develop a concatenated random forest framework with a tailored contextual
+In this work, we develop a concatenated random forest framework with a
 feature image set (both spatial and intensity-based)
-for segmenting WMHs in a large TBI cohort.  Additionally,
-the entire framework is provided publicly through the well-known open-source
+for segmenting WMHs in a large TBI cohort.
+The entire framework is built on the well-known open-source
 Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.

### 2016-06-24 methods.Rmd
ENH:  More responses.

@@
 In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
+
+\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
+subjects.
+J.
+R.
+S.
+is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
+All
+multi-modal MR dicom image slices were converted to the nifti file format} [^1].
+\textcolor{blue}{Given the
+utility of the FLAIR sequence in detecting white matter lesions in TBI}  [@Marquez-de-la-Plata:2007aa]\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
+The normalized MRI volumes were then provided to
+J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
+\textcolor{blue}{
+which has multi-image overlay capabilities for visualizing all modalities in all three
+canonical views.}
+
+[^1]: http://nifti.nimh.nih.gov/nifti-1
 
@@
 
+\textcolor{blue}{
 Crucial to these supervised segmentation approaches are the creation and selection of
-"features" as input in conjunction with
-expertly identified structures of interest for model construction.  For the targeted application in this work (i.e., WMHs),
+``features'' as input (i.e., feature images constructed from the training data)
+in conjunction with expertly identified structures of interest
+(i.e., WMHs) for model construction.}  For the targeted application in this work,
 tissue classification is performed at the voxelwise level.  In other words, each voxel
@@
 FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
+[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
 (N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
@@
 As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
-acquisitions.  The FLAIR and T2 images are rigidly registered to the T1 image using
+acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
 the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
@@
 extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
-Any effect of pathology is minimal because WMHs are not commonly seen in T1-weighted MRI.
-This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.
-Once the
-brain has been extracted, we apply a Bayesian-based segmentation algorithm
-with a template-based prior probability strategy to segment the
-parenchymal tissue types.  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
+  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
 CSF, gray matter, white matter, deep gray matter, brain stem, and
-cerebellum.  These provide the expertly annotated labels for the first six tissue labels
-given above.  The WMHs were manually identified by one of the authors
-(J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
-the ANTs Atropos tool [@Avants:2011aa] and
-multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
-MMRR data set  [@landman2011] (cf Figure 3).
+cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
+first six tissue labels given above.  Tissue prior probability maps for segmentation
+are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 3).
 
 <!--
-
 ![Canonical views of the mutlivariate, bilaterally symmetric template constructed
@@
 features.](Figures/MMRR.png)
-
 -->
 
-To model the intensity information the first set of images simply includes the
-preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.  We also calculate a set
-of neighborhood statistics (mean, standard deviation, and skewness) feature images using
+Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
+We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
 a Manhattan radius of one voxel given the typical size
-of individual WMHs. For each of the normalized
+of individual WMHs. For each of the preprocessed
 images, we calculate the difference in intensities with the corresponding warped template
@@
 from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 1.  We employ the
-same strategy here.  For example, the template
+same strategy here.
+
+<!--
+For example, the template
 difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
@@
 which maps from the individual subject space to the
-template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
-to take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
+template space and $T_{FLAIR}$ is the FLAIR template component.
+-->
+
+To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
 we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature.  For the FLAIR component, this contralateral
+differences as an additional intensity feature.
+
+<!--
+For the FLAIR component, this contralateral
 difference image is calculated from
@@
 the symmetric template.
+-->
 
-The segmentation probability images described above are used as feature images to provide a
-spatial context for the random forest model prediction step.  Additional spatial contextual
+The segmentation probability images described above are used as feature images to
+provide a spatial context for the random forest model prediction step.  Additional spatial contextual
 feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
@@
 matter interface.  A third set of images are based on the voxel location within the
-space of the template.  The T1 image of the subject is registered to the T1 template
-component using a B-spline variant [@Tustison:2013ac] of the well-known ANTs Symmetric
-Normalization (SyN) algorithm [@Avants:2011ab].  Since the inverse transform is also
-derived as part of the registration process, we can warp the voxel index locations
-back to the space of the individual subject which motivates similar work by others [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
+space of the template.
+\textcolor{blue}{Similar feature images were used in}
+[@Anbeek:2004aa]
+\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
+normalization to the standard
 coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.
+across all subjects.  To generate these images, the T1 image of each subject is
+registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
+\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
 
@@
 are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the Atropos maximum aposteriori algorithm with an
-additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
-the second stage we use all three aligned preprocessed images for a multivariate
-segmentation.  The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
+tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
+additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
+\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
+all three aligned preprocessed images for multivariate segmentation during the
+second stage.}
+The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
 
-### Code and data availability
+### Implementation
 
-As pointed out in a recent comprehensive multiple sclerosis lesion segmentation review
+As pointed out in a recent comprehensive lesion segmentation review
 [@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
@@
 
-In addition, similar to our previous offering [^1], we plan on creating
-a self-encapsulated example to showcase the proposed methodology.  The fact that
+In addition, similar to our previous offering [^2], we plan on creating
+a self-encapsulated example to showcase the proposed methodology
+\textcolor{blue}{which will also be available on github.}[^3]  The fact that
 the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
@@
 
-[^1]: https://github.com/ntustison/ANTsAndArboles
+[^2]: https://github.com/ntustison/ANTsAndArboles
+
+[^3]: https://github.com/ntustison/WatchMeHyperventilate
 
@@
 In order to evaluate the protocol described, we performed a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.  Initial processing included the creation
-of all Stage 1 feature images for all subjects.  The initial brain segmentation of each T1 image and the
-manual white matter hyperintensity tracings were combined to provide the truth labels
-for the training data.  The "truth" labels are the seven anatomical regions given above.
+the data acquired from the 24 subjects described above.
+Initial processing included the creation of all Stage 1 feature images for all subjects.
+The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
+The ``truth'' labels are the seven anatomical regions given above.
 
@@
     + the previous step produces the Stage 1 voting maps for all seven labels.
-    + for each of the 23 subjects, perform a Bayesian-based segmentation with an MRF spatial
+    + for each of the 23 subjects, perform a Bayesian-based segmentation with an
+      MRF spatial
       prior using the seven voting maps as additional tissue priors.

### 2016-06-24 results.Rmd
ENH:  More responses.

@@
 
+<!--
 One can also use these measurements as a type of sanity check.  For example, from the Stage
@@
 template.
+-->
 
@@
 current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.  On the left of Figure 7 are the improvement in Dice values [@tustison2009],
+in training data.
+
+<!--
 i.e.,
@@
 $$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$
+-->
 
+On the left of Figure 7 are the improvement in Dice values [@tustison2009],
 over all white matter hyperintensities when comparing the segmentations between the two stages

### 2016-06-24 stitched.Rmd
ENH:  More responses.

@@
 techniques for automated segmentation workflows.  Herein, we develop a
-concatenated random forest framework with tailored features for
+concatenated random forest framework with image features for
 segmenting WMHs in a TBI cohort. The framework is provided publicly

### 2016-06-24 stitched.tex
ENH:  More responses.

@@
 techniques for automated segmentation workflows. Herein, we develop a
-concatenated random forest framework with tailored features for
-segmenting WMHs in a TBI cohort. The framework is provided publicly
-through the Advanced Normalization Tools (ANTs) and ANTsR toolkits. MR
-(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
-veterans scanned in the Chronic Effects of Neurotrauma Consortium's
-(CENC) observational study were acquired. Manual annotations were
-employed for both training and evaluation using a leave-one-out
-strategy. Lesion load and overlap evaluative comparisons are
-complimented by feature rankings which showcase the utility of the
-concatenated approach. Our findings suggest supervised learning methods
-may be applied to quantify WMHs on routine brain imaging. Paired with
-correlative outcome data, supervised learning methods may allow for
-identification of imaging features predictive of diagnosis and prognosis
-in individual TBI patients.
+concatenated random forest framework with image features for segmenting
+WMHs in a TBI cohort. The framework is provided publicly through the
+Advanced Normalization Tools (ANTs) and ANTsR toolkits. MR (3D FLAIR,
+T2-, and T1-weighted) images from 24 service members and veterans
+scanned in the Chronic Effects of Neurotrauma Consortium's (CENC)
+observational study were acquired. Manual annotations were employed for
+both training and evaluation using a leave-one-out strategy. Lesion load
+and overlap evaluative comparisons are complimented by feature rankings
+which showcase the utility of the concatenated approach. Our findings
+suggest supervised learning methods may be applied to quantify WMHs on
+routine brain imaging. Paired with correlative outcome data, supervised
+learning methods may allow for identification of imaging features
+predictive of diagnosis and prognosis in individual TBI patients.
 
@@
 
-White matter hyperintensities (WMHs) are foci of abnormally increased
-signal intensity seen within white matter regions within the cerebrum
-and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic
-resonance imaging (MRI) sequences. These lesions are distinguished from
-prominent perivascular spaces seen on T2-weighted imaging by the lack of
-fluid suppression of signal on FLAIR sequences. WMHs in the
-periventricular and deep brain regions are associated with normal aging
-and neurological conditions including hypertension and stroke. WMHs are
-also a frequent finding following traumatic brain injury (TBI) and have
-been correlated with functional outcome and injury severity in both
-pediatric\textsuperscript{1,2} and adult\textsuperscript{3--6}. Further,
-the regional distribution and volume of WMHs have been shown to possess
-prognostic value in the TBI patient\textsuperscript{2,6--8}.
-Specifically, lesion volume in corpus callosum correlates with
-functional scores in the acute phase following injury, while lesion
-volume in frontal lobes correlates with scores at 1 year following
-injury\textsuperscript{6}. Further, volume of FLAIR lesions within the
-corpus callosum, brainstem, and thalamus in patients with severe TBI
-correlates with Glasgow Outcome-Extended (GOS-E)
-scores\textsuperscript{4}. Additionally, in patients who are comatose
-following severe TBI the regional distribution of FLAIR lesions within
-the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal,
-occipital lobes, and insula along with the observation of grasping or
-chewing behavior are associated with poor outcome\textsuperscript{7}.
-
-Despite the above findings, outside of multiple sclerosis, WMHs are not
-routinely employed as a diagnostic measure in clinical practice. Their
-presence within asymptomatic patients or in association with a variety
-of conditions, such as stroke, dementia, neuroinflammatory conditions,
-and TBI challenge their utility in narrowing a radiological differential
-diagnosis. Further, performing a comprehensive manual counting of number
-and distribution of lesions in the clinical setting is simply not
-practical. Despite the limited inclusion of WMH observations in routine
-radiological reports, two large meta-analyses demonstrated an
-association between WMHs, cognitive function, increased risk of stroke,
-dementia, and death\textsuperscript{9,10}. As such, the development of
-automated methods for the rapid identification and quantification of
-WMHs within individual patients may allow for identification of
-correlative patterns between WMH number, volume, distribution, and
-disease state. Further, the development of such lesion quantification
-approaches may allow for the practical inclusion of this type of
-information within routine radiological practice.
+\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
+WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric}\textsuperscript{1,2}
+\textcolor{blue}{and adult}\textsuperscript{3--6}
+\textcolor{blue}{populations.}
+
+\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient}\textsuperscript{2,6--8}\textcolor{blue}{.
+Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury}\textsuperscript{6}\textcolor{blue}{.
+Further, volume of FLAIR lesions within the corpus
+callosum, brainstem, and thalamus in patients with severe TBI correlates
+with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
+classify ``outcome'' in TBI patients where ``outcome'' refers to the
+spectrum of possible prognoses from death to disability to recovery.}\textsuperscript{4}\textcolor{blue}{. Additionally,
+in patients who are comatose following severe TBI
+the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome}\textsuperscript{7}.
+
+\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
+Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
+As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state.
+Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
+In this work, we present an automated framework for quantification of WMHs in
+multi-modal MRI using the random forest machine learning technique.}
 
@@
 
-Machine learning and pattern recognition techniques have seen increased
-application for various medical image analysis workflows (see, for
-example, the annual Workshop on Machine Learning in Medical Imaging held
-in conjunction with the Medical Image Computing and Computer-Aided
-Intervention (MICCAI) international meeting). Popular techniques such as
-support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as supervised
-image segmentation (e.g.,\textsuperscript{11}) and diagnostic prediction
-(e.g.,\textsuperscript{12,13}). Facilitating the current employment of
-such techniques are the number of available imaging data
-sets\textsuperscript{14} and the public availability of data science
-packages such as SciPy\textsuperscript{15} and the R project for
-statistical computing\textsuperscript{16} and their associated
-extensions.
-
-The random forests framework\textsuperscript{17} is a popular machine
+The random forests framework\textsuperscript{9} is a popular machine
 learning technique that has demonstrated significant utility for
 supervised segmentation tasks (e.g., normal human brain
-segmentation\textsuperscript{18}) and other computer vision applications
-(e.g., human gait detection\textsuperscript{19}). In the context of
-neuropathology, random forest-based paradigms have been employed in the
-delineation of multiple sclerosis lesions\textsuperscript{20}, stroke
-lesions\textsuperscript{21}, and brain tumors\textsuperscript{22--25}
-for both single and multi-modal acquisition protocols. Of note, these
-latter random forest approaches for brain tumor segmentation have
-performed well in recent international competitions established in
-response to the lack of objective comparisons between segmentation
-algorithms (i.e., the Multimodal Brain Tumor Segmentation (BRATS)
-challenge was initiated in 2012\textsuperscript{26}.
-
-Random forests are conceptually straightforward\textsuperscript{17}.
-They consist of ensembles of decision trees that are built from training
+segmentation\textsuperscript{10}) and other computer vision applications
+(e.g.,\textsuperscript{11}).
+\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
+delineation of other neuropathologies}\textsuperscript{12--17}
+\textcolor{blue}{for both single and multi-modal acquisition protocols.}
+
+Random forests are conceptually straightforward\textsuperscript{9}. They
+consist of ensembles of decision trees that are built from training
 data. Once \textcolor{blue}{the ensemble of decision trees is}
@@
 collections of such weak learners for boosting machine learning
-performance (e.g., AdaBoost\textsuperscript{27,28}) influenced the
+performance (e.g., AdaBoost\textsuperscript{18,19}) influenced the
 similarly sytled conglomeration of decision trees into ``forests'' with
-randomized node optimization\textsuperscript{29,30}. Finally,
-Breiman\textsuperscript{17} improved accuracy by random sampling of
+randomized node optimization\textsuperscript{20,21}. Finally,
+Breiman\textsuperscript{9} improved accuracy by random sampling of
 training data (i.e., ``bagging'') resulting in the current random forest
@@
 In this work, we develop a concatenated random forest framework with a
-tailored contextual feature image set (both spatial and intensity-based)
-for segmenting WMHs in a large TBI cohort. Additionally, the entire
-framework is provided publicly through the well-known open-source
-Advanced Normalization Tools (ANTs)\footnote{\url{https://github.com/stnava/ANTs}}
+feature image set (both spatial and intensity-based) for segmenting WMHs
+in a large TBI cohort. The entire framework is built on the well-known
+open-source Advanced Normalization Tools (ANTs)\footnote{\url{https://github.com/stnava/ANTs}}
 and ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
@@
 
+\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
+subjects.
+J.
+R.
+S.
+is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
+All
+multi-modal MR dicom image slices were converted to the nifti file format}.\footnote{\url{http://nifti.nimh.nih.gov/nifti-1}}
+\textcolor{blue}{Given the
+utility of the FLAIR sequence in detecting white matter lesions in TBI}\textsuperscript{3}\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software}\textsuperscript{22}
+\textcolor{blue}{.  The normalized MRI volumes were then provided to
+J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{23}
+\textcolor{blue}{
+which has multi-image overlay capabilities for visualizing all modalities in all three
+canonical views.}
+
 \subsection{Quantitative analysis}\label{quantitative-analysis}
@@
 
-Crucial to these supervised segmentation approaches are the creation and
-selection of ``features'' as input in conjunction with expertly
-identified structures of interest for model construction. For the
-targeted application in this work (i.e., WMHs), tissue classification is
-performed at the voxelwise level. In other words, each voxel within the
-region of interest is sent through the ensemble of decision trees and
-receives a set of classification votes from each tree thus permitting a
-regression or classification solution. Since this procedure is performed
-at the voxelwise level, intensity information alone is insufficient for
-good segmentation performance due to the lack of spatial context. For
-example, as pointed out in\textsuperscript{31}, higher intensities can
-be found at the periventricular caps in normal subjects which often
-confounds automated lesion detection algorithms. Other potential
-confounds include MR signal inhomogeneity and noise. Therefore, even
-though machine learning and pattern recognition techniques are extremely
-powerful and have significant potential, just as crucial to outcome is
-the creative construction and deployment of salient feature images which
-we detail below.
+\textcolor{blue}{
+Crucial to these supervised segmentation approaches are the creation and selection of
+``features'' as input (i.e., feature images constructed from the training data)
+in conjunction with expertly identified structures of interest
+(i.e., WMHs) for model construction.} For the targeted application in
+this work, tissue classification is performed at the voxelwise level. In
+other words, each voxel within the region of interest is sent through
+the ensemble of decision trees and receives a set of classification
+votes from each tree thus permitting a regression or classification
+solution. Since this procedure is performed at the voxelwise level,
+intensity information alone is insufficient for good segmentation
+performance due to the lack of spatial context. For example, as pointed
+out in\textsuperscript{24}, higher intensities can be found at the
+periventricular caps in normal subjects which often confounds automated
+lesion detection algorithms. Other potential confounds include MR signal
+inhomogeneity and noise. Therefore, even though machine learning and
+pattern recognition techniques are extremely powerful and have
+significant potential, just as crucial to outcome is the creative
+construction and deployment of salient feature images which we detail
+below.
 
@@
 As mentioned previously, input for each subject comprises FLAIR, T1-,
-and T2-weighted acquisitions. The FLAIR and T2 images are rigidly
-registered to the T1 image using the open-source Advanced Normalization
-Tools (ANTs)\textsuperscript{32}. The aligned images are then
-preprocessed using the denoising algorithm of\textsuperscript{33}
-followed by N4 bias correction\textsuperscript{34} which are then
+and T2-weighted acquisitions. The T1 and T2 images are rigidly
+registered to the FLAIR image using the open-source Advanced
+Normalization Tools (ANTs)\textsuperscript{22}. The aligned images are
+then preprocessed using the denoising algorithm of\textsuperscript{25}
+followed by N4 bias correction\textsuperscript{26} which are then
 normalized to the intensity range \([0,1]\). Although we could have used
 an alternative intensity standardization algorithm
-(e.g.,\textsuperscript{35}), we found that a simple linear rescaling
-produced better results similar to previous work\textsuperscript{25}.
+(e.g.,\textsuperscript{27}), we found that a simple linear rescaling
+produced better results similar to previous work\textsuperscript{17}.
 
 The T1 image is then processed via the ANTs brain extraction and normal
-tissue segmentation pipelines\textsuperscript{36}. Any effect of
-pathology is minimal because WMHs are not commonly seen in T1-weighted
-MRI. This protocol involves preprocessing using N4 bias correction
-followed by a template-based strategy for brain extraction. Once the
-brain has been extracted, we apply a Bayesian-based segmentation
-algorithm with a template-based prior probability strategy to segment
-the parenchymal tissue types. The result is a mask delineating the brain
-parenchyma and probabilistic estimates of the CSF, gray matter, white
-matter, deep gray matter, brain stem, and cerebellum. These provide the
-expertly annotated labels for the first six tissue labels given above.
-The WMHs were manually identified by one of the authors (J. R. S.) using
-the ITK-SNAP tool\textsuperscript{37}. Segmentation is performed using
-the ANTs Atropos tool\textsuperscript{38} and multi-model optimal
-symmetric shape/intensity templates\textsuperscript{25} created from the
-public MMRR data set\textsuperscript{39} (cf Figure 3).
-
-To model the intensity information the first set of images simply
-includes the preprocessed and normalized intensity FLAIR, T1, and T2
-image voxel values. We also calculate a set of neighborhood statistics
-(mean, standard deviation, and skewness) feature images using a
-Manhattan radius of one voxel given the typical size of individual WMHs.
-For each of the normalized images, we calculate the difference in
-intensities with the corresponding warped template component. Previous
-success in the international brain tumor segmentation
-competition\textsuperscript{26} was based on an important set of
+tissue segmentation pipelines\textsuperscript{28}. The result is a mask
+delineating the brain parenchyma and probabilistic estimates of the CSF,
+gray matter, white matter, deep gray matter, brain stem, and
+cerebellum\textsuperscript{29}. These provide the expertly annotated
+labels for the first six tissue labels given above. Tissue prior
+probability maps for segmentation are from multi-model optimal symmetric
+shape/intensity templates\textsuperscript{17} created from the public
+MMRR data set\textsuperscript{30} (cf Figure 3).
+
+Feature values include the preprocessed FLAIR, T1, and T2 image voxel
+intensities. We also calculate a set of neighborhood statistics (mean,
+standard deviation, and skewness) feature images using a Manhattan
+radius of one voxel given the typical size of individual WMHs. For each
+of the preprocessed images, we calculate the difference in intensities
+with the corresponding warped template component. Previous success in
+the international brain tumor segmentation
+competition\textsuperscript{31} was based on an important set of
 intensity features that were created from multi-modal templates
-mentioned previously\textsuperscript{25} and listed in Table 1. We
-employ the same strategy here. For example, the template difference
-feature image for the FLAIR image, \(S_{FLAIR}\) is calculated as:
-
-\[S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)\]
-
-where
-
-\[\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T\]
-
-is the transform which maps from the individual subject space to the
-template space and \(T_{FLAIR}\) is the FLAIR template component. Also,
-to take advantage of the gross bilateral symmetry of the normal brain
+mentioned previously\textsuperscript{17} and listed in Table 1. We
+employ the same strategy here.
+
+To take advantage of the gross bilateral symmetry of the normal brain
 (in terms of both shape and intensity), and the fact that WMHs do not
@@
 symmetric templates to compute the contralateral intensity differences
-as an additional intensity feature. For the FLAIR component, this
-contralateral difference image is calculated from
-
-\[S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)\]
-
-where \(\phi_R\) denotes a horizontal reflection perpendicular to the
-mid-sagittal plane of the symmetric template.
+as an additional intensity feature.
 
@@
 prediction step. Additional spatial contextual feature images include
-the distance maps\textsuperscript{40} based on the csf, gray matter, and
+the distance maps\textsuperscript{32} based on the csf, gray matter, and
 deep gray matter images. These latter images are intended to help
@@
 of images are based on the voxel location within the space of the
-template. The T1 image of the subject is registered to the T1 template
-component using a B-spline variant\textsuperscript{41} of the well-known
-ANTs Symmetric Normalization (SyN) algorithm\textsuperscript{42}. Since
-the inverse transform is also derived as part of the registration
-process, we can warp the voxel index locations back to the space of the
-individual subject which motivates similar work by
-others\textsuperscript{43}. However, this previous work lacks the
-normalization to the standard coordinate system provided by the template
-to dramatically improve spatial specificity across all subjects.
+template.
+\textcolor{blue}{Similar feature images were used in}\textsuperscript{33}
+\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
+normalization to the standard
+coordinate system provided by the template to dramatically improve spatial specificity
+across all subjects.  To generate these images, the T1 image of each subject is
+registered to the T1 template component using a B-spline variant}\textsuperscript{34}
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{35}.
+\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
 
@@
 
-In previous brain tumor segmentation work\textsuperscript{25}, it was
+In previous brain tumor segmentation work\textsuperscript{17}, it was
 demonstrated that a concatenated supervised approach, whereby the
@@
 estimates of the tissue spatial locations) for a second application of
-the Atropos maximum aposteriori algorithm with an additional Markov
-Random Field spatial prior (MAP-MRF)\textsuperscript{38}. However, for
-the second stage we use all three aligned preprocessed images for a
-multivariate segmentation. The resulting seven posterior probability
-images constitute a third additional feature image set for Stage 2.
-
-\subsubsection{Code and data
-availability}\label{code-and-data-availability}
-
-As pointed out in a recent comprehensive multiple sclerosis lesion
-segmentation review\textsuperscript{44}, although the number of
-algorithms reported in the literature is quite extensive, there were
-only four publicly available segmentation algorithms at the time of
-writing this article. In contrast to the current work, none are based on
-supervised learning. As we did for our brain tumor segmentation
-algorithm\textsuperscript{25}, all of the code described in this work is
+the \(6\)-tissue segmentation algorithm with an additional Markov Random
+Field spatial prior (MAP-MRF)\textsuperscript{29}.
+\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
+all three aligned preprocessed images for multivariate segmentation during the
+second stage.} The resulting seven posterior probability images
+constitute a third additional feature image set for Stage 2.
+
+\subsubsection{Implementation}\label{implementation}
+
+As pointed out in a recent comprehensive lesion segmentation
+review\textsuperscript{36}, although the number of algorithms reported
+in the literature is quite extensive, there were only four publicly
+available segmentation algorithms at the time of writing this article.
+In contrast to the current work, none are based on supervised learning.
+As we did for our brain tumor segmentation
+algorithm\textsuperscript{17}, all of the code described in this work is
 publicly available through the open-source ANTs/ANTsR toolkits. Through
@@
 statistical project) we use the \emph{randomForest}
-package\textsuperscript{45} using the default settings with 2000 trees
+package\textsuperscript{37} using the default settings with 2000 trees
 per model and 500 randomly selected samples per label per image. Note
@@
 we plan on creating a self-encapsulated example to showcase the proposed
-methodology. The fact that the data will also be made available through
-the Federal Interagency Traumatic Brain Injury Research (FITBIR)
-repository along with the manual labelings will facilitate
-reproducibility on the part of the reader as well as any interest in
-extending the proposed framework to other data sets.
+methodology
+\textcolor{blue}{which will also be available on github.}\footnote{\url{https://github.com/ntustison/WatchMeHyperventilate}}
+The fact that the data will also be made available through the Federal
+Interagency Traumatic Brain Injury Research (FITBIR) repository along
+with the manual labelings will facilitate reproducibility on the part of
+the reader as well as any interest in extending the proposed framework
+to other data sets.
 
@@
 feature types and names.
-
-One can also use these measurements as a type of sanity check. For
-example, from the Stage 1 plot, one can see that the
-\emph{MeanDecreaseAccuracy} values for the location indices in the
-anterior-posterior direction (i.e., \emph{TemplateIndicesWarped1}) are
-greater than those for either the inferior-superior (i.e.,
-\emph{TemplateIndicesWarped2}) or the left-right (i.e.,
-\emph{TemplateIndicesWarped0}) directions in the space of the symmetric
-template.
 
@@
 characteristic with manual labelings in related
-fields\textsuperscript{44,46,47}, such labelings are characteristic of
+fields\textsuperscript{36,38,39}, such labelings are characteristic of
 current clinical practices and the methodology proposed herein is
-readily adapted to refinements in training data. On the left of Figure 7
-are the improvement in Dice values\textsuperscript{48}, i.e.,
-
-\[ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } \]
-
-over all white matter hyperintensities when comparing the segmentations
-between the two stages where the sum is taken over all individually
-labeled manual, \(T_r\), and automated, \(S_r\), lesions and \(\cap\)
-represents the intersection between the manual/automated lesion pair.
-Performing the second round of supervised learning improves these Dice
-values. One can also note from the right side of Figure 6 that the total
-lesion load volume illustrates a few subjects that are severe outliers
-in terms of the number of false positives. The second round helps to
-correct this issue.
+readily adapted to refinements in training data.
+
+On the left of Figure 7 are the improvement in Dice
+values\textsuperscript{40}, over all white matter hyperintensities when
+comparing the segmentations between the two stages where the sum is
+taken over all individually labeled manual, \(T_r\), and automated,
+\(S_r\), lesions and \(\cap\) represents the intersection between the
+manual/automated lesion pair. Performing the second round of supervised
+learning improves these Dice values. One can also note from the right
+side of Figure 6 that the total lesion load volume illustrates a few
+subjects that are severe outliers in terms of the number of false
+positives. The second round helps to correct this issue.
 
@@
 latter. As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available. Our
-framework is also available as open-source as part of well-known
-neuroimaging tools which easily allows for additions/extensions but is
-also, as far as we know, the first random forest-based technique
-available for such application.
+extensive with a handful of techniques being publicly available.
 
 Two major meta-analyses of WMHs have been published covering the periods
-prior to 2010\textsuperscript{9} and after\textsuperscript{10}. The
+prior to 2010\textsuperscript{41} and after\textsuperscript{42}. The
 earlier meta-analysis covered 53 longitudinal studies that included
@@
 samples representative of the general population are more relevant to
-the focus of the present paper. Debette \& Markus\textsuperscript{9}
+the focus of the present paper. Debette \& Markus\textsuperscript{41}
 found that the presence of WMHs was related to subsequent cognitive
@@
 
-The more recent meta-analysis by Kloppenborg et al.\textsuperscript{10}
+The more recent meta-analysis by Kloppenborg et al.\textsuperscript{42}
 of 23 cross-sectional studies reporting MRI and concurrent
@@
 across a wide variety of settings for the rapid identification of
-predictive imaging features\textsuperscript{49--52}. Automobile
+predictive imaging features\textsuperscript{43--46}. Automobile
 manufacturers utilize these types of approaches to equip self-driving
@@
 through the identification of visual information sufficiently similar to
-previously assimilated training data\textsuperscript{53,54}. Similarly,
+previously assimilated training data\textsuperscript{47,48}. Similarly,
 in the context of the neuroimaging assessments, deep learning approaches
@@
 applied to the segmentation of macroscopically visible
-structures\textsuperscript{49--52}. Additionally, these approaches might
+structures\textsuperscript{43--46}. Additionally, these approaches might
 be applied to the interrogation of imaging data in the individual
@@
 analytical tools to assess the information in the setting of the
-individual patient\textsuperscript{55}. Application of deep learning
+individual patient\textsuperscript{49}. Application of deep learning
 approaches in the context of data with primary quantitative outputs will
@@
 01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
-pre-aligned\textsuperscript{32} to the space of the T1 image. The three
+pre-aligned\textsuperscript{22} to the space of the T1 image. The three
 modality images are then preprocessed (N4 bias
-correction\textsuperscript{34} and adaptive
-denoising\textsuperscript{33}) followed by application of standard ANTs
+correction\textsuperscript{26} and adaptive
+denoising\textsuperscript{25}) followed by application of standard ANTs
 brain extraction and \(n\)-tissue segmentation protocols using the MMRR
-symmetric template and corresponding priors\textsuperscript{36} applied
+symmetric template and corresponding priors\textsuperscript{28} applied
 to the T1 image. The feature images are then generated for voxelwise
@@
 symmetric template constructed from the MMRR data
-set\textsuperscript{39} (only shown are the FLAIR, T1, and T2
+set\textsuperscript{30} (only shown are the FLAIR, T1, and T2
 modalities--- the components relevant for this work). Template
-construction is detailed in\textsuperscript{25}. These images are
+construction is detailed in\textsuperscript{17}. These images are
 important for specific intensity-based features.
@@
 
-\hypertarget{ref-Debette:2010aa}{}
-9. Debette S, Markus HS. The clinical importance of white matter
-hyperintensities on brain magnetic resonance imaging: Systematic review
-and meta-analysis. BMJ. 2010;341:c3666.
-
-\hypertarget{ref-Kloppenborg:2014aa}{}
-10. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
-Presence and progression of white matter hyperintensities and cognition:
-A meta-analysis. Neurology. 2014;82(23):2127--38.
-
-\hypertarget{ref-Bauer:2011aa}{}
-11. Bauer S, Nolte L-P, Reyes M. Fully automatic segmentation of brain
-tumor images using support vector machine classification in combination
-with hierarchical conditional random field regularization. Med Image
-Comput Comput Assist Interv. 2011;14(Pt 3):354--61.
-
-\hypertarget{ref-Tong:2014aa}{}
-12. Tong T, Wolz R, Gao Q, Guerrero R, Hajnal JV, Rueckert D,
-Alzheimer's Disease Neuroimaging Initiative. Multiple instance learning
-for classification of dementia in brain mRI. Med Image Anal.
-2014;18(5):808--18.
-
-\hypertarget{ref-Liu:2013aa}{}
-13. Liu X, Tosun D, Weiner MW, Schuff N, Alzheimer's Disease
-Neuroimaging Initiative. Locally linear embedding (lLE) for mRI based
-alzheimer's disease classification. Neuroimage. 2013;83:148--57.
-
-\hypertarget{ref-Van-Horn:2014aa}{}
-14. Van Horn JD, Toga AW. Human neuroimaging as a ``big data'' science.
-Brain Imaging Behav. 2014;8(2):323--31.
-
-\hypertarget{ref-scipy}{}
-15. Jones E, Oliphant T, Peterson P, others. SciPy: Open source
-scientific tools for Python. 2001--2001-\/-.
-
-\hypertarget{ref-R}{}
-16. R Core Team. R: A Language and Environment for Statistical
-Computing. Vienna, Austria: R Foundation for Statistical Computing;
-2016.
-
 \hypertarget{ref-breiman2001}{}
-17. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
+9. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
 
 \hypertarget{ref-yi2009}{}
-18. Yi Z, Criminisi A, Shotton J, Blake A. Discriminative, semantic
+10. Yi Z, Criminisi A, Shotton J, Blake A. Discriminative, semantic
 segmentation of brain tissue in MR images. Med Image Comput Comput
@@
 \hypertarget{ref-viola2005}{}
-19. Viola P, Jones M, Snow D. Detecting pedestrians using patterns of
+11. Viola P, Jones M, Snow D. Detecting pedestrians using patterns of
 motion and appearance. International Journal of Computer Vision.
@@
 \hypertarget{ref-geremia2011}{}
-20. Geremia E, Clatz O, Menze BH, Konukoglu E, Criminisi A, Ayache N.
+12. Geremia E, Clatz O, Menze BH, Konukoglu E, Criminisi A, Ayache N.
 Spatial decision forests for MS lesion segmentation in multi-channel
@@
 \hypertarget{ref-Pustina:2016aa}{}
-21. Pustina D, Coslett HB, Turkeltaub PE, Tustison N, Schwartz MF,
+13. Pustina D, Coslett HB, Turkeltaub PE, Tustison N, Schwartz MF,
 Avants B. Automated segmentation of chronic stroke lesions using lINDA:
@@
 \hypertarget{ref-geremia2012}{}
-22. Geremia E, Menze BH, Ayache N. Spatial decision forests for glioma
+14. Geremia E, Menze BH, Ayache N. Spatial decision forests for glioma
 segmentation in multi-channel MR images. In: Proceedings of MICCAI-BRATS
@@
 \hypertarget{ref-bauer2012}{}
-23. Bauer S, Fejes T, Slotboom J, Wiest R, Nolte L-P, Reyes M.
+15. Bauer S, Fejes T, Slotboom J, Wiest R, Nolte L-P, Reyes M.
 Segmentation of brain tumor images based on integrated hierarchical
@@
 \hypertarget{ref-zikic2012}{}
-24. Zikic D, Glocker B, Konukoglu E, Shotton J, Criminisi A, Ye DH,
+16. Zikic D, Glocker B, Konukoglu E, Shotton J, Criminisi A, Ye DH,
 Demiralp C, Thomas OM, Das T, Jena R, et al. Context-sensitive
@@
 \hypertarget{ref-Tustison:2015aa}{}
-25. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee
+17. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee
 JC, Grossman MC, Avants BB. Optimal symmetric multimodal templates and
@@
 
+\hypertarget{ref-schapire1990}{}
+18. Schapire R. The strength of weak learnability. Machine Learning.
+1990;5:197--227.
+
+\hypertarget{ref-freund1997}{}
+19. Freund Y, Schapire R. A decision-theoretic generalization of on-line
+learning and an application to boosting. Journal of Computer and System
+Sciences. 1997;55:119--139.
+
+\hypertarget{ref-ho1995}{}
+20. Ho TK. Random decision forests. In: Document analysis and
+recognition, 1995., proceedings of the third international conference
+on. Vol. 1. 1995. pp. 278--282 vol.1.
+
+\hypertarget{ref-amit1997}{}
+21. Amit Y, Geman D. Shape quantization and recognition with randomized
+trees. Neural Computation. 1997;9:1545--1588.
+
+\hypertarget{ref-Avants:2014aa}{}
+22. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The
+Insight ToolKit image registration framework. Front Neuroinform.
+2014;8:44.
+
+\hypertarget{ref-Yushkevich:2006aa}{}
+23. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
+User-guided 3D active contour segmentation of anatomical structures:
+Significantly improved efficiency and reliability. Neuroimage.
+2006;31(3):1116--28.
+
+\hypertarget{ref-Neema:2009aa}{}
+24. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
+Normal findings on brain fluid-attenuated inversion recovery mR images
+at 3T. AJNR Am J Neuroradiol. 2009;30(5):911--6.
+
+\hypertarget{ref-Manjon:2010aa}{}
+25. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
+non-local means denoising of mR images with spatially varying noise
+levels. J Magn Reson Imaging. 2010;31(1):192--203.
+
+\hypertarget{ref-Tustison:2010ac}{}
+26. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
+JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
+2010;29(6):1310--20.
+
+\hypertarget{ref-nyul2000}{}
+27. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
+standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
+
+\hypertarget{ref-Tustison:2014ab}{}
+28. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
+Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
+and freeSurfer cortical thickness measurements. Neuroimage.
+2014;99:166--79.
+
+\hypertarget{ref-Avants:2011aa}{}
+29. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+multivariate framework for \(n\)-tissue segmentation with evaluation on
+public data. Neuroinformatics. 2011;9(4):381--400.
+
+\hypertarget{ref-landman2011}{}
+30. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
+reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
+
 \hypertarget{ref-Menze:2015aa}{}
-26. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
+31. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
 Burren Y, Porz N, Slotboom J, Wiest R, et al. The multimodal brain tumor
@@
 
-\hypertarget{ref-schapire1990}{}
-27. Schapire R. The strength of weak learnability. Machine Learning.
-1990;5:197--227.
-
-\hypertarget{ref-freund1997}{}
-28. Freund Y, Schapire R. A decision-theoretic generalization of on-line
-learning and an application to boosting. Journal of Computer and System
-Sciences. 1997;55:119--139.
-
-\hypertarget{ref-ho1995}{}
-29. Ho TK. Random decision forests. In: Document analysis and
-recognition, 1995., proceedings of the third international conference
-on. Vol. 1. 1995. pp. 278--282 vol.1.
-
-\hypertarget{ref-amit1997}{}
-30. Amit Y, Geman D. Shape quantization and recognition with randomized
-trees. Neural Computation. 1997;9:1545--1588.
-
-\hypertarget{ref-Neema:2009aa}{}
-31. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
-Normal findings on brain fluid-attenuated inversion recovery mR images
-at 3T. AJNR Am J Neuroradiol. 2009;30(5):911--6.
-
-\hypertarget{ref-Avants:2014aa}{}
-32. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The
-Insight ToolKit image registration framework. Front Neuroinform.
-2014;8:44.
-
-\hypertarget{ref-Manjon:2010aa}{}
-33. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
-non-local means denoising of mR images with spatially varying noise
-levels. J Magn Reson Imaging. 2010;31(1):192--203.
-
-\hypertarget{ref-Tustison:2010ac}{}
-34. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
-JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
-2010;29(6):1310--20.
-
-\hypertarget{ref-nyul2000}{}
-35. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
-standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
-
-\hypertarget{ref-Tustison:2014ab}{}
-36. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
-Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
-and freeSurfer cortical thickness measurements. Neuroimage.
-2014;99:166--79.
-
-\hypertarget{ref-Yushkevich:2006aa}{}
-37. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
-User-guided 3D active contour segmentation of anatomical structures:
-Significantly improved efficiency and reliability. Neuroimage.
-2006;31(3):1116--28.
-
-\hypertarget{ref-Avants:2011aa}{}
-38. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
-multivariate framework for \(n\)-tissue segmentation with evaluation on
-public data. Neuroinformatics. 2011;9(4):381--400.
-
-\hypertarget{ref-landman2011}{}
-39. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
-Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
-reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
-
 \hypertarget{ref-maurer2003}{}
-40. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
+32. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
 computing exact Euclidean distance transforms of binary images in
@@
 
+\hypertarget{ref-Anbeek:2004aa}{}
+33. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+Probabilistic segmentation of white matter lesions in mR imaging.
+Neuroimage. 2004;21(3):1037--44.
+
 \hypertarget{ref-Tustison:2013ac}{}
-41. Tustison NJ, Avants BB. Explicit B-spline regularization in
+34. Tustison NJ, Avants BB. Explicit B-spline regularization in
 diffeomorphic image registration. Front Neuroinform. 2013;7:39.
@@
 \hypertarget{ref-Avants:2011ab}{}
-42. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+35. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
 reproducible evaluation of ANTs similarity metric performance in brain
@@
 
-\hypertarget{ref-Anbeek:2004aa}{}
-43. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
-Probabilistic segmentation of white matter lesions in mR imaging.
-Neuroimage. 2004;21(3):1037--44.
-
 \hypertarget{ref-Garcia-Lorenzo:2013aa}{}
-44. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
+36. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
 Review of automatic segmentation methods of multiple sclerosis white
@@
 \hypertarget{ref-liaw2002}{}
-45. Liaw A, Wiener M. Classification and regression by randomForest. R
+37. Liaw A, Wiener M. Classification and regression by randomForest. R
 News. 2002;2/3:18--22.
@@
 \hypertarget{ref-Grimaud:1996aa}{}
-46. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
+38. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
 DL, Tofts PS, McDonald WI, Miller DH. Quantification of mRI lesion load
@@
 \hypertarget{ref-styner2008}{}
-47. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
+39. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
 S, Jewells V, Warfield S, editors. Special Issue on 2008 MICCAI Workshop
@@
 \hypertarget{ref-tustison2009}{}
-48. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
+40. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
 overlap measures to ITK. Insight Journal. 2009.
 
+\hypertarget{ref-Debette:2010aa}{}
+41. Debette S, Markus HS. The clinical importance of white matter
+hyperintensities on brain magnetic resonance imaging: Systematic review
+and meta-analysis. BMJ. 2010;341:c3666.
+
+\hypertarget{ref-Kloppenborg:2014aa}{}
+42. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
+Presence and progression of white matter hyperintensities and cognition:
+A meta-analysis. Neurology. 2014;82(23):2127--38.
+
 \hypertarget{ref-Plis:2014aa}{}
-49. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
+43. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
 Johnson HJ, Paulsen JS, Turner JA, Calhoun VD. Deep learning for
@@
 \hypertarget{ref-Suk:2015aa}{}
-50. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
+44. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
 Initiative. Deep sparse multi-task learning for feature selection in
@@
 \hypertarget{ref-Li:2014aa}{}
-51. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
+45. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
 based imaging data completion for improved brain disease diagnosis. Med
@@
 \hypertarget{ref-Liu:2015aa}{}
-52. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
+46. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
 ADNI. Multimodal neuroimaging feature learning for multiclass diagnosis
@@
 \hypertarget{ref-HadsellSBESKML09}{}
-53. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
+47. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
 Muller U, LeCun Y. Learning long-range vision for autonomous off-road
@@
 \hypertarget{ref-FarabetCNL12}{}
-54. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
+48. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
 multiscale feature learning, purity trees, and optimal covers. In:
@@
 \hypertarget{ref-Mayer:2014aa}{}
-55. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
+49. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
 identifying subject-specific abnormalities in neuroimaging data. Hum

### 2016-06-27 responseToReviewers.Rmd
ENH:  More response items.

@@
 
-2. Second, the manuscript has no results. All the results section is [sic] focused
+2. Second, the manuscript has no results. All the results section is focused
 on talking about feature ranking, which by the way is explained in the
@@
 
-> _\textcolor{red}{Here.}_
+> _\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
+>   It can be seen below in response to item 2 of Reviewer 2.  We reference
+>   this figure numerous times in the text as a visual explanatory aid.}_
+> _\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
+>   which gives an overview of the methods section including an description of Stage 1 and Stage2.}_
+>
+>   \textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
+include offline generation of symmetric multimodal templates, the creation of
+feature images from the training data which are then employed for statistical
+prediction using a concatenated random forest framework.  This framework involves
+the use of two random forest models where the outcome (i.e., the tissue
+probability estimates) of the first RF model application,
+which we denote as ``Stage1 '', is used as input (along with the original set of
+feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
+Once these offline steps
+are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+>
+> _\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}_
+>
+> \textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
+> serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
+> For this work we use the multi-modal data available from the public MMRR data set}
+> [@landman2011]\textcolor{blue}{.
+We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
+> results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
+>
+> _\textcolor{blue}{These changes complement the edits made in response to other review items.}_
 
@@
 
-> \textcolor{red}{We confess that we are also ignorant concerning the
-definition of an ``identification tree.''}
+> _\textcolor{blue}{
+We revised the explanation of random forests as follows:}_
+
+> \textcolor{blue}{The basic component of
+the random forest paradigm is the ``decision tree'' often represented by a flowchart or
+graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
+The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
+For the proposed application, individual
+voxels (and their corresponding feature values) are introduced at the root of a particular
+decision tree and traverse the edges and internal nodes ultimately ending up at one of
+the classification nodes according to the tests at each internal node.  A single random
+forest model will consist of many such trees (often refered to as an ``ensemble'').}
+> \textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
+(e.g., AdaBoost)} [@schapire1990;@freund1997]
+>  \textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
+\textcolor{blue}{Finally, Breiman} [@breiman2001]
+\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
+in the current random forest technique applied here.}
+> \textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
+each voxel location.}
 
@@
 
+> _\textcolor{blue}{We thank the reviewer for the positive comments regarding our work.}_
+
 1. However the authors do little to establish clinical utility of the
@@
 
-> _\textcolor{red}{Similar concerns were voiced by Reviewer 1.  We appreciate
+> _\textcolor{blue}{We appreciate
 > the opportunity to clarify the precise nature of our contribution which is methodological
-> (vs. clinical).  Given the saliency of WMHs as a potential biomarker for TBI
-> (as established by previous research), we believe that the automatic identification
-> of WMHs is of sufficient merit for publication.  However, our follow-up research plans
-> concern the application of our work to a much larger and investigation into the clinical
-> correlates of WMH load.}_
+> (vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
+clinical research involving WMHs and then pointing to the need for automated
+techniques to minimize the variability and time requirements of manual
+segmentation of WMHs.  This would allow for large-scale studies and potentially more accurate
+and reproducible quantitation.
+}_
 

### 2016-06-27 responseToReviewers.tex
ENH:  More response items.

@@
   Second, the manuscript has no results. All the results section is
-  {[}sic{]} focused on talking about feature ranking, which by the way
-  is explained in the Results section instead of the Methods section.
-  The most important results (the prediction accuracy of automated
+  focused on talking about feature ranking, which by the way is
+  explained in the Results section instead of the Methods section. The
+  most important results (the prediction accuracy of automated
   segmentation), are not mentioned in the manuscript but are only
@@
 \begin{quote}
-\emph{\textcolor{red}{Here.}}
+\emph{\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
+  It can be seen below in response to item 2 of Reviewer 2.  We reference
+  this figure numerous times in the text as a visual explanatory aid.}}
+\emph{\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
+  which gives an overview of the methods section including an description of Stage 1 and Stage2.}}
+
+\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
+include offline generation of symmetric multimodal templates, the creation of
+feature images from the training data which are then employed for statistical
+prediction using a concatenated random forest framework.  This framework involves
+the use of two random forest models where the outcome (i.e., the tissue
+probability estimates) of the first RF model application,
+which we denote as ``Stage1 '', is used as input (along with the original set of
+feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
+Once these offline steps
+are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+
+\emph{\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template.}}
+
+\textcolor{blue}{Following}\textsuperscript{4}
+\textcolor{blue}{and}\textsuperscript{5},
+\textcolor{blue}{optimally derived templates
+serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
+For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{6}\textcolor{blue}{.
+We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{7}
+\textcolor{blue}{which
+results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
 \end{quote}
@@
 parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
-deep gray matter, brain stem, and cerebellum}}\textsuperscript{5}
+deep gray matter, brain stem, and cerebellum}}\textsuperscript{8}
 \emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
 first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{6}
-\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{7}
+are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{5}
+\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{6}
 \emph{\textcolor{blue}{(cf Figure 3).}}
@@
 \begin{quote}
-\textcolor{red}{We confess that we are also ignorant concerning the
-definition of an ``identification tree.''}
+\emph{\textcolor{blue}{
+We revised the explanation of random forests as follows:}}
+\end{quote}
+
+\begin{quote}
+\textcolor{blue}{The basic component of
+the random forest paradigm is the ``decision tree'' often represented by a flowchart or
+graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
+The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
+For the proposed application, individual
+voxels (and their corresponding feature values) are introduced at the root of a particular
+decision tree and traverse the edges and internal nodes ultimately ending up at one of
+the classification nodes according to the tests at each internal node.  A single random
+forest model will consist of many such trees (often refered to as an ``ensemble'').}
+\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
+(e.g., AdaBoost)}\textsuperscript{9,10}
+\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{11,12}.
+\textcolor{blue}{Finally, Breiman}\textsuperscript{13}
+\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
+in the current random forest technique applied here.}
+\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
+each voxel location.}
 \end{quote}
@@
 callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{8}
+with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{14}
 \end{quote}
@@
 callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{8}
+with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{14}
 \textcolor{blue}{---a numeric groupwise assessment used to
@@
 \begin{quote}
-\textcolor{blue}{Similar feature images were used in}\textsuperscript{9}
+\textcolor{blue}{Similar feature images were used in}\textsuperscript{15}
 \textcolor{blue}{although, unlike the proposed framework, this previous work lacks
@@
 across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant}\textsuperscript{10}
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{11}.
+registered to the T1 template component using a B-spline variant}\textsuperscript{16}
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{17}.
 \textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
@@
 
+\begin{quote}
+\emph{\textcolor{blue}{We thank the reviewer for the positive comments regarding our work.}}
+\end{quote}
+
 \begin{enumerate}
@@
 \begin{quote}
-\emph{\textcolor{red}{Similar concerns were voiced by Reviewer 1.  We appreciate
+\emph{\textcolor{blue}{We appreciate
 the opportunity to clarify the precise nature of our contribution which is methodological
-(vs. clinical).  Given the saliency of WMHs as a potential biomarker for TBI
-(as established by previous research), we believe that the automatic identification
-of WMHs is of sufficient merit for publication.  However, our follow-up research plans
-concern the application of our work to a much larger and investigation into the clinical
-correlates of WMH load.}}
+(vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
+clinical research involving WMHs and then pointing to the need for automated
+techniques to minimize the variability and time requirements of manual
+segmentation of WMHs.  This would allow for large-scale studies and potentially more accurate
+and reproducible quantitation.
+}}
 \end{quote}
@@
 
-\hypertarget{ref-Avants:2011aa}{}
-5. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
-multivariate framework for \(n\)-tissue segmentation with evaluation on
-public data. Neuroinformatics. 2011;9(4):381--400.
-
 \hypertarget{ref-Tustison:2015aa}{}
-6. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
+5. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
 Grossman MC, Avants BB. Optimal symmetric multimodal templates and
@@
 \hypertarget{ref-landman2011}{}
-7. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+6. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
 Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
@@
 
+\hypertarget{ref-Avants:2010aa}{}
+7. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre J,
+Gee JC. The optimal template effect in hippocampus studies of diseased
+populations. Neuroimage. 2010;49(3):2457--66.
+
+\hypertarget{ref-Avants:2011aa}{}
+8. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+multivariate framework for \(n\)-tissue segmentation with evaluation on
+public data. Neuroinformatics. 2011;9(4):381--400.
+
+\hypertarget{ref-schapire1990}{}
+9. Schapire R. The strength of weak learnability. Machine Learning.
+1990;5:197--227.
+
+\hypertarget{ref-freund1997}{}
+10. Freund Y, Schapire R. A decision-theoretic generalization of on-line
+learning and an application to boosting. Journal of Computer and System
+Sciences. 1997;55:119--139.
+
+\hypertarget{ref-ho1995}{}
+11. Ho TK. Random decision forests. In: Document analysis and
+recognition, 1995., proceedings of the third international conference
+on. Vol. 1. 1995. pp. 278--282 vol.1.
+
+\hypertarget{ref-amit1997}{}
+12. Amit Y, Geman D. Shape quantization and recognition with randomized
+trees. Neural Computation. 1997;9:1545--1588.
+
+\hypertarget{ref-breiman2001}{}
+13. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
+
 \hypertarget{ref-Moen:2014aa}{}
-8. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A. Traumatic
-axonal injury: The prognostic value of lesion load in corpus callosum,
-brain stem, and thalamus in different magnetic resonance imaging
-sequences. J Neurotrauma. 2014;31(17):1486--96.
+14. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A.
+Traumatic axonal injury: The prognostic value of lesion load in corpus
+callosum, brain stem, and thalamus in different magnetic resonance
+imaging sequences. J Neurotrauma. 2014;31(17):1486--96.
 
 \hypertarget{ref-Anbeek:2004aa}{}
-9. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+15. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
 Probabilistic segmentation of white matter lesions in mR imaging.
@@
 \hypertarget{ref-Tustison:2013ac}{}
-10. Tustison NJ, Avants BB. Explicit B-spline regularization in
+16. Tustison NJ, Avants BB. Explicit B-spline regularization in
 diffeomorphic image registration. Front Neuroinform. 2013;7:39.
@@
 \hypertarget{ref-Avants:2011ab}{}
-11. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+17. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
 reproducible evaluation of ANTs similarity metric performance in brain

### 2016-06-27 abstract.Rmd
ENH:  More responses.
ENH:  More response items.

@@
 techniques for automated segmentation workflows.  Herein, we develop a
-concatenated random forest framework with tailored features for
-segmenting WMHs in a TBI cohort. The framework is provided publicly
-through the Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
+concatenated random forest framework with image features for
+segmenting WMHs in a TBI cohort. The framework is built upon the
+Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
 (3D FLAIR, T2-, and T1-weighted) images from 24 service members and

### 2016-06-27 figureCaptions.Rmd
ENH:  More response items.

@@
 
-__Figure X:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+__Figure 1:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
 input MRI for a single subject, using the multi-modal symmetric template, results in
@@
 
-__Figure 1:__  Representation of Stage 1 feature images for subject 01C1019.  The
+__Figure 2:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
+from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
+the components relevant for this work).  Template construction is detailed in
+[@Tustison:2015aa].  These images are important for asymmetry-based
+features.
+
+__Figure 3:__  Representation of Stage 1 feature images for subject 01C1019.  The
 FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
@@
 
-__Figure 2:__  Sample FLAIR acquisition image slices showing both manual and random forest
+__Figure 4:__  Sample FLAIR acquisition image slices showing both manual and random forest
 segmentations for both stages obtained during the leave-one-out evaluation.  Manual
@@
 
-__Figure 3:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific intensity-based
-features.
-
-__Figure 4:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
+__Figure 5:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
 forest models for Stage 1 during the leave-one-out evaluation.  These plots
@@
 
-__Figure 5:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
+__Figure 6:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
 forest models for Stage 2 during the leave-one-out evaluation.  These plots
@@
 
-__Figure 6:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
+__Figure 7:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
 The region around the WMHs is enlarged (b) in the original FLAIR and the
@@
 
-__Figure 7:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
+__Figure 8:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
 On the left are the

### 2016-06-27 intro.Rmd
ENH:  More response items.

@@
 
-Random forests are conceptually straightforward [@breiman2001].  They consist of ensembles of decision trees
-that are built from training data.  Once \textcolor{blue}{the ensemble of decision trees is} constructed, data to be classified is "pushed"
-through each decision tree resulting in a single classification "vote" per tree.  These
-votes are then used for regression or classification of the data.  Although decision
-trees had been extensively studied, the success of employing collections of
-such weak learners for boosting machine learning performance
-(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
-of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
-Finally, Breiman [@breiman2001] improved accuracy by random sampling of training
-data (i.e., "bagging") resulting in the current random forest technique applied here.
+Random forests are conceptually straightforward [@breiman2001].
+\textcolor{blue}{The basic component of
+the random forest paradigm is the ``decision tree'' often represented by a flowchart or
+graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
+The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
+For the proposed application, individual
+voxels (and their corresponding feature values) are introduced at the root of a particular
+decision tree and traverse the edges and internal nodes ultimately ending up at one of
+the classification nodes according to the tests at each internal node.  A single random
+forest model will consist of many such trees (often refered to as an ``ensemble'').}
+\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
+(e.g., AdaBoost)} [@schapire1990;@freund1997]
+\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
+\textcolor{blue}{Finally, Breiman} [@breiman2001]
+\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
+in the current random forest technique applied here.}
+\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
+each voxel location.}
 

### 2016-06-27 methods.Rmd
ENH:  More response items.

@@
 
-\textcolor{blue}{Figure X provides a graphical overview of the proposed workflow.  The major components
+\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
 include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then used to model the statistical
-prediction using a concatenated random forest framework.  Once these offline steps
+feature images from the training data which are then employed for statistical
+prediction using a concatenated random forest framework.  This framework involves
+the use of two random forest models where the outcome (i.e., the tissue
+probability estimates) of the first RF model application,
+which we denote as ``Stage1 '', is used as input (along with the original set of
+feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
+Once these offline steps
 are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
@@
 same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
+
+-->
+
+### Symmetric multi-modal templates
+
+\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
+serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
+For this work we use the multi-modal data available from the public MMRR data set}
+[@landman2011]\textcolor{blue}{.
+We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
+results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
+
+<!--
+![Canonical views of the mutlivariate, bilaterally symmetric template constructed
+from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
+the components relevant for this work).  Template construction is detailed in
+[@Tustison:2015aa].  These images are important for specific asymmetry-based
+features.](Figures/MMRR.png)
 
@@
 of multiple random forest stages below) segmentation of the WMHs.
-In addition Figure 1 provides a representation of a set of feature images for a single
+In addition Figure 3 provides a representation of a set of feature images for a single
 subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
@@
 first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 3).
-
-<!--
-![Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific intensity-based
-features.](Figures/MMRR.png)
--->
+are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 2).
 

### 2016-06-27 results.Rmd
ENH:  More response items.

@@
 
-The resulting rankings for both Stages are given in Figures 4 and 5 where the values for the
+The resulting rankings for both Stages are given in Figures 5 and 6 where the values for the
 separate stages are averaged over the entire corresponding model set.  In addition, we
@@
 The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 6).  This accords with the known clinical relevance of
+evaluation random forest models (see Figure 7).  This accords with the known clinical relevance of
 FLAIR images for identifying white matter hyperintensities and the fact that such
@@
 1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 5, the voting maps are labeled as
+an additional MRF spatial prior.  In Figure 6, the voting maps are labeled as
 "_RFStage1VotingMaps_" where the final numeral is associated with the brain
@@
 
-In Figure 7 are the segmentation comparisons derived from manual segmentations of the same
+In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
 data.  Despite the large variability characteristic with manual labelings in related fields
@@
 
-On the left of Figure 7 are the improvement in Dice values [@tustison2009],
+On the left of Figure 8 are the improvement in Dice values [@tustison2009],
 over all white matter hyperintensities when comparing the segmentations between the two stages
@@
 the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 6 that the total lesion load volume illustrates a few subjects that are
+right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
 severe outliers in terms of the number of false positives.  The second round helps to

### 2016-06-27 stitched.Rmd
ENH:  More response items.

@@
 concatenated random forest framework with image features for
-segmenting WMHs in a TBI cohort. The framework is provided publicly
-through the Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
+segmenting WMHs in a TBI cohort. The framework is built upon the
+Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
 (3D FLAIR, T2-, and T1-weighted) images from 24 service members and

### 2016-06-27 stitched.tex
ENH:  More response items.

@@
 concatenated random forest framework with image features for segmenting
-WMHs in a TBI cohort. The framework is provided publicly through the
-Advanced Normalization Tools (ANTs) and ANTsR toolkits. MR (3D FLAIR,
-T2-, and T1-weighted) images from 24 service members and veterans
-scanned in the Chronic Effects of Neurotrauma Consortium's (CENC)
-observational study were acquired. Manual annotations were employed for
-both training and evaluation using a leave-one-out strategy. Lesion load
-and overlap evaluative comparisons are complimented by feature rankings
-which showcase the utility of the concatenated approach. Our findings
-suggest supervised learning methods may be applied to quantify WMHs on
-routine brain imaging. Paired with correlative outcome data, supervised
-learning methods may allow for identification of imaging features
-predictive of diagnosis and prognosis in individual TBI patients.
+WMHs in a TBI cohort. The framework is built upon the Advanced
+Normalization Tools (ANTs) and ANTsR toolkits. MR (3D FLAIR, T2-, and
+T1-weighted) images from 24 service members and veterans scanned in the
+Chronic Effects of Neurotrauma Consortium's (CENC) observational study
+were acquired. Manual annotations were employed for both training and
+evaluation using a leave-one-out strategy. Lesion load and overlap
+evaluative comparisons are complimented by feature rankings which
+showcase the utility of the concatenated approach. Our findings suggest
+supervised learning methods may be applied to quantify WMHs on routine
+brain imaging. Paired with correlative outcome data, supervised learning
+methods may allow for identification of imaging features predictive of
+diagnosis and prognosis in individual TBI patients.
 
@@
 
-Random forests are conceptually straightforward\textsuperscript{9}. They
-consist of ensembles of decision trees that are built from training
-data. Once \textcolor{blue}{the ensemble of decision trees is}
-constructed, data to be classified is ``pushed'' through each decision
-tree resulting in a single classification ``vote'' per tree. These votes
-are then used for regression or classification of the data. Although
-decision trees had been extensively studied, the success of employing
-collections of such weak learners for boosting machine learning
-performance (e.g., AdaBoost\textsuperscript{18,19}) influenced the
-similarly sytled conglomeration of decision trees into ``forests'' with
-randomized node optimization\textsuperscript{20,21}. Finally,
-Breiman\textsuperscript{9} improved accuracy by random sampling of
-training data (i.e., ``bagging'') resulting in the current random forest
-technique applied here.
+Random forests are conceptually straightforward\textsuperscript{9}.
+\textcolor{blue}{The basic component of
+the random forest paradigm is the ``decision tree'' often represented by a flowchart or
+graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
+The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
+For the proposed application, individual
+voxels (and their corresponding feature values) are introduced at the root of a particular
+decision tree and traverse the edges and internal nodes ultimately ending up at one of
+the classification nodes according to the tests at each internal node.  A single random
+forest model will consist of many such trees (often refered to as an ``ensemble'').}
+\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
+(e.g., AdaBoost)}\textsuperscript{18,19}
+\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{20,21}.
+\textcolor{blue}{Finally, Breiman}\textsuperscript{9}
+\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
+in the current random forest technique applied here.}
+\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
+each voxel location.}
 
@@
 
-\textcolor{blue}{Figure X provides a graphical overview of the proposed workflow.  The major components
+\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
 include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then used to model the statistical
-prediction using a concatenated random forest framework.  Once these offline steps
+feature images from the training data which are then employed for statistical
+prediction using a concatenated random forest framework.  This framework involves
+the use of two random forest models where the outcome (i.e., the tissue
+probability estimates) of the first RF model application,
+which we denote as ``Stage1 '', is used as input (along with the original set of
+feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
+Once these offline steps
 are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+
+\subsubsection{Symmetric multi-modal
+templates}\label{symmetric-multi-modal-templates}
+
+\textcolor{blue}{Following}\textsuperscript{24}
+\textcolor{blue}{and}\textsuperscript{17},
+\textcolor{blue}{optimally derived templates
+serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
+For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{25}\textcolor{blue}{.
+We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{26}
+\textcolor{blue}{which
+results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
 
@@
 performance due to the lack of spatial context. For example, as pointed
-out in\textsuperscript{24}, higher intensities can be found at the
+out in\textsuperscript{27}, higher intensities can be found at the
 periventricular caps in normal subjects which often confounds automated
@@
 random forest stages below) segmentation of the WMHs. In addition Figure
-1 provides a representation of a set of feature images for a single
+3 provides a representation of a set of feature images for a single
 subject analyzed in this work. Note that in this work we categorize the
@@
 Normalization Tools (ANTs)\textsuperscript{22}. The aligned images are
-then preprocessed using the denoising algorithm of\textsuperscript{25}
-followed by N4 bias correction\textsuperscript{26} which are then
+then preprocessed using the denoising algorithm of\textsuperscript{28}
+followed by N4 bias correction\textsuperscript{29} which are then
 normalized to the intensity range \([0,1]\). Although we could have used
 an alternative intensity standardization algorithm
-(e.g.,\textsuperscript{27}), we found that a simple linear rescaling
+(e.g.,\textsuperscript{30}), we found that a simple linear rescaling
 produced better results similar to previous work\textsuperscript{17}.
@@
 The T1 image is then processed via the ANTs brain extraction and normal
-tissue segmentation pipelines\textsuperscript{28}. The result is a mask
+tissue segmentation pipelines\textsuperscript{24}. The result is a mask
 delineating the brain parenchyma and probabilistic estimates of the CSF,
 gray matter, white matter, deep gray matter, brain stem, and
-cerebellum\textsuperscript{29}. These provide the expertly annotated
+cerebellum\textsuperscript{31}. These provide the expertly annotated
 labels for the first six tissue labels given above. Tissue prior
@@
 shape/intensity templates\textsuperscript{17} created from the public
-MMRR data set\textsuperscript{30} (cf Figure 3).
+MMRR data set\textsuperscript{25} (cf Figure 2).
 
@@
 the international brain tumor segmentation
-competition\textsuperscript{31} was based on an important set of
+competition\textsuperscript{32} was based on an important set of
 intensity features that were created from multi-modal templates
@@
 prediction step. Additional spatial contextual feature images include
-the distance maps\textsuperscript{32} based on the csf, gray matter, and
+the distance maps\textsuperscript{33} based on the csf, gray matter, and
 deep gray matter images. These latter images are intended to help
@@
 template.
-\textcolor{blue}{Similar feature images were used in}\textsuperscript{33}
+\textcolor{blue}{Similar feature images were used in}\textsuperscript{34}
 \textcolor{blue}{although, unlike the proposed framework, this previous work lacks
@@
 across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant}\textsuperscript{34}
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{35}.
+registered to the T1 template component using a B-spline variant}\textsuperscript{35}
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{36}.
 \textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
@@
 the \(6\)-tissue segmentation algorithm with an additional Markov Random
-Field spatial prior (MAP-MRF)\textsuperscript{29}.
+Field spatial prior (MAP-MRF)\textsuperscript{31}.
 \textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
@@
 As pointed out in a recent comprehensive lesion segmentation
-review\textsuperscript{36}, although the number of algorithms reported
+review\textsuperscript{37}, although the number of algorithms reported
 in the literature is quite extensive, there were only four publicly
@@
 statistical project) we use the \emph{randomForest}
-package\textsuperscript{37} using the default settings with 2000 trees
+package\textsuperscript{38} using the default settings with 2000 trees
 per model and 500 randomly selected samples per label per image. Note
@@
 
-The resulting rankings for both Stages are given in Figures 4 and 5
+The resulting rankings for both Stages are given in Figures 5 and 6
 where the values for the separate stages are averaged over the entire
@@
 discriminative over the set of evaluation random forest models (see
-Figure 6). This accords with the known clinical relevance of FLAIR
+Figure 7). This accords with the known clinical relevance of FLAIR
 images for identifying white matter hyperintensities and the fact that
@@
 input to a Bayesian-based segmentation with an additional MRF spatial
-prior. In Figure 5, the voting maps are labeled as
+prior. In Figure 6, the voting maps are labeled as
 ``\emph{RFStage1VotingMaps}'' where the final numeral is associated with
@@
 
-In Figure 7 are the segmentation comparisons derived from manual
+In Figure 8 are the segmentation comparisons derived from manual
 segmentations of the same data. Despite the large variability
 characteristic with manual labelings in related
-fields\textsuperscript{36,38,39}, such labelings are characteristic of
+fields\textsuperscript{37,39,40}, such labelings are characteristic of
 current clinical practices and the methodology proposed herein is
@@
 
-On the left of Figure 7 are the improvement in Dice
-values\textsuperscript{40}, over all white matter hyperintensities when
+On the left of Figure 8 are the improvement in Dice
+values\textsuperscript{41}, over all white matter hyperintensities when
 comparing the segmentations between the two stages where the sum is
@@
 learning improves these Dice values. One can also note from the right
-side of Figure 6 that the total lesion load volume illustrates a few
+side of Figure 7 that the total lesion load volume illustrates a few
 subjects that are severe outliers in terms of the number of false
@@
 Two major meta-analyses of WMHs have been published covering the periods
-prior to 2010\textsuperscript{41} and after\textsuperscript{42}. The
+prior to 2010\textsuperscript{42} and after\textsuperscript{43}. The
 earlier meta-analysis covered 53 longitudinal studies that included
@@
 samples representative of the general population are more relevant to
-the focus of the present paper. Debette \& Markus\textsuperscript{41}
+the focus of the present paper. Debette \& Markus\textsuperscript{42}
 found that the presence of WMHs was related to subsequent cognitive
@@
 
-The more recent meta-analysis by Kloppenborg et al.\textsuperscript{42}
+The more recent meta-analysis by Kloppenborg et al.\textsuperscript{43}
 of 23 cross-sectional studies reporting MRI and concurrent
@@
 across a wide variety of settings for the rapid identification of
-predictive imaging features\textsuperscript{43--46}. Automobile
+predictive imaging features\textsuperscript{44--47}. Automobile
 manufacturers utilize these types of approaches to equip self-driving
@@
 through the identification of visual information sufficiently similar to
-previously assimilated training data\textsuperscript{47,48}. Similarly,
+previously assimilated training data\textsuperscript{48,49}. Similarly,
 in the context of the neuroimaging assessments, deep learning approaches
@@
 applied to the segmentation of macroscopically visible
-structures\textsuperscript{43--46}. Additionally, these approaches might
+structures\textsuperscript{44--47}. Additionally, these approaches might
 be applied to the interrogation of imaging data in the individual
@@
 analytical tools to assess the information in the setting of the
-individual patient\textsuperscript{49}. Application of deep learning
+individual patient\textsuperscript{50}. Application of deep learning
 approaches in the context of data with primary quantitative outputs will
@@
 
-\textbf{Figure X:}
+\textbf{Figure 1:}
 \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
@@
 
-\textbf{Figure 1:} Representation of Stage 1 feature images for subject
+\textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
+symmetric template constructed from the MMRR data
+set\textsuperscript{25} (only shown are the FLAIR, T1, and T2
+modalities--- the components relevant for this work). Template
+construction is detailed in\textsuperscript{17}. These images are
+important for asymmetry-based features.
+
+\textbf{Figure 3:} Representation of Stage 1 feature images for subject
 01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
@@
 modality images are then preprocessed (N4 bias
-correction\textsuperscript{26} and adaptive
-denoising\textsuperscript{25}) followed by application of standard ANTs
+correction\textsuperscript{29} and adaptive
+denoising\textsuperscript{28}) followed by application of standard ANTs
 brain extraction and \(n\)-tissue segmentation protocols using the MMRR
-symmetric template and corresponding priors\textsuperscript{28} applied
+symmetric template and corresponding priors\textsuperscript{24} applied
 to the T1 image. The feature images are then generated for voxelwise
@@
 
-\textbf{Figure 2:} Sample FLAIR acquisition image slices showing both
+\textbf{Figure 4:} Sample FLAIR acquisition image slices showing both
 manual and random forest segmentations for both stages obtained during
@@
 
-\textbf{Figure 3:} Canonical views of the mutlivariate, bilaterally
-symmetric template constructed from the MMRR data
-set\textsuperscript{30} (only shown are the FLAIR, T1, and T2
-modalities--- the components relevant for this work). Template
-construction is detailed in\textsuperscript{17}. These images are
-important for specific intensity-based features.
-
-\textbf{Figure 4:} Average \emph{MeanDecreaseAccuracy} plots generated
+\textbf{Figure 5:} Average \emph{MeanDecreaseAccuracy} plots generated
 from the creation of all 24 random forest models for Stage 1 during the
@@
 
-\textbf{Figure 5:} Average \emph{MeanDecreaseAccuracy} plots generated
+\textbf{Figure 6:} Average \emph{MeanDecreaseAccuracy} plots generated
 from the creation of all 24 random forest models for Stage 2 during the
@@
 
-\textbf{Figure 6:} (a) FLAIR image slice illustrating WMHs which have
+\textbf{Figure 7:} (a) FLAIR image slice illustrating WMHs which have
 been manually delineated. The region around the WMHs is enlarged (b) in
@@
 
-\textbf{Figure 7:} Voxelwise comparison with manual delineation of white
+\textbf{Figure 8:} Voxelwise comparison with manual delineation of white
 matter hyperintensities. On the left are the calculated Dice values over
@@
 
-\hypertarget{ref-Neema:2009aa}{}
-24. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
-Normal findings on brain fluid-attenuated inversion recovery mR images
-at 3T. AJNR Am J Neuroradiol. 2009;30(5):911--6.
-
-\hypertarget{ref-Manjon:2010aa}{}
-25. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
-non-local means denoising of mR images with spatially varying noise
-levels. J Magn Reson Imaging. 2010;31(1):192--203.
-
-\hypertarget{ref-Tustison:2010ac}{}
-26. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
-JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
-2010;29(6):1310--20.
-
-\hypertarget{ref-nyul2000}{}
-27. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
-standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
-
 \hypertarget{ref-Tustison:2014ab}{}
-28. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
+24. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
 Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
@@
 
+\hypertarget{ref-landman2011}{}
+25. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
+reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
+
+\hypertarget{ref-Avants:2010aa}{}
+26. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre
+J, Gee JC. The optimal template effect in hippocampus studies of
+diseased populations. Neuroimage. 2010;49(3):2457--66.
+
+\hypertarget{ref-Neema:2009aa}{}
+27. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
+Normal findings on brain fluid-attenuated inversion recovery mR images
+at 3T. AJNR Am J Neuroradiol. 2009;30(5):911--6.
+
+\hypertarget{ref-Manjon:2010aa}{}
+28. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
+non-local means denoising of mR images with spatially varying noise
+levels. J Magn Reson Imaging. 2010;31(1):192--203.
+
+\hypertarget{ref-Tustison:2010ac}{}
+29. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
+JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
+2010;29(6):1310--20.
+
+\hypertarget{ref-nyul2000}{}
+30. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
+standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
+
 \hypertarget{ref-Avants:2011aa}{}
-29. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+31. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
 multivariate framework for \(n\)-tissue segmentation with evaluation on
@@
 
-\hypertarget{ref-landman2011}{}
-30. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
-Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
-reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
-
 \hypertarget{ref-Menze:2015aa}{}
-31. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
+32. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
 Burren Y, Porz N, Slotboom J, Wiest R, et al. The multimodal brain tumor
@@
 \hypertarget{ref-maurer2003}{}
-32. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
+33. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
 computing exact Euclidean distance transforms of binary images in
@@
 \hypertarget{ref-Anbeek:2004aa}{}
-33. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+34. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
 Probabilistic segmentation of white matter lesions in mR imaging.
@@
 \hypertarget{ref-Tustison:2013ac}{}
-34. Tustison NJ, Avants BB. Explicit B-spline regularization in
+35. Tustison NJ, Avants BB. Explicit B-spline regularization in
 diffeomorphic image registration. Front Neuroinform. 2013;7:39.
@@
 \hypertarget{ref-Avants:2011ab}{}
-35. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+36. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
 reproducible evaluation of ANTs similarity metric performance in brain
@@
 \hypertarget{ref-Garcia-Lorenzo:2013aa}{}
-36. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
+37. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
 Review of automatic segmentation methods of multiple sclerosis white
@@
 \hypertarget{ref-liaw2002}{}
-37. Liaw A, Wiener M. Classification and regression by randomForest. R
+38. Liaw A, Wiener M. Classification and regression by randomForest. R
 News. 2002;2/3:18--22.
@@
 \hypertarget{ref-Grimaud:1996aa}{}
-38. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
+39. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
 DL, Tofts PS, McDonald WI, Miller DH. Quantification of mRI lesion load
@@
 \hypertarget{ref-styner2008}{}
-39. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
+40. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
 S, Jewells V, Warfield S, editors. Special Issue on 2008 MICCAI Workshop
@@
 \hypertarget{ref-tustison2009}{}
-40. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
+41. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
 overlap measures to ITK. Insight Journal. 2009.
@@
 \hypertarget{ref-Debette:2010aa}{}
-41. Debette S, Markus HS. The clinical importance of white matter
+42. Debette S, Markus HS. The clinical importance of white matter
 hyperintensities on brain magnetic resonance imaging: Systematic review
@@
 \hypertarget{ref-Kloppenborg:2014aa}{}
-42. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
+43. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
 Presence and progression of white matter hyperintensities and cognition:
@@
 \hypertarget{ref-Plis:2014aa}{}
-43. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
+44. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
 Johnson HJ, Paulsen JS, Turner JA, Calhoun VD. Deep learning for
@@
 \hypertarget{ref-Suk:2015aa}{}
-44. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
+45. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
 Initiative. Deep sparse multi-task learning for feature selection in
@@
 \hypertarget{ref-Li:2014aa}{}
-45. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
+46. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
 based imaging data completion for improved brain disease diagnosis. Med
@@
 \hypertarget{ref-Liu:2015aa}{}
-46. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
+47. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
 ADNI. Multimodal neuroimaging feature learning for multiclass diagnosis
@@
 \hypertarget{ref-HadsellSBESKML09}{}
-47. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
+48. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
 Muller U, LeCun Y. Learning long-range vision for autonomous off-road
@@
 \hypertarget{ref-FarabetCNL12}{}
-48. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
+49. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
 multiscale feature learning, purity trees, and optimal covers. In:
@@
 \hypertarget{ref-Mayer:2014aa}{}
-49. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
+50. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
 identifying subject-specific abnormalities in neuroimaging data. Hum

### 2016-06-27 responseToReviewers.Rmd
ENH:  More response.

@@
 
-> _\textcolor{red}{Nick---Add a table for each patient and get all this
-information before doing the evaluation.}_
+> _\textcolor{blue}{We added the following sentence to the opening paragraph of the
+Imaging subsection of the Methods section:}_
+
+> \textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
+Table 1 provides a descriptive statistical summary of the variation in lesion load across the selected cohort.}
+
+> _\textcolor{blue}{as well as the corresponding table:}_
+
+\input{trainingData.tex}
+
+\clearpage
 

### 2016-06-27 responseToReviewers.tex
ENH:  More response.

@@
 \begin{quote}
-\emph{\textcolor{red}{Nick---Add a table for each patient and get all this
-information before doing the evaluation.}}
+\emph{\textcolor{blue}{We added the following sentence to the opening paragraph of the
+Imaging subsection of the Methods section:}}
 \end{quote}
+
+\begin{quote}
+\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
+Table 1 provides a descriptive statistical summary of the variation in lesion load across the selected cohort.}
+\end{quote}
+
+\begin{quote}
+\emph{\textcolor{blue}{as well as the corresponding table:}}
+\end{quote}
+
+\input{trainingData.tex}
+
+\clearpage
 
@@
 
-\emph{\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template.}}
+\emph{\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}}
 
@@
 results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
+
+\emph{\textcolor{blue}{These changes complement the edits made in response to other review items.}}
 \end{quote}

### 2016-06-27 format.Rmd
ENH:  More response.

@@
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{2}
+   - \linespread{1.2}
    - \usepackage[compact]{titlesec}

### 2016-06-27 methods.Rmd
ENH:  More response.

@@
 Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
-Each of the participants that were selected from the larger cohort of participants in this study demonstrated at least one white matter hyperintensity (but as many as 20) on FLAIR imaging.
+\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
+Table 1 provides a descriptive statistical summary of the variation in lesion
+load across the selected cohort.}
+
+<!--
+
+\input{trainingData.tex}
+
+-->
 
@@
 Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.  In Table 1, we provide a list and basic
+are used to identify the regions of interest.  In Table 2, we provide a list and basic
 categorization of the feature images used for the initial (i.e., Stage 1---more on the use
@@
 competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 1.  We employ the
+from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 2.  We employ the
 same strategy here.

### 2016-06-27 stitched.tex
ENH:  More response.

@@
 \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-\linespread{2}
+\linespread{1.2}
 \usepackage[compact]{titlesec}
@@
 concussive events (PCE) interview process described in detail in Walker
-et al., this issue). Each of the participants that were selected from
-the larger cohort of participants in this study demonstrated at least
-one white matter hyperintensity (but as many as 20) on FLAIR imaging.
+et al., this issue).
+\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.  Table 1 provides a descriptive statistical summary of the variation in lesion
+load across the selected cohort.}
 
@@
 feature images that are used to identify the regions of interest. In
-Table 1, we provide a list and basic categorization of the feature
+Table 2, we provide a list and basic categorization of the feature
 images used for the initial (i.e., Stage 1---more on the use of multiple
@@
 intensity features that were created from multi-modal templates
-mentioned previously\textsuperscript{17} and listed in Table 1. We
+mentioned previously\textsuperscript{17} and listed in Table 2. We
 employ the same strategy here.

### 2016-06-29 responseToReviewers.Rmd
ENH:  Evaluation results.

@@
 
-> _\textcolor{red}{Here.}_
+> _\textcolor{blue}{We removed the sentence in question.  Also, as mentioned in
+response to item 3, we added Table 1.}_
 
@@
 
-> _\textcolor{red}{Get rid of the paragraphs in question.  Need to redo evaluation
-to discuss.}_
+> _\textcolor{red}{We removed the paragraphs in question and added the following
+comparison discussion:}_
 

### 2016-06-29 responseToReviewers.tex
ENH:  Evaluation results.

@@
 \begin{quote}
-\emph{\textcolor{red}{Here.}}
+\emph{\textcolor{blue}{We removed the sentence in question.  Also, as mentioned in
+response to item 3, we added Table 1.}}
 \end{quote}
@@
 \begin{quote}
-\emph{\textcolor{red}{Get rid of the paragraphs in question.  Need to redo evaluation
-to discuss.}}
+\emph{\textcolor{red}{We removed the paragraphs in question and added the following
+comparison discussion:}}
 \end{quote}

### 2016-06-29 discussion.Rmd
ENH:  Evaluation results.

@@
 Two major meta-analyses of WMHs have been published covering the periods
-prior to 2010 [@Debette:2010aa] and after [@Kloppenborg:2014aa]. The earlier meta-analysis
-covered 53 longitudinal studies that included samples of high-risk
-populations, i.e., patients selected for a specific disease or condition
-such as hypertension, whereas other studies recruited samples of the
-general population. Longitudinal studies of samples representative of
-the general population are more relevant to the focus of the present
-paper. Debette & Markus [@Debette:2010aa] found that the presence of
+prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
+Debette & Markus [@Debette:2010aa] found that the presence of
 WMHs was related to subsequent cognitive decline, a higher risk of
 developing dementia, stroke, and of mortality. Lesion volume at baseline
-was also predictive of cognitive decline. Limitations of this
-meta-analysis include heterogeneity in the method of measuring WMHs;
-some studies used automated volumetric measurement, whereas others used
-a visual rating scale. The studies analyzed by Debette & Markus were
-limited to the occurrence of one of the aforementioned conditions which
-they analyzed by hazard ratios.
-
-The more recent meta-analysis by Kloppenborg et al.
+was also predictive of cognitive decline. Kloppenborg et al.
 [@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
@@
 that WMHs were associated with cognitive deficit (effect size of -0.10,
-95% CI: -0.13 to -0.08) after controlling for age. These studies also
-differed in the metric used to measure the WMHs, including volume, % of
-total intracranial volume, and a visual rating score. The effect size
-for the association with cognitive deficit in these cross-sectional
-studies did not differ significantly across various cognitive domains or
-the method of measuring lesion volume.  Among eight longitudinal studies
-analyzed by Kloppenborg et al that included a follow-up MRI and also
-controlled for age, the effect size for the association of progression
-in WMHs and cognitive impairment was -0.16 (95% CI:-0.27 to -0.09). This
-association was stronger for attention and executive function than for
-memory and processing speed. Although baseline WMHs were predictive of
-cognitive deficit at follow-up in the seven studies which did not repeat
-MRI, the effect size was smaller [-0.10 (95% CI: 0.13 to -0.05) than in
-the longitudinal studies that calculated progression in WMHs. In
-summary, progression of WMHs seen on repeat MRI has a stronger relation
-to cognitive deficit than concurrent imaging findings. These
-meta-analyses support the rationale for repeating an MRI in patients
-younger than 50 years whose initial scan shows WMHs.
+95% CI: -0.13 to -0.08) after controlling for age.
 
-Despite the
-above-described associations between WMHs, cognitive decline, increased
-risk of developing dementia, and mortality, these lesions receive little
+Despite the potential clinical significance of WMHs
+these lesions receive little
 attention in current clinical workflows. When reported in a standard
@@
 application to such large-scale datasets.
+
+<!--
 
@@
 
+-->
+
 \clearpage

### 2016-06-29 format.Rmd
ENH:  Evaluation results.

--- 2016-02-25 format.Rmd
@@
     keep_tex: yes
+  html_document:
+    toc: false
+
 header-includes:

### 2016-06-29 intro.Rmd
ENH:  Evaluation results.

@@
 Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state.
-Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
+The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
 In this work, we present an automated framework for quantification of WMHs in

### 2016-06-29 methods.Rmd
ENH:  Evaluation results.

@@
       regions.
+
+\textcolor{blue}{Several measures have been employed in the literature for evaluating
+automated white matter lesion segmentation involving such quantities as lesion load,
+voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
+For this work, due to the relatively small-size distribution}[^4] \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
+four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
+The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
+It should be noted that the number of true negatives is not readily
+incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
+The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
+
+$$ sensitivity =  \frac{ TP }{TP + FN}, $$
+
+\textcolor{blue}{and the positive predictive value (PPV),}
+
+$$ PPV =  \frac{ TP }{TP + FP}, $$
+
+\textcolor{blue}{such that $F_1$ is given by}
+
+$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
+
+\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
+the manual and predicted lesion volumes:}
+
+$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
+
+\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
+These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
+
+[^4]: \textcolor{blue}{For comparison, the training data set of
+the MS Lesion Segmentation challenge
+associated with the international MICCAI 2008 conference has a mean lesion load of
+204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
+in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}

### 2016-06-29 results.Rmd
ENH:  Evaluation results.

@@
 # Results
+
+## White matter hyperintensity segmentation evaluation
+
+\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
+of the previously described TBI data.  These performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
+size ranges over which these measures are computed to illustrate the variation
+in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
+to identify which is why the sensitivity is widely varied over the cohort vs. the largest
+set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
+entire lesions.  The relative volume difference provides a direct assessment of how accurate
+the volumetric estimate is when comparing the manually identified lesions versus the
+automatically predicted lesions.}
+
+<!--
+
+In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
+data.  Despite the large variability characteristic with manual labelings in related fields
+[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
+current clinical practices and the methodology proposed herein is readily adapted to refinements
+in training data.
+
+On the left of Figure 8 are the improvement in Dice values [@tustison2009],
+over all white matter hyperintensities when comparing the segmentations between the two stages
+where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
+represents the intersection between the manual/automated lesion pair.  Performing
+the second round of supervised learning improves these Dice values.  One can also note from the
+right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
+severe outliers in terms of the number of false positives.  The second round helps to
+correct this issue.
+
+-->
+
+<!--
+
+![Evaluation measures for the leave-one-out protocol of the described protocol in the Methods
+section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
+volume difference.   The three quantile ranges include the following lesion volume ranges:
+$[3-12)$, $[12-28)$, and $[28-551]$ based on the manually-derived lesion volumes.  These
+measures are computed for both Stages.  Overall improvement is seen as the
+second Stage RF model is applied for all three quantile ranges.](Figures/evaluationResults.png)
+
+-->
 
@@
 alone it is intuitive that such features would be important.
-
-## White matter hyperintensity segmentation evaluation
-
-In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-<!--
-i.e.,
-
-$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$
--->
-
-On the left of Figure 8 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
-<!--
-
-![Voxelwise comparison with manual delineation of white matter hyperintensities.  On the left are the
-calculated Dice values over all white matter hyperintensities.  Note the improvement
-in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
-(Right) Similar results can be seen by comparing the total lesion load volume between manual
-and automated detection strategies.  Although some outliers are found after the Stage 2
-processing in a couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.](Figures/llvAndDice.png)
-
--->

### 2016-06-29 stitched.Rmd
ENH:  Evaluation results.

@@
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.2}
+   - \linespread{2}
    - \usepackage[compact]{titlesec}
@@
 
+\textcolor{blue}{Several measures have been employed in the literature for evaluating
+automated white matter lesion segmentation involving such quantities as lesion load,
+voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
+For this work, due to the relatively small-size distribution}[^4] \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
+four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
+The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
+It should be noted that the number of true negatives is not readily
+incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
+The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
+
+$$ sensitivity =  \frac{ TP }{TP + FN}, $$
+
+\textcolor{blue}{and the positive predictive value (PPV),}
+
+$$ PPV =  \frac{ TP }{TP + FP}, $$
+
+\textcolor{blue}{such that $F_1$ is given by}
+
+$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
+
+\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
+the manual and predicted lesion volumes:}
+
+$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
+
+\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
+These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
+
+[^4]: \textcolor{blue}{For comparison, the training data set of
+the MS Lesion Segmentation challenge
+associated with the international MICCAI 2008 conference has a mean lesion load of
+204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
+in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}
+
 # Results
+
+## White matter hyperintensity segmentation evaluation
+
+\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
+of the previously described TBI data.  These performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
+size ranges over which these measures are computed to illustrate the variation
+in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
+to identify which is why the sensitivity is widely varied over the cohort vs. the largest
+set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
+entire lesions.  The relative volume difference provides a direct assessment of how accurate
+the volumetric estimate is when comparing the manually identified lesions versus the
+automatically predicted lesions.}
+
+<!--
+
+In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
+data.  Despite the large variability characteristic with manual labelings in related fields
+[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
+current clinical practices and the methodology proposed herein is readily adapted to refinements
+in training data.
+
+On the left of Figure 8 are the improvement in Dice values [@tustison2009],
+over all white matter hyperintensities when comparing the segmentations between the two stages
+where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
+represents the intersection between the manual/automated lesion pair.  Performing
+the second round of supervised learning improves these Dice values.  One can also note from the
+right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
+severe outliers in terms of the number of false positives.  The second round helps to
+correct this issue.
+
+-->
+
+<!--
+
+![Evaluation measures for the leave-one-out protocol of the described protocol in the Methods
+section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
+volume difference.   The three quantile ranges include the following lesion volume ranges:
+$[3-12)$, $[12-28)$, and $[28-551]$ based on the manually-derived lesion volumes.  These
+measures are computed for both Stages.  Overall improvement is seen as the
+second Stage RF model is applied for all three quantile ranges.](Figures/evaluationResults.png)
+
+-->
 
@@
 
-## White matter hyperintensity segmentation evaluation
-
-In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-<!--
-i.e.,
-
-$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$
--->
-
-On the left of Figure 8 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
-<!--
-
-![Voxelwise comparison with manual delineation of white matter hyperintensities.  On the left are the
-calculated Dice values over all white matter hyperintensities.  Note the improvement
-in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
-(Right) Similar results can be seen by comparing the total lesion load volume between manual
-and automated detection strategies.  Although some outliers are found after the Stage 2
-processing in a couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.](Figures/llvAndDice.png)
-
--->
-
 # Discussion
@@
 this disease process.
+
+-->
 

### 2016-06-29 stitched.tex
ENH:  Evaluation results.

@@
 \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-\linespread{1.2}
+\linespread{2}
 \usepackage[compact]{titlesec}
@@
 
+\textcolor{blue}{Several measures have been employed in the literature for evaluating
+automated white matter lesion segmentation involving such quantities as lesion load,
+voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.
+For this work, due to the relatively small-size distribution}\footnote{\textcolor{blue}{For comparison, the training data set of
+  the MS Lesion Segmentation challenge
+  associated with the international MICCAI 2008 conference has a mean lesion load of
+  204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
+  in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}}
+\textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
+four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
+The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
+It should be noted that the number of true negatives is not readily
+incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
+The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
+
+\[ sensitivity =  \frac{ TP }{TP + FN}, \]
+
+\textcolor{blue}{and the positive predictive value (PPV),}
+
+\[ PPV =  \frac{ TP }{TP + FP}, \]
+
+\textcolor{blue}{such that $F_1$ is given by}
+
+\[ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. \]
+
+\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
+the manual and predicted lesion volumes:}
+
+\[ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. \]
+
+\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
+These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
+
 \section{Results}\label{results}
+
+\subsection{White matter hyperintensity segmentation
+evaluation}\label{white-matter-hyperintensity-segmentation-evaluation}
+
+\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
+of the previously described TBI data.  These performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
+size ranges over which these measures are computed to illustrate the variation
+in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
+to identify which is why the sensitivity is widely varied over the cohort vs. the largest
+set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
+entire lesions.  The relative volume difference provides a direct assessment of how accurate
+the volumetric estimate is when comparing the manually identified lesions versus the
+automatically predicted lesions.}
 
@@
 
-\subsection{White matter hyperintensity segmentation
-evaluation}\label{white-matter-hyperintensity-segmentation-evaluation}
-
-In Figure 8 are the segmentation comparisons derived from manual
-segmentations of the same data. Despite the large variability
-characteristic with manual labelings in related
-fields\textsuperscript{37,39,40}, such labelings are characteristic of
-current clinical practices and the methodology proposed herein is
-readily adapted to refinements in training data.
-
-On the left of Figure 8 are the improvement in Dice
-values\textsuperscript{41}, over all white matter hyperintensities when
-comparing the segmentations between the two stages where the sum is
-taken over all individually labeled manual, \(T_r\), and automated,
-\(S_r\), lesions and \(\cap\) represents the intersection between the
-manual/automated lesion pair. Performing the second round of supervised
-learning improves these Dice values. One can also note from the right
-side of Figure 7 that the total lesion load volume illustrates a few
-subjects that are severe outliers in terms of the number of false
-positives. The second round helps to correct this issue.
-
 \section{Discussion}\label{discussion}
@@
 Two major meta-analyses of WMHs have been published covering the periods
-prior to 2010\textsuperscript{42} and after\textsuperscript{43}. The
-earlier meta-analysis covered 53 longitudinal studies that included
-samples of high-risk populations, i.e., patients selected for a specific
-disease or condition such as hypertension, whereas other studies
-recruited samples of the general population. Longitudinal studies of
-samples representative of the general population are more relevant to
-the focus of the present paper. Debette \& Markus\textsuperscript{42}
-found that the presence of WMHs was related to subsequent cognitive
-decline, a higher risk of developing dementia, stroke, and of mortality.
-Lesion volume at baseline was also predictive of cognitive decline.
-Limitations of this meta-analysis include heterogeneity in the method of
-measuring WMHs; some studies used automated volumetric measurement,
-whereas others used a visual rating scale. The studies analyzed by
-Debette \& Markus were limited to the occurrence of one of the
-aforementioned conditions which they analyzed by hazard ratios.
-
-The more recent meta-analysis by Kloppenborg et al.\textsuperscript{43}
+prior to\textsuperscript{39} and after 2010\textsuperscript{40}. Debette
+\& Markus\textsuperscript{39} found that the presence of WMHs was
+related to subsequent cognitive decline, a higher risk of developing
+dementia, stroke, and of mortality. Lesion volume at baseline was also
+predictive of cognitive decline. Kloppenborg et al.\textsuperscript{40}
 of 23 cross-sectional studies reporting MRI and concurrent
@@
 associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
-to -0.08) after controlling for age. These studies also differed in the
-metric used to measure the WMHs, including volume, \% of total
-intracranial volume, and a visual rating score. The effect size for the
-association with cognitive deficit in these cross-sectional studies did
-not differ significantly across various cognitive domains or the method
-of measuring lesion volume. Among eight longitudinal studies analyzed by
-Kloppenborg et al that included a follow-up MRI and also controlled for
-age, the effect size for the association of progression in WMHs and
-cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
-association was stronger for attention and executive function than for
-memory and processing speed. Although baseline WMHs were predictive of
-cognitive deficit at follow-up in the seven studies which did not repeat
-MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
-in the longitudinal studies that calculated progression in WMHs. In
-summary, progression of WMHs seen on repeat MRI has a stronger relation
-to cognitive deficit than concurrent imaging findings. These
-meta-analyses support the rationale for repeating an MRI in patients
-younger than 50 years whose initial scan shows WMHs.
-
-Despite the above-described associations between WMHs, cognitive
-decline, increased risk of developing dementia, and mortality, these
-lesions receive little attention in current clinical workflows. When
-reported in a standard neuroradiologist interpretation, they are
-typically handled as incidental findings and are assigned little
-clinical significance. This likely reflects the impracticality of
-performing a detailed assessment of number, volume, and distribution
-within a qualitative neuroradiologist interpretation as well as the lack
-of correlative information on how the presence and distribution of these
-lesions may inform a diagnosis and prognosis in the appropriate clinical
-setting. To date, automated or semi-automated tools for the detection of
-WMHs have lacked the specificity and efficiency for the mining of
-large-scale datasets to generate highly granular data on whether these
-lesions possess any true diagnostic or prognostic value in the setting
-of a specific disease process. The present communication describes a
+to -0.08) after controlling for age.
+
+Despite the potential clinical significance of WMHs these lesions
+receive little attention in current clinical workflows. When reported in
+a standard neuroradiologist interpretation, they are typically handled
+as incidental findings and are assigned little clinical significance.
+This likely reflects the impracticality of performing a detailed
+assessment of number, volume, and distribution within a qualitative
+neuroradiologist interpretation as well as the lack of correlative
+information on how the presence and distribution of these lesions may
+inform a diagnosis and prognosis in the appropriate clinical setting. To
+date, automated or semi-automated tools for the detection of WMHs have
+lacked the specificity and efficiency for the mining of large-scale
+datasets to generate highly granular data on whether these lesions
+possess any true diagnostic or prognostic value in the setting of a
+specific disease process. The present communication describes a
 supervised statistical learning tool that is appropriate for the
 application to such large-scale datasets.
-
-The currently described tool is just one example of how ``supervised
-learning'' algorithms might be applied to aid in the diagnosis of TBI
-and other disease processes through the specific identification of
-features predictive of a given disease state. It is an important
-demonstration of the potential power of these analytical approaches in
-the rapid but comprehensive mining of information from neuroimaging
-examinations. Supervised learning algorithms are presently employed
-across a wide variety of settings for the rapid identification of
-predictive imaging features\textsuperscript{44--47}. Automobile
-manufacturers utilize these types of approaches to equip self-driving
-vehicles to recognize and respond to unique external surroundings
-through the identification of visual information sufficiently similar to
-previously assimilated training data\textsuperscript{48,49}. Similarly,
-in the context of the neuroimaging assessments, deep learning approaches
-may allow for the rapid identification of information predictive of
-disease state in an individual patient. These approaches have been
-applied to the segmentation of macroscopically visible
-structures\textsuperscript{44--47}. Additionally, these approaches might
-be applied to the interrogation of imaging data in the individual
-patient with a primary quantitative output metrics to include sequences
-such as diffusion tensor imaging (DTI) and its variants, functional
-connectivity, perfusion weighted imaging, and cortical thickness
-assessments. At present, these advanced neuroimaging sequences are
-confined to cohort-based research studies due to the lack of available
-analytical tools to assess the information in the setting of the
-individual patient\textsuperscript{50}. Application of deep learning
-approaches in the context of data with primary quantitative outputs will
-require large scale normative and disease specific databases. Building
-these large scale imaging libraries is resource intensive and requires a
-multi-center approach with harmonized scanners between sites and
-correlative non-imaging clinical data. Large scale TBI data is becoming
-increasingly available through activities such as the Chronic Effects of
-Neurotrauma Consortium (CENC), Transforming Research and Clinical
-Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma
-Effectiveness Research in TBI (CENTER-TBI), Department of Defense
-Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and other data
-being consolidated through FITBIR. In concert with any available high
-quality normative neuroimaging data, deep learning algorithms may be
-well positioned to help transform how neuroimaging is interpreted for
-the clinical management of patients with this disease process.
 
@@
 
-\hypertarget{ref-Grimaud:1996aa}{}
-39. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
-DL, Tofts PS, McDonald WI, Miller DH. Quantification of mRI lesion load
-in multiple sclerosis: A comparison of three computer-assisted
-techniques. Magn Reson Imaging. 1996;14(5):495--505.
-
-\hypertarget{ref-styner2008}{}
-40. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
-S, Jewells V, Warfield S, editors. Special Issue on 2008 MICCAI Workshop
-- MS Lesion Segmentation. MIDAS J; 2008.
-
-\hypertarget{ref-tustison2009}{}
-41. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
-overlap measures to ITK. Insight Journal. 2009.
-
 \hypertarget{ref-Debette:2010aa}{}
-42. Debette S, Markus HS. The clinical importance of white matter
+39. Debette S, Markus HS. The clinical importance of white matter
 hyperintensities on brain magnetic resonance imaging: Systematic review
@@
 \hypertarget{ref-Kloppenborg:2014aa}{}
-43. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
+40. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
 Presence and progression of white matter hyperintensities and cognition:
@@
 
-\hypertarget{ref-Plis:2014aa}{}
-44. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
-Johnson HJ, Paulsen JS, Turner JA, Calhoun VD. Deep learning for
-neuroimaging: A validation study. Front Neurosci. 2014;8:229.
-
-\hypertarget{ref-Suk:2015aa}{}
-45. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
-Initiative. Deep sparse multi-task learning for feature selection in
-alzheimer's disease diagnosis. Brain Struct Funct. 2015 May.
-
-\hypertarget{ref-Li:2014aa}{}
-46. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
-based imaging data completion for improved brain disease diagnosis. Med
-Image Comput Comput Assist Interv. 2014;17(Pt 3):305--12.
-
-\hypertarget{ref-Liu:2015aa}{}
-47. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
-ADNI. Multimodal neuroimaging feature learning for multiclass diagnosis
-of alzheimer's disease. IEEE Trans Biomed Eng. 2015;62(4):1132--40.
-
-\hypertarget{ref-HadsellSBESKML09}{}
-48. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
-Muller U, LeCun Y. Learning long-range vision for autonomous off-road
-driving. J. Field Robotics. 2009;26(2):120--144.
-
-\hypertarget{ref-FarabetCNL12}{}
-49. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
-multiscale feature learning, purity trees, and optimal covers. In:
-Proceedings of the 29th international conference on machine learning,
-ICML 2012, edinburgh, scotland, uK, june 26 - july 1, 2012. icml.cc /
-Omnipress; 2012.
-
-\hypertarget{ref-Mayer:2014aa}{}
-50. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
-identifying subject-specific abnormalities in neuroimaging data. Hum
-Brain Mapp. 2014;35(11):5457--70.
-
 \end{document}

### 2016-06-29 Evaluation/evaluation2.R

@@ -1,0 +1,115 @@
+library( ggplot2 )
+
+whichEval <- 'ResultsOriginal_500Samples'
+
+td <- read.csv( "trainingData.csv" )
+ev <- read.csv( paste0( 'Evaluation/', whichEval, '/evaluation', whichEval, '.csv' ) )
+
+tdQ <- quantile( td$Volume, c( 0.3333333, 0.66666666, 1.0 ) )
+
+types <- c( "FalseNegative", "FalsePositive", "TruePositive" )
+
+boxPlotDataFrame <- data.frame( Subject = character( 0 ), Stage = character( 0 ),
+                                Quantile = character( 0 ),
+                                MeasureType = character( 0 ), Value = numeric( 0 ) )
+
+for( i in 0:23 )
+  {
+  subjectIndex <- i
+
+  for( stage in 1:2 )
+    {
+    for( j in 1:length( tdQ ) )
+      {
+      quantLeft <- 3
+      if( j > 1 )
+        {
+        quantLeft <- tdQ[j-1]
+        }
+      quantRight <- tdQ[j]
+
+      indicesFN <- which( ev$Type == "FalseNegative" & ev$SubjectIndex == subjectIndex & ev$Stage == stage & ev$TrueVolume >= quantLeft & ev$TrueVolume < quantRight )
+      indicesFP <- which( ev$Type == "FalsePositive" & ev$SubjectIndex == subjectIndex & ev$Stage == stage & ev$PredictedVolume >= quantLeft & ev$PredictedVolume < quantRight )
+      indicesTP <- which( ev$Type == "TruePositive" & ev$SubjectIndex == subjectIndex & ev$Stage == stage & ev$TrueVolume >= quantLeft & ev$TrueVolume < quantRight )
+
+      numberOfFalseNegatives <- length( ev$TrueVolume[indicesFN] )
+      numberOfFalsePositives <- length( ev$TrueVolume[indicesFP] )
+      numberOfTruePositives <- length( ev$TrueVolume[indicesTP] )
+
+      f1 <- 2 * numberOfTruePositives / ( 2 * numberOfTruePositives + numberOfFalseNegatives + numberOfFalsePositives )
+      sensitivity <- numberOfTruePositives / ( numberOfTruePositives + numberOfFalseNegatives )
+      ppv <- numberOfTruePositives / ( numberOfTruePositives + numberOfFalsePositives )
+
[…]

### 2016-06-29 Evaluation/getStatsForMiccaiMsLesionData.pl

@@ -1,0 +1,59 @@
+#/usr/bin/perl -w
+
+use strict;
+
+use Cwd 'realpath';
+use File::Find;
+use File::Basename;
+use File::Path;
+use File::Spec;
+use FindBin qw($Bin);
+
+my $usage = qq{
+  Usage:
+ };
+
+my @miccaiDirectories = ();
+
+$miccaiDirectories[0] = '/Users/ntustison/Downloads/CHB_train_Part1/';
+$miccaiDirectories[1] = '/Users/ntustison/Downloads/CHB_train_Part2/';
+$miccaiDirectories[2] = '/Users/ntustison/Downloads/UNC_train_Part1/';
+$miccaiDirectories[3] = '/Users/ntustison/Downloads/UNC_train_Part2/';
+
+my $statsFile = '/Users/ntustison/Desktop/miccaiMsLesionStats.csv';
+open( FILE, ">${statsFile}" );
+
+print FILE "SubjectId,LesionLabel,PixelVolume,Volume\n";
+
+for( my $i = 0; $i < @miccaiDirectories; $i++ )
+  {
+  my @lesionImages = <${miccaiDirectories[$i]}/*//*lesion.nhdr>;
+
+  for( my $j = 0; $j < @lesionImages; $j++ )
+    {
+    print "Processing ${lesionImages[$j]}\n";
+
+    ( my $tmpLesion = ${lesionImages[$j]} ) =~ s/\.nhdr/\Tmp\.nii\.gz/;
+
+    my ( $filename, $path, $suffix ) = fileparse( $lesionImages[$j], ".nhdr" );
+
+    my $spacingString = `GetImageInformation 3 $lesionImages[$j] 1`;
+    my @spacing = split( 'x', $spacingString );
+    my $pixelVolume = $spacing[0] * $spacing[1] * $spacing[2];
+
+    `GetConnectedComponents 3 $lesionImages[$j] $tmpLesion`;
+    my @out = `LabelGeometryMeasures 3 $tmpLesion`;
+
+    for( my $k = 1; $k < @out; $k++ )
+      {
+      my @stats = split( ' ', ${out[$k]} );
+
+      print "  Label ${stats[0]}\n";
+
+      print FILE "${filename},${stats[0]},$pixelVolume,${stats[1]}\n";
+      }
+    }
+  }
+
+
+close( FILE )

### 2016-06-29 Evaluation/trainingData.R

@@ -1,0 +1,9 @@
+library( ggplot2 )
+
+td <- read.csv( "trainingData.csv" )
+
+myPlot <- ggplot( data = td, aes( Volume ) ) +
+          geom_histogram( binwidth = 1 )
+ggsave( "myPlot.pdf", plot = myPlot, width = 10, height = 5 )
+
+tdHist <- hist( td$Volume, breaks = 200 )

### 2016-06-29 Evaluation/volumeDifference.R

@@ -1,0 +1,95 @@
+library( ggplot2 )
+
+# whichEval <- 'Results_500Samples'
+# whichEval <- 'ResultsMorphology_500Samples'
+# whichEval <- 'ResultsOriginal_500Samples'
+whichEval <- 'ResultsMorphologyOriginal_500Samples'
+
+td <- read.csv( "trainingData.csv" )
+ev <- read.csv( paste0( '~/Desktop/wmhEvaluation/', whichEval, '/evaluation', whichEval, '.csv' ) )
+
+tdQ <- quantile( td$Volume, c( 0.3333333, 0.66666666, 1.0 ) )
+
+types <- c( "FalseNegative", "FalsePositive", "TruePositive" )
+
+boxPlotDataFrame <- data.frame( Subject = character( 0 ), Stage = character( 0 ),
+                                Quantile = character( 0 ), Value = numeric( 0 ) )
+
+for( i in 0:23 )
+  {
+  subjectIndex <- i
+
+  for( stage in 1:2 )
+    {
+    for( j in 1:length( tdQ ) )
+      {
+      quantLeft <- 3
+      if( j > 1 )
+        {
+        quantLeft <- tdQ[j-1]
+        }
+      quantRight <- tdQ[j]
+
+      indicesFN <- which( ev$Type == "FalseNegative" & ev$SubjectIndex == subjectIndex & ev$Stage == stage & ev$TrueVolume >= quantLeft & ev$TrueVolume < quantRight )
+      indicesFP <- which( ev$Type == "FalsePositive" & ev$SubjectIndex == subjectIndex & ev$Stage == stage & ev$PredictedVolume >= quantLeft & ev$PredictedVolume < quantRight )
+      indicesTP <- which( ev$Type == "TruePositive" & ev$SubjectIndex == subjectIndex & ev$Stage == stage & ev$TrueVolume >= quantLeft & ev$TrueVolume < quantRight )
+
+      numberOfFalseNegatives <- length( ev$TrueVolume[indicesFN] )
+      numberOfFalsePositives <- length( ev$TrueVolume[indicesFP] )
+      numberOfTruePositives <- length( ev$TrueVolume[indicesTP] )
+
+      f1 <- 2 * numberOfTruePositives / ( 2 * numberOfTruePositives + numberOfFalseNegatives + numberOfFalsePositives )
+      sensitivity <- numberOfTruePositives / ( numberOfTruePositives + numberOfFalseNegatives )
[…]

### 2016-06-30 responseToReviewers.Rmd
ENH:  More revisions.

@@
 
-> _\textcolor{red}{Nick}_
+> _\textcolor{blue}{We completely reorganized the results and discussion
+section to reflect the new evaluation.  In addition, we added a synopsis
+of these results to the abstract.}_
 
@@
 All multi-modal MR dicom image slices were converted to the nifti file format.
-Given the utility of the FLAIR sequence in detecting white matter lesions in TBI}  [@Marquez-de-la-Plata:2007aa]\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
+All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
 The normalized MRI volumes were then provided to J.
@@
 be posted to the github repository associated with this manuscript
-(https://github.com/ntustison/WatchMeHyperventilate) for future readers.
+(https://github.com/ntustison/WatchMeHyperventilate), which we mention in the text,
+for future readers.
 }_
@@
 Once these offline steps
-are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
 >
@@
 
-> _\textcolor{red}{We removed the paragraphs in question and added the following
-comparison discussion:}_
+> _\textcolor{blue}{We removed the paragraphs in question and added a
+comparison discussion.}_
 

### 2016-06-30 responseToReviewers.tex
ENH:  More revisions.

@@
 \begin{quote}
-\emph{\textcolor{red}{Nick}}
+\emph{\textcolor{blue}{We completely reorganized the results and discussion
+section to reflect the new evaluation.  In addition, we added a synopsis
+of these results to the abstract.}}
 \end{quote}
@@
 All multi-modal MR dicom image slices were converted to the nifti file format.
-Given the utility of the FLAIR sequence in detecting white matter lesions in TBI}\textsuperscript{1}\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software}\textsuperscript{2}
-\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{3}
+All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}\textsuperscript{1}
+\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{2}
 \textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
@@
 be posted to the github repository associated with this manuscript
-(https://github.com/ntustison/WatchMeHyperventilate) for future readers.
+(https://github.com/ntustison/WatchMeHyperventilate), which we mention in the text,
+for future readers.
 }}
@@
 Once these offline steps
-are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
 
@@
 
-\textcolor{blue}{Following}\textsuperscript{4}
-\textcolor{blue}{and}\textsuperscript{5},
+\textcolor{blue}{Following}\textsuperscript{3}
+\textcolor{blue}{and}\textsuperscript{4},
 \textcolor{blue}{optimally derived templates
 serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{6}\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{7}
+For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{5}\textcolor{blue}{.
+We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{6}
 \textcolor{blue}{which
@@
 \emph{\textcolor{blue}{The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines}}\textsuperscript{4}
+extraction and normal tissue segmentation pipelines}}\textsuperscript{3}
 \emph{\textcolor{blue}{.  The result is a mask delineating the brain
 parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
-deep gray matter, brain stem, and cerebellum}}\textsuperscript{8}
+deep gray matter, brain stem, and cerebellum}}\textsuperscript{7}
 \emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
 first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{5}
-\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{6}
+are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{4}
+\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{5}
 \emph{\textcolor{blue}{(cf Figure 3).}}
@@
 \textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)}\textsuperscript{9,10}
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{11,12}.
-\textcolor{blue}{Finally, Breiman}\textsuperscript{13}
+(e.g., AdaBoost)}\textsuperscript{8,9}
+\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{10,11}.
+\textcolor{blue}{Finally, Breiman}\textsuperscript{12}
 \textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
@@
 callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{14}
+with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{13}
 \end{quote}
@@
 callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{14}
+with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{13}
 \textcolor{blue}{---a numeric groupwise assessment used to
@@
 \begin{quote}
-\textcolor{blue}{Similar feature images were used in}\textsuperscript{15}
+\textcolor{blue}{Similar feature images were used in}\textsuperscript{14}
 \textcolor{blue}{although, unlike the proposed framework, this previous work lacks
@@
 across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant}\textsuperscript{16}
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{17}.
+registered to the T1 template component using a B-spline variant}\textsuperscript{15}
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{16}.
 \textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
@@
 \begin{quote}
-\emph{\textcolor{red}{We removed the paragraphs in question and added the following
-comparison discussion:}}
+\emph{\textcolor{blue}{We removed the paragraphs in question and added a
+comparison discussion.}}
 \end{quote}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Marquez-de-la-Plata:2007aa}{}
-1. Marquez de la Plata C, Ardelean A, Koovakkattu D, Srinivasan P,
-Miller A, Phuong V, Harper C, Moore C, Whittemore A, Madden C, et al.
-Magnetic resonance imaging of diffuse axonal injury: Quantitative
-assessment of white matter lesion volume. J Neurotrauma.
-2007;24(4):591--8.
-
 \hypertarget{ref-Avants:2014aa}{}
-2. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
+1. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
 ToolKit image registration framework. Front Neuroinform. 2014;8:44.
@@
 \hypertarget{ref-Yushkevich:2006aa}{}
-3. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
+2. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
 User-guided 3D active contour segmentation of anatomical structures:
@@
 \hypertarget{ref-Tustison:2014ab}{}
-4. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
+3. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
 Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
@@
 \hypertarget{ref-Tustison:2015aa}{}
-5. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
+4. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
 Grossman MC, Avants BB. Optimal symmetric multimodal templates and
@@
 \hypertarget{ref-landman2011}{}
-6. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+5. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
 Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
@@
 \hypertarget{ref-Avants:2010aa}{}
-7. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre J,
+6. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre J,
 Gee JC. The optimal template effect in hippocampus studies of diseased
@@
 \hypertarget{ref-Avants:2011aa}{}
-8. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
+7. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
 multivariate framework for \(n\)-tissue segmentation with evaluation on
@@
 \hypertarget{ref-schapire1990}{}
-9. Schapire R. The strength of weak learnability. Machine Learning.
+8. Schapire R. The strength of weak learnability. Machine Learning.
 1990;5:197--227.
@@
 \hypertarget{ref-freund1997}{}
-10. Freund Y, Schapire R. A decision-theoretic generalization of on-line
+9. Freund Y, Schapire R. A decision-theoretic generalization of on-line
 learning and an application to boosting. Journal of Computer and System
@@
 \hypertarget{ref-ho1995}{}
-11. Ho TK. Random decision forests. In: Document analysis and
+10. Ho TK. Random decision forests. In: Document analysis and
 recognition, 1995., proceedings of the third international conference
@@
 \hypertarget{ref-amit1997}{}
-12. Amit Y, Geman D. Shape quantization and recognition with randomized
+11. Amit Y, Geman D. Shape quantization and recognition with randomized
 trees. Neural Computation. 1997;9:1545--1588.
@@
 \hypertarget{ref-breiman2001}{}
-13. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
+12. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
 
 \hypertarget{ref-Moen:2014aa}{}
-14. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A.
+13. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A.
 Traumatic axonal injury: The prognostic value of lesion load in corpus
@@
 \hypertarget{ref-Anbeek:2004aa}{}
-15. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
+14. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
 Probabilistic segmentation of white matter lesions in mR imaging.
@@
 \hypertarget{ref-Tustison:2013ac}{}
-16. Tustison NJ, Avants BB. Explicit B-spline regularization in
+15. Tustison NJ, Avants BB. Explicit B-spline regularization in
 diffeomorphic image registration. Front Neuroinform. 2013;7:39.
@@
 \hypertarget{ref-Avants:2011ab}{}
-17. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
+16. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
 reproducible evaluation of ANTs similarity metric performance in brain

### 2016-06-30 abstract.Rmd
ENH:  More revisions.

@@
 employed for both training and evaluation using a leave-one-out
-strategy.  Lesion load and overlap evaluative comparisons are
-complimented by feature rankings which showcase the utility of the
-concatenated approach.  Our findings suggest supervised learning methods
-may be applied to quantify WMHs on routine brain imaging. Paired with
+strategy.  \textcolor{blue}{Performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.
+Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
+predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
+relative volume difference} = 43 $\pm$ 26\%.  In addition, three
+lesion size ranges are selected to illustrate the variation in
+performance with lesion size.}  Paired with
 correlative outcome data, supervised learning methods may allow for

### 2016-06-30 discussion.Rmd
ENH:  More revisions.

@@
 to other automated or semi automated approaches for WMH detection? -->
+
+\textcolor{blue}{
+In evaluating the segmentation results, there are limited methodologies described in the
+literature for automatic segmentation of WMHs despite the numerous clinical studies}
+[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
+However, we can extrapolate from similar application areas where such methodologies have a
+much longer history of development such as white matter lesion segmentation in
+multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
+were found during a literature search and evaluated which concern automated segmentation of white
+matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
+of two cohorts from two different sites used in
+the MS Lesion segmentation challenge associated with the international MICCAI 2008
+conference.  For comparison, this challenge training data set
+has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
+our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
+40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
+almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
+Note that our relative volume difference numbers are comparable to the
+relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
+[@vanginneken:2007aa].
+
+Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
+The contralateral difference FLAIR image is  highly discriminative over the set of
+evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
+FLAIR images for identifying white matter hyperintensities and the fact that such
+pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
+posterior maps for the deep gray matter are extremely important for accurate white matter
+hyperintensity segmentation.  Perhaps the spatial specification of deep
+gray matter aids in the removal of false positives.  Inspection of the bottom of the
+plots demonstrates the lack of discriminating features associated with the
+T1 image which is also well-known in the clinical literature.
+
+![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
+The region around the WMHs is enlarged (b) in the original FLAIR and the
+(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
+
+As described earlier, for Stage 2, we used the output random forest voting maps from Stage
+1 as both features themselves and as priors for input to a Bayesian-based segmentation with
+an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
+"_RFStage1VotingMaps_" where the final numeral is associated with the brain
+parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
+feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
+The Stage 2 feature importance plot follows similar
+trends as that for Stage 1 with the T1 images not contributing much to the identification
+of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
+important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
+3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
+alone it is intuitive that such features would be important.
 
@@
 -->
+
+# Conclusions
 

### 2016-06-30 figureCaptions.Rmd
ENH:  More revisions.

@@
+---
+output:
+  word_document:
+    fig_caption: true
+  pdf_document:
+    fig_caption: true
+    latex_engine: xelatex
+    keep_tex: yes
+header-includes:
+   - \usepackage{booktabs}
+   - \usepackage[final]{changes}
+   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
+   - \linespread{2}
+   - \usepackage[compact]{titlesec}
+   - \usepackage{enumitem}
+   - \usepackage{tikz}
+   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
+   - \setlist{nolistsep}
+   - \titlespacing{\section}{2pt}{*0}{*0}
+   - \titlespacing{\subsection}{2pt}{*0}{*0}
+   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
+   - \setlength{\parskip}{3pt}
+   - \setremarkmarkup{(#2)}
+bibliography: references.bib
+csl: council-of-science-editors.csl
+fontsize: 12pt
+mainfont: Georgia
+---
+
+<style type="text/css">
+body,
+code.bash{
+  font-size: 8px;
+}
+pre {
+  font-size: 8px
+}
+</style>
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set( cache=TRUE )
+```
+
 # Figure Captions
@@
 
-__Figure 5:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
+__Figure 5:__ \textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
+volume difference.   These quantitative assessments are given for three quantile ranges
+spanning the range of the manually-derived lesion volumes.    Overall improvement in all
+three whole lesion-based measuers is seen as the
+second Stage RF model is applied for all three quantile ranges.
+The relative volume difference corresponding to the Stage 2 results tend to predict a
+decreased predicted volume over the Stage 1 results.}
+
+__Figure 6:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
 forest models for Stage 1 during the leave-one-out evaluation.  These plots
@@
 
-__Figure 6:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
+__Figure 7:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
 forest models for Stage 2 during the leave-one-out evaluation.  These plots
@@
 
-__Figure 7:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
+__Figure 8:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
 The region around the WMHs is enlarged (b) in the original FLAIR and the
@@
 
-__Figure 8:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
-On the left are the
-calculated Dice values over all white matter hyperintensities.  Note the improvement
-in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
-(Right) Similar results can be seen by comparing the total lesion load volume between manual
-and automated detection strategies.  Although some outliers are found after the Stage 2
-processing in a couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.
-
 \clearpage

### 2016-06-30 figureCaptions.tex
ENH:  More revisions.

--- 2016-06-29 stitched.tex
@@
-\documentclass[11pt,]{article}
+\documentclass[12pt,]{article}
 \usepackage{lmodern}
@@
 }{}
-\usepackage[margin=1.0in]{geometry}
+\usepackage[margin=1in]{geometry}
 \ifxetex
@@
 \setremarkmarkup{(#2)}
-\definechangesauthor[name={Nick Tustison}, color=red]{nt}
-\definechangesauthor[name={James Stone}, color=magenta]{js}
-\definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
-\definechangesauthor[name={Andy Mayer}, color=green]{am}
-\definechangesauthor[name={Harvey Levin}, color=blue]{hl}
-\definechangesauthor[name={Brian Taylor}, color=purple]{bt}
-\definechangesauthor[name={David Tate}, color=orange]{dt}
 
@@
 
-\pagenumbering{gobble}
-
-\section{Abstract}\label{abstract}
-
-White matter hyperintensities (WMHs) are foci of abnormal signal
-intensity in white matter regions seen with magnetic resonance imaging
-(MRI). WMHs are associated with normal aging and have shown prognostic
-value in neurological conditions such as traumatic brain injury (TBI).
-The impracticality of manually quantifying these lesions limits their
-clinical utility and motivates the utilization of machine learning
-techniques for automated segmentation workflows. Herein, we develop a
-concatenated random forest framework with image features for segmenting
-WMHs in a TBI cohort. The framework is built upon the Advanced
-Normalization Tools (ANTs) and ANTsR toolkits. MR (3D FLAIR, T2-, and
-T1-weighted) images from 24 service members and veterans scanned in the
-Chronic Effects of Neurotrauma Consortium's (CENC) observational study
-were acquired. Manual annotations were employed for both training and
-evaluation using a leave-one-out strategy. Lesion load and overlap
-evaluative comparisons are complimented by feature rankings which
-showcase the utility of the concatenated approach. Our findings suggest
-supervised learning methods may be applied to quantify WMHs on routine
-brain imaging. Paired with correlative outcome data, supervised learning
-methods may allow for identification of imaging features predictive of
-diagnosis and prognosis in individual TBI patients.
-
-\clearpage
-
-\section{Introduction}\label{introduction}
-
-\subsection{White matter hyperintensities in
-TBI}\label{white-matter-hyperintensities-in-tbi}
-
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric}\textsuperscript{1,2}
-\textcolor{blue}{and adult}\textsuperscript{3--6}
-\textcolor{blue}{populations.}
-
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient}\textsuperscript{2,6--8}\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury}\textsuperscript{6}\textcolor{blue}{.
-Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery.}\textsuperscript{4}\textcolor{blue}{. Additionally,
-in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome}\textsuperscript{7}.
-
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
-Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
-In this work, we present an automated framework for quantification of WMHs in
-multi-modal MRI using the random forest machine learning technique.}
-
-\subsection{Random forests for WMH
-segmentation}\label{random-forests-for-wmh-segmentation}
-
-The random forests framework\textsuperscript{9} is a popular machine
-learning technique that has demonstrated significant utility for
-supervised segmentation tasks (e.g., normal human brain
-segmentation\textsuperscript{10}) and other computer vision applications
-(e.g.,\textsuperscript{11}).
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
-delineation of other neuropathologies}\textsuperscript{12--17}
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
-
-Random forests are conceptually straightforward\textsuperscript{9}.
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
-For the proposed application, individual
-voxels (and their corresponding feature values) are introduced at the root of a particular
-decision tree and traverse the edges and internal nodes ultimately ending up at one of
-the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)}\textsuperscript{18,19}
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{20,21}.
-\textcolor{blue}{Finally, Breiman}\textsuperscript{9}
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
-
-In this work, we develop a concatenated random forest framework with a
-feature image set (both spatial and intensity-based) for segmenting WMHs
-in a large TBI cohort. The entire framework is built on the well-known
-open-source Advanced Normalization Tools (ANTs)\footnote{\url{https://github.com/stnava/ANTs}}
-and ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
-Further motivating this research is the availability of several large
-publicly available imaging data sets that permits testing
-reproducibility of this automated routine for WMH segmentation and
-quantification.
-
-\section{Materials and Methods}\label{materials-and-methods}
-
-\subsection{Imaging}\label{imaging}
-
-MR images utilized for this initial report were acquired from a single
-scanner involved in the Chronic Effects of Neurotrauma Consortium's
-(CENC) observational study (see Walker et al., this issue). Briefly,
-participants were Operation Iraqi Freedom/Operation Enduring Freedom
-(OIF/OEF) era Service Members and Veterans between the ages of 18-60
-years with prior combat exposure and deployment(s). The feature images
-\textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged
-39.6 \(\pm\) 8.1 years (range 28--58 years). Within this cohort, 24
-(92\%) were considered positive for TBI based upon the potential
-concussive events (PCE) interview process described in detail in Walker
-et al., this issue).
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.  Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
-
-Images were acquired on a Philips 3.0T Ingenia system with an 8-channel
-SENSE head coil (Philips Medical Systems, Best, Netherlands). 3D FLAIR
-sequences were acquired with a turbo spin echo inversion recovery
-sequence with the following parameters: repetition time (TR) = 4800 ms,
-echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal
-slices with a 1.2 mm slice thickness, 256 \(\times\) 256 acquisition
-matrix, and 256 \(\times\) 256 mm FOV. 3D T1-weighted sequences were
-acquired with a fast field echo (FFE) sequence with the following
-parameters: TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240;
-Flip angle = 9\(^\circ\), 170 sagittal slices with a 1.2 mm slice
-thickness, 256x240 acquisition matrix, and 256 \(\times\) 256 mm FOV. In
-addition, 3D T2-weighted images were acquired with a turbo spin echo
-sequence with the following parameters: TR = 2500 ms, TE = 245 ms, ET: =
-133; 170 sagittal slices with a 1.2 mm slice thickness, 256 \(\times\)
-256 acquisition matrix, and 256 x 256 mm FOV.
-
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
-subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All
-multi-modal MR dicom image slices were converted to the nifti file format}.\footnote{\url{http://nifti.nimh.nih.gov/nifti-1}}
-\textcolor{blue}{Given the
-utility of the FLAIR sequence in detecting white matter lesions in TBI}\textsuperscript{3}\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software}\textsuperscript{22}
-\textcolor{blue}{.  The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{23}
-\textcolor{blue}{
-which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
-
-\subsection{Quantitative analysis}\label{quantitative-analysis}
-
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1 '', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
-
-\subsubsection{Symmetric multi-modal
-templates}\label{symmetric-multi-modal-templates}
-
-\textcolor{blue}{Following}\textsuperscript{24}
-\textcolor{blue}{and}\textsuperscript{17},
-\textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{25}\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{26}
-\textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
-
-\subsubsection{Feature images for WMH
-segmentation}\label{feature-images-for-wmh-segmentation}
-
-\textcolor{blue}{
-Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.} For the targeted application in
-this work, tissue classification is performed at the voxelwise level. In
-other words, each voxel within the region of interest is sent through
-the ensemble of decision trees and receives a set of classification
-votes from each tree thus permitting a regression or classification
-solution. Since this procedure is performed at the voxelwise level,
-intensity information alone is insufficient for good segmentation
-performance due to the lack of spatial context. For example, as pointed
-out in\textsuperscript{27}, higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated
-lesion detection algorithms. Other potential confounds include MR signal
-inhomogeneity and noise. Therefore, even though machine learning and
-pattern recognition techniques are extremely powerful and have
-significant potential, just as crucial to outcome is the creative
-construction and deployment of salient feature images which we detail
-below.
-
-Supervised methodologies are uniquely characterized, in part, by the
-feature images that are used to identify the regions of interest. In
-Table 2, we provide a list and basic categorization of the feature
-images used for the initial (i.e., Stage 1---more on the use of multiple
-random forest stages below) segmentation of the WMHs. In addition Figure
-3 provides a representation of a set of feature images for a single
-subject analyzed in this work. Note that in this work we categorize the
-brain parenchyma with seven labels:
-
-\begin{itemize}
-\tightlist
-\item
-  cerebrospinal fluid (label 1),
-\item
-  gray matter (label 2),
-\item
-  white matter (label 3),
-\item
-  deep gray matter (label 4),
-\item
-  brain stem (label 5),
-\item
-  cerebellum (label 6), and
-\item
-  white matter hyperintensities (label 7).
-\end{itemize}
-
-As mentioned previously, input for each subject comprises FLAIR, T1-,
-and T2-weighted acquisitions. The T1 and T2 images are rigidly
-registered to the FLAIR image using the open-source Advanced
-Normalization Tools (ANTs)\textsuperscript{22}. The aligned images are
-then preprocessed using the denoising algorithm of\textsuperscript{28}
-followed by N4 bias correction\textsuperscript{29} which are then
-normalized to the intensity range \([0,1]\). Although we could have used
-an alternative intensity standardization algorithm
-(e.g.,\textsuperscript{30}), we found that a simple linear rescaling
-produced better results similar to previous work\textsuperscript{17}.
-
-The T1 image is then processed via the ANTs brain extraction and normal
-tissue segmentation pipelines\textsuperscript{24}. The result is a mask
-delineating the brain parenchyma and probabilistic estimates of the CSF,
-gray matter, white matter, deep gray matter, brain stem, and
-cerebellum\textsuperscript{31}. These provide the expertly annotated
-labels for the first six tissue labels given above. Tissue prior
-probability maps for segmentation are from multi-model optimal symmetric
-shape/intensity templates\textsuperscript{17} created from the public
-MMRR data set\textsuperscript{25} (cf Figure 2).
-
-Feature values include the preprocessed FLAIR, T1, and T2 image voxel
-intensities. We also calculate a set of neighborhood statistics (mean,
-standard deviation, and skewness) feature images using a Manhattan
-radius of one voxel given the typical size of individual WMHs. For each
-of the preprocessed images, we calculate the difference in intensities
-with the corresponding warped template component. Previous success in
-the international brain tumor segmentation
-competition\textsuperscript{32} was based on an important set of
-intensity features that were created from multi-modal templates
-mentioned previously\textsuperscript{17} and listed in Table 2. We
-employ the same strategy here.
-
-To take advantage of the gross bilateral symmetry of the normal brain
-(in terms of both shape and intensity), and the fact that WMHs do not
-generally manifest symmetrically across hemispheres, we use the
-symmetric templates to compute the contralateral intensity differences
-as an additional intensity feature.
-
-The segmentation probability images described above are used as feature
-images to provide a spatial context for the random forest model
-prediction step. Additional spatial contextual feature images include
-the distance maps\textsuperscript{33} based on the csf, gray matter, and
-deep gray matter images. These latter images are intended to help
-distinguish white matter hyperintensities from false positives induced
-by the partial voluming at the gray/white matter interface. A third set
-of images are based on the voxel location within the space of the
-template.
-\textcolor{blue}{Similar feature images were used in}\textsuperscript{34}
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant}\textsuperscript{35}
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{36}.
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
-
-\subsubsection{Stacked (concatenated) random forests for improved
-segmentation
-performance}\label{stacked-concatenated-random-forests-for-improved-segmentation-performance}
-
-In previous brain tumor segmentation work\textsuperscript{17}, it was
-demonstrated that a concatenated supervised approach, whereby the
-prediction output from the first random forest model serves as partial
-input for a second random forest model, can significantly improve
-segmentation performance. We do the same thing for the work described
-here where we employ two stacked random forests (or two ``stages''). The
-Stage 1 feature images of the training data (as described previously)
-are used to construct the Stage 1 model. The training data Stage 1
-features are then used to produce the voxelwise ``voting maps'' (i.e.,
-the classification count of each decision tree for each tissue label)
-via the Stage 1 random forest model. All the Stage 1 features plus the
-Stage 1 voting maps are used as input to the Stage 2 model. In addition,
-we use the Stage 1 voting maps as tissue priors (i.e., probabilistic
-estimates of the tissue spatial locations) for a second application of
-the \(6\)-tissue segmentation algorithm with an additional Markov Random
-Field spatial prior (MAP-MRF)\textsuperscript{31}.
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
-all three aligned preprocessed images for multivariate segmentation during the
-second stage.} The resulting seven posterior probability images
-constitute a third additional feature image set for Stage 2.
-
-\subsubsection{Implementation}\label{implementation}
-
-As pointed out in a recent comprehensive lesion segmentation
-review\textsuperscript{37}, although the number of algorithms reported
-in the literature is quite extensive, there were only four publicly
-available segmentation algorithms at the time of writing this article.
-In contrast to the current work, none are based on supervised learning.
-As we did for our brain tumor segmentation
-algorithm\textsuperscript{17}, all of the code described in this work is
-publicly available through the open-source ANTs/ANTsR toolkits. Through
-ANTsR (an add-on toolkit which, in part, bridges ANTs and the R
-statistical project) we use the \emph{randomForest}
-package\textsuperscript{38} using the default settings with 2000 trees
-per model and 500 randomly selected samples per label per image. Note
-that we saw little variation in performance when these parameters were
-changed (i.e.~up to 1000 random samples and as little as 1000 trees)
-which is consistent with our previous experience.
-
-In addition, similar to our previous offering,\footnote{\url{https://github.com/ntustison/ANTsAndArboles}}
-we plan on creating a self-encapsulated example to showcase the proposed
-methodology
-\textcolor{blue}{which will also be available on github.}\footnote{\url{https://github.com/ntustison/WatchMeHyperventilate}}
-The fact that the data will also be made available through the Federal
-Interagency Traumatic Brain Injury Research (FITBIR) repository along
-with the manual labelings will facilitate reproducibility on the part of
-the reader as well as any interest in extending the proposed framework
-to other data sets.
-
-\subsubsection{Evaluation protocol
-overview}\label{evaluation-protocol-overview}
-
-In order to evaluate the protocol described, we performed a
-leave-one-out evaluation using the data acquired from the 24 subjects
-described above. Initial processing included the creation of all Stage 1
-feature images for all subjects. The initial brain segmentation of each
-T1 image and the manual white matter hyperintensity tracings were
-combined to provide the truth labels for the training data. The
-``truth'' labels are the seven anatomical regions given above.
-
-The leave-one-out procedure is as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  Create Stage 1 feature images for all 24 subjects.
-\item
-  For each of the 24 subjects:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    sequester the current subject and corresponding feature images.
-  \item
-    construct the Stage 1 random forest model from the remaining 23
-    subjects.
-  \item
-    apply the Stage 1 random forest model to the feature images of the
-    23 training subjects.
-  \item
-    the previous step produces the Stage 1 voting maps for all seven
-    labels.
-  \item
-    for each of the 23 subjects, perform a Bayesian-based segmentation
-    with an MRF spatial prior using the seven voting maps as additional
-    tissue priors.
-  \item
-    construct the Stage 2 random forest model from all the Stage 1
-    feature images, seven voting maps, and seven posterior probability
-    maps from the previous step.
-  \item
-    send the sequestered subject through the random forest models for
-    both stages.
-  \item
-    compare the final results with the manually-defined white matter
-    hyperintensity regions.
-  \end{itemize}
-\end{itemize}
-
-\textcolor{blue}{Several measures have been employed in the literature for evaluating
-automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.
-For this work, due to the relatively small-size distribution}\footnote{\textcolor{blue}{For comparison, the training data set of
-  the MS Lesion Segmentation challenge
-  associated with the international MICCAI 2008 conference has a mean lesion load of
-  204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
-  in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}}
-\textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
-four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
-incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
-The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
-
-\[ sensitivity =  \frac{ TP }{TP + FN}, \]
-
-\textcolor{blue}{and the positive predictive value (PPV),}
-
-\[ PPV =  \frac{ TP }{TP + FP}, \]
-
-\textcolor{blue}{such that $F_1$ is given by}
-
-\[ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. \]
-
-\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
-the manual and predicted lesion volumes:}
-
-\[ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. \]
-
-\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
-These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
-
-\section{Results}\label{results}
-
-\subsection{White matter hyperintensity segmentation
-evaluation}\label{white-matter-hyperintensity-segmentation-evaluation}
-
-\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data.  These performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed to illustrate the variation
-in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity is widely varied over the cohort vs. the largest
-set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of how accurate
-the volumetric estimate is when comparing the manually identified lesions versus the
-automatically predicted lesions.}
-
-\subsection{Ranking feature
-importance}\label{ranking-feature-importance}
-
-After performing the leave-one-out evaluation, we calculated the
-\emph{MeanDecreaseAccuracy} feature values for each of the 24 subjects
-\(\times\) 2 models per subject \(=48\) total models. This measure (per
-feature, per model) is calculated during the out-of-bag phase of the
-random forest model construction and quantifies the decrease in
-prediction accuracy from omitting the specified feature. In other words,
-this quantity helps determine the importance of a particular feature
-and, although we save such efforts for future work, this information
-provides us with guidance for future feature pruning and/or additions.
-
-The resulting rankings for both Stages are given in Figures 5 and 6
-where the values for the separate stages are averaged over the entire
-corresponding model set. In addition, we track the variance for each
-feature over all models to illustrate the stability of the chosen
-features during the evaluation. This latter information is illustrated
-as horizontal errors bars providing the \(95^{th}\) percentile Note that
-the reader can cross reference Table 1 for identifying corresponding
-feature types and names.
-
-Additionally, it is interesting to note some of the other top performing
-features for Stage 1. The contralateral difference FLAIR image is highly
-discriminative over the set of evaluation random forest models (see
-Figure 7). This accords with the known clinical relevance of FLAIR
-images for identifying white matter hyperintensities and the fact that
-such pathology does not typically manifest symmetrically in both
-hemispheres. Interestingly, the posterior maps for the deep gray matter
-are extremely important for accurate white matter hyperintensity
-segmentation. Perhaps the spatial specification of deep gray matter aids
-in the removal of false positives. Inspection of the bottom of the plots
-demonstrates the lack of discriminating features associated with the T1
-image which is also well-known in the clinical literature.
-
-As described earlier, for Stage 2, we used the output random forest
-voting maps from Stage 1 as both features themselves and as priors for
-input to a Bayesian-based segmentation with an additional MRF spatial
-prior. In Figure 6, the voting maps are labeled as
-``\emph{RFStage1VotingMaps}'' where the final numeral is associated with
-the brain parenchymal labeling given previously. Similarly, the
-additional RF prior segmentation feature probability maps are labeled as
-``\emph{RFBrainSegmentationPosteriors}''. The Stage 2 feature importance
-plot follows similar trends as that for Stage 1 with the T1 images not
-contributing much to the identification of white matter hyperintensity
-voxels. The initial voting maps from Stage 1 are extremely important
-with the top 3 being the estimated locations of the 1) gray matter, 2)
-white matter, and 3) white matter hyperintensities. Since these tissue
-type can be conflated based on intensity alone it is intuitive that such
-features would be important.
-
-\section{Discussion}\label{discussion}
-
-The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation
-of WMHs from FLAIR images to help build two-stage ensembles of decision
-trees for the automated identification of these lesions. Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology. Also, we
-recognize that only a single site was used for evaluating the proposed
-framework. However, we are currently processing other site data with the
-models developed for this work and the results look promising since the
-developed features are site-agnostic.
-
-As far as we know, this is the first report utilizing a novel random
-forest approach to identify WMHs in a cohort of TBI patients. TBI WMHs
-tend to be more difficult to segment than MS lesions as the former tend
-to be smaller with an overall smaller lesion load. Also, enhancement
-protocols with the former tend to be less successful than with the
-latter. As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.
-
-Two major meta-analyses of WMHs have been published covering the periods
-prior to\textsuperscript{39} and after 2010\textsuperscript{40}. Debette
-\& Markus\textsuperscript{39} found that the presence of WMHs was
-related to subsequent cognitive decline, a higher risk of developing
-dementia, stroke, and of mortality. Lesion volume at baseline was also
-predictive of cognitive decline. Kloppenborg et al.\textsuperscript{40}
-of 23 cross-sectional studies reporting MRI and concurrent
-neuropsychological results in patients with heterogeneous diagnoses but
-without previously diagnosed cognitive impairment, found that WMHs were
-associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
-to -0.08) after controlling for age.
-
-Despite the potential clinical significance of WMHs these lesions
-receive little attention in current clinical workflows. When reported in
-a standard neuroradiologist interpretation, they are typically handled
-as incidental findings and are assigned little clinical significance.
-This likely reflects the impracticality of performing a detailed
-assessment of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
-supervised statistical learning tool that is appropriate for the
-application to such large-scale datasets.
-
-\clearpage
-
-\subsection{Acknowledgements}\label{acknowledgements}
-
-The authors wish to acknowledge all other members of the CENC
-Neuroimaging Steering Committee and CENC leadership (Drs. David X. Cifu,
-Ramon Diaz-Arrastia, and Rick Williams) for their support. We also
-gratefully acknowledge the assistance of Tracy Nolen, Chris Siege and
-Kevin Wilson. We would also like to thank the study participants and
-their family members. This project was jointly supported by the
-Department of Defense (W81XWH-13-2-0095), the U.S. Department of
-Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
-Grant HU 0001-08-0001.
-
-\subsection{Declaration of
-Interest/Disclaimer}\label{declaration-of-interestdisclaimer}
-
-The authors report no financial disclosures or conflicts of interest.
-The views expressed here are those of the authors and do not necessarily
-reflect the official policy of position of the Department of the Navy,
-Department of Defense, nor the U.S. Government. This work was prepared
-as a part of official duties; Title 17 USC §105 provides that Copyright
-protection under this title is not available for any work of the U.S.
-Government. Title 17 USC §101 defines a US Government work as a work
-prepared by a military service member of employee of the US Government
-as part of that person's official duties.
-
-\clearpage
-
-\section{References}\label{references}
-
 \section{Figure Captions}\label{figure-captions}
@@
 \textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
-symmetric template constructed from the MMRR data
-set\textsuperscript{25} (only shown are the FLAIR, T1, and T2
-modalities--- the components relevant for this work). Template
-construction is detailed in\textsuperscript{17}. These images are
-important for asymmetry-based features.
+symmetric template constructed from the MMRR data set\textsuperscript{1}
+(only shown are the FLAIR, T1, and T2 modalities--- the components
+relevant for this work). Template construction is detailed
+in\textsuperscript{2}. These images are important for asymmetry-based
+features.
 
@@
 01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
-pre-aligned\textsuperscript{22} to the space of the T1 image. The three
+pre-aligned\textsuperscript{3} to the space of the T1 image. The three
 modality images are then preprocessed (N4 bias
-correction\textsuperscript{29} and adaptive
-denoising\textsuperscript{28}) followed by application of standard ANTs
-brain extraction and \(n\)-tissue segmentation protocols using the MMRR
-symmetric template and corresponding priors\textsuperscript{24} applied
-to the T1 image. The feature images are then generated for voxelwise
-input to the RF model which results in the voting maps illustrated on
-the right. This gives a probabilistic classification of tissue type. Not
-shown are the probability and voting images for the brain stem and
-cerebellum.
+correction\textsuperscript{4} and adaptive denoising\textsuperscript{5})
+followed by application of standard ANTs brain extraction and
+\(n\)-tissue segmentation protocols using the MMRR symmetric template
+and corresponding priors\textsuperscript{6} applied to the T1 image. The
+feature images are then generated for voxelwise input to the RF model
+which results in the voting maps illustrated on the right. This gives a
+probabilistic classification of tissue type. Not shown are the
+probability and voting images for the brain stem and cerebellum.
 
@@
 
-\textbf{Figure 5:} Average \emph{MeanDecreaseAccuracy} plots generated
+\textbf{Figure 5:}
+\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
+volume difference.   These quantitative assessments are given for three quantile ranges
+spanning the range of the manually-derived lesion volumes.    Overall improvement in all
+three whole lesion-based measuers is seen as the
+second Stage RF model is applied for all three quantile ranges.
+The relative volume difference corresponding to the Stage 2 results tend to predict a
+decreased predicted volume over the Stage 1 results.}
+
+\textbf{Figure 6:} Average \emph{MeanDecreaseAccuracy} plots generated
 from the creation of all 24 random forest models for Stage 1 during the
@@
 
-\textbf{Figure 6:} Average \emph{MeanDecreaseAccuracy} plots generated
+\textbf{Figure 7:} Average \emph{MeanDecreaseAccuracy} plots generated
 from the creation of all 24 random forest models for Stage 2 during the
@@
 
-\textbf{Figure 7:} (a) FLAIR image slice illustrating WMHs which have
+\textbf{Figure 8:} (a) FLAIR image slice illustrating WMHs which have
 been manually delineated. The region around the WMHs is enlarged (b) in
 the original FLAIR and the (c) contralateral FLAIR difference image.
-
-\textbf{Figure 8:} Voxelwise comparison with manual delineation of white
-matter hyperintensities. On the left are the calculated Dice values over
-all white matter hyperintensities. Note the improvement in the Dice
-metric from the employment of the Stage 2 component of the processing
-pipeline. (Right) Similar results can be seen by comparing the total
-lesion load volume between manual and automated detection strategies.
-Although some outliers are found after the Stage 2 processing in a
-couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Bigler:2013aa}{}
-1. Bigler ED, Abildskov TJ, Petrie J, Farrer TJ, Dennis M, Simic N,
-Taylor HG, Rubin KH, Vannatta K, Gerhardt CA, et al. Heterogeneity of
-brain lesions in pediatric traumatic brain injury. Neuropsychology.
-2013;27(4):438--51.
-
-\hypertarget{ref-Smitherman:2016aa}{}
-2. Smitherman E, Hernandez A, Stavinoha PL, Huang R, Kernie SG,
-Diaz-Arrastia R, Miles DK. Predicting outcome after pediatric traumatic
-brain injury by early magnetic resonance imaging lesion location and
-volume. J Neurotrauma. 2016;33(1):35--48.
-
-\hypertarget{ref-Marquez-de-la-Plata:2007aa}{}
-3. Marquez de la Plata C, Ardelean A, Koovakkattu D, Srinivasan P,
-Miller A, Phuong V, Harper C, Moore C, Whittemore A, Madden C, et al.
-Magnetic resonance imaging of diffuse axonal injury: Quantitative
-assessment of white matter lesion volume. J Neurotrauma.
-2007;24(4):591--8.
-
-\hypertarget{ref-Moen:2014aa}{}
-4. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A. Traumatic
-axonal injury: The prognostic value of lesion load in corpus callosum,
-brain stem, and thalamus in different magnetic resonance imaging
-sequences. J Neurotrauma. 2014;31(17):1486--96.
-
-\hypertarget{ref-Ding:2008aa}{}
-5. Ding K, Marquez de la Plata C, Wang JY, Mumphrey M, Moore C, Harper
-C, Madden CJ, McColl R, Whittemore A, Devous MD, et al. Cerebral atrophy
-after traumatic white matter injury: Correlation with acute neuroimaging
-and outcome. J Neurotrauma. 2008;25(12):1433--40.
-
-\hypertarget{ref-Pierallini:2000aa}{}
-6. Pierallini A, Pantano P, Fantozzi LM, Bonamini M, Vichi R, Zylberman
-R, Pisarri F, Colonnese C, Bozzao L. Correlation between mRI findings
-and long-term outcome in patients with severe brain trauma.
-Neuroradiology. 2000;42(12):860--7.
-
-\hypertarget{ref-Weiss:2008aa}{}
-7. Weiss N, Galanaud D, Carpentier A, Tezenas de Montcel S, Naccache L,
-Coriat P, Puybasset L. A combined clinical and mRI approach for outcome
-assessment of traumatic head injured comatose patients. J Neurol.
-2008;255(2):217--23.
-
-\hypertarget{ref-Levin:1988aa}{}
-8. Levin HS, Williams D, Crofford MJ, High WM Jr, Eisenberg HM, Amparo
-EG, Guinto FC Jr, Kalisky Z, Handel SF, Goldman AM. Relationship of
-depth of brain lesions to consciousness and outcome after closed head
-injury. J Neurosurg. 1988;69(6):861--6.
-
-\hypertarget{ref-breiman2001}{}
-9. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
-
-\hypertarget{ref-yi2009}{}
-10. Yi Z, Criminisi A, Shotton J, Blake A. Discriminative, semantic
-segmentation of brain tissue in MR images. Med Image Comput Comput
-Assist Interv. 2009;12(Pt 2):558--65.
-
-\hypertarget{ref-viola2005}{}
-11. Viola P, Jones M, Snow D. Detecting pedestrians using patterns of
-motion and appearance. International Journal of Computer Vision.
-2005;63:153--161.
-
-\hypertarget{ref-geremia2011}{}
-12. Geremia E, Clatz O, Menze BH, Konukoglu E, Criminisi A, Ayache N.
-Spatial decision forests for MS lesion segmentation in multi-channel
-magnetic resonance images. Neuroimage. 2011;57(2):378--90.
-
-\hypertarget{ref-Pustina:2016aa}{}
-13. Pustina D, Coslett HB, Turkeltaub PE, Tustison N, Schwartz MF,
-Avants B. Automated segmentation of chronic stroke lesions using lINDA:
-Lesion identification with neighborhood data analysis. Hum Brain Mapp.
-2016 Jan.
-
-\hypertarget{ref-geremia2012}{}
-14. Geremia E, Menze BH, Ayache N. Spatial decision forests for glioma
-segmentation in multi-channel MR images. In: Proceedings of MICCAI-BRATS
-2012. 2012.
-
-\hypertarget{ref-bauer2012}{}
-15. Bauer S, Fejes T, Slotboom J, Wiest R, Nolte L-P, Reyes M.
-Segmentation of brain tumor images based on integrated hierarchical
-classification and regularization. In: Proceedings of MICCAI-BRATS 2012.
-2012. pp. 10--13.
-
-\hypertarget{ref-zikic2012}{}
-16. Zikic D, Glocker B, Konukoglu E, Shotton J, Criminisi A, Ye DH,
-Demiralp C, Thomas OM, Das T, Jena R, et al. Context-sensitive
-classification forests for segmentation of brain tumor tissues. In:
-Proceedings of MICCAI-BRATS 2012. 2012. pp. 1--9.
+\hypertarget{ref-landman2011}{}
+1. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
+Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
+reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
 
 \hypertarget{ref-Tustison:2015aa}{}
-17. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee
-JC, Grossman MC, Avants BB. Optimal symmetric multimodal templates and
+2. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
+Grossman MC, Avants BB. Optimal symmetric multimodal templates and
 concatenated random forests for supervised brain tumor segmentation
@@
 
-\hypertarget{ref-schapire1990}{}
-18. Schapire R. The strength of weak learnability. Machine Learning.
-1990;5:197--227.
+\hypertarget{ref-Avants:2014aa}{}
+3. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
+ToolKit image registration framework. Front Neuroinform. 2014;8:44.
 
-\hypertarget{ref-freund1997}{}
-19. Freund Y, Schapire R. A decision-theoretic generalization of on-line
-learning and an application to boosting. Journal of Computer and System
-Sciences. 1997;55:119--139.
+\hypertarget{ref-Tustison:2010ac}{}
+4. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
+JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
+2010;29(6):1310--20.
 
-\hypertarget{ref-ho1995}{}
-20. Ho TK. Random decision forests. In: Document analysis and
-recognition, 1995., proceedings of the third international conference
-on. Vol. 1. 1995. pp. 278--282 vol.1.
-
-\hypertarget{ref-amit1997}{}
-21. Amit Y, Geman D. Shape quantization and recognition with randomized
-trees. Neural Computation. 1997;9:1545--1588.
-
-\hypertarget{ref-Avants:2014aa}{}
-22. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The
-Insight ToolKit image registration framework. Front Neuroinform.
-2014;8:44.
-
-\hypertarget{ref-Yushkevich:2006aa}{}
-23. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
-User-guided 3D active contour segmentation of anatomical structures:
-Significantly improved efficiency and reliability. Neuroimage.
-2006;31(3):1116--28.
+\hypertarget{ref-Manjon:2010aa}{}
+5. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
+non-local means denoising of mR images with spatially varying noise
+levels. J Magn Reson Imaging. 2010;31(1):192--203.
 
 \hypertarget{ref-Tustison:2014ab}{}
-24. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
+6. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
 Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
@@
 
-\hypertarget{ref-landman2011}{}
-25. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
-Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
-reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
-
-\hypertarget{ref-Avants:2010aa}{}
-26. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre
-J, Gee JC. The optimal template effect in hippocampus studies of
-diseased populations. Neuroimage. 2010;49(3):2457--66.
-
-\hypertarget{ref-Neema:2009aa}{}
-27. Neema M, Guss ZD, Stankiewicz JM, Arora A, Healy BC, Bakshi R.
-Normal findings on brain fluid-attenuated inversion recovery mR images
-at 3T. AJNR Am J Neuroradiol. 2009;30(5):911--6.
-
-\hypertarget{ref-Manjon:2010aa}{}
-28. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
-non-local means denoising of mR images with spatially varying noise
-levels. J Magn Reson Imaging. 2010;31(1):192--203.
-
-\hypertarget{ref-Tustison:2010ac}{}
-29. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
-JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
-2010;29(6):1310--20.
-
-\hypertarget{ref-nyul2000}{}
-30. Nyúl LG, Udupa JK, Zhang X. New variants of a method of MRI scale
-standardization. IEEE Trans Med Imaging. 2000;19(2):143--50.
-
-\hypertarget{ref-Avants:2011aa}{}
-31. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
-multivariate framework for \(n\)-tissue segmentation with evaluation on
-public data. Neuroinformatics. 2011;9(4):381--400.
-
-\hypertarget{ref-Menze:2015aa}{}
-32. Menze BH, Jakab A, Bauer S, Kalpathy-Cramer J, Farahani K, Kirby J,
-Burren Y, Porz N, Slotboom J, Wiest R, et al. The multimodal brain tumor
-image segmentation benchmark (bRATS). IEEE Trans Med Imaging.
-2015;34(10):1993--2024.
-
-\hypertarget{ref-maurer2003}{}
-33. Maurer CR, Rensheng Q, Raghavan V. A linear time algorithm for
-computing exact Euclidean distance transforms of binary images in
-arbitrary dimensions. Pattern Analysis and Machine Intelligence, IEEE
-Transactions on. 2003;25(2):265--270.
-
-\hypertarget{ref-Anbeek:2004aa}{}
-34. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
-Probabilistic segmentation of white matter lesions in mR imaging.
-Neuroimage. 2004;21(3):1037--44.
-
-\hypertarget{ref-Tustison:2013ac}{}
-35. Tustison NJ, Avants BB. Explicit B-spline regularization in
-diffeomorphic image registration. Front Neuroinform. 2013;7:39.
-
-\hypertarget{ref-Avants:2011ab}{}
-36. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
-reproducible evaluation of ANTs similarity metric performance in brain
-image registration. Neuroimage. 2011;54(3):2033--44.
-
-\hypertarget{ref-Garcia-Lorenzo:2013aa}{}
-37. García-Lorenzo D, Francis S, Narayanan S, Arnold DL, Collins DL.
-Review of automatic segmentation methods of multiple sclerosis white
-matter lesions on conventional magnetic resonance imaging. Med Image
-Anal. 2013;17(1):1--18.
-
-\hypertarget{ref-liaw2002}{}
-38. Liaw A, Wiener M. Classification and regression by randomForest. R
-News. 2002;2/3:18--22.
-
-\hypertarget{ref-Debette:2010aa}{}
-39. Debette S, Markus HS. The clinical importance of white matter
-hyperintensities on brain magnetic resonance imaging: Systematic review
-and meta-analysis. BMJ. 2010;341:c3666.
-
-\hypertarget{ref-Kloppenborg:2014aa}{}
-40. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
-Presence and progression of white matter hyperintensities and cognition:
-A meta-analysis. Neurology. 2014;82(23):2127--38.
-
 \end{document}

### 2016-06-30 intro.Rmd
ENH:  More revisions.

@@
-<!--
-A couple notes in response to comments:
-
-[David Tate, Comment 15]:  Stupid question, but why not show the same subject and slice for both for a better comparison?
-
-NT:  I see what he is saying and probably agree.
-
-[David Tate, Comment 28]:  Maybe this could go in a Figure as a flow chart??
-
-NT:  Flow charts are great and I included a couple as Figure 5 in my tumor paper.
-However, they take a long time to make correctly and I do not want to commit to producing one unless it is absolutely necessary.
-
-[Andy Mayer, Comment 30]: It will be difficult to see this font for labeling on Y axis at current size/resolution.
-You may also consider alerting your reader directly that the magnitudes of the scales are very different for stage 1 versus stage 2 in the legend or scale them to be similar.
-
-NT:  Comparing these two plots in Figure 4 in terms of their absolute MeanDecreaseAccuracy
-is meaningless so coordinating the scales is not important.
-
-[Andy Mayer, Comment 34]:  A reviewer is likely to ask for some sort of reliability rating on the labeling of the lesions.
-
-NT:  I agree with this but unless somebody else is willing to manually segment the WMHs,
-then there is not really much we can do other than point to other papers which have looked
-at reliability.
-
-[David Tate, Comment 37]:  Seems like we are missing a picture.
-That is the football player in me, but it seems like we need a picture that shows the manual compared to the automated at each stage if possible.
-
-NT:  We could do this but I do not think that such images in publications are that meaningful as they typically result from heavy selection bias.  I did do something similar
-for my brain segmentation paper where I showed multiple slices (skipped every 10, or something like that) for each subject in the challenge cohort but tumors are obviously much bigger and can be displayed in this manner.
-
-[David Tate, Comment 39]:  I would would move these comments down to a limitation like paragraph, otherwise it minimizes the cool findings.
-
-NT:  I have no opinion on this.  Do what you think is best.
-
-[Andy Mayer, Comment 40]:  I do not know if you can include any of this data (even one subject).
-It would be great to demonstrate that your algorithm worked on data collected on another scanner since the intensities are likely to be very different.
-
-NT:  As I mentioned to you previously, I think this is a great idea and I would be happy to include it but how do we showcase this unless somebody is willing to annotate data from these other sites.
-We could include a picture of a single subject but how persuasive that is will be reviewer-dependent.
-Personally, I would not be persuaded as the showing the results of a single subject is very prone to selection bias.
-
--->
-
 # Introduction
@@
 classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery.} [@Moen:2014aa]\textcolor{blue}{. Additionally,
+spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
 in patients who are comatose following severe TBI

### 2016-06-30 methods.Rmd
ENH:  More revisions.

@@
 
-<!--
-
 \input{trainingData.tex}
-
--->
 
@@
 multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{Given the
-utility of the FLAIR sequence in detecting white matter lesions in TBI}  [@Marquez-de-la-Plata:2007aa]\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
+\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
 The normalized MRI volumes were then provided to
@@
 probability estimates) of the first RF model application,
-which we denote as ``Stage1 '', is used as input (along with the original set of
+which we denote as ``Stage1'', is used as input (along with the original set of
 feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
 Once these offline steps
-are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
-
-<!--
+are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
 
@@
 
--->
-
 ### Symmetric multi-modal templates
@@
 
-<!--
 ![Canonical views of the mutlivariate, bilaterally symmetric template constructed
@@
 features.](Figures/MMRR.png)
-
--->
 
@@
 
-<!--
-
 \input{featureImagesTable.tex}
-
--->
 
@@
 
-<!--
-
 ![Representation of Stage 1 feature images for subject 01C1019.  The
@@
 
--->
-
 As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
@@
 
-<!--
-
 ![Sample FLAIR acquisition image slices showing both manual and random forest
@@
 truth WMH labels for training the random forest models.](Figures/sampleResults.png)
-
--->
 
@@
 voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
-For this work, due to the relatively small-size distribution}[^4] \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
+For this work, due to the relatively small-size distribution} \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
 four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
@@
 \textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
-These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
 
-[^4]: \textcolor{blue}{For comparison, the training data set of
-the MS Lesion Segmentation challenge
-associated with the international MICCAI 2008 conference has a mean lesion load of
-204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
-in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}
+* \textcolor{blue}{Quantile 1: $[1-12)$,}
+
+* \textcolor{blue}{Quantile 2: $[12-28)$, and}
+
+* \textcolor{blue}{Quantile 3: $[28-551]$}.
+
+\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}

### 2016-06-30 results.Rmd
ENH:  More revisions.

--- 2016-06-27 results.Rmd
@@
 # Results
+
+## White matter hyperintensity segmentation evaluation
+
+\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
+of the previously described TBI data over the three lesion volume ranges.
+These performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
+size ranges over which these measures are computed are meant to illustrate the variation
+in performance with lesion size.}
+
+\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
+to identify which is why the sensitivity for this range is more varied compared with the largest
+set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
+entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
+the volumetric estimate when comparing the manually identified lesions versus the
+automatically predicted lesions.}
+
+\textcolor{blue}{
+We averaged these measures over all lesions and all subjects which resulted in the
+following values:}
+
+* \textcolor{blue}{Stage 1}
+
+    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
+
+    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
+
+    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
+
+    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
+
+* \textcolor{blue}{Stage 1}
+
+    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
+
+    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
+
+    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
+
+    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
+
+<!--
+
+In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
+data.  Despite the large variability characteristic with manual labelings in related fields
+[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
+current clinical practices and the methodology proposed herein is readily adapted to refinements
+in training data.
+
+On the left of Figure 5 are the improvement in Dice values [@tustison2009],
+over all white matter hyperintensities when comparing the segmentations between the two stages
+where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
+represents the intersection between the manual/automated lesion pair.  Performing
+the second round of supervised learning improves these Dice values.  One can also note from the
+right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
+severe outliers in terms of the number of false positives.  The second round helps to
+correct this issue.
+
+-->
+
+![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
+volume difference.   These quantitative assessments are given for three quantile ranges
+spanning the range of the manually-derived lesion volumes.    Overall improvement in all
+three whole lesion-based measuers is seen as the
+second Stage RF model is applied for all three quantile ranges.
+The relative volume difference corresponding to the Stage 2 results tend to predict a
+decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
 
@@
 pruning and/or additions.
-
-<!--
 
@@
 
--->
-
-The resulting rankings for both Stages are given in Figures 5 and 6 where the values for the
+The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the
 separate stages are averaged over the entire corresponding model set.  In addition, we
@@
 and names.
-
-<!--
-One can also use these measurements as a type of sanity check.  For example, from the Stage
-1  plot, one can see that the _MeanDecreaseAccuracy_ values for the location indices in the
-anterior-posterior direction (i.e., _TemplateIndicesWarped1_) are greater than
-those for either the inferior-superior (i.e., _TemplateIndicesWarped2_) or the
-left-right (i.e., _TemplateIndicesWarped0_) directions in the space of the symmetric
-template.
--->
-
-Additionally, it is interesting to note some of the other top performing features for Stage 1.
-The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 7).  This accords with the known clinical relevance of
-FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
-posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives.  Inspection of the bottom of the
-plots demonstrates the lack of discriminating features associated with the
-T1 image which is also well-known in the clinical literature.
-
-<!--
-
-![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
--->
-
-As described earlier, for Stage 2, we used the output random forest voting maps from Stage
-1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 6, the voting maps are labeled as
-"_RFStage1VotingMaps_" where the final numeral is associated with the brain
-parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
-The Stage 2 feature importance plot follows similar
-trends as that for Stage 1 with the T1 images not contributing much to the identification
-of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
-important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
-3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
-alone it is intuitive that such features would be important.
-
-## White matter hyperintensity segmentation evaluation
-
-In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-<!--
-i.e.,
-
-$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$
--->
-
-On the left of Figure 8 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
-<!--
-
-![Voxelwise comparison with manual delineation of white matter hyperintensities.  On the left are the
-calculated Dice values over all white matter hyperintensities.  Note the improvement
-in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
-(Right) Similar results can be seen by comparing the total lesion load volume between manual
-and automated detection strategies.  Although some outliers are found after the Stage 2
-processing in a couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.](Figures/llvAndDice.png)
-
--->

### 2016-06-30 stitched.Rmd
ENH:  More revisions.

@@
 \clearpage
-<!--
-A couple notes in response to comments:
-
-[David Tate, Comment 15]:  Stupid question, but why not show the same subject and slice for both for a better comparison?
-
-NT:  I see what he is saying and probably agree.
-
-[David Tate, Comment 28]:  Maybe this could go in a Figure as a flow chart??
-
-NT:  Flow charts are great and I included a couple as Figure 5 in my tumor paper.
-However, they take a long time to make correctly and I do not want to commit to producing one unless it is absolutely necessary.
-
-[Andy Mayer, Comment 30]: It will be difficult to see this font for labeling on Y axis at current size/resolution.
-You may also consider alerting your reader directly that the magnitudes of the scales are very different for stage 1 versus stage 2 in the legend or scale them to be similar.
-
-NT:  Comparing these two plots in Figure 4 in terms of their absolute MeanDecreaseAccuracy
-is meaningless so coordinating the scales is not important.
-
-[Andy Mayer, Comment 34]:  A reviewer is likely to ask for some sort of reliability rating on the labeling of the lesions.
-
-NT:  I agree with this but unless somebody else is willing to manually segment the WMHs,
-then there is not really much we can do other than point to other papers which have looked
-at reliability.
-
-[David Tate, Comment 37]:  Seems like we are missing a picture.
-That is the football player in me, but it seems like we need a picture that shows the manual compared to the automated at each stage if possible.
-
-NT:  We could do this but I do not think that such images in publications are that meaningful as they typically result from heavy selection bias.  I did do something similar
-for my brain segmentation paper where I showed multiple slices (skipped every 10, or something like that) for each subject in the challenge cohort but tumors are obviously much bigger and can be displayed in this manner.
-
-[David Tate, Comment 39]:  I would would move these comments down to a limitation like paragraph, otherwise it minimizes the cool findings.
-
-NT:  I have no opinion on this.  Do what you think is best.
-
-[Andy Mayer, Comment 40]:  I do not know if you can include any of this data (even one subject).
-It would be great to demonstrate that your algorithm worked on data collected on another scanner since the intensities are likely to be very different.
-
-NT:  As I mentioned to you previously, I think this is a great idea and I would be happy to include it but how do we showcase this unless somebody is willing to annotate data from these other sites.
-We could include a picture of a single subject but how persuasive that is will be reviewer-dependent.
-Personally, I would not be persuaded as the showing the results of a single subject is very prone to selection bias.
-
--->
 
@@
 \textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
-These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
 
-[^4]: \textcolor{blue}{For comparison, the training data set of
-the MS Lesion Segmentation challenge
-associated with the international MICCAI 2008 conference has a mean lesion load of
-204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
-in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}
+* \textcolor{blue}{Quantile 1: $[1-12)$,}
+
+* \textcolor{blue}{Quantile 2: $[12-28)$, and}
+
+* \textcolor{blue}{Quantile 3: $[28-551]$}.
+
+\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
 
@@
 
-\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data.  These performance measures include sensitivity,
+\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
+of the previously described TBI data over the three lesion volume ranges.
+These performance measures include sensitivity,
 positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed to illustrate the variation
-in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity is widely varied over the cohort vs. the largest
-set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of how accurate
-the volumetric estimate is when comparing the manually identified lesions versus the
+size ranges over which these measures are computed are meant to illustrate the variation
+in performance with lesion size.}
+
+\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
+to identify which is why the sensitivity for this range is more varied compared with the largest
+set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
+entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
+the volumetric estimate when comparing the manually identified lesions versus the
 automatically predicted lesions.}
+
+\textcolor{blue}{
+We averaged these measures over all lesions and all subjects which resulted in the
+following values:}
+
+* \textcolor{blue}{Stage 1}
+
+    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
+
+    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
+
+    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
+
+    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
+
+* \textcolor{blue}{Stage 1}
+
+    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
+
+    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
+
+    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
+
+    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
 
@@
 
-In Figure 8 are the segmentation comparisons derived from manual segmentations of the same
+In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
 data.  Despite the large variability characteristic with manual labelings in related fields
@@
 
-On the left of Figure 8 are the improvement in Dice values [@tustison2009],
+On the left of Figure 5 are the improvement in Dice values [@tustison2009],
 over all white matter hyperintensities when comparing the segmentations between the two stages
@@
 the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 7 that the total lesion load volume illustrates a few subjects that are
+right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
 severe outliers in terms of the number of false positives.  The second round helps to
@@
 
-<!--
-
-![Evaluation measures for the leave-one-out protocol of the described protocol in the Methods
+![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
 section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
-volume difference.   The three quantile ranges include the following lesion volume ranges:
-$[3-12)$, $[12-28)$, and $[28-551]$ based on the manually-derived lesion volumes.  These
-measures are computed for both Stages.  Overall improvement is seen as the
-second Stage RF model is applied for all three quantile ranges.](Figures/evaluationResults.png)
-
--->
+volume difference.   These quantitative assessments are given for three quantile ranges
+spanning the range of the manually-derived lesion volumes.    Overall improvement in all
+three whole lesion-based measuers is seen as the
+second Stage RF model is applied for all three quantile ranges.
+The relative volume difference corresponding to the Stage 2 results tend to predict a
+decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
 
@@
 
-<!--
-One can also use these measurements as a type of sanity check.  For example, from the Stage
-1  plot, one can see that the _MeanDecreaseAccuracy_ values for the location indices in the
-anterior-posterior direction (i.e., _TemplateIndicesWarped1_) are greater than
-those for either the inferior-superior (i.e., _TemplateIndicesWarped2_) or the
-left-right (i.e., _TemplateIndicesWarped0_) directions in the space of the symmetric
-template.
--->
+# Discussion
 
-Additionally, it is interesting to note some of the other top performing features for Stage 1.
+<!-- Nick could you add some detail here on how this technique compares
+to other automated or semi automated approaches for WMH detection? -->
+
+\textcolor{blue}{
+In evaluating the segmentation results, there are limited methodologies described in the
+literature for automatic segmentation of WMHs despite the numerous clinical studies}
+[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
+However, we can extrapolate from similar application areas where such methodologies have a
+much longer history of development such as white matter lesion segmentation in
+multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
+were found during a literature search and evaluated which concern automated segmentation of white
+matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
+of two cohorts from two different sites used in
+the MS Lesion segmentation challenge associated with the international MICCAI 2008
+conference.  For comparison, this challenge training data set
+has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
+our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
+40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
+almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
+Note that our relative volume difference numbers are comparable to the
+relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
+[@vanginneken:2007aa].
+
+Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
 The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 7).  This accords with the known clinical relevance of
+evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
 FLAIR images for identifying white matter hyperintensities and the fact that such
@@
 
-<!--
-
 ![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
@@
 
--->
-
 As described earlier, for Stage 2, we used the output random forest voting maps from Stage
 1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 6, the voting maps are labeled as
+an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
 "_RFStage1VotingMaps_" where the final numeral is associated with the brain
@@
 alone it is intuitive that such features would be important.
-
-# Discussion
-
-<!-- Nick could you add some detail here on how this technique compares
-to other automated or semi automated approaches for WMH detection? -->
 
@@
 # References
-
-# Figure Captions
-
-__Figure 1:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}
-
-__Figure 2:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for asymmetry-based
-features.
-
-__Figure 3:__  Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
-application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
-The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right.  This gives a probabilistic classification of
-tissue type.  Not shown are the probability and voting images for the brain stem and
-cerebellum.
-
-__Figure 4:__  Sample FLAIR acquisition image slices showing both manual and random forest
-segmentations for both stages obtained during the leave-one-out evaluation.  Manual
-segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.
-
-__Figure 5:__ Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 1 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile
-and illustrate the
-stability of the feature importance across the leave-one-out models.
-At this initial stage only 31 feature images are used.
-
-__Figure 6:__  Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile
- and illustrate the
-stability of the feature importance across the leave-one-out models.
-We augment the 31 feature images from the first stage by adding an additional
-seven voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.
-
-__Figure 7:__ (a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.
-
-__Figure 8:__ Voxelwise comparison with manual delineation of white matter hyperintensities.
-On the left are the
-calculated Dice values over all white matter hyperintensities.  Note the improvement
-in the Dice metric from the employment of the Stage 2 component of the processing pipeline.
-(Right) Similar results can be seen by comparing the total lesion load volume between manual
-and automated detection strategies.  Although some outliers are found after the Stage 2
-processing in a couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.
-
-\clearpage

### 2016-06-30 stitched.tex
ENH:  More revisions.

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \setlength{\parindent}{0pt}
@@
 were acquired. Manual annotations were employed for both training and
-evaluation using a leave-one-out strategy. Lesion load and overlap
-evaluative comparisons are complimented by feature rankings which
-showcase the utility of the concatenated approach. Our findings suggest
-supervised learning methods may be applied to quantify WMHs on routine
-brain imaging. Paired with correlative outcome data, supervised learning
+evaluation using a leave-one-out strategy.
+\textcolor{blue}{Performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.
+Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
+predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
+relative volume difference} = 43 \(\pm\) 26\%. In addition, three lesion
+size ranges are selected to illustrate the variation in performance with
+lesion size.\} Paired with correlative outcome data, supervised learning
 methods may allow for identification of imaging features predictive of
@@
 classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery.}\textsuperscript{4}\textcolor{blue}{. Additionally,
+spectrum of possible prognoses from death to disability to recovery}\textsuperscript{4}\textcolor{blue}{. Additionally,
 in patients who are comatose following severe TBI
@@
 
+\input{trainingData.tex}
+
 Images were acquired on a Philips 3.0T Ingenia system with an 8-channel
@@
 multi-modal MR dicom image slices were converted to the nifti file format}.\footnote{\url{http://nifti.nimh.nih.gov/nifti-1}}
-\textcolor{blue}{Given the
-utility of the FLAIR sequence in detecting white matter lesions in TBI}\textsuperscript{3}\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software}\textsuperscript{22}
+\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}\textsuperscript{22}
 \textcolor{blue}{.  The normalized MRI volumes were then provided to
@@
 probability estimates) of the first RF model application,
-which we denote as ``Stage1 '', is used as input (along with the original set of
+which we denote as ``Stage1'', is used as input (along with the original set of
 feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
 Once these offline steps
-are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/wmhPipeline.png}
+\caption{\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+input MRI for a single subject, using the multi-modal symmetric template, results in
+the generation of the feature images.  These feature images are used as input to the
+Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
+voting maps, the original feature images, and the Stage 2 RF model result in the
+final voting maps which includes the WMH probability estimate.  Note that the RF models
+are constructed once from a set of training data which are processed using the
+same feature-construction pipeline as the single-subject input MRI.}}
+\end{figure}
 
@@
 results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/MMRR.png}
+\caption{Canonical views of the mutlivariate, bilaterally symmetric
+template constructed from the MMRR data set\textsuperscript{25} (only
+shown are the FLAIR, T1, and T2 modalities--- the components relevant
+for this work). Template construction is detailed
+in\textsuperscript{17}. These images are important for specific
+asymmetry-based features.}
+\end{figure}
 
@@
 
+\input{featureImagesTable.tex}
+
 Supervised methodologies are uniquely characterized, in part, by the
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/featureImages.png}
+\caption{Representation of Stage 1 feature images for subject 01C1019.
+The FLAIR, T1-, and T2-weighted images are rigidly
+pre-aligned\textsuperscript{22} to the space of the FLAIR image. The
+three modality images are then preprocessed (N4 bias
+correction\textsuperscript{28} and adaptive
+denoising\textsuperscript{29}) followed by application of standard ANTs
+brain extraction and \(n\)-tissue segmentation protocols using the MMRR
+symmetric template and corresponding priors\textsuperscript{24} applied
+to the T1 image. The feature images are then generated for voxelwise
+input to the RF model which results in the voting maps illustrated on
+the right. This gives a probabilistic classification of tissue type. Not
+shown are the probability and voting images for the brain stem and
+cerebellum.}
+\end{figure}
+
 As mentioned previously, input for each subject comprises FLAIR, T1-,
@@
 Normalization Tools (ANTs)\textsuperscript{22}. The aligned images are
-then preprocessed using the denoising algorithm of\textsuperscript{28}
-followed by N4 bias correction\textsuperscript{29} which are then
+then preprocessed using the denoising algorithm of\textsuperscript{29}
+followed by N4 bias correction\textsuperscript{28} which are then
 normalized to the intensity range \([0,1]\). Although we could have used
@@
 produced better results similar to previous work\textsuperscript{17}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/sampleResults.png}
+\caption{Sample FLAIR acquisition image slices showing both manual and
+random forest segmentations for both stages obtained during the
+leave-one-out evaluation. Manual segmentations were performed by one of
+the authors and provided the ground truth WMH labels for training the
+random forest models.}
+\end{figure}
 
@@
 automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.
-For this work, due to the relatively small-size distribution}\footnote{\textcolor{blue}{For comparison, the training data set of
-  the MS Lesion Segmentation challenge
-  associated with the international MICCAI 2008 conference has a mean lesion load of
-  204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
-  in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}}
+voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.  For this work, due to the relatively small-size distribution}
 \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
@@
 \textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.
-These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
+
+\begin{itemize}
+\item
+  \textcolor{blue}{Quantile 1: $[1-12)$,}
+\item
+  \textcolor{blue}{Quantile 2: $[12-28)$, and}
+\item
+  \textcolor{blue}{Quantile 3: $[28-551]$}.
+\end{itemize}
+
+\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
 
@@
 
-\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data.  These performance measures include sensitivity,
+\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
+of the previously described TBI data over the three lesion volume ranges.
+These performance measures include sensitivity,
 positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed to illustrate the variation
-in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity is widely varied over the cohort vs. the largest
-set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of how accurate
-the volumetric estimate is when comparing the manually identified lesions versus the
+size ranges over which these measures are computed are meant to illustrate the variation
+in performance with lesion size.}
+
+\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
+to identify which is why the sensitivity for this range is more varied compared with the largest
+set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
+entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
+the volumetric estimate when comparing the manually identified lesions versus the
 automatically predicted lesions.}
+
+\textcolor{blue}{
+We averaged these measures over all lesions and all subjects which resulted in the
+following values:}
+
+\begin{itemize}
+\item
+  \textcolor{blue}{Stage 1}
+
+  \begin{itemize}
+  \item
+    \textcolor{blue}{sensitivity} = 0.70 \(\pm\) 0.34
+  \item
+    \textcolor{blue}{PPV} = 0.42 \(\pm\) 0.36
+  \item
+    \textcolor{blue}{$F_1$} = 0.47 \(\pm\) 0.36
+  \item
+    \textcolor{blue}{relative volume difference} = 43 \(\pm\) 38\%
+  \end{itemize}
+\item
+  \textcolor{blue}{Stage 1}
+
+  \begin{itemize}
+  \item
+    \textcolor{blue}{sensitivity} = 0.68 \(\pm\) 0.38
+  \item
+    \textcolor{blue}{PPV} = 0.51 \(\pm\) 0.40
+  \item
+    \textcolor{blue}{$F_1$} = 0.52 \(\pm\) 0.36
+  \item
+    \textcolor{blue}{relative volume difference} = 43 \(\pm\) 26\%
+  \end{itemize}
+\end{itemize}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/evaluationResults.png}
+\caption{\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
+volume difference.   These quantitative assessments are given for three quantile ranges
+spanning the range of the manually-derived lesion volumes.    Overall improvement in all
+three whole lesion-based measuers is seen as the
+second Stage RF model is applied for all three quantile ranges.
+The relative volume difference corresponding to the Stage 2 results tend to predict a
+decreased predicted volume over the Stage 1 results.}}
+\end{figure}
 
@@
 
-The resulting rankings for both Stages are given in Figures 5 and 6
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/averageLeaveOneOutStage1.png}
+\caption{Average \emph{MeanDecreaseAccuracy} plots generated from the
+creation of all 24 random forest models for Stage 1 during the
+leave-one-out evaluation. These plots are useful in providing a
+quantitative assessment of the predictive importance of each feature.
+Features are ranked in descending order of importance. The horizontal
+error bars provide the \(95^{th}\) percentile and illustrate the
+stability of the feature importance across the leave-one-out models. At
+this initial stage only 31 features images are used.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/averageLeaveOneOutStage2.png}
+\caption{Average \emph{MeanDecreaseAccuracy} plots generated from the
+creation of all 24 random forest models for Stage 2 during the
+leave-one-out evaluation. These plots are useful in providing a
+quantitative assessment of the predictive importance of each feature.
+Features are ranked in descending order of importance. The horizontal
+error bars provide the \(95^{th}\) percentile and illustrate the
+stability of the feature importance across the leave-one-out models. We
+augment the 31 feature images from the first stage by adding an
+additional 7 voting maps and 7 segmentation posteriors from application
+of the Bayesian-based segmentation for a total of 45 images for the
+second stage.}
+\end{figure}
+
+The resulting rankings for both Stages are given in Figures 6 and 7
 where the values for the separate stages are averaged over the entire
@@
 
-Additionally, it is interesting to note some of the other top performing
-features for Stage 1. The contralateral difference FLAIR image is highly
-discriminative over the set of evaluation random forest models (see
-Figure 7). This accords with the known clinical relevance of FLAIR
-images for identifying white matter hyperintensities and the fact that
-such pathology does not typically manifest symmetrically in both
-hemispheres. Interestingly, the posterior maps for the deep gray matter
-are extremely important for accurate white matter hyperintensity
-segmentation. Perhaps the spatial specification of deep gray matter aids
-in the removal of false positives. Inspection of the bottom of the plots
-demonstrates the lack of discriminating features associated with the T1
-image which is also well-known in the clinical literature.
+\section{Discussion}\label{discussion}
+
+\textcolor{blue}{
+In evaluating the segmentation results, there are limited methodologies described in the
+literature for automatic segmentation of WMHs despite the numerous clinical studies}\textsuperscript{39,40}
+\textcolor{blue}{exploring the connections between WMHs and TBI.
+However, we can extrapolate from similar application areas where such methodologies have a
+much longer history of development such as white matter lesion segmentation in
+multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
+were found during a literature search and evaluated which concern automated segmentation of white
+matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
+of two cohorts from two different sites used in
+the MS Lesion segmentation challenge associated with the international MICCAI 2008
+conference.  For comparison, this challenge training data set
+has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
+our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
+40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
+almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
+Note that our relative volume difference numbers are comparable to the
+relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}\textsuperscript{41}.
+
+Regarding the feature rankings, it is interesting to note some of the
+other top performing features for Stage 1. The contralateral difference
+FLAIR image is highly discriminative over the set of evaluation random
+forest models (see Figure 8). This accords with the known clinical
+relevance of FLAIR images for identifying white matter hyperintensities
+and the fact that such pathology does not typically manifest
+symmetrically in both hemispheres. Interestingly, the posterior maps for
+the deep gray matter are extremely important for accurate white matter
+hyperintensity segmentation. Perhaps the spatial specification of deep
+gray matter aids in the removal of false positives. Inspection of the
+bottom of the plots demonstrates the lack of discriminating features
+associated with the T1 image which is also well-known in the clinical
+literature.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figures/FLAIRcontralaleteralWithLesionsAlternate.png}
+\caption{(a) FLAIR image slice illustrating WMHs which have been
+manually delineated. The region around the WMHs is enlarged (b) in the
+original FLAIR and the (c) contralateral FLAIR difference image.}
+\end{figure}
 
@@
 input to a Bayesian-based segmentation with an additional MRF spatial
-prior. In Figure 6, the voting maps are labeled as
+prior. In Figure 7, the voting maps are labeled as
 ``\emph{RFStage1VotingMaps}'' where the final numeral is associated with
@@
 
-\section{Discussion}\label{discussion}
+\section{Conclusions}\label{conclusions}
 
@@
 
-\section{References}\label{references}
-
-\section{Figure Captions}\label{figure-captions}
-
-\textbf{Figure 1:}
-\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}
-
-\textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
-symmetric template constructed from the MMRR data
-set\textsuperscript{25} (only shown are the FLAIR, T1, and T2
-modalities--- the components relevant for this work). Template
-construction is detailed in\textsuperscript{17}. These images are
-important for asymmetry-based features.
-
-\textbf{Figure 3:} Representation of Stage 1 feature images for subject
-01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
-pre-aligned\textsuperscript{22} to the space of the T1 image. The three
-modality images are then preprocessed (N4 bias
-correction\textsuperscript{29} and adaptive
-denoising\textsuperscript{28}) followed by application of standard ANTs
-brain extraction and \(n\)-tissue segmentation protocols using the MMRR
-symmetric template and corresponding priors\textsuperscript{24} applied
-to the T1 image. The feature images are then generated for voxelwise
-input to the RF model which results in the voting maps illustrated on
-the right. This gives a probabilistic classification of tissue type. Not
-shown are the probability and voting images for the brain stem and
-cerebellum.
-
-\textbf{Figure 4:} Sample FLAIR acquisition image slices showing both
-manual and random forest segmentations for both stages obtained during
-the leave-one-out evaluation. Manual segmentations were performed by one
-of the authors and provided the ground truth WMH labels for training the
-random forest models.
-
-\textbf{Figure 5:} Average \emph{MeanDecreaseAccuracy} plots generated
-from the creation of all 24 random forest models for Stage 1 during the
-leave-one-out evaluation. These plots are useful in providing a
-quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance. The horizontal
-error bars provide the \(95^{th}\) percentile and illustrate the
-stability of the feature importance across the leave-one-out models. At
-this initial stage only 31 feature images are used.
-
-\textbf{Figure 6:} Average \emph{MeanDecreaseAccuracy} plots generated
-from the creation of all 24 random forest models for Stage 2 during the
-leave-one-out evaluation. These plots are useful in providing a
-quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance. The horizontal
-error bars provide the \(95^{th}\) percentile and illustrate the
-stability of the feature importance across the leave-one-out models. We
-augment the 31 feature images from the first stage by adding an
-additional seven voting maps and 7 segmentation posteriors from
-application of the Bayesian-based segmentation for a total of 45 images
-for the second stage.
-
-\textbf{Figure 7:} (a) FLAIR image slice illustrating WMHs which have
-been manually delineated. The region around the WMHs is enlarged (b) in
-the original FLAIR and the (c) contralateral FLAIR difference image.
-
-\textbf{Figure 8:} Voxelwise comparison with manual delineation of white
-matter hyperintensities. On the left are the calculated Dice values over
-all white matter hyperintensities. Note the improvement in the Dice
-metric from the employment of the Stage 2 component of the processing
-pipeline. (Right) Similar results can be seen by comparing the total
-lesion load volume between manual and automated detection strategies.
-Although some outliers are found after the Stage 2 processing in a
-couple subjects, the number of outliers caused by false positives is
-decreased significantly with the second stage processing.
-
-\clearpage
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
@@
 
+\hypertarget{ref-Tustison:2010ac}{}
+28. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
+JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
+2010;29(6):1310--20.
+
 \hypertarget{ref-Manjon:2010aa}{}
-28. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
+29. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
 non-local means denoising of mR images with spatially varying noise
 levels. J Magn Reson Imaging. 2010;31(1):192--203.
-
-\hypertarget{ref-Tustison:2010ac}{}
-29. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
-JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
-2010;29(6):1310--20.
 
@@
 
+\hypertarget{ref-vanginneken:2007aa}{}
+41. Ginneken BV, Heimann T, Styner M. 3D segmentation in the clinic: A
+grand challenge. In: 3D segmentation in the clinic: A grand challenge.
+2007. pp. 7--15.
+
 \end{document}

### 2016-06-30 Evaluation/evaluation2.R

@@ -112,4 +112,21 @@ ggsave( filename = paste( 'Evaluation/', whichEval, '/RelativeVolumeDifference.p
 
 
 
+for( stage in 1:2 )
+  {
+  meanF1 <- mean( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "F1" & boxPlotDataFrame$Stage == stage )], na.rm = TRUE )
+  meanSensitivity <- mean( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "Sensitivity" & boxPlotDataFrame$Stage == stage )], na.rm = TRUE )
+  meanPPV <- mean( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "PPV" & boxPlotDataFrame$Stage == stage )], na.rm = TRUE )
+  meanRVD <- mean( abs( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "RelativeVolumeDifference" & boxPlotDataFrame$Stage == stage )] ), na.rm = TRUE )
 
+  cat( "Stage ", stage, ": mean F1 = ", meanF1, ", mean Sens. = ", meanSensitivity, ", mean PPV = ", meanPPV, ", mean RVD = ", meanRVD, "\n", sep = '' )
+
+  sdF1 <- sd( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "F1" & boxPlotDataFrame$Stage == stage )], na.rm = TRUE )
+  sdSensitivity <- sd( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "Sensitivity" & boxPlotDataFrame$Stage == stage )], na.rm = TRUE )
+  sdPPV <- sd( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "PPV" & boxPlotDataFrame$Stage == stage )], na.rm = TRUE )
+  sdRVD <- sd( abs( boxPlotDataFrame$Value[which( boxPlotDataFrame$MeasureType == "RelativeVolumeDifference" & boxPlotDataFrame$Stage == stage )] ), na.rm = TRUE )
+
+  cat( "Stage ", stage, ": sd F1 = ", sdF1, ", sd Sens. = ", sdSensitivity, ", sd PPV = ", sdPPV, ", sd RVD = ", sdRVD, "\n", sep = '' )
+
+
+  }

### 2016-06-30 README.R

@@ -12,8 +12,7 @@ rmdFiles <- c( formatFile,
                "methods.Rmd",
                "results.Rmd",
                "discussion.Rmd",
-               "acknowledgementsDisclaimers.Rmd",
-               "figureCaptions.Rmd"
+               "acknowledgementsDisclaimers.Rmd"
    )
 
 for( i in 1:length( rmdFiles ) )
@@ -32,7 +31,8 @@ render( stitchedFile, output_format = "all" )
 
 
 singleRmdFiles <- c(
-               "responseToReviewers.Rmd"
+               "responseToReviewers.Rmd",
+               "figureCaptions.Rmd"
    )
 
 for( i in 1:length( singleRmdFiles ) )

### 2016-06-30 responseToReviewers.Rmd
ENH:  More revisions.

@@
 ``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
-through each decision tree resulting in a single classification ``vote'' per tree.``}_
+through each decision tree resulting in a single classification ``vote'' per tree.''}_
 

### 2016-06-30 responseToReviewers.tex
ENH:  More revisions.

@@
 ``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
-through each decision tree resulting in a single classification ``vote'' per tree.``}}
+through each decision tree resulting in a single classification ``vote'' per tree.''}}
 \end{quote}

### 2016-06-30 abstract.Rmd
ENH:  Submission files.

--- 2016-06-27 abstract.Rmd
@@
 employed for both training and evaluation using a leave-one-out
-strategy.  Lesion load and overlap evaluative comparisons are
-complimented by feature rankings which showcase the utility of the
-concatenated approach.  Our findings suggest supervised learning methods
-may be applied to quantify WMHs on routine brain imaging. Paired with
+strategy.  Performance measures include sensitivity,
+positive predictive value, $F_1$ score, and relative volume difference.
+Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
+predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
+relative volume difference} = 43 $\pm$ 26\%.  In addition, three
+lesion size ranges are selected to illustrate the variation in
+performance with lesion size.  Paired with
 correlative outcome data, supervised learning methods may allow for

### 2016-06-30 discussion.Rmd
ENH:  Submission files.

@@
 
-\textcolor{blue}{
 In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
+literature for automatic segmentation of WMHs despite the numerous clinical studies
+[@Debette:2010aa;@Kloppenborg:2014aa] exploring the connections between WMHs and TBI.
 However, we can extrapolate from similar application areas where such methodologies have a
 much longer history of development such as white matter lesion segmentation in
-multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
+multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
 were found during a literature search and evaluated which concern automated segmentation of white
@@
 conference.  For comparison, this challenge training data set
-has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
-our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
-40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
+has a mean lesion load of 204 ($\pm$ 752) $mm^3$ per lesion (compared with
+our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 $mm^3$ per voxel =
+40.56 ($\pm$ 66.48) $mm^3$ per lesion) and the resolution is
 almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
 Note that our relative volume difference numbers are comparable to the
-relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
+relative volume difference _between raters_ of 68 percent (cf Figure 5 (d))
 [@vanginneken:2007aa].
@@
 
+<!--
+
 ![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
@@
 (c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
+
+-->
 

### 2016-06-30 discussionX.Rmd
ENH:  Submission files.

--- 2016-06-30 discussion.Rmd
@@
 
+\textcolor{blue}{
 In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies
-[@Debette:2010aa;@Kloppenborg:2014aa] exploring the connections between WMHs and TBI.
+literature for automatic segmentation of WMHs despite the numerous clinical studies}
+[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
 However, we can extrapolate from similar application areas where such methodologies have a
@@
 Note that our relative volume difference numbers are comparable to the
-relative volume difference _between raters_ of 68 percent (cf Figure 5 (d))
+relative volume difference between raters of 68 percent (cf Figure 5 (d))}
 [@vanginneken:2007aa].
@@
 (c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
--->
 

### 2016-06-30 figureCaptions.Rmd
ENH:  Submission files.

@@
 
-__Figure 1:__ \textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+__Figure 1:__ Workflow illustration for the proposed pipeline.  Processing of the multi-modal
 input MRI for a single subject, using the multi-modal symmetric template, results in
@@
 are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}
+same feature-construction pipeline as the single-subject input MRI.
 
 __Figure 2:__  Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for asymmetry-based
+from the MMRR data set (only shown are the FLAIR, T1, and T2 modalities---
+the components relevant for this work).  These images are important for asymmetry-based
 features.
@@
 FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
+ to the space of the T1 image.  The three modality images are then preprocessed
+(N4 bias correction and adaptive denoising) followed by
 application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
+the MMRR symmetric template and corresponding priors applied to the T1 image.
 The feature images are then generated for voxelwise input to the RF model which results in
@@
 
-__Figure 5:__ \textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+__Figure 5:__ Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
 section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
@@
 The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}
+decreased predicted volume over the Stage 1 results.
 

### 2016-06-30 figureCaptions.tex
ENH:  Submission files.

@@
 
-\textbf{Figure 1:}
-\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}
+\textbf{Figure 1:} Workflow illustration for the proposed pipeline.
+Processing of the multi-modal input MRI for a single subject, using the
+multi-modal symmetric template, results in the generation of the feature
+images. These feature images are used as input to the Stage 1 RF model
+producing the initial RF probability map estimates. The Stage 1 voting
+maps, the original feature images, and the Stage 2 RF model result in
+the final voting maps which includes the WMH probability estimate. Note
+that the RF models are constructed once from a set of training data
+which are processed using the same feature-construction pipeline as the
+single-subject input MRI.
 
 \textbf{Figure 2:} Canonical views of the mutlivariate, bilaterally
-symmetric template constructed from the MMRR data set\textsuperscript{1}
-(only shown are the FLAIR, T1, and T2 modalities--- the components
-relevant for this work). Template construction is detailed
-in\textsuperscript{2}. These images are important for asymmetry-based
-features.
+symmetric template constructed from the MMRR data set (only shown are
+the FLAIR, T1, and T2 modalities--- the components relevant for this
+work). These images are important for asymmetry-based features.
 
 \textbf{Figure 3:} Representation of Stage 1 feature images for subject
-01C1019. The FLAIR, T1-, and T2-weighted images are rigidly
-pre-aligned\textsuperscript{3} to the space of the T1 image. The three
-modality images are then preprocessed (N4 bias
-correction\textsuperscript{4} and adaptive denoising\textsuperscript{5})
-followed by application of standard ANTs brain extraction and
-\(n\)-tissue segmentation protocols using the MMRR symmetric template
-and corresponding priors\textsuperscript{6} applied to the T1 image. The
-feature images are then generated for voxelwise input to the RF model
-which results in the voting maps illustrated on the right. This gives a
-probabilistic classification of tissue type. Not shown are the
-probability and voting images for the brain stem and cerebellum.
+01C1019. The FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
+to the space of the T1 image. The three modality images are then
+preprocessed (N4 bias correction and adaptive denoising) followed by
+application of standard ANTs brain extraction and \(n\)-tissue
+segmentation protocols using the MMRR symmetric template and
+corresponding priors applied to the T1 image. The feature images are
+then generated for voxelwise input to the RF model which results in the
+voting maps illustrated on the right. This gives a probabilistic
+classification of tissue type. Not shown are the probability and voting
+images for the brain stem and cerebellum.
 
@@
 
-\textbf{Figure 5:}
-\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
-section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
-volume difference.   These quantitative assessments are given for three quantile ranges
-spanning the range of the manually-derived lesion volumes.    Overall improvement in all
-three whole lesion-based measuers is seen as the
-second Stage RF model is applied for all three quantile ranges.
-The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}
+\textbf{Figure 5:} Evaluation measures for both Stages of the
+leave-one-out protocol of the described protocol in the Methods section:
+(a) sensitivity, (b) positive predictive value, (c) \(F_1\) score, and
+(d) relative volume difference. These quantitative assessments are given
+for three quantile ranges spanning the range of the manually-derived
+lesion volumes. Overall improvement in all three whole lesion-based
+measuers is seen as the second Stage RF model is applied for all three
+quantile ranges. The relative volume difference corresponding to the
+Stage 2 results tend to predict a decreased predicted volume over the
+Stage 1 results.
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-landman2011}{}
-1. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
-Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
-reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
-
-\hypertarget{ref-Tustison:2015aa}{}
-2. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
-Grossman MC, Avants BB. Optimal symmetric multimodal templates and
-concatenated random forests for supervised brain tumor segmentation
-(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.
-
-\hypertarget{ref-Avants:2014aa}{}
-3. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
-ToolKit image registration framework. Front Neuroinform. 2014;8:44.
-
-\hypertarget{ref-Tustison:2010ac}{}
-4. Tustison NJ, Avants BB, Cook PA, Zheng Y, Egan A, Yushkevich PA, Gee
-JC. N4ITK: Improved N3 bias correction. IEEE Trans Med Imaging.
-2010;29(6):1310--20.
-
-\hypertarget{ref-Manjon:2010aa}{}
-5. Manjón JV, Coupé P, Martí-Bonmatí L, Collins DL, Robles M. Adaptive
-non-local means denoising of mR images with spatially varying noise
-levels. J Magn Reson Imaging. 2010;31(1):192--203.
-
-\hypertarget{ref-Tustison:2014ab}{}
-6. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
-Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
-and freeSurfer cortical thickness measurements. Neuroimage.
-2014;99:166--79.
 

### 2016-06-30 format.Rmd
ENH:  Submission files.

@@
    - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{2}
+   - \linespread{1.5}
    - \usepackage[compact]{titlesec}

### 2016-06-30 formatX.Rmd
ENH:  Submission files.

--- 2016-06-30 stitched.Rmd
@@
 \pagenumbering{gobble}
-
-# Abstract
-
-White matter hyperintensities (WMHs) are foci of abnormal signal
-intensity in white matter regions seen with magnetic resonance imaging
-(MRI). WMHs are associated with normal aging and have shown prognostic
-value in neurological conditions such as traumatic brain injury (TBI).
-The impracticality of manually quantifying these lesions limits their
-clinical utility and motivates the utilization of machine learning
-techniques for automated segmentation workflows.  Herein, we develop a
-concatenated random forest framework with image features for
-segmenting WMHs in a TBI cohort. The framework is built upon the
-Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
-(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
-veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
-(CENC) observational study were acquired. Manual annotations were
-employed for both training and evaluation using a leave-one-out
-strategy.  \textcolor{blue}{Performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.
-Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
-predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
-relative volume difference} = 43 $\pm$ 26\%.  In addition, three
-lesion size ranges are selected to illustrate the variation in
-performance with lesion size.}  Paired with
-correlative outcome data, supervised learning methods may allow for
-identification of imaging features predictive of diagnosis and prognosis
-in individual TBI patients.
-
-\clearpage
-
-# Introduction
-
-## White matter hyperintensities in TBI
-
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
-
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
-Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
-in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].
-
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
-Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
-In this work, we present an automated framework for quantification of WMHs in
-multi-modal MRI using the random forest machine learning technique.}
-
-## Random forests for WMH segmentation
-
-<!--
-Machine learning and pattern recognition techniques have seen increased application
-for various medical image analysis workflows (see, for example, the annual
-Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
-Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as supervised image segmentation
-(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
-Facilitating the current employment of such techniques are the number of available imaging
-data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
-SciPy [@scipy] and the R project for statistical computing [@R] and their associated
-extensions.
--->
-
-The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
-normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., [@viola2005]).
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
-delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
-
-<!--
-Of note, these
-latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
--->
-
-Random forests are conceptually straightforward [@breiman2001].
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
-For the proposed application, individual
-voxels (and their corresponding feature values) are introduced at the root of a particular
-decision tree and traverse the edges and internal nodes ultimately ending up at one of
-the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
-\textcolor{blue}{Finally, Breiman} [@breiman2001]
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
-
-In this work, we develop a concatenated random forest framework with a
-feature image set (both spatial and intensity-based)
-for segmenting WMHs in a large TBI cohort.
-The entire framework is built on the well-known open-source
-Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
-Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.
-
-[^i1]: https://github.com/stnava/ANTs
-[^i2]: https://github.com/stnava/ANTsR
-
-# Materials and Methods
-
-## Imaging
-
-MR images utilized for this initial report were acquired from a single scanner involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (see Walker et al., this issue).
-Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
-Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
-
-\input{trainingData.tex}
-
-Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
-3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
-
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
-subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All
-multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
-The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
-\textcolor{blue}{
-which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
-
-[^1]: http://nifti.nimh.nih.gov/nifti-1
-
-## Quantitative analysis
-
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1'', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
-
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
-
-### Symmetric multi-modal templates
-
-\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}
-[@landman2011]\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
-
-![Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific asymmetry-based
-features.](Figures/MMRR.png)
-
-### Feature images for WMH segmentation
-
-\textcolor{blue}{
-Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}  For the targeted application in this work,
-tissue classification is performed at the voxelwise level.  In other words, each voxel
-within the region of interest is sent through the ensemble of decision trees and receives
-a set of classification votes from each tree thus permitting a regression or classification
-solution.  Since this procedure is performed at the voxelwise level, intensity information
-alone is insufficient for good segmentation performance due to the lack of spatial context.
-For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated lesion detection
-algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
-even though machine learning and pattern recognition techniques are extremely powerful and
-have significant potential, just as crucial to outcome is the creative construction and
-deployment of salient feature images which we detail below.
-
-\input{featureImagesTable.tex}
-
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.  In Table 2, we provide a list and basic
-categorization of the feature images used for the initial (i.e., Stage 1---more on the use
-of multiple random forest stages below) segmentation of the WMHs.
-In addition Figure 3 provides a representation of a set of feature images for a single
-subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
-with seven labels:
-
-* cerebrospinal fluid (label 1),
-* gray matter (label 2),
-* white matter (label 3),
-* deep gray matter (label 4),
-* brain stem (label 5),
-* cerebellum (label 6), and
-* white matter hyperintensities (label 7).
-
-![Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
-application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
-The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right.  This gives a probabilistic classification of
-tissue type.  Not shown are the probability and voting images for the brain stem and
-cerebellum.](Figures/featureImages.png)
-
-As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
-acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
-the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
-are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
-N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
-range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].
-
-![Sample FLAIR acquisition image slices showing both manual and random forest
-segmentations for both stages obtained during the leave-one-out evaluation.  Manual
-segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.](Figures/sampleResults.png)
-
-The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
-  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
-CSF, gray matter, white matter, deep gray matter, brain stem, and
-cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
-first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 2).
-
-Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
-We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a Manhattan radius of one voxel given the typical size
-of individual WMHs. For each of the preprocessed
-images, we calculate the difference in intensities with the corresponding warped template
-component.  Previous success in the international brain tumor segmentation
-competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 2.  We employ the
-same strategy here.
-
-<!--
-For example, the template
-difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
-
-$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$
-
-where
-
-$$\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$$
-
-is the transform
-which maps from the individual subject space to the
-template space and $T_{FLAIR}$ is the FLAIR template component.
--->
-
-To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
-we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature.
-
-<!--
-For the FLAIR component, this contralateral
-difference image is calculated from
-
-$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$
-
-where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
-the symmetric template.
--->
-
-The segmentation probability images described above are used as feature images to
-provide a spatial context for the random forest model prediction step.  Additional spatial contextual
-feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
-deep gray matter images.  These latter images are intended to help distinguish white matter
-hyperintensities from false positives induced by the partial voluming at the gray/white
-matter interface.  A third set of images are based on the voxel location within the
-space of the template.
-\textcolor{blue}{Similar feature images were used in}
-[@Anbeek:2004aa]
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
-
-### Stacked (concatenated) random forests for improved segmentation performance
-
-In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
-concatenated supervised approach, whereby the prediction output from the first random forest
-model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
-feature images of the training data (as described previously) are used to construct the
-Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
-are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
-additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
-all three aligned preprocessed images for multivariate segmentation during the
-second stage.}
-The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
-
-### Implementation
-
-As pointed out in a recent comprehensive lesion segmentation review
-[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
-is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
-our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
-this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
-ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
-use the _randomForest_ package [@liaw2002] using the default settings with
-2000 trees per model and 500 randomly selected samples per label per image.
-Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.
-
-In addition, similar to our previous offering [^2], we plan on creating
-a self-encapsulated example to showcase the proposed methodology
-\textcolor{blue}{which will also be available on github.}[^3]  The fact that
-the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
-with the manual labelings will facilitate reproducibility on the part of the reader
-as well as any interest in extending the proposed framework to other data sets.
-
-[^2]: https://github.com/ntustison/ANTsAndArboles
-
-[^3]: https://github.com/ntustison/WatchMeHyperventilate
-
-### Evaluation protocol overview
-
-In order to evaluate the protocol described, we performed a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.
-Initial processing included the creation of all Stage 1 feature images for all subjects.
-The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
-The ``truth'' labels are the seven anatomical regions given above.
-
-The leave-one-out procedure is as follows:
-
-* Create Stage 1 feature images for all 24 subjects.
-* For each of the 24 subjects:
-    + sequester the current subject and corresponding feature images.
-    + construct the Stage 1 random forest model from the remaining 23 subjects.
-    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
-    + the previous step produces the Stage 1 voting maps for all seven labels.
-    + for each of the 23 subjects, perform a Bayesian-based segmentation with an
-      MRF spatial
-      prior using the seven voting maps as additional tissue priors.
-    + construct the Stage 2 random forest model from all the Stage 1 feature images,
-      seven voting maps, and seven posterior probability maps from the previous step.
-    + send the sequestered subject through the random forest models for both stages.
-    + compare the final results with the manually-defined white matter hyperintensity
-      regions.
-
-\textcolor{blue}{Several measures have been employed in the literature for evaluating
-automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
-For this work, due to the relatively small-size distribution} \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
-four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
-incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
-The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
-
-$$ sensitivity =  \frac{ TP }{TP + FN}, $$
-
-\textcolor{blue}{and the positive predictive value (PPV),}
-
-$$ PPV =  \frac{ TP }{TP + FP}, $$
-
-\textcolor{blue}{such that $F_1$ is given by}
-
-$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
-
-\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
-the manual and predicted lesion volumes:}
-
-$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
-
-\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
-
-* \textcolor{blue}{Quantile 1: $[1-12)$,}
-
-* \textcolor{blue}{Quantile 2: $[12-28)$, and}
-
-* \textcolor{blue}{Quantile 3: $[28-551]$}.
-
-\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
-
-# Results
-
-## White matter hyperintensity segmentation evaluation
-
-\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data over the three lesion volume ranges.
-These performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed are meant to illustrate the variation
-in performance with lesion size.}
-
-\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity for this range is more varied compared with the largest
-set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
-the volumetric estimate when comparing the manually identified lesions versus the
-automatically predicted lesions.}
-
-\textcolor{blue}{
-We averaged these measures over all lesions and all subjects which resulted in the
-following values:}
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
-
-    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
-
-    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
-
-    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
-
-    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
-
-<!--
-
-In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-On the left of Figure 5 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
--->
-
-![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
-section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
-volume difference.   These quantitative assessments are given for three quantile ranges
-spanning the range of the manually-derived lesion volumes.    Overall improvement in all
-three whole lesion-based measuers is seen as the
-second Stage RF model is applied for all three quantile ranges.
-The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
-
-## Ranking feature importance
-
-After performing the leave-one-out evaluation,
-we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
-2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
-during the out-of-bag phase of the random forest model construction and quantifies the decrease
-in prediction accuracy from omitting the specified feature.  In other words, this quantity
-helps determine the importance of a particular feature and, although we save such
-efforts for future work, this information provides us with guidance for future feature
-pruning and/or additions.
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 1 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile
- and illustrate the
-stability of the feature importance across the leave-one-out models.
-At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile and illustrate the
-stability of the feature importance across the leave-one-out models.
-We augment the 31 feature images from the first stage by adding an additional
-7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.](Figures/averageLeaveOneOutStage2.png)
-
-The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the
-separate stages are averaged over the entire corresponding model set.  In addition, we
-track the variance for each feature over all models to illustrate the stability of
-the chosen features during the evaluation.  This latter information is illustrated as
-horizontal errors bars providing the $95^{th}$ percentile
-Note that the reader can cross reference Table 1 for identifying corresponding feature types
-and names.
-
-# Discussion
-
-<!-- Nick could you add some detail here on how this technique compares
-to other automated or semi automated approaches for WMH detection? -->
-
-\textcolor{blue}{
-In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
-However, we can extrapolate from similar application areas where such methodologies have a
-much longer history of development such as white matter lesion segmentation in
-multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
-matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
-of two cohorts from two different sites used in
-the MS Lesion segmentation challenge associated with the international MICCAI 2008
-conference.  For comparison, this challenge training data set
-has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
-our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
-40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
-almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
-Note that our relative volume difference numbers are comparable to the
-relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
-[@vanginneken:2007aa].
-
-Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
-The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
-FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
-posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives.  Inspection of the bottom of the
-plots demonstrates the lack of discriminating features associated with the
-T1 image which is also well-known in the clinical literature.
-
-![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
-As described earlier, for Stage 2, we used the output random forest voting maps from Stage
-1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
-"_RFStage1VotingMaps_" where the final numeral is associated with the brain
-parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
-The Stage 2 feature importance plot follows similar
-trends as that for Stage 1 with the T1 images not contributing much to the identification
-of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
-important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
-3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
-alone it is intuitive that such features would be important.
-
-<!--
-
-* Evaluation * Proper context * We only have manual labelings from a
-single radiologist ( * TBI white matter hyperintensities are more
-difficult to segment than MS lesions * individual lesions are smaller *
-lesion load is typically smaller * not prone to enhancement * less
-distinct from surrounding tissue ? * Despite the doubts regarding the
-efficacy of a typical segmentation validation, our contribution is
-useful because * the total segmentation framework is publicly available
-within the ANTs/ANTsR framework * although only a single site is
-analyzed, the feature images are site-agnostic * we plan to apply the
-current RF models directly to the other site data * we can then build
-the training database further by enlisting other experts * Is this the
-first work exploring segmentation of white matter hyperintensities in
-TBI? * The data (including manual tracings) will be made available as
-part of the FITBIR effort.
-
--->
-
-# Conclusions
-
-The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation
-of WMHs from FLAIR images to help build two-stage ensembles of decision
-trees for the automated identification of these lesions. Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology.  Also, we
-recognize that only a single site was used for evaluating the proposed
-framework.  However, we are currently processing other site data with
-the models developed for this work and the results look promising since
-the developed features are site-agnostic.
-
-As far as we know, this is the first report utilizing a novel random
-forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
-tend to be more difficult to segment than MS lesions as the former tend
-to be smaller with an overall smaller lesion load.  Also, enhancement
-protocols with the former tend to be less successful than with the
-latter.  As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.
-
-<!--
-Our
-framework is also available as open-source as part of well-known
-neuroimaging tools which easily allows for additions/extensions but is
-also, as far as we know, the first random forest-based technique
-available for such application.
--->
-
-Two major meta-analyses of WMHs have been published covering the periods
-prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
-Debette & Markus [@Debette:2010aa] found that the presence of
-WMHs was related to subsequent cognitive decline, a higher risk of
-developing dementia, stroke, and of mortality. Lesion volume at baseline
-was also predictive of cognitive decline. Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
-concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
-that WMHs were associated with cognitive deficit (effect size of -0.10,
-95% CI: -0.13 to -0.08) after controlling for age.
-
-Despite the potential clinical significance of WMHs
-these lesions receive little
-attention in current clinical workflows. When reported in a standard
-neuroradiologist interpretation, they are typically handled as
-incidental findings and are assigned little clinical significance. This
-likely reflects the impracticality of performing a detailed assessment
-of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
-supervised statistical learning tool that is appropriate for the
-application to such large-scale datasets.
-
-<!--
-
-The currently described tool is just one example of how "supervised
-learning" algorithms might be applied to aid in the diagnosis of TBI and
-other disease processes through the specific identification of features
-predictive of a given disease state. It is an important demonstration of
-the potential power of these analytical approaches in the rapid but
-comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide
-variety of settings for the rapid identification of predictive imaging
-features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
-manufacturers utilize these types of approaches to equip self-driving
-vehicles to recognize and respond to unique external surroundings
-through the identification of visual information sufficiently similar to
-previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning
-approaches may allow for the rapid identification of information
-predictive of disease state in an individual patient. These approaches
-have been applied to the segmentation of macroscopically visible
-structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of
-imaging data in the individual patient with a primary quantitative
-output metrics to include sequences such as diffusion tensor imaging
-(DTI) and its variants, functional connectivity, perfusion weighted
-imaging, and cortical thickness assessments. At present, these advanced
-neuroimaging sequences are confined to cohort-based research studies due
-to the lack of available analytical tools to assess the information in
-the setting of the individual patient [@Mayer:2014aa]. Application of
-deep learning approaches in the context of data with primary
-quantitative outputs will require large scale normative and disease
-specific databases. Building these large scale imaging libraries is
-resource intensive and requires a multi-center approach with harmonized
-scanners between sites and correlative non-imaging clinical data. Large
-scale TBI data is becoming increasingly available through activities
-such as the Chronic Effects of Neurotrauma Consortium (CENC),
-Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
-Collaborative European Neurotrauma Effectiveness Research in TBI
-(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
-Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data,
-deep learning algorithms may be well positioned to help transform how
-neuroimaging is interpreted for the clinical management of patients with
-this disease process.
-
--->
-
-\clearpage
-
-## Acknowledgements
-
-The authors wish to acknowledge all other members of
-the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
-X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
-also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
-and Kevin Wilson. We would also like to thank the study participants and
-their family members. This project was jointly supported by the
-Department of Defense (W81XWH-13-2-0095), the U.S. Department of
-Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
-Grant HU 0001-08-0001.
-
-## Declaration of Interest/Disclaimer
-
-The authors report no financial
-disclosures or conflicts of interest. The views expressed here are those
-of the authors and do not necessarily reflect the official policy of
-position of the Department of the Navy, Department of Defense, nor the
-U.S. Government. This work was prepared as a part of official duties;
-Title 17 USC §105 provides that Copyright protection under this title is
-not available for any work of the U.S. Government. Title 17 USC §101
-defines a US Government work as a work prepared by a military service
-member of employee of the US Government as part of that person’s
-official duties.
-
-\clearpage
-
-<!-- \listofchanges
-
-\clearpage -->
-
-# References

### 2016-06-30 intro.Rmd
ENH:  Submission files.

@@
 
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
+White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
+WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] populations.
 
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
+Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa].
+Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa].
 Further, volume of FLAIR lesions within the corpus
@@
 with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
+classify "outcome" in TBI patients where "outcome" refers to the
+spectrum of possible prognoses from death to disability to recovery [@Moen:2014aa]. Additionally,
 in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].
+the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].
 
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
+Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
 Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
@@
 In this work, we present an automated framework for quantification of WMHs in
-multi-modal MRI using the random forest machine learning technique.}
+multi-modal MRI using the random forest machine learning technique.
 
@@
 (e.g., [@viola2005]).
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
-delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
+Random forest-based paradigms have been successfully employed in the
+delineation of other neuropathologies [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
+for both single and multi-modal acquisition protocols.
 
@@
 Random forests are conceptually straightforward [@breiman2001].
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
+The basic component of
+the random forest paradigm is the "decision tree" often represented by a flowchart or
+graph where internal nodes represent "tests", or decisions, and the edges represent the outcome of those tests.
 The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
@@
 the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
-\textcolor{blue}{Finally, Breiman} [@breiman2001]
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
+forest model will consist of many such trees (often refered to as an "ensemble").
+Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
+(e.g., AdaBoost) [@schapire1990;@freund1997]
+influenced the similarly sytled conglomeration of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
+Finally, Breiman [@breiman2001]
+improved accuracy by random sampling of training data (i.e., "bagging") resulting
+in the current random forest technique applied here.
+As voxels and their feature values are "pushed" through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
+each voxel location.
 

### 2016-06-30 methods.Rmd
ENH:  Submission files.

--- 2016-06-27 methods.Rmd
@@
 Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
+The feature images were derived from MR acquisitions of 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
 Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
+All lesions were isolated in the white matter of the cerebrum.
 Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
+load across the selected cohort.
 
@@
 3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
+3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
 In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
 
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
+The first author (J. R. S.) performed the manual WMH tracings for all 26
 subjects.
@@
 All
-multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{Given the
-utility of the FLAIR sequence in detecting white matter lesions in TBI}  [@Marquez-de-la-Plata:2007aa]\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
-The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
-\textcolor{blue}{
+multi-modal MR dicom image slices were converted to the nifti file format [^1].
+All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software [@Avants:2014aa].  The normalized MRI volumes were then provided to
+J. R. S. who traced each lesion using the ITK-SNAP tool [@Yushkevich:2006aa]
 which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
+canonical views.
 
@@
 
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
+Figure 1 provides a graphical overview of the proposed workflow.  The major components
 include offline generation of symmetric multimodal templates, the creation of
@@
 probability estimates) of the first RF model application,
-which we denote as ``Stage1 '', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
+which we denote as "Stage1", is used as input (along with the original set of
+feature images) to a refinement RF model segmentation which we denote as "Stage 2".
 Once these offline steps
-are performed, a new, unsegmented subject can then be processed using the proposed pipeline.}
+are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.
 
@@
 
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
+![Workflow illustration for the proposed pipeline.  Processing of the multi-modal
 input MRI for a single subject, using the multi-modal symmetric template, results in
@@
 are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
+same feature-construction pipeline as the single-subject input MRI.](Figures/wmhPipeline.png)
 
@@
 
-\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
+Following [@Tustison:2014ab] and [@Tustison:2015aa], optimally derived templates
 serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}
-[@landman2011]\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
+For this work we use the multi-modal data available from the public MMRR data set
+[@landman2011].  We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in [@Avants:2010aa] which
+results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.
 
 <!--
+
 ![Canonical views of the mutlivariate, bilaterally symmetric template constructed
@@
 
-\textcolor{blue}{
 Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
+"features" as input (i.e., feature images constructed from the training data)
 in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}  For the targeted application in this work,
+(i.e., WMHs) for model construction.  For the targeted application in this work,
 tissue classification is performed at the voxelwise level.  In other words, each voxel
@@
 space of the template.
-\textcolor{blue}{Similar feature images were used in}
+Similar feature images were used in
 [@Anbeek:2004aa]
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
+although, unlike the proposed framework, this previous work lacks
 normalization to the standard
@@
 across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
+registered to the T1 template component using a B-spline variant [@Tustison:2013ac]
+of the well-known ANTs Symmetric Normalization (SyN) algorithm [@Avants:2011ab].
+Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.
 
@@
 additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
+In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
 all three aligned preprocessed images for multivariate segmentation during the
-second stage.}
+second stage.
 The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
@@
 a self-encapsulated example to showcase the proposed methodology
-\textcolor{blue}{which will also be available on github.}[^3]  The fact that
+which will also be available on github.[^3]  The fact that
 the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
@@
 The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
-The ``truth'' labels are the seven anatomical regions given above.
+The "truth" labels are the seven anatomical regions given above.
 
@@
       regions.
+
+Several measures have been employed in the literature for evaluating
+automated white matter lesion segmentation involving such quantities as lesion load,
+voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.[@Garcia-Lorenzo:2013aa].
+For this work, due to the relatively small-size distribution of the lesion load in our data cohort (see Table 1), we used
+four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
+The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
+It should be noted that the number of true negatives is not readily
+incorporated into measures of accuracy as the quantity of "true negatives" (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
+The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,
+
+$$ sensitivity =  \frac{ TP }{TP + FN}, $$
+
+and the positive predictive value (PPV),
+
+$$ PPV =  \frac{ TP }{TP + FP}, $$
+
+such that $F_1$ is given by
+
+$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
+
+The relative volume difference is calculated for each of the true positive lesions using
+the manual and predicted lesion volumes:
+
+$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
+
+To illustrate how the performance of our framework varies with lesion size, we calculated
+the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:
+
+* Quantile 1: $[1-12)$,
+
+* Quantile 2: $[12-28)$, and
+
+* Quantile 3: $[28-551]$.
+
+These quantiles are used to showcase the performance (cf Figure 5).

### 2016-06-30 methodsX.Rmd
ENH:  Submission files.

--- 2016-06-30 stitched.Rmd
@@
----
-output:
-  word_document:
-    fig_caption: true
-  pdf_document:
-    fig_caption: true
-    latex_engine: xelatex
-    keep_tex: yes
-  html_document:
-    toc: false
-
-header-includes:
-   - \usepackage{booktabs}
-   - \usepackage[final]{changes}
-   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.2}
-   - \usepackage[compact]{titlesec}
-   - \usepackage{enumitem}
-   - \usepackage{tikz}
-   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-   - \setlist{nolistsep}
-   - \titlespacing{\section}{2pt}{*0}{*0}
-   - \titlespacing{\subsection}{2pt}{*0}{*0}
-   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
-   - \setlength{\parskip}{3pt}
-   - \setremarkmarkup{(#2)}
-   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
-   - \definechangesauthor[name={James Stone}, color=magenta]{js}
-   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
-   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
-   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
-   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
-   - \definechangesauthor[name={David Tate}, color=orange]{dt}
-bibliography: references.bib
-csl: council-of-science-editors.csl
-fontsize: 11pt
-mainfont: Georgia
-geometry: margin=1.0in
----
-
-<!--
-   - \setlength{\parskip}{3pt}
-   - \setlength{\topsep}{0pt}
-   - \setlength{\partopsep}{0pt}
-   - \setlength{\itemsep}{0pt}
-   - \setlength{\floatsep}{0pt}
-   - \setlength{\intextsep}{2pt}
-   - \setlength{\abovecaptionskip}{2pt}
-   - \setlength{\belowcaptionskip}{0pt}
--->
-
-<style type="text/css">
-body,
-code.bash{
-  font-size: 8px;
-}
-pre {
-  font-size: 8px
-}
-</style>
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set( cache=TRUE )
-```
-
-\pagenumbering{gobble}
-
-# Abstract
-
-White matter hyperintensities (WMHs) are foci of abnormal signal
-intensity in white matter regions seen with magnetic resonance imaging
-(MRI). WMHs are associated with normal aging and have shown prognostic
-value in neurological conditions such as traumatic brain injury (TBI).
-The impracticality of manually quantifying these lesions limits their
-clinical utility and motivates the utilization of machine learning
-techniques for automated segmentation workflows.  Herein, we develop a
-concatenated random forest framework with image features for
-segmenting WMHs in a TBI cohort. The framework is built upon the
-Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
-(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
-veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
-(CENC) observational study were acquired. Manual annotations were
-employed for both training and evaluation using a leave-one-out
-strategy.  \textcolor{blue}{Performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.
-Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
-predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
-relative volume difference} = 43 $\pm$ 26\%.  In addition, three
-lesion size ranges are selected to illustrate the variation in
-performance with lesion size.}  Paired with
-correlative outcome data, supervised learning methods may allow for
-identification of imaging features predictive of diagnosis and prognosis
-in individual TBI patients.
-
-\clearpage
-
-# Introduction
-
-## White matter hyperintensities in TBI
-
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
-
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
-Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
-in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].
-
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
-Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
-In this work, we present an automated framework for quantification of WMHs in
-multi-modal MRI using the random forest machine learning technique.}
-
-## Random forests for WMH segmentation
-
-<!--
-Machine learning and pattern recognition techniques have seen increased application
-for various medical image analysis workflows (see, for example, the annual
-Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
-Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as supervised image segmentation
-(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
-Facilitating the current employment of such techniques are the number of available imaging
-data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
-SciPy [@scipy] and the R project for statistical computing [@R] and their associated
-extensions.
--->
-
-The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
-normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., [@viola2005]).
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
-delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
-
-<!--
-Of note, these
-latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
--->
-
-Random forests are conceptually straightforward [@breiman2001].
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
-For the proposed application, individual
-voxels (and their corresponding feature values) are introduced at the root of a particular
-decision tree and traverse the edges and internal nodes ultimately ending up at one of
-the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
-\textcolor{blue}{Finally, Breiman} [@breiman2001]
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
-
-In this work, we develop a concatenated random forest framework with a
-feature image set (both spatial and intensity-based)
-for segmenting WMHs in a large TBI cohort.
-The entire framework is built on the well-known open-source
-Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
-Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.
-
-[^i1]: https://github.com/stnava/ANTs
-[^i2]: https://github.com/stnava/ANTsR
-
 # Materials and Methods
@@
 3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
+3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
 In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
@@
 \textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
-
-# Results
-
-## White matter hyperintensity segmentation evaluation
-
-\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data over the three lesion volume ranges.
-These performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed are meant to illustrate the variation
-in performance with lesion size.}
-
-\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity for this range is more varied compared with the largest
-set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
-the volumetric estimate when comparing the manually identified lesions versus the
-automatically predicted lesions.}
-
-\textcolor{blue}{
-We averaged these measures over all lesions and all subjects which resulted in the
-following values:}
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
-
-    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
-
-    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
-
-    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
-
-    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
-
-<!--
-
-In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-On the left of Figure 5 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
--->
-
-![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
-section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
-volume difference.   These quantitative assessments are given for three quantile ranges
-spanning the range of the manually-derived lesion volumes.    Overall improvement in all
-three whole lesion-based measuers is seen as the
-second Stage RF model is applied for all three quantile ranges.
-The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
-
-## Ranking feature importance
-
-After performing the leave-one-out evaluation,
-we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
-2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
-during the out-of-bag phase of the random forest model construction and quantifies the decrease
-in prediction accuracy from omitting the specified feature.  In other words, this quantity
-helps determine the importance of a particular feature and, although we save such
-efforts for future work, this information provides us with guidance for future feature
-pruning and/or additions.
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 1 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile
- and illustrate the
-stability of the feature importance across the leave-one-out models.
-At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile and illustrate the
-stability of the feature importance across the leave-one-out models.
-We augment the 31 feature images from the first stage by adding an additional
-7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.](Figures/averageLeaveOneOutStage2.png)
-
-The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the
-separate stages are averaged over the entire corresponding model set.  In addition, we
-track the variance for each feature over all models to illustrate the stability of
-the chosen features during the evaluation.  This latter information is illustrated as
-horizontal errors bars providing the $95^{th}$ percentile
-Note that the reader can cross reference Table 1 for identifying corresponding feature types
-and names.
-
-# Discussion
-
-<!-- Nick could you add some detail here on how this technique compares
-to other automated or semi automated approaches for WMH detection? -->
-
-\textcolor{blue}{
-In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
-However, we can extrapolate from similar application areas where such methodologies have a
-much longer history of development such as white matter lesion segmentation in
-multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
-matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
-of two cohorts from two different sites used in
-the MS Lesion segmentation challenge associated with the international MICCAI 2008
-conference.  For comparison, this challenge training data set
-has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
-our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
-40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
-almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
-Note that our relative volume difference numbers are comparable to the
-relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
-[@vanginneken:2007aa].
-
-Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
-The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
-FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
-posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives.  Inspection of the bottom of the
-plots demonstrates the lack of discriminating features associated with the
-T1 image which is also well-known in the clinical literature.
-
-![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
-As described earlier, for Stage 2, we used the output random forest voting maps from Stage
-1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
-"_RFStage1VotingMaps_" where the final numeral is associated with the brain
-parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
-The Stage 2 feature importance plot follows similar
-trends as that for Stage 1 with the T1 images not contributing much to the identification
-of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
-important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
-3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
-alone it is intuitive that such features would be important.
-
-<!--
-
-* Evaluation * Proper context * We only have manual labelings from a
-single radiologist ( * TBI white matter hyperintensities are more
-difficult to segment than MS lesions * individual lesions are smaller *
-lesion load is typically smaller * not prone to enhancement * less
-distinct from surrounding tissue ? * Despite the doubts regarding the
-efficacy of a typical segmentation validation, our contribution is
-useful because * the total segmentation framework is publicly available
-within the ANTs/ANTsR framework * although only a single site is
-analyzed, the feature images are site-agnostic * we plan to apply the
-current RF models directly to the other site data * we can then build
-the training database further by enlisting other experts * Is this the
-first work exploring segmentation of white matter hyperintensities in
-TBI? * The data (including manual tracings) will be made available as
-part of the FITBIR effort.
-
--->
-
-# Conclusions
-
-The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation
-of WMHs from FLAIR images to help build two-stage ensembles of decision
-trees for the automated identification of these lesions. Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology.  Also, we
-recognize that only a single site was used for evaluating the proposed
-framework.  However, we are currently processing other site data with
-the models developed for this work and the results look promising since
-the developed features are site-agnostic.
-
-As far as we know, this is the first report utilizing a novel random
-forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
-tend to be more difficult to segment than MS lesions as the former tend
-to be smaller with an overall smaller lesion load.  Also, enhancement
-protocols with the former tend to be less successful than with the
-latter.  As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.
-
-<!--
-Our
-framework is also available as open-source as part of well-known
-neuroimaging tools which easily allows for additions/extensions but is
-also, as far as we know, the first random forest-based technique
-available for such application.
--->
-
-Two major meta-analyses of WMHs have been published covering the periods
-prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
-Debette & Markus [@Debette:2010aa] found that the presence of
-WMHs was related to subsequent cognitive decline, a higher risk of
-developing dementia, stroke, and of mortality. Lesion volume at baseline
-was also predictive of cognitive decline. Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
-concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
-that WMHs were associated with cognitive deficit (effect size of -0.10,
-95% CI: -0.13 to -0.08) after controlling for age.
-
-Despite the potential clinical significance of WMHs
-these lesions receive little
-attention in current clinical workflows. When reported in a standard
-neuroradiologist interpretation, they are typically handled as
-incidental findings and are assigned little clinical significance. This
-likely reflects the impracticality of performing a detailed assessment
-of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
-supervised statistical learning tool that is appropriate for the
-application to such large-scale datasets.
-
-<!--
-
-The currently described tool is just one example of how "supervised
-learning" algorithms might be applied to aid in the diagnosis of TBI and
-other disease processes through the specific identification of features
-predictive of a given disease state. It is an important demonstration of
-the potential power of these analytical approaches in the rapid but
-comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide
-variety of settings for the rapid identification of predictive imaging
-features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
-manufacturers utilize these types of approaches to equip self-driving
-vehicles to recognize and respond to unique external surroundings
-through the identification of visual information sufficiently similar to
-previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning
-approaches may allow for the rapid identification of information
-predictive of disease state in an individual patient. These approaches
-have been applied to the segmentation of macroscopically visible
-structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of
-imaging data in the individual patient with a primary quantitative
-output metrics to include sequences such as diffusion tensor imaging
-(DTI) and its variants, functional connectivity, perfusion weighted
-imaging, and cortical thickness assessments. At present, these advanced
-neuroimaging sequences are confined to cohort-based research studies due
-to the lack of available analytical tools to assess the information in
-the setting of the individual patient [@Mayer:2014aa]. Application of
-deep learning approaches in the context of data with primary
-quantitative outputs will require large scale normative and disease
-specific databases. Building these large scale imaging libraries is
-resource intensive and requires a multi-center approach with harmonized
-scanners between sites and correlative non-imaging clinical data. Large
-scale TBI data is becoming increasingly available through activities
-such as the Chronic Effects of Neurotrauma Consortium (CENC),
-Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
-Collaborative European Neurotrauma Effectiveness Research in TBI
-(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
-Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data,
-deep learning algorithms may be well positioned to help transform how
-neuroimaging is interpreted for the clinical management of patients with
-this disease process.
-
--->
-
-\clearpage
-
-## Acknowledgements
-
-The authors wish to acknowledge all other members of
-the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
-X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
-also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
-and Kevin Wilson. We would also like to thank the study participants and
-their family members. This project was jointly supported by the
-Department of Defense (W81XWH-13-2-0095), the U.S. Department of
-Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
-Grant HU 0001-08-0001.
-
-## Declaration of Interest/Disclaimer
-
-The authors report no financial
-disclosures or conflicts of interest. The views expressed here are those
-of the authors and do not necessarily reflect the official policy of
-position of the Department of the Navy, Department of Defense, nor the
-U.S. Government. This work was prepared as a part of official duties;
-Title 17 USC §105 provides that Copyright protection under this title is
-not available for any work of the U.S. Government. Title 17 USC §101
-defines a US Government work as a work prepared by a military service
-member of employee of the US Government as part of that person’s
-official duties.
-
-\clearpage
-
-<!-- \listofchanges
-
-\clearpage -->
-
-# References

### 2016-06-30 results.Rmd
ENH:  Submission files.

@@
 
-\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
+In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
 of the previously described TBI data over the three lesion volume ranges.
@@
 size ranges over which these measures are computed are meant to illustrate the variation
-in performance with lesion size.}
+in performance with lesion size.
 
-\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
+Smaller lesions ($< 12$ voxels) are more difficult
 to identify which is why the sensitivity for this range is more varied compared with the largest
@@
 the volumetric estimate when comparing the manually identified lesions versus the
-automatically predicted lesions.}
+automatically predicted lesions.
 
-\textcolor{blue}{
 We averaged these measures over all lesions and all subjects which resulted in the
-following values:}
+following values:
 
-* \textcolor{blue}{Stage 1}
+* Stage 1
 
-    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
+    * sensitivity = 0.70 $\pm$ 0.34
 
-    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
+    * PPV = 0.42 $\pm$ 0.36
 
-    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
+    * $F_1$ = 0.47 $\pm$ 0.36
 
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
+    * relative volume difference = 43 $\pm$ 38\%
 
-* \textcolor{blue}{Stage 1}
+* Stage 2
 
-    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
+    * sensitivity = 0.68 $\pm$ 0.38
 
-    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
+    * PPV = 0.51 $\pm$ 0.40
 
-    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
+    * $F_1$ = 0.52 $\pm$ 0.36
 
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
+    * relative volume difference = 43 $\pm$ 26\%
 
@@
 
-![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+<!--
+
+![Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
 section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
@@
 The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
+decreased predicted volume over the Stage 1 results.](Figures/evaluationResults.png)
+
+-->
 
@@
 
+<!--
+
 ![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
@@
 At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)
+
+-->
+
+<!--
 
@@
 
+-->
+
 The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the

### 2016-06-30 resultsX.Rmd
ENH:  Submission files.

--- 2016-06-30 stitched.Rmd
@@
----
-output:
-  word_document:
-    fig_caption: true
-  pdf_document:
-    fig_caption: true
-    latex_engine: xelatex
-    keep_tex: yes
-  html_document:
-    toc: false
-
-header-includes:
-   - \usepackage{booktabs}
-   - \usepackage[final]{changes}
-   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.2}
-   - \usepackage[compact]{titlesec}
-   - \usepackage{enumitem}
-   - \usepackage{tikz}
-   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-   - \setlist{nolistsep}
-   - \titlespacing{\section}{2pt}{*0}{*0}
-   - \titlespacing{\subsection}{2pt}{*0}{*0}
-   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
-   - \setlength{\parskip}{3pt}
-   - \setremarkmarkup{(#2)}
-   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
-   - \definechangesauthor[name={James Stone}, color=magenta]{js}
-   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
-   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
-   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
-   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
-   - \definechangesauthor[name={David Tate}, color=orange]{dt}
-bibliography: references.bib
-csl: council-of-science-editors.csl
-fontsize: 11pt
-mainfont: Georgia
-geometry: margin=1.0in
----
-
-<!--
-   - \setlength{\parskip}{3pt}
-   - \setlength{\topsep}{0pt}
-   - \setlength{\partopsep}{0pt}
-   - \setlength{\itemsep}{0pt}
-   - \setlength{\floatsep}{0pt}
-   - \setlength{\intextsep}{2pt}
-   - \setlength{\abovecaptionskip}{2pt}
-   - \setlength{\belowcaptionskip}{0pt}
--->
-
-<style type="text/css">
-body,
-code.bash{
-  font-size: 8px;
-}
-pre {
-  font-size: 8px
-}
-</style>
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set( cache=TRUE )
-```
-
-\pagenumbering{gobble}
-
-# Abstract
-
-White matter hyperintensities (WMHs) are foci of abnormal signal
-intensity in white matter regions seen with magnetic resonance imaging
-(MRI). WMHs are associated with normal aging and have shown prognostic
-value in neurological conditions such as traumatic brain injury (TBI).
-The impracticality of manually quantifying these lesions limits their
-clinical utility and motivates the utilization of machine learning
-techniques for automated segmentation workflows.  Herein, we develop a
-concatenated random forest framework with image features for
-segmenting WMHs in a TBI cohort. The framework is built upon the
-Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
-(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
-veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
-(CENC) observational study were acquired. Manual annotations were
-employed for both training and evaluation using a leave-one-out
-strategy.  \textcolor{blue}{Performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.
-Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
-predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
-relative volume difference} = 43 $\pm$ 26\%.  In addition, three
-lesion size ranges are selected to illustrate the variation in
-performance with lesion size.}  Paired with
-correlative outcome data, supervised learning methods may allow for
-identification of imaging features predictive of diagnosis and prognosis
-in individual TBI patients.
-
-\clearpage
-
-# Introduction
-
-## White matter hyperintensities in TBI
-
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
-
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
-Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
-in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].
-
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
-Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
-In this work, we present an automated framework for quantification of WMHs in
-multi-modal MRI using the random forest machine learning technique.}
-
-## Random forests for WMH segmentation
-
-<!--
-Machine learning and pattern recognition techniques have seen increased application
-for various medical image analysis workflows (see, for example, the annual
-Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
-Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as supervised image segmentation
-(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
-Facilitating the current employment of such techniques are the number of available imaging
-data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
-SciPy [@scipy] and the R project for statistical computing [@R] and their associated
-extensions.
--->
-
-The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
-normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., [@viola2005]).
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
-delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
-
-<!--
-Of note, these
-latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
--->
-
-Random forests are conceptually straightforward [@breiman2001].
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
-For the proposed application, individual
-voxels (and their corresponding feature values) are introduced at the root of a particular
-decision tree and traverse the edges and internal nodes ultimately ending up at one of
-the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
-\textcolor{blue}{Finally, Breiman} [@breiman2001]
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
-
-In this work, we develop a concatenated random forest framework with a
-feature image set (both spatial and intensity-based)
-for segmenting WMHs in a large TBI cohort.
-The entire framework is built on the well-known open-source
-Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
-Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.
-
-[^i1]: https://github.com/stnava/ANTs
-[^i2]: https://github.com/stnava/ANTsR
-
-# Materials and Methods
-
-## Imaging
-
-MR images utilized for this initial report were acquired from a single scanner involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (see Walker et al., this issue).
-Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
-Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
-
-\input{trainingData.tex}
-
-Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
-3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
-
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
-subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All
-multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
-The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
-\textcolor{blue}{
-which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
-
-[^1]: http://nifti.nimh.nih.gov/nifti-1
-
-## Quantitative analysis
-
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1'', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
-
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
-
-### Symmetric multi-modal templates
-
-\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}
-[@landman2011]\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
-
-![Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific asymmetry-based
-features.](Figures/MMRR.png)
-
-### Feature images for WMH segmentation
-
-\textcolor{blue}{
-Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}  For the targeted application in this work,
-tissue classification is performed at the voxelwise level.  In other words, each voxel
-within the region of interest is sent through the ensemble of decision trees and receives
-a set of classification votes from each tree thus permitting a regression or classification
-solution.  Since this procedure is performed at the voxelwise level, intensity information
-alone is insufficient for good segmentation performance due to the lack of spatial context.
-For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated lesion detection
-algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
-even though machine learning and pattern recognition techniques are extremely powerful and
-have significant potential, just as crucial to outcome is the creative construction and
-deployment of salient feature images which we detail below.
-
-\input{featureImagesTable.tex}
-
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.  In Table 2, we provide a list and basic
-categorization of the feature images used for the initial (i.e., Stage 1---more on the use
-of multiple random forest stages below) segmentation of the WMHs.
-In addition Figure 3 provides a representation of a set of feature images for a single
-subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
-with seven labels:
-
-* cerebrospinal fluid (label 1),
-* gray matter (label 2),
-* white matter (label 3),
-* deep gray matter (label 4),
-* brain stem (label 5),
-* cerebellum (label 6), and
-* white matter hyperintensities (label 7).
-
-![Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
-application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
-The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right.  This gives a probabilistic classification of
-tissue type.  Not shown are the probability and voting images for the brain stem and
-cerebellum.](Figures/featureImages.png)
-
-As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
-acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
-the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
-are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
-N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
-range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].
-
-![Sample FLAIR acquisition image slices showing both manual and random forest
-segmentations for both stages obtained during the leave-one-out evaluation.  Manual
-segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.](Figures/sampleResults.png)
-
-The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
-  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
-CSF, gray matter, white matter, deep gray matter, brain stem, and
-cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
-first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 2).
-
-Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
-We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a Manhattan radius of one voxel given the typical size
-of individual WMHs. For each of the preprocessed
-images, we calculate the difference in intensities with the corresponding warped template
-component.  Previous success in the international brain tumor segmentation
-competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 2.  We employ the
-same strategy here.
-
-<!--
-For example, the template
-difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
-
-$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$
-
-where
-
-$$\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$$
-
-is the transform
-which maps from the individual subject space to the
-template space and $T_{FLAIR}$ is the FLAIR template component.
--->
-
-To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
-we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature.
-
-<!--
-For the FLAIR component, this contralateral
-difference image is calculated from
-
-$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$
-
-where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
-the symmetric template.
--->
-
-The segmentation probability images described above are used as feature images to
-provide a spatial context for the random forest model prediction step.  Additional spatial contextual
-feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
-deep gray matter images.  These latter images are intended to help distinguish white matter
-hyperintensities from false positives induced by the partial voluming at the gray/white
-matter interface.  A third set of images are based on the voxel location within the
-space of the template.
-\textcolor{blue}{Similar feature images were used in}
-[@Anbeek:2004aa]
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
-
-### Stacked (concatenated) random forests for improved segmentation performance
-
-In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
-concatenated supervised approach, whereby the prediction output from the first random forest
-model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
-feature images of the training data (as described previously) are used to construct the
-Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
-are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
-additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
-all three aligned preprocessed images for multivariate segmentation during the
-second stage.}
-The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
-
-### Implementation
-
-As pointed out in a recent comprehensive lesion segmentation review
-[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
-is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
-our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
-this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
-ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
-use the _randomForest_ package [@liaw2002] using the default settings with
-2000 trees per model and 500 randomly selected samples per label per image.
-Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.
-
-In addition, similar to our previous offering [^2], we plan on creating
-a self-encapsulated example to showcase the proposed methodology
-\textcolor{blue}{which will also be available on github.}[^3]  The fact that
-the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
-with the manual labelings will facilitate reproducibility on the part of the reader
-as well as any interest in extending the proposed framework to other data sets.
-
-[^2]: https://github.com/ntustison/ANTsAndArboles
-
-[^3]: https://github.com/ntustison/WatchMeHyperventilate
-
-### Evaluation protocol overview
-
-In order to evaluate the protocol described, we performed a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.
-Initial processing included the creation of all Stage 1 feature images for all subjects.
-The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
-The ``truth'' labels are the seven anatomical regions given above.
-
-The leave-one-out procedure is as follows:
-
-* Create Stage 1 feature images for all 24 subjects.
-* For each of the 24 subjects:
-    + sequester the current subject and corresponding feature images.
-    + construct the Stage 1 random forest model from the remaining 23 subjects.
-    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
-    + the previous step produces the Stage 1 voting maps for all seven labels.
-    + for each of the 23 subjects, perform a Bayesian-based segmentation with an
-      MRF spatial
-      prior using the seven voting maps as additional tissue priors.
-    + construct the Stage 2 random forest model from all the Stage 1 feature images,
-      seven voting maps, and seven posterior probability maps from the previous step.
-    + send the sequestered subject through the random forest models for both stages.
-    + compare the final results with the manually-defined white matter hyperintensity
-      regions.
-
-\textcolor{blue}{Several measures have been employed in the literature for evaluating
-automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
-For this work, due to the relatively small-size distribution} \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
-four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
-incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
-The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
-
-$$ sensitivity =  \frac{ TP }{TP + FN}, $$
-
-\textcolor{blue}{and the positive predictive value (PPV),}
-
-$$ PPV =  \frac{ TP }{TP + FP}, $$
-
-\textcolor{blue}{such that $F_1$ is given by}
-
-$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
-
-\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
-the manual and predicted lesion volumes:}
-
-$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
-
-\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
-
-* \textcolor{blue}{Quantile 1: $[1-12)$,}
-
-* \textcolor{blue}{Quantile 2: $[12-28)$, and}
-
-* \textcolor{blue}{Quantile 3: $[28-551]$}.
-
-\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
-
 # Results
@@
 
-* \textcolor{blue}{Stage 1}
+* \textcolor{blue}{Stage 2}
 
@@
 and names.
-
-# Discussion
-
-<!-- Nick could you add some detail here on how this technique compares
-to other automated or semi automated approaches for WMH detection? -->
-
-\textcolor{blue}{
-In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
-However, we can extrapolate from similar application areas where such methodologies have a
-much longer history of development such as white matter lesion segmentation in
-multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
-matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
-of two cohorts from two different sites used in
-the MS Lesion segmentation challenge associated with the international MICCAI 2008
-conference.  For comparison, this challenge training data set
-has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
-our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
-40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
-almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
-Note that our relative volume difference numbers are comparable to the
-relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
-[@vanginneken:2007aa].
-
-Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
-The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
-FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
-posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives.  Inspection of the bottom of the
-plots demonstrates the lack of discriminating features associated with the
-T1 image which is also well-known in the clinical literature.
-
-![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
-As described earlier, for Stage 2, we used the output random forest voting maps from Stage
-1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
-"_RFStage1VotingMaps_" where the final numeral is associated with the brain
-parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
-The Stage 2 feature importance plot follows similar
-trends as that for Stage 1 with the T1 images not contributing much to the identification
-of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
-important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
-3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
-alone it is intuitive that such features would be important.
-
-<!--
-
-* Evaluation * Proper context * We only have manual labelings from a
-single radiologist ( * TBI white matter hyperintensities are more
-difficult to segment than MS lesions * individual lesions are smaller *
-lesion load is typically smaller * not prone to enhancement * less
-distinct from surrounding tissue ? * Despite the doubts regarding the
-efficacy of a typical segmentation validation, our contribution is
-useful because * the total segmentation framework is publicly available
-within the ANTs/ANTsR framework * although only a single site is
-analyzed, the feature images are site-agnostic * we plan to apply the
-current RF models directly to the other site data * we can then build
-the training database further by enlisting other experts * Is this the
-first work exploring segmentation of white matter hyperintensities in
-TBI? * The data (including manual tracings) will be made available as
-part of the FITBIR effort.
-
--->
-
-# Conclusions
-
-The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation
-of WMHs from FLAIR images to help build two-stage ensembles of decision
-trees for the automated identification of these lesions. Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology.  Also, we
-recognize that only a single site was used for evaluating the proposed
-framework.  However, we are currently processing other site data with
-the models developed for this work and the results look promising since
-the developed features are site-agnostic.
-
-As far as we know, this is the first report utilizing a novel random
-forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
-tend to be more difficult to segment than MS lesions as the former tend
-to be smaller with an overall smaller lesion load.  Also, enhancement
-protocols with the former tend to be less successful than with the
-latter.  As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.
-
-<!--
-Our
-framework is also available as open-source as part of well-known
-neuroimaging tools which easily allows for additions/extensions but is
-also, as far as we know, the first random forest-based technique
-available for such application.
--->
-
-Two major meta-analyses of WMHs have been published covering the periods
-prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
-Debette & Markus [@Debette:2010aa] found that the presence of
-WMHs was related to subsequent cognitive decline, a higher risk of
-developing dementia, stroke, and of mortality. Lesion volume at baseline
-was also predictive of cognitive decline. Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
-concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
-that WMHs were associated with cognitive deficit (effect size of -0.10,
-95% CI: -0.13 to -0.08) after controlling for age.
-
-Despite the potential clinical significance of WMHs
-these lesions receive little
-attention in current clinical workflows. When reported in a standard
-neuroradiologist interpretation, they are typically handled as
-incidental findings and are assigned little clinical significance. This
-likely reflects the impracticality of performing a detailed assessment
-of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
-supervised statistical learning tool that is appropriate for the
-application to such large-scale datasets.
-
-<!--
-
-The currently described tool is just one example of how "supervised
-learning" algorithms might be applied to aid in the diagnosis of TBI and
-other disease processes through the specific identification of features
-predictive of a given disease state. It is an important demonstration of
-the potential power of these analytical approaches in the rapid but
-comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide
-variety of settings for the rapid identification of predictive imaging
-features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
-manufacturers utilize these types of approaches to equip self-driving
-vehicles to recognize and respond to unique external surroundings
-through the identification of visual information sufficiently similar to
-previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning
-approaches may allow for the rapid identification of information
-predictive of disease state in an individual patient. These approaches
-have been applied to the segmentation of macroscopically visible
-structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of
-imaging data in the individual patient with a primary quantitative
-output metrics to include sequences such as diffusion tensor imaging
-(DTI) and its variants, functional connectivity, perfusion weighted
-imaging, and cortical thickness assessments. At present, these advanced
-neuroimaging sequences are confined to cohort-based research studies due
-to the lack of available analytical tools to assess the information in
-the setting of the individual patient [@Mayer:2014aa]. Application of
-deep learning approaches in the context of data with primary
-quantitative outputs will require large scale normative and disease
-specific databases. Building these large scale imaging libraries is
-resource intensive and requires a multi-center approach with harmonized
-scanners between sites and correlative non-imaging clinical data. Large
-scale TBI data is becoming increasingly available through activities
-such as the Chronic Effects of Neurotrauma Consortium (CENC),
-Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
-Collaborative European Neurotrauma Effectiveness Research in TBI
-(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
-Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data,
-deep learning algorithms may be well positioned to help transform how
-neuroimaging is interpreted for the clinical management of patients with
-this disease process.
-
--->
-
-\clearpage
-
-## Acknowledgements
-
-The authors wish to acknowledge all other members of
-the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
-X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
-also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
-and Kevin Wilson. We would also like to thank the study participants and
-their family members. This project was jointly supported by the
-Department of Defense (W81XWH-13-2-0095), the U.S. Department of
-Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
-Grant HU 0001-08-0001.
-
-## Declaration of Interest/Disclaimer
-
-The authors report no financial
-disclosures or conflicts of interest. The views expressed here are those
-of the authors and do not necessarily reflect the official policy of
-position of the Department of the Navy, Department of Defense, nor the
-U.S. Government. This work was prepared as a part of official duties;
-Title 17 USC §105 provides that Copyright protection under this title is
-not available for any work of the U.S. Government. Title 17 USC §101
-defines a US Government work as a work prepared by a military service
-member of employee of the US Government as part of that person’s
-official duties.
-
-\clearpage
-
-<!-- \listofchanges
-
-\clearpage -->
-
-# References

### 2016-06-30 README.R

@@ -2,9 +2,13 @@ library( rmarkdown )
 library( ggplot2 )
 library( RISmed )
 
-stitchedFile <- "stitched.Rmd";
-
 formatFile <- "format.Rmd"
+
+######################################
+#
+#  Clean
+#
+stitchedFile <- "stitched.Rmd";
 
 rmdFiles <- c( formatFile,
                "abstract.Rmd",
@@ -27,7 +31,36 @@ for( i in 1:length( rmdFiles ) )
   }
 
 cat( '\n Pandoc rendering', stitchedFile, '\n' )
-render( stitchedFile, output_format = "all" )
+render( stitchedFile, output_format = c( "pdf_document", "word_document" ) )
+
+######################################
+#
+#  Annotated
+#
+stitchedFile <- "stitchedX.Rmd";
+
+rmdFiles <- c( formatFile,
+               "abstractX.Rmd",
+               "introX.Rmd",
+               "methodsX.Rmd",
+               "resultsX.Rmd",
+               "discussionX.Rmd",
+               "acknowledgementsDisclaimers.Rmd"
+   )
+
+for( i in 1:length( rmdFiles ) )
+  {
+  if( i == 1 )
+    {
+    cmd <- paste( "cat", rmdFiles[i], ">", stitchedFile )
+    } else {
+    cmd <- paste( "cat", rmdFiles[i], ">>", stitchedFile )
+    }
+  system( cmd )
+  }
+
+cat( '\n Pandoc rendering', stitchedFile, '\n' )
+render( stitchedFile, output_format = "pdf_document" )
 
 
 singleRmdFiles <- c(
@@ -38,7 +71,7 @@ singleRmdFiles <- c(
 for( i in 1:length( singleRmdFiles ) )
   {
   cat( '\n Pandoc rendering', singleRmdFiles[i], '\n' )
-  render( singleRmdFiles[i], output_format = "all" )
+  render( singleRmdFiles[i], output_format = c( "pdf_document", "word_document" ) )
   }
 
 

### 2016-07-07 responseToReviewers.Rmd
ENH:  Second round.

@@
 bibliography: references.bib
-csl: council-of-science-editors.csl
+csl: national-science-foundation-grant-proposals.csl
 fontsize: 12pt
@@
 probability estimates) of the first RF model application,
-which we denote as ``Stage1 '', is used as input (along with the original set of
+which we denote as ``Stage 1'', is used as input (along with the original set of
 feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
@@
 graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
+The final, or end, nodes represent the various classifications produced by traversal through the decision tree.
 For the proposed application, individual
@@
 > (vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
-clinical research involving WMHs and then pointing to the need for automated
+clinical research involving WMHs and then point to the need for automated
 techniques to minimize the variability and time requirements of manual

### 2016-07-07 responseToReviewers.tex
ENH:  Second round.

@@
 All multi-modal MR dicom image slices were converted to the nifti file format.
-All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}\textsuperscript{1}
-\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{2}
+All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}
+{[}1{]}
+\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}
+{[}2{]}
 \textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
@@
 
-\textcolor{blue}{Following}\textsuperscript{3}
-\textcolor{blue}{and}\textsuperscript{4},
+\textcolor{blue}{Following} {[}3{]} \textcolor{blue}{and} {[}4{]},
 \textcolor{blue}{optimally derived templates
 serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{5}\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{6}
-\textcolor{blue}{which
+For this work we use the multi-modal data available from the public MMRR data set}
+{[}5{]}\textcolor{blue}{.  We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}
+{[}6{]} \textcolor{blue}{which
 results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
@@
 \emph{\textcolor{blue}{The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines}}\textsuperscript{3}
+extraction and normal tissue segmentation pipelines}} {[}3{]}
 \emph{\textcolor{blue}{.  The result is a mask delineating the brain
 parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
-deep gray matter, brain stem, and cerebellum}}\textsuperscript{7}
+deep gray matter, brain stem, and cerebellum}} {[}7{]}
 \emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
 first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{4}
-\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{5}
-\emph{\textcolor{blue}{(cf Figure 3).}}
+are from multi-model optimal symmetric shape/intensity templates}}
+{[}4{]} \emph{\textcolor{blue}{created from the public MMRR data set}}
+{[}5{]} \emph{\textcolor{blue}{(cf Figure 3).}}
 \end{quote}
@@
 \textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)}\textsuperscript{8,9}
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{10,11}.
-\textcolor{blue}{Finally, Breiman}\textsuperscript{12}
+(e.g., AdaBoost)} {[}8, 9{]}
+\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}
+{[}10, 11{]}. \textcolor{blue}{Finally, Breiman} {[}12{]}
 \textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
@@
 callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{13}
+with Glasgow Outcome-Extended (GOS-E) scores.} {[}13{]}
 \end{quote}
@@
 callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{13}
+with Glasgow Outcome-Extended (GOS-E) scores} {[}13{]}
 \textcolor{blue}{---a numeric groupwise assessment used to
@@
 \begin{quote}
-\textcolor{blue}{Similar feature images were used in}\textsuperscript{14}
+\textcolor{blue}{Similar feature images were used in} {[}14{]}
 \textcolor{blue}{although, unlike the proposed framework, this previous work lacks
@@
 across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant}\textsuperscript{15}
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{16}.
+registered to the T1 template component using a B-spline variant}
+{[}15{]}
+\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}
+{[}16{]}.
 \textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
@@
 (vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
-clinical research involving WMHs and then pointing to the need for automated
+clinical research involving WMHs and then point to the need for automated
 techniques to minimize the variability and time requirements of manual
@@
 \hypertarget{ref-Avants:2014aa}{}
-1. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
-ToolKit image registration framework. Front Neuroinform. 2014;8:44.
+1. Avants, B. B., Tustison, N. J., Stauffer, M., Song, G., Wu, B., and
+Gee, J. C. ``\textbf{The Insight ToolKit Image Registration Framework}''
+\emph{Front Neuroinform} 8, (2014): 44.
+doi:\href{https://doi.org/10.3389/fninf.2014.00044}{10.3389/fninf.2014.00044}
 
 \hypertarget{ref-Yushkevich:2006aa}{}
-2. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
-User-guided 3D active contour segmentation of anatomical structures:
-Significantly improved efficiency and reliability. Neuroimage.
-2006;31(3):1116--28.
+2. Yushkevich, P. A., Piven, J., Hazlett, H. C., Smith, R. G., Ho, S.,
+Gee, J. C., and Gerig, G. ``\textbf{User-Guided 3D Active Contour
+Segmentation of Anatomical Structures: Significantly Improved Efficiency
+and Reliability}'' \emph{Neuroimage} 31, no. 3 (2006): 1116--28.
+doi:\href{https://doi.org/10.1016/j.neuroimage.2006.01.015}{10.1016/j.neuroimage.2006.01.015}
 
 \hypertarget{ref-Tustison:2014ab}{}
-3. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
-Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
-and freeSurfer cortical thickness measurements. Neuroimage.
-2014;99:166--79.
+3. Tustison, N. J., Cook, P. A., Klein, A., Song, G., Das, S. R., Duda,
+J. T., Kandel, B. M., Strien, N. van, Stone, J. R., Gee, J. C., and
+Avants, B. B. ``\textbf{Large-Scale Evaluation of ANTs and FreeSurfer
+Cortical Thickness Measurements}'' \emph{Neuroimage} 99, (2014):
+166--79.
+doi:\href{https://doi.org/10.1016/j.neuroimage.2014.05.044}{10.1016/j.neuroimage.2014.05.044}
 
 \hypertarget{ref-Tustison:2015aa}{}
-4. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
-Grossman MC, Avants BB. Optimal symmetric multimodal templates and
-concatenated random forests for supervised brain tumor segmentation
-(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.
+4. Tustison, N. J., Shrinidhi, K. L., Wintermark, M., Durst, C. R.,
+Kandel, B. M., Gee, J. C., Grossman, M. C., and Avants, B. B.
+``\textbf{Optimal Symmetric Multimodal Templates and Concatenated Random
+Forests for Supervised Brain Tumor Segmentation (Simplified) with
+ANTsR}'' \emph{Neuroinformatics} 13, no. 2 (2015): 209--25.
+doi:\href{https://doi.org/10.1007/s12021-014-9245-2}{10.1007/s12021-014-9245-2}
 
 \hypertarget{ref-landman2011}{}
-5. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
-Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
-reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
+5. Landman, B. A., Huang, A. J., Gifford, A., Vikram, D. S., Lim, I. A.
+L., Farrell, J. A. D., Bogovic, J. A., Hua, J., Chen, M., Jarso, S.,
+Smith, S. A., Joel, S., Mori, S., Pekar, J. J., Barker, P. B., Prince,
+J. L., and Zijl, P. C. M. van. ``\textbf{Multi-Parametric Neuroimaging
+Reproducibility: A 3-T Resource Study}'' \emph{Neuroimage} 54, no. 4
+(2011): 2854--66.
+doi:\href{https://doi.org/10.1016/j.neuroimage.2010.11.047}{10.1016/j.neuroimage.2010.11.047}
 
 \hypertarget{ref-Avants:2010aa}{}
-6. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre J,
-Gee JC. The optimal template effect in hippocampus studies of diseased
-populations. Neuroimage. 2010;49(3):2457--66.
+6. Avants, B. B., Yushkevich, P., Pluta, J., Minkoff, D., Korczykowski,
+M., Detre, J., and Gee, J. C. ``\textbf{The Optimal Template Effect in
+Hippocampus Studies of Diseased Populations}'' \emph{Neuroimage} 49, no.
+3 (2010): 2457--66.
+doi:\href{https://doi.org/10.1016/j.neuroimage.2009.09.062}{10.1016/j.neuroimage.2009.09.062}
 
 \hypertarget{ref-Avants:2011aa}{}
-7. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
-multivariate framework for \(n\)-tissue segmentation with evaluation on
-public data. Neuroinformatics. 2011;9(4):381--400.
+7. Avants, B. B., Tustison, N. J., Wu, J., Cook, P. A., and Gee, J. C.
+``\textbf{An Open Source Multivariate Framework for \(n\)-Tissue
+Segmentation with Evaluation on Public Data}'' \emph{Neuroinformatics}
+9, no. 4 (2011): 381--400.
+doi:\href{https://doi.org/10.1007/s12021-011-9109-y}{10.1007/s12021-011-9109-y}
 
 \hypertarget{ref-schapire1990}{}
-8. Schapire R. The strength of weak learnability. Machine Learning.
-1990;5:197--227.
+8. Schapire, R. ``\textbf{The Strength of Weak Learnability}''
+\emph{Machine Learning} 5, (1990): 197--227.
 
 \hypertarget{ref-freund1997}{}
-9. Freund Y, Schapire R. A decision-theoretic generalization of on-line
-learning and an application to boosting. Journal of Computer and System
-Sciences. 1997;55:119--139.
+9. Freund, Y. and Schapire, R. ``\textbf{A Decision-Theoretic
+Generalization of on-Line Learning and an Application to Boosting}''
+\emph{Journal of Computer and System Sciences} 55, (1997): 119--139.
 
 \hypertarget{ref-ho1995}{}
-10. Ho TK. Random decision forests. In: Document analysis and
-recognition, 1995., proceedings of the third international conference
-on. Vol. 1. 1995. pp. 278--282 vol.1.
+10. Ho, T. K. ``\textbf{Random Decision Forests}'' \emph{Document
+analysis and recognition, 1995., proceedings of the third international
+conference on} 1, (1995): 278--282 vol.1.
+doi:\href{https://doi.org/10.1109/ICDAR.1995.598994}{10.1109/ICDAR.1995.598994}
 
 \hypertarget{ref-amit1997}{}
-11. Amit Y, Geman D. Shape quantization and recognition with randomized
-trees. Neural Computation. 1997;9:1545--1588.
+11. Amit, Y. and Geman, D. ``\textbf{Shape Quantization and Recognition
+with Randomized Trees}'' \emph{Neural Computation} 9, (1997):
+1545--1588.
 
 \hypertarget{ref-breiman2001}{}
-12. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
+12. Breiman, L. ``\textbf{Random Forests}'' \emph{Machine learning}
+(2001): 5--32.
 
 \hypertarget{ref-Moen:2014aa}{}
-13. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A.
-Traumatic axonal injury: The prognostic value of lesion load in corpus
-callosum, brain stem, and thalamus in different magnetic resonance
-imaging sequences. J Neurotrauma. 2014;31(17):1486--96.
+13. Moen, K. G., Brezova, V., Skandsen, T., Håberg, A. K., Folvik, M.,
+and Vik, A. ``\textbf{Traumatic Axonal Injury: The Prognostic Value of
+Lesion Load in Corpus Callosum, Brain Stem, and Thalamus in Different
+Magnetic Resonance Imaging Sequences}'' \emph{J Neurotrauma} 31, no. 17
+(2014): 1486--96.
+doi:\href{https://doi.org/10.1089/neu.2013.3258}{10.1089/neu.2013.3258}
 
 \hypertarget{ref-Anbeek:2004aa}{}
-14. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
-Probabilistic segmentation of white matter lesions in mR imaging.
-Neuroimage. 2004;21(3):1037--44.
+14. Anbeek, P., Vincken, K. L., Osch, M. J. P. van, Bisschops, R. H. C.,
+and Grond, J. van der. ``\textbf{Probabilistic Segmentation of White
+Matter Lesions in MR Imaging}'' \emph{Neuroimage} 21, no. 3 (2004):
+1037--44.
+doi:\href{https://doi.org/10.1016/j.neuroimage.2003.10.012}{10.1016/j.neuroimage.2003.10.012}
 
 \hypertarget{ref-Tustison:2013ac}{}
-15. Tustison NJ, Avants BB. Explicit B-spline regularization in
-diffeomorphic image registration. Front Neuroinform. 2013;7:39.
+15. Tustison, N. J. and Avants, B. B. ``\textbf{Explicit B-Spline
+Regularization in Diffeomorphic Image Registration}'' \emph{Front
+Neuroinform} 7, (2013): 39.
+doi:\href{https://doi.org/10.3389/fninf.2013.00039}{10.3389/fninf.2013.00039}
 
 \hypertarget{ref-Avants:2011ab}{}
-16. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
-reproducible evaluation of ANTs similarity metric performance in brain
-image registration. Neuroimage. 2011;54(3):2033--44.
+16. Avants, B. B., Tustison, N. J., Song, G., Cook, P. A., Klein, A.,
+and Gee, J. C. ``\textbf{A Reproducible Evaluation of ANTs Similarity
+Metric Performance in Brain Image Registration}'' \emph{Neuroimage} 54,
+no. 3 (2011): 2033--44.
+doi:\href{https://doi.org/10.1016/j.neuroimage.2010.09.025}{10.1016/j.neuroimage.2010.09.025}
 

### 2016-07-07 abstract.Rmd
ENH:  Second round.

@@
 predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
-relative volume difference} = 43 $\pm$ 26\%.  In addition, three
+relative volume difference = 43 $\pm$ 26\%.  In addition, three
 lesion size ranges are selected to illustrate the variation in

### 2016-07-07 discussion.Rmd
ENH:  Second round.

@@
 much longer history of development such as white matter lesion segmentation in
-multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
+multiple sclerosis (MS).  In a recent literature review [@Garcia-Lorenzo:2013aa], 47 papers
+involving  automated segmentation of white
 matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
@@
 The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
+methodology for identifying WMHs within multimodal MR brain imaging.
 This effort utilized information acquired from the manual segmentation
@@
 was also predictive of cognitive decline. Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
+[@Kloppenborg:2014aa] reviewed 23 cross-sectional studies reporting MRI and
 concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
+diagnoses but without previously diagnosed cognitive impairment.  This review found
 that WMHs were associated with cognitive deficit (effect size of -0.10,

### 2016-07-07 discussionX.Rmd
ENH:  Second round.

@@
 much longer history of development such as white matter lesion segmentation in
-multiple sclerosis (MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
+multiple sclerosis (MS).  In a recent literature review [@Garcia-Lorenzo:2013aa], 47 papers
+involving automated segmentation of white
 matter lesions in MS.  The most widely-used data for algorithmic evaluation consists

### 2016-07-07 format.Rmd
ENH:  Second round.

--- 2016-06-29 format.Rmd
@@
 bibliography: references.bib
-csl: council-of-science-editors.csl
+csl: national-science-foundation-grant-proposals.csl
 fontsize: 11pt
@@
 
+<!--
+
 \pagenumbering{gobble}
+
+-->

### 2016-07-07 intro.Rmd
ENH:  Second round.

@@
 normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., [@viola2005]).
+[@viola2005].
 Random forest-based paradigms have been successfully employed in the
@@
 graph where internal nodes represent "tests", or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
+The final, or end, nodes represent the various classifications produced by traversal through the decision tree.
 For the proposed application, individual
@@
 (e.g., AdaBoost) [@schapire1990;@freund1997]
-influenced the similarly sytled conglomeration of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
+influenced the similarly styled conglomeration of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
 Finally, Breiman [@breiman2001]

### 2016-07-07 introX.Rmd
ENH:  Second round.

--- 2016-06-30 stitched.Rmd
@@
----
-output:
-  word_document:
-    fig_caption: true
-  pdf_document:
-    fig_caption: true
-    latex_engine: xelatex
-    keep_tex: yes
-  html_document:
-    toc: false
-
-header-includes:
-   - \usepackage{booktabs}
-   - \usepackage[final]{changes}
-   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.2}
-   - \usepackage[compact]{titlesec}
-   - \usepackage{enumitem}
-   - \usepackage{tikz}
-   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-   - \setlist{nolistsep}
-   - \titlespacing{\section}{2pt}{*0}{*0}
-   - \titlespacing{\subsection}{2pt}{*0}{*0}
-   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
-   - \setlength{\parskip}{3pt}
-   - \setremarkmarkup{(#2)}
-   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
-   - \definechangesauthor[name={James Stone}, color=magenta]{js}
-   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
-   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
-   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
-   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
-   - \definechangesauthor[name={David Tate}, color=orange]{dt}
-bibliography: references.bib
-csl: council-of-science-editors.csl
-fontsize: 11pt
-mainfont: Georgia
-geometry: margin=1.0in
----
-
-<!--
-   - \setlength{\parskip}{3pt}
-   - \setlength{\topsep}{0pt}
-   - \setlength{\partopsep}{0pt}
-   - \setlength{\itemsep}{0pt}
-   - \setlength{\floatsep}{0pt}
-   - \setlength{\intextsep}{2pt}
-   - \setlength{\abovecaptionskip}{2pt}
-   - \setlength{\belowcaptionskip}{0pt}
--->
-
-<style type="text/css">
-body,
-code.bash{
-  font-size: 8px;
-}
-pre {
-  font-size: 8px
-}
-</style>
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set( cache=TRUE )
-```
-
-\pagenumbering{gobble}
-
-# Abstract
-
-White matter hyperintensities (WMHs) are foci of abnormal signal
-intensity in white matter regions seen with magnetic resonance imaging
-(MRI). WMHs are associated with normal aging and have shown prognostic
-value in neurological conditions such as traumatic brain injury (TBI).
-The impracticality of manually quantifying these lesions limits their
-clinical utility and motivates the utilization of machine learning
-techniques for automated segmentation workflows.  Herein, we develop a
-concatenated random forest framework with image features for
-segmenting WMHs in a TBI cohort. The framework is built upon the
-Advanced Normalization Tools (ANTs)  and ANTsR toolkits.  MR
-(3D FLAIR, T2-, and T1-weighted) images from 24 service members and
-veterans scanned in the Chronic Effects of Neurotrauma Consortium’s
-(CENC) observational study were acquired. Manual annotations were
-employed for both training and evaluation using a leave-one-out
-strategy.  \textcolor{blue}{Performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.
-Final average results were: sensitivity = 0.68 $\pm$ 0.38, positive
-predictive value = 0.51 $\pm$ 0.40, $F_1$ = 0.52 $\pm$ 0.36,
-relative volume difference} = 43 $\pm$ 26\%.  In addition, three
-lesion size ranges are selected to illustrate the variation in
-performance with lesion size.}  Paired with
-correlative outcome data, supervised learning methods may allow for
-identification of imaging features predictive of diagnosis and prognosis
-in individual TBI patients.
-
-\clearpage
-
 # Introduction
@@
 normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., [@viola2005]).
+[@viola2005].
 \textcolor{blue}{Random forest-based paradigms have been successfully employed in the
@@
 (e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
+\textcolor{blue}{influenced the similarly styled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
 \textcolor{blue}{Finally, Breiman} [@breiman2001]
@@
 [^i2]: https://github.com/stnava/ANTsR
-
-# Materials and Methods
-
-## Imaging
-
-MR images utilized for this initial report were acquired from a single scanner involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (see Walker et al., this issue).
-Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
-Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
-
-\input{trainingData.tex}
-
-Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
-3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
-
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
-subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All
-multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
-The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
-\textcolor{blue}{
-which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
-
-[^1]: http://nifti.nimh.nih.gov/nifti-1
-
-## Quantitative analysis
-
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1'', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
-
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
-
-### Symmetric multi-modal templates
-
-\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}
-[@landman2011]\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
-
-![Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific asymmetry-based
-features.](Figures/MMRR.png)
-
-### Feature images for WMH segmentation
-
-\textcolor{blue}{
-Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}  For the targeted application in this work,
-tissue classification is performed at the voxelwise level.  In other words, each voxel
-within the region of interest is sent through the ensemble of decision trees and receives
-a set of classification votes from each tree thus permitting a regression or classification
-solution.  Since this procedure is performed at the voxelwise level, intensity information
-alone is insufficient for good segmentation performance due to the lack of spatial context.
-For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated lesion detection
-algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
-even though machine learning and pattern recognition techniques are extremely powerful and
-have significant potential, just as crucial to outcome is the creative construction and
-deployment of salient feature images which we detail below.
-
-\input{featureImagesTable.tex}
-
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.  In Table 2, we provide a list and basic
-categorization of the feature images used for the initial (i.e., Stage 1---more on the use
-of multiple random forest stages below) segmentation of the WMHs.
-In addition Figure 3 provides a representation of a set of feature images for a single
-subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
-with seven labels:
-
-* cerebrospinal fluid (label 1),
-* gray matter (label 2),
-* white matter (label 3),
-* deep gray matter (label 4),
-* brain stem (label 5),
-* cerebellum (label 6), and
-* white matter hyperintensities (label 7).
-
-![Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
-application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
-The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right.  This gives a probabilistic classification of
-tissue type.  Not shown are the probability and voting images for the brain stem and
-cerebellum.](Figures/featureImages.png)
-
-As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
-acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
-the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
-are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
-N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
-range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].
-
-![Sample FLAIR acquisition image slices showing both manual and random forest
-segmentations for both stages obtained during the leave-one-out evaluation.  Manual
-segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.](Figures/sampleResults.png)
-
-The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
-  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
-CSF, gray matter, white matter, deep gray matter, brain stem, and
-cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
-first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 2).
-
-Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
-We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a Manhattan radius of one voxel given the typical size
-of individual WMHs. For each of the preprocessed
-images, we calculate the difference in intensities with the corresponding warped template
-component.  Previous success in the international brain tumor segmentation
-competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 2.  We employ the
-same strategy here.
-
-<!--
-For example, the template
-difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
-
-$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$
-
-where
-
-$$\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$$
-
-is the transform
-which maps from the individual subject space to the
-template space and $T_{FLAIR}$ is the FLAIR template component.
--->
-
-To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
-we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature.
-
-<!--
-For the FLAIR component, this contralateral
-difference image is calculated from
-
-$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$
-
-where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
-the symmetric template.
--->
-
-The segmentation probability images described above are used as feature images to
-provide a spatial context for the random forest model prediction step.  Additional spatial contextual
-feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
-deep gray matter images.  These latter images are intended to help distinguish white matter
-hyperintensities from false positives induced by the partial voluming at the gray/white
-matter interface.  A third set of images are based on the voxel location within the
-space of the template.
-\textcolor{blue}{Similar feature images were used in}
-[@Anbeek:2004aa]
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
-
-### Stacked (concatenated) random forests for improved segmentation performance
-
-In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
-concatenated supervised approach, whereby the prediction output from the first random forest
-model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
-feature images of the training data (as described previously) are used to construct the
-Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
-are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
-additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
-all three aligned preprocessed images for multivariate segmentation during the
-second stage.}
-The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
-
-### Implementation
-
-As pointed out in a recent comprehensive lesion segmentation review
-[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
-is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
-our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
-this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
-ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
-use the _randomForest_ package [@liaw2002] using the default settings with
-2000 trees per model and 500 randomly selected samples per label per image.
-Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.
-
-In addition, similar to our previous offering [^2], we plan on creating
-a self-encapsulated example to showcase the proposed methodology
-\textcolor{blue}{which will also be available on github.}[^3]  The fact that
-the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
-with the manual labelings will facilitate reproducibility on the part of the reader
-as well as any interest in extending the proposed framework to other data sets.
-
-[^2]: https://github.com/ntustison/ANTsAndArboles
-
-[^3]: https://github.com/ntustison/WatchMeHyperventilate
-
-### Evaluation protocol overview
-
-In order to evaluate the protocol described, we performed a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.
-Initial processing included the creation of all Stage 1 feature images for all subjects.
-The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
-The ``truth'' labels are the seven anatomical regions given above.
-
-The leave-one-out procedure is as follows:
-
-* Create Stage 1 feature images for all 24 subjects.
-* For each of the 24 subjects:
-    + sequester the current subject and corresponding feature images.
-    + construct the Stage 1 random forest model from the remaining 23 subjects.
-    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
-    + the previous step produces the Stage 1 voting maps for all seven labels.
-    + for each of the 23 subjects, perform a Bayesian-based segmentation with an
-      MRF spatial
-      prior using the seven voting maps as additional tissue priors.
-    + construct the Stage 2 random forest model from all the Stage 1 feature images,
-      seven voting maps, and seven posterior probability maps from the previous step.
-    + send the sequestered subject through the random forest models for both stages.
-    + compare the final results with the manually-defined white matter hyperintensity
-      regions.
-
-\textcolor{blue}{Several measures have been employed in the literature for evaluating
-automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
-For this work, due to the relatively small-size distribution} \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
-four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
-incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
-The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
-
-$$ sensitivity =  \frac{ TP }{TP + FN}, $$
-
-\textcolor{blue}{and the positive predictive value (PPV),}
-
-$$ PPV =  \frac{ TP }{TP + FP}, $$
-
-\textcolor{blue}{such that $F_1$ is given by}
-
-$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
-
-\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
-the manual and predicted lesion volumes:}
-
-$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
-
-\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
-
-* \textcolor{blue}{Quantile 1: $[1-12)$,}
-
-* \textcolor{blue}{Quantile 2: $[12-28)$, and}
-
-* \textcolor{blue}{Quantile 3: $[28-551]$}.
-
-\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
-
-# Results
-
-## White matter hyperintensity segmentation evaluation
-
-\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data over the three lesion volume ranges.
-These performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed are meant to illustrate the variation
-in performance with lesion size.}
-
-\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity for this range is more varied compared with the largest
-set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
-the volumetric estimate when comparing the manually identified lesions versus the
-automatically predicted lesions.}
-
-\textcolor{blue}{
-We averaged these measures over all lesions and all subjects which resulted in the
-following values:}
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
-
-    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
-
-    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
-
-    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
-
-    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
-
-<!--
-
-In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-On the left of Figure 5 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
--->
-
-![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
-section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
-volume difference.   These quantitative assessments are given for three quantile ranges
-spanning the range of the manually-derived lesion volumes.    Overall improvement in all
-three whole lesion-based measuers is seen as the
-second Stage RF model is applied for all three quantile ranges.
-The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
-
-## Ranking feature importance
-
-After performing the leave-one-out evaluation,
-we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
-2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
-during the out-of-bag phase of the random forest model construction and quantifies the decrease
-in prediction accuracy from omitting the specified feature.  In other words, this quantity
-helps determine the importance of a particular feature and, although we save such
-efforts for future work, this information provides us with guidance for future feature
-pruning and/or additions.
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 1 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile
- and illustrate the
-stability of the feature importance across the leave-one-out models.
-At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile and illustrate the
-stability of the feature importance across the leave-one-out models.
-We augment the 31 feature images from the first stage by adding an additional
-7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.](Figures/averageLeaveOneOutStage2.png)
-
-The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the
-separate stages are averaged over the entire corresponding model set.  In addition, we
-track the variance for each feature over all models to illustrate the stability of
-the chosen features during the evaluation.  This latter information is illustrated as
-horizontal errors bars providing the $95^{th}$ percentile
-Note that the reader can cross reference Table 1 for identifying corresponding feature types
-and names.
-
-# Discussion
-
-<!-- Nick could you add some detail here on how this technique compares
-to other automated or semi automated approaches for WMH detection? -->
-
-\textcolor{blue}{
-In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
-However, we can extrapolate from similar application areas where such methodologies have a
-much longer history of development such as white matter lesion segmentation in
-multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
-matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
-of two cohorts from two different sites used in
-the MS Lesion segmentation challenge associated with the international MICCAI 2008
-conference.  For comparison, this challenge training data set
-has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
-our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
-40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
-almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
-Note that our relative volume difference numbers are comparable to the
-relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
-[@vanginneken:2007aa].
-
-Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
-The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
-FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
-posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives.  Inspection of the bottom of the
-plots demonstrates the lack of discriminating features associated with the
-T1 image which is also well-known in the clinical literature.
-
-![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
-As described earlier, for Stage 2, we used the output random forest voting maps from Stage
-1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
-"_RFStage1VotingMaps_" where the final numeral is associated with the brain
-parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
-The Stage 2 feature importance plot follows similar
-trends as that for Stage 1 with the T1 images not contributing much to the identification
-of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
-important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
-3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
-alone it is intuitive that such features would be important.
-
-<!--
-
-* Evaluation * Proper context * We only have manual labelings from a
-single radiologist ( * TBI white matter hyperintensities are more
-difficult to segment than MS lesions * individual lesions are smaller *
-lesion load is typically smaller * not prone to enhancement * less
-distinct from surrounding tissue ? * Despite the doubts regarding the
-efficacy of a typical segmentation validation, our contribution is
-useful because * the total segmentation framework is publicly available
-within the ANTs/ANTsR framework * although only a single site is
-analyzed, the feature images are site-agnostic * we plan to apply the
-current RF models directly to the other site data * we can then build
-the training database further by enlisting other experts * Is this the
-first work exploring segmentation of white matter hyperintensities in
-TBI? * The data (including manual tracings) will be made available as
-part of the FITBIR effort.
-
--->
-
-# Conclusions
-
-The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation
-of WMHs from FLAIR images to help build two-stage ensembles of decision
-trees for the automated identification of these lesions. Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology.  Also, we
-recognize that only a single site was used for evaluating the proposed
-framework.  However, we are currently processing other site data with
-the models developed for this work and the results look promising since
-the developed features are site-agnostic.
-
-As far as we know, this is the first report utilizing a novel random
-forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
-tend to be more difficult to segment than MS lesions as the former tend
-to be smaller with an overall smaller lesion load.  Also, enhancement
-protocols with the former tend to be less successful than with the
-latter.  As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.
-
-<!--
-Our
-framework is also available as open-source as part of well-known
-neuroimaging tools which easily allows for additions/extensions but is
-also, as far as we know, the first random forest-based technique
-available for such application.
--->
-
-Two major meta-analyses of WMHs have been published covering the periods
-prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
-Debette & Markus [@Debette:2010aa] found that the presence of
-WMHs was related to subsequent cognitive decline, a higher risk of
-developing dementia, stroke, and of mortality. Lesion volume at baseline
-was also predictive of cognitive decline. Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
-concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
-that WMHs were associated with cognitive deficit (effect size of -0.10,
-95% CI: -0.13 to -0.08) after controlling for age.
-
-Despite the potential clinical significance of WMHs
-these lesions receive little
-attention in current clinical workflows. When reported in a standard
-neuroradiologist interpretation, they are typically handled as
-incidental findings and are assigned little clinical significance. This
-likely reflects the impracticality of performing a detailed assessment
-of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
-supervised statistical learning tool that is appropriate for the
-application to such large-scale datasets.
-
-<!--
-
-The currently described tool is just one example of how "supervised
-learning" algorithms might be applied to aid in the diagnosis of TBI and
-other disease processes through the specific identification of features
-predictive of a given disease state. It is an important demonstration of
-the potential power of these analytical approaches in the rapid but
-comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide
-variety of settings for the rapid identification of predictive imaging
-features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
-manufacturers utilize these types of approaches to equip self-driving
-vehicles to recognize and respond to unique external surroundings
-through the identification of visual information sufficiently similar to
-previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning
-approaches may allow for the rapid identification of information
-predictive of disease state in an individual patient. These approaches
-have been applied to the segmentation of macroscopically visible
-structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of
-imaging data in the individual patient with a primary quantitative
-output metrics to include sequences such as diffusion tensor imaging
-(DTI) and its variants, functional connectivity, perfusion weighted
-imaging, and cortical thickness assessments. At present, these advanced
-neuroimaging sequences are confined to cohort-based research studies due
-to the lack of available analytical tools to assess the information in
-the setting of the individual patient [@Mayer:2014aa]. Application of
-deep learning approaches in the context of data with primary
-quantitative outputs will require large scale normative and disease
-specific databases. Building these large scale imaging libraries is
-resource intensive and requires a multi-center approach with harmonized
-scanners between sites and correlative non-imaging clinical data. Large
-scale TBI data is becoming increasingly available through activities
-such as the Chronic Effects of Neurotrauma Consortium (CENC),
-Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
-Collaborative European Neurotrauma Effectiveness Research in TBI
-(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
-Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data,
-deep learning algorithms may be well positioned to help transform how
-neuroimaging is interpreted for the clinical management of patients with
-this disease process.
-
--->
-
-\clearpage
-
-## Acknowledgements
-
-The authors wish to acknowledge all other members of
-the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
-X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
-also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
-and Kevin Wilson. We would also like to thank the study participants and
-their family members. This project was jointly supported by the
-Department of Defense (W81XWH-13-2-0095), the U.S. Department of
-Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
-Grant HU 0001-08-0001.
-
-## Declaration of Interest/Disclaimer
-
-The authors report no financial
-disclosures or conflicts of interest. The views expressed here are those
-of the authors and do not necessarily reflect the official policy of
-position of the Department of the Navy, Department of Defense, nor the
-U.S. Government. This work was prepared as a part of official duties;
-Title 17 USC §105 provides that Copyright protection under this title is
-not available for any work of the U.S. Government. Title 17 USC §101
-defines a US Government work as a work prepared by a military service
-member of employee of the US Government as part of that person’s
-official duties.
-
-\clearpage
-
-<!-- \listofchanges
-
-\clearpage -->
-
-# References

### 2016-07-07 methods.Rmd
BUG:  Switch subject ids.
ENH:  Second round.

@@
 The feature images were derived from MR acquisitions of 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
+Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue.
 All lesions were isolated in the white matter of the cerebrum.
@@
 3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
+3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
+In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ETL: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
 
@@
 
-![Representation of Stage 1 feature images for subject 01C1019.  The
+![Representation of Stage 1 feature images for subject 9.  The
 FLAIR, T1-, and T2-weighted images are rigidly pre-aligned

### 2016-07-07 methodsX.Rmd
BUG:  Switch subject ids.
ENH:  Second round.

@@
 The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
+Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue.
 \textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.

### 2016-12-02 responseToReviewers2.Rmd
ENH:  Final commit.

---
output:
  word_document:
    fig_caption: true
  pdf_document:
    fig_caption: true
    latex_engine: xelatex
    keep_tex: yes
header-includes:
   - \usepackage{booktabs}
   - \usepackage[final]{changes}
   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
   - \linespread{1.2}
   - \usepackage[compact]{titlesec}
   - \usepackage{enumitem}
   - \usepackage{tikz}
   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
   - \setlist{nolistsep}
   - \titlespacing{\section}{2pt}{*0}{*0}
   - \titlespacing{\subsection}{2pt}{*0}{*0}
   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
   - \setlength{\parskip}{3pt}
   - \setremarkmarkup{(#2)}
bibliography: references.bib
csl: national-science-foundation-grant-proposals.csl
fontsize: 12pt
mainfont: Georgia
---

<style type="text/css">
body,
code.bash{
  font-size: 8px;
}
pre {
  font-size: 8px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set( cache=TRUE )
```

# Response to reviewers (Round 2)

_\textcolor{blue}{We appreciate the additional time spent by the editors and
reviewers in evaluating the revised version of our manuscript.
Please see below for a point-by-point response to the minor issues raised
during this round.}_

## Reviewer 1

The revised manuscript is significantly improved and much clearer. The
authors have addressed all my comments.

A couple of minor notes, Figures 7 and 8 seem to have the descriptions swapped.

> _\textcolor{blue}{The figure captions had been incorrectly assigned during the
manuscript upload process which has since been corrected.}_

Also, Table 1 shows Subject 2 with 0 lesions. Was there a
specific reason for including this subject in the study?

> _\textcolor{blue}{Yes.  We included all usable data from the source study including
the subject with 0 lesions.  In future work, we imagine using this tool in a typical
workflow which would include subjects without lesions and believed that an evaluation would
best be conducted if we included as much representative data as possible.
}_

## Reviewer: 2

After reviewing the revised manuscript I believe it can now be accepted.  Two minor issues:
pp20-21:  The formulas have errors i.e. false positive (FP) versus true negative (TN).

> _\textcolor{blue}{We assume that the reviewer is referring to the definitions for
sensitivity and PPV on pp. 20--21.  We consulted a number of sources and do not believe
the given equations are incorrect.  Perhaps the reviewer is referring to the previously
incomplete text given previous to the definition of sensitivity. This has since been
repaired in the revised manuscript.}_

Also, the legends for figures 7 and 8 are swapped.

> _\textcolor{blue}{The figure captions had been incorrectly assigned during the
manuscript upload process which has since been corrected.}_

## Reviewer: 3

Regarding the “imaging" methodology, 3rd para, authors need to remove
the information about JRS, and instead just indicate ‘how’ the work was
performed, since authorship already implies these people did the work.
For example, instead of "The first author (J. R. S.) performed the
manual WMH tracings…” it probably should read something like “Manual WMH
tracings were performed….”.

> _\textcolor{blue}{We believe the current text provides important precision in specifying
how the manual tracings were performed, specifically in terms of qualifications and experience
of the person(s) involved.    Additionally, authorship does not necessarily imply performance of the manual tracings as
there are numerous labeled image data sets which have been made publicly available and used
by third-parties in the development of methodologies.}_

### 2016-12-02 responseToReviewers2.tex
ENH:  Final commit.

--- 2016-06-30 responseToReviewers.tex
@@
 
-\section{Response to reviewers}\label{response-to-reviewers}
+\section{Response to reviewers (Round
+2)}\label{response-to-reviewers-round-2}
 
-\emph{\textcolor{blue}{We appreciate the time spent by the editors and reviewers in
-assessing our manuscript.  Please see below for a point-by-point response to the
-issues raised.}}
+\emph{\textcolor{blue}{We appreciate the additional time spent by the editors and
+reviewers in evaluating the revised version of our manuscript.
+Please see below for a point-by-point response to the minor issues raised
+during this round.}}
 
@@
 
-Thank you for the opportunity to review the paper ``Supervised learning
-technique for the automated identification of white matter
-hyperintensities in traumatic brain injury''. In this study, the authors
-present a machine learning algorithm for automated segmentation of white
-matter hyperintensities, a form of white matter degeneration visible in
-FLAIR-MRI that is observed in a wide range of clinical syndromes, and
-even in healthy controls. The method proposed by the authors relies on
-training random forests to learn the relationship between voxel
-intensities in a set of feature images derived from multimodal MRI (T1,
-T2, FLAIR) and lesions drawn manually by an expert. Since the method
-uses examples to learn what constitutes a lesion, this is correctly
-considered a supervised method. The method is applied on a set of 24
-patients with clinically confirmed traumatic brain injury, and results
-are obtained with a leave-one-out cross validation.
+The revised manuscript is significantly improved and much clearer. The
+authors have addressed all my comments.
 
-Overall, the method proposed by the authors is innovative and
-thoughtful, and has great potential for bridging the gap between
-research findings and clinical needs. The method uses a series of
-cutting edge solutions to the segmentation problem: (1) multimodal
-imaging inputs, which is known to increase segmentation accuracy, (2)
-random forests, which are an excellent tool for finding multivariate
-nonlinear relationships, (3) voxel neighborhood information, which
-provide contextual information for each voxel, and (4) an algorithm that
-improves the segmentation in stages. However, the manuscript itself, to
-my opinion, is not well focused and needs significant improvement. Below
-are my specific comments in support of this.
+A couple of minor notes, Figures 7 and 8 seem to have the descriptions
+swapped.
 
 \begin{quote}
-\emph{\textcolor{blue}{We thank the reviewer for the generally positive assessment of our work.  We also
-appreciate the detailed critique below to help us improve specific aspects of the
-presentation.}}
+\emph{\textcolor{blue}{The figure captions had been incorrectly assigned during the
+manuscript upload process which has since been corrected.}}
 \end{quote}
 
-Major comments:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  The biggest issue I had while reading the manuscript is the lack of a
-  coherent theme of what this paper is about. The authors span their
-  exposé from multiple sclerosis to the automobile industry, without
-  stating why these links to other diseases (or industries) have any
-  relevance. It appears in some parts that the closest disease with
-  similar white matter lesions is multiple sclerosis, but this is not
-  clearly stated, nor the two types of lesions are compared. It is,
-  therefore, unclear why counting how many MS lesion segmentation
-  methods are publicly available is of any importance. When explaining
-  methods, the authors mention the utility of support vector machines,
-  or the existence of a python package called SciPy. Yet, it is unclear
-  why these things have any importance, unless the authors make
-  connections to their topic/method.
-\end{enumerate}
+Also, Table 1 shows Subject 2 with 0 lesions. Was there a specific
+reason for including this subject in the study?
 
 \begin{quote}
-\emph{\textcolor{blue}{We reduced the introduction to limit the clinical motivation to a
-TBI context and removed mention of multiple sclerosis to remove any assumption of
-disease similarity (although the segmentation approaches can be similar).  We also
-removed the paragraph dealing with publicly available software for neuroimaging analysis.
+\emph{\textcolor{blue}{Yes.  We included all usable data from the source study including
+the subject with 0 lesions.  In future work, we imagine using this tool in a typical
+workflow which would include subjects without lesions and believed that an evaluation would
+best be conducted if we included as much representative data as possible.
 }}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Second, the manuscript has no results. All the results section is
-  focused on talking about feature ranking, which by the way is
-  explained in the Results section instead of the Methods section. The
-  most important results (the prediction accuracy of automated
-  segmentation), are not mentioned in the manuscript but are only
-  plotted in Figure 7. My advice is to put the focus of the results
-  section on accuracy, i.e., give precise dice values and perhaps add
-  other measures (metric displacement, Hausdorf distance, sensitivity,
-  specificity, etc.). Also, the gain in accuracy from stage 1 to stage 2
-  cannot be guessed by looking at a plot, the authors must provide a
-  statistical comparison. When investigating accuracy, the error in
-  lesion size should be reported, both in terms of volume and in terms
-  of individual lesions counts (i.e.~out of 10 lesions 9 were found, 1
-  was missed, etc.). This information is crucial to clinicians, if the
-  authors foresee the use of the method in a clinical setting, as it
-  seems from the discussion. Beside adding results to the manuscript, it
-  is equally important to give some results in the abstract. The
-  abstract must be self-sufficient in reporting all the major points of
-  the paper.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We completely reorganized the results and discussion
-section to reflect the new evaluation.  In addition, we added a synopsis
-of these results to the abstract.}}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  Third, there is no information about the manual lesion tracing
-  procedure. Were lesion drawn on FLAIR or were other modalities used as
-  well? If FLAIR was used as reference it is not surprising that FLAIR
-  emerges as one of the top predictive modalities.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We added the following to the Methods section:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{The first author (J.
-R.
-S.) performed the manual WMH tracings for all 26  subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All multi-modal MR dicom image slices were converted to the nifti file format.
-All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software}\textsuperscript{1}
-\textcolor{blue}{.  The normalized MRI volumes were then provided to J. R. S. who traced each lesion using the ITK-SNAP tool}\textsuperscript{2}
-\textcolor{blue}{ which has multi-image overlay capabilities for visualizing all modalities in all three canonical views.}
-\end{quote}
-
-\begin{quote}
-Descriptive statistics on the lesions are also necessary: how big were
-the lesions, how many lesions were found in each patient (the range 1-20
-is not sufficiently informative), where were the lesions located? This
-information helps understand why, for example, brain stem or cerebellum
-labels were not that important for achieving good lesion predictions.
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{We added the following sentence to the opening paragraph of the
-Imaging subsection of the Methods section:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
-Table 1 provides a descriptive statistical summary of the variation in lesion load across the selected cohort.}
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{as well as the corresponding table:}}
-\end{quote}
-
-\input{trainingData.tex}
-
-\clearpage
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
-\item
-  Fourth, although the authors claim the method is available, I couldn't
-  find any link for the public. Note, it would be equally important if
-  the publication of the method online is associated with proper
-  documentation on how a new (naive) user can apply it.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{Our methodological publication history demonstrates that we are
-very supportive of open science and have released both code and data in conjunction
-with these publications in public forums, such as github.  However, permission to
-publicly release the data for this data has not yet been obtained and so we removed
-the mention of public availability.  However, when we do get permission, it will
-be posted to the github repository associated with this manuscript
-(https://github.com/ntustison/WatchMeHyperventilate), which we mention in the text,
-for future readers.
-}}
-\end{quote}
-
-Minor comments:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
-\tightlist
-\item
-  In the abstract, the authors mention the creation of tailored features
-  for segmenting WMHs. My understanding is that the method proposed by
-  the authors has no tailoring (if we think of tailoring as custom
-  cutting to WMH needs).
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We removed the word ``tailor'' in its various forms in the
-manuscript to avoid the misreading described by the reviewer.}}
-\end{quote}
-
-\begin{quote}
-There was no feature selection process all the features were included as
-predictors.
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{Although feature selection is mentioned in the manuscript, nowhere
-do we mention or imply any type of computational feature selection.  There was
-a feature selection process in that we selected ones from the myriad feature images that
-one can possibly produce based on our targeted application, i.e. WMHs.}}
-\end{quote}
-
-\begin{quote}
-A similar pitfall if found on pg. 6 ln 56, where the manuscript reads
-``Crucial to these supervised segmentation approaches are the creation
-and selection of features as input in conjunction with expertly
-identified structures.'' What are the structures identified by the
-expert in this study?
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{For this particular application, white matter hyperintensities
-are the structure of interest.  We have re-emphasized this point by writing:
-}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{5}
-\tightlist
-\item
-  The methods section is very confusing. The section ``Feature images
-  for WMH segmentation'' starts with no explanation of what is Stage 1.
-  The logical flow would require template registration explained first,
-  and all feature computation performed later, and finally RF stages can
-  be explained with the available features.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We added Figure 1 which gives an overview of the basic workflow.
-  It can be seen below in response to item 2 of Reviewer 2.  We reference
-  this figure numerous times in the text as a visual explanatory aid.}}
-\emph{\textcolor{blue}{We also added a paragraph at the beginning of the "Quantitative analysis" subsection
-  which gives an overview of the methods section including an description of Stage 1 and Stage2.}}
-
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1 '', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
-
-\emph{\textcolor{blue}{We also added a subsection explaining the symmetric multimodal template:}}
-
-\textcolor{blue}{Following}\textsuperscript{3}
-\textcolor{blue}{and}\textsuperscript{4},
-\textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}\textsuperscript{5}\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in}\textsuperscript{6}
-\textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
-
-\emph{\textcolor{blue}{These changes complement the edits made in response to other review items.}}
-\end{quote}
-
-\begin{quote}
-The pipeline also seem to be repetitive, the N4 correction is mentioned
-in ln. 26 and then again in ln. 43, which sounds like it was performed
-twice. Tissue segmentation is mentioned in ln. 39, then it is stated is
-a Bayesian method (ln 45), and then it is referred to as Atropos in the
-next page.
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{The two mentions of N4 are due to the fact that they are associated
-with two distinct pipelines:  1) preparation of the images for extracting feature values
-and 2) brain extration and $n$-tissue segmentation.  We rewrote the section in question
-to avoid any confusing redundancy:}}
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines}}\textsuperscript{3}
-\emph{\textcolor{blue}{.  The result is a mask delineating the brain
-parenchyma and probabilistic estimates of the CSF, gray matter, white matter,
-deep gray matter, brain stem, and cerebellum}}\textsuperscript{7}
-\emph{\textcolor{blue}{.  These provide the expertly annotated labels for the
-first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates}}\textsuperscript{4}
-\emph{\textcolor{blue}{created from the public MMRR data set}}\textsuperscript{5}
-\emph{\textcolor{blue}{(cf Figure 3).}}
-\end{quote}
-
-\begin{quote}
-It is unclear what is ``the first set of images'' in pg. 9, or if there
-is a second set.
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{We clarified this confusion by changing the sentence in question from}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{To model the intensity information the first set of images simply includes the
-preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.}
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{to}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.}
-\end{quote}
-
-\begin{quote}
-Also, mathematical formulas seem superfluous, particularly since it
-takes more space to explain the formula than to say in plain English
-that transformation matrices were applied to bring subject's images in
-template space.
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{Done.  We removed the mathematical formulas.}}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{6}
-\tightlist
-\item
-  Random forests are explained in a way that can be understood only by
-  readers who already know how they work; a new reader (especially a
-  clinician) would not be able to understand the method. For example,
-  data pushed onto a tree is does not mean much, unless the reader knows
-  what is an identification tree.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{
-We revised the explanation of random forests as follows:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
-For the proposed application, individual
-voxels (and their corresponding feature values) are introduced at the root of a particular
-decision tree and traverse the edges and internal nodes ultimately ending up at one of
-the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)}\textsuperscript{8,9}
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization}\textsuperscript{10,11}.
-\textcolor{blue}{Finally, Breiman}\textsuperscript{12}
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{7}
-\tightlist
-\item
-  The relationship between WMH and GOS-E is mentioned in pg. 2 but the
-  reader is not informed what is GOS-E. A similar issue is met few lines
-  after, where the ``outcome'' is mentioned, but there is no information
-  what is referred to as outcome.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{The sentence in question reads:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores.}\textsuperscript{13}
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{Although the full phrase corresponding to the acronym
-in question is provided prior to the acronym in the text (along with an
-explanatory reference), we have added the following clarification:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores}\textsuperscript{13}
-\textcolor{blue}{---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery.}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{8}
-\tightlist
-\item
-  In pg. 3 the authors propose that automated methods ``may allow for
-  identification of correlative patterns between WMH number, volume,
-  distribution, and disease state''. Some of these correlations have
-  been already identified in the literature, as the authors mention few
-  lines before. Moreover, the authors do not report number, volume, or
-  distribution of the lesions, to back up the utility of their method in
-  this regard.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We removed the sentence in question.  Also, as mentioned in
-response to item 3, we added Table 1.}}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{9}
-\tightlist
-\item
-  In pg. 4 ln 43, reads that the data is ``constructed.'' I don't think
-  this is the right verb.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We agree that the sentence in question could be read this way.
-We have, therefore, clarified the sentence in question to read
-``Once the ensemble of decision trees is constructed, data to be classified is ``pushed''
-through each decision tree resulting in a single classification ``vote'' per tree.''}}
-\end{quote}
-
-\begin{quote}
-A similar issue is in pg 5 ln 52 where it is stated that ``feature
-images consisted of 26 subjects.'' It is not ethical to consider
-participants as feature images.
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{This was a grammatical mistake, not a question of
-ethics.  It has been fixed: ``The feature images
-were derived from MR acquisitions of 26 subjects\ldots''}}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{10}
-\tightlist
-\item
-  It is unclear to me the meaning of the sentence: ``Since the inverse
-  transform is also derived as part of the registration process, we can
-  warp the voxel index locations back to the space of the individual
-  subject which motivates similar work by others {[}43{]}.''
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{The sentences were rewritten as follows:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{Similar feature images were used in}\textsuperscript{14}
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant}\textsuperscript{15}
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm}\textsuperscript{16}.
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{11}
-\tightlist
-\item
-  Why was Stage 1 segmentation based only on T1, while Stage 2
-  segmentation was based on multivariate segmentation using all three
-  modalities?
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We added the following clarification:}}
-\end{quote}
-
-\begin{quote}
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation
-process following the voxelwise RF classification of Stage 1, we use all three
-aligned preprocessed images for multivariate segmentation during the second stage.}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{12}
-\tightlist
-\item
-  Similar to the feature ranking method, which is explain in Results,
-  Dice is also explained in Results.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We removed the equation for the Dice metric and described the
-evaluation procedure in the Methods section.}}
-\end{quote}
-
-\begin{quote}
-Also, it is unclear why finding which direction of space is informative
-for the segmentation process would be a ``sanity check.''
-\end{quote}
-
-\begin{quote}
-\emph{\textcolor{blue}{We removed this section to avoid unnecessary confusion.}}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{13}
-\tightlist
-\item
-  The discussion starts with an ``intent to further refine the proposed
-  paradigm,'' and proceeds by mentioning other work currently being
-  performed which looks promising. The opening of the discussion with
-  such statements truly make the manuscript sound like an incomplete
-  effort that is not ready for publication.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We appreciate the positive view of our current work.
-The full sentence in question is ``Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology.''
-In our opinion, simply mentioning specific details regarding future work
-does not mean that the manuscript ``sounds like an incomplete effort.''
-Rather, it is simply an admission, similar to much of what is reported
-in the literature, that our research program is ongoing.}}
-\end{quote}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{14}
-\tightlist
-\item
-  The discussion also spends several paragraphs on defining literature
-  that shows how WMHs are related to cognitive performance, or how WMHs
-  can be applied in diagnosing TBI. Yet, this study makes no effort to
-  test these efforts in these 24 patients. Consequently, those
-  paragraphs sound wishful and speculative. Given the methodological
-  nature of this paper, a more appropriate discussion would perhaps
-  focus on the comparison with other methods in terms of accuracy,
-  similarities, differences, etc.
-\end{enumerate}
-
-\begin{quote}
-\emph{\textcolor{blue}{We removed the paragraphs in question and added a
-comparison discussion.}}
 \end{quote}
@@
 
-Interesting paper which offers those without a current background in
-state-of-the-art brain analysis tools. A good introduction to their
-capabilities.
+After reviewing the revised manuscript I believe it can now be accepted.
+Two minor issues: pp20-21: The formulas have errors i.e.~false positive
+(FP) versus true negative (TN).
 
 \begin{quote}
-\emph{\textcolor{blue}{We thank the reviewer for the positive comments regarding our work.}}
+\emph{\textcolor{blue}{We assume that the reviewer is referring to the definitions for
+sensitivity and PPV on pp. 20--21.  We consulted a number of sources and do not believe
+the given equations are incorrect.  Perhaps the reviewer is referring to the previously
+incomplete text given previous to the definition of sensitivity. This has since been
+repaired in the revised manuscript.}}
 \end{quote}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  However the authors do little to establish clinical utility of the
-  algorithm.
-\end{enumerate}
+Also, the legends for figures 7 and 8 are swapped.
 
 \begin{quote}
-\emph{\textcolor{blue}{We appreciate
-the opportunity to clarify the precise nature of our contribution which is methodological
-(vs. clinical).  We have shortened the introduction to focus specifically on previous TBI
-clinical research involving WMHs and then pointing to the need for automated
-techniques to minimize the variability and time requirements of manual
-segmentation of WMHs.  This would allow for large-scale studies and potentially more accurate
-and reproducible quantitation.
-}}
+\emph{\textcolor{blue}{The figure captions had been incorrectly assigned during the
+manuscript upload process which has since been corrected.}}
 \end{quote}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  A graphical description of the pipeline would help most readers to
-  understand how this might get into the clinic.
-\end{enumerate}
+\subsection{Reviewer: 3}\label{reviewer-3}
+
+Regarding the ``imaging'' methodology, 3rd para, authors need to remove
+the information about JRS, and instead just indicate `how' the work was
+performed, since authorship already implies these people did the work.
+For example, instead of ``The first author (J. R. S.) performed the
+manual WMH tracings\ldots{}'' it probably should read something like
+``Manual WMH tracings were performed\ldots{}.''.
 
 \begin{quote}
-\emph{\textcolor{blue}{Done.  We added the following figure as a graphical description of
-the pipeline to the manuscript:}}
+\emph{\textcolor{blue}{We believe the current text provides important precision in specifying
+how the manual tracings were performed, specifically in terms of qualifications and experience
+of the person(s) involved.    Additionally, authorship does not necessarily imply performance of the manual tracings as
+there are numerous labeled image data sets which have been made publicly available and used
+by third-parties in the development of methodologies.}}
 \end{quote}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figures/wmhPipeline.png}
-\caption{\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}}
-\end{figure}
-
-\clearpage
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
-
 \hypertarget{refs}{}
-\hypertarget{ref-Avants:2014aa}{}
-1. Avants BB, Tustison NJ, Stauffer M, Song G, Wu B, Gee JC. The Insight
-ToolKit image registration framework. Front Neuroinform. 2014;8:44.
-
-\hypertarget{ref-Yushkevich:2006aa}{}
-2. Yushkevich PA, Piven J, Hazlett HC, Smith RG, Ho S, Gee JC, Gerig G.
-User-guided 3D active contour segmentation of anatomical structures:
-Significantly improved efficiency and reliability. Neuroimage.
-2006;31(3):1116--28.
-
-\hypertarget{ref-Tustison:2014ab}{}
-3. Tustison NJ, Cook PA, Klein A, Song G, Das SR, Duda JT, Kandel BM,
-Strien N van, Stone JR, Gee JC, et al. Large-scale evaluation of aNTs
-and freeSurfer cortical thickness measurements. Neuroimage.
-2014;99:166--79.
-
-\hypertarget{ref-Tustison:2015aa}{}
-4. Tustison NJ, Shrinidhi KL, Wintermark M, Durst CR, Kandel BM, Gee JC,
-Grossman MC, Avants BB. Optimal symmetric multimodal templates and
-concatenated random forests for supervised brain tumor segmentation
-(simplified) with aNTsR. Neuroinformatics. 2015;13(2):209--25.
-
-\hypertarget{ref-landman2011}{}
-5. Landman BA, Huang AJ, Gifford A, Vikram DS, Lim IAL, Farrell JAD,
-Bogovic JA, Hua J, Chen M, Jarso S, et al. Multi-parametric neuroimaging
-reproducibility: A 3-T resource study. Neuroimage. 2011;54(4):2854--66.
-
-\hypertarget{ref-Avants:2010aa}{}
-6. Avants BB, Yushkevich P, Pluta J, Minkoff D, Korczykowski M, Detre J,
-Gee JC. The optimal template effect in hippocampus studies of diseased
-populations. Neuroimage. 2010;49(3):2457--66.
-
-\hypertarget{ref-Avants:2011aa}{}
-7. Avants BB, Tustison NJ, Wu J, Cook PA, Gee JC. An open source
-multivariate framework for \(n\)-tissue segmentation with evaluation on
-public data. Neuroinformatics. 2011;9(4):381--400.
-
-\hypertarget{ref-schapire1990}{}
-8. Schapire R. The strength of weak learnability. Machine Learning.
-1990;5:197--227.
-
-\hypertarget{ref-freund1997}{}
-9. Freund Y, Schapire R. A decision-theoretic generalization of on-line
-learning and an application to boosting. Journal of Computer and System
-Sciences. 1997;55:119--139.
-
-\hypertarget{ref-ho1995}{}
-10. Ho TK. Random decision forests. In: Document analysis and
-recognition, 1995., proceedings of the third international conference
-on. Vol. 1. 1995. pp. 278--282 vol.1.
-
-\hypertarget{ref-amit1997}{}
-11. Amit Y, Geman D. Shape quantization and recognition with randomized
-trees. Neural Computation. 1997;9:1545--1588.
-
-\hypertarget{ref-breiman2001}{}
-12. Breiman L. Random forests. In: Machine learning. 2001. pp. 5--32.
-
-\hypertarget{ref-Moen:2014aa}{}
-13. Moen KG, Brezova V, Skandsen T, Håberg AK, Folvik M, Vik A.
-Traumatic axonal injury: The prognostic value of lesion load in corpus
-callosum, brain stem, and thalamus in different magnetic resonance
-imaging sequences. J Neurotrauma. 2014;31(17):1486--96.
-
-\hypertarget{ref-Anbeek:2004aa}{}
-14. Anbeek P, Vincken KL, Osch MJP van, Bisschops RHC, Grond J van der.
-Probabilistic segmentation of white matter lesions in mR imaging.
-Neuroimage. 2004;21(3):1037--44.
-
-\hypertarget{ref-Tustison:2013ac}{}
-15. Tustison NJ, Avants BB. Explicit B-spline regularization in
-diffeomorphic image registration. Front Neuroinform. 2013;7:39.
-
-\hypertarget{ref-Avants:2011ab}{}
-16. Avants BB, Tustison NJ, Song G, Cook PA, Klein A, Gee JC. A
-reproducible evaluation of ANTs similarity metric performance in brain
-image registration. Neuroimage. 2011;54(3):2033--44.
 

### 2016-12-02 abstractX.Rmd
ENH:  Final commit.

--- 2016-06-30 stitched.Rmd
@@
----
-output:
-  word_document:
-    fig_caption: true
-  pdf_document:
-    fig_caption: true
-    latex_engine: xelatex
-    keep_tex: yes
-  html_document:
-    toc: false
-
-header-includes:
-   - \usepackage{booktabs}
-   - \usepackage[final]{changes}
-   - \usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
-   - \linespread{1.2}
-   - \usepackage[compact]{titlesec}
-   - \usepackage{enumitem}
-   - \usepackage{tikz}
-   - \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
-   - \setlist{nolistsep}
-   - \titlespacing{\section}{2pt}{*0}{*0}
-   - \titlespacing{\subsection}{2pt}{*0}{*0}
-   - \titlespacing{\subsubsection}{2pt}{*0}{*0}
-   - \setlength{\parskip}{3pt}
-   - \setremarkmarkup{(#2)}
-   - \definechangesauthor[name={Nick Tustison}, color=red]{nt}
-   - \definechangesauthor[name={James Stone}, color=magenta]{js}
-   - \definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
-   - \definechangesauthor[name={Andy Mayer}, color=green]{am}
-   - \definechangesauthor[name={Harvey Levin}, color=blue]{hl}
-   - \definechangesauthor[name={Brian Taylor}, color=purple]{bt}
-   - \definechangesauthor[name={David Tate}, color=orange]{dt}
-bibliography: references.bib
-csl: council-of-science-editors.csl
-fontsize: 11pt
-mainfont: Georgia
-geometry: margin=1.0in
----
-
-<!--
-   - \setlength{\parskip}{3pt}
-   - \setlength{\topsep}{0pt}
-   - \setlength{\partopsep}{0pt}
-   - \setlength{\itemsep}{0pt}
-   - \setlength{\floatsep}{0pt}
-   - \setlength{\intextsep}{2pt}
-   - \setlength{\abovecaptionskip}{2pt}
-   - \setlength{\belowcaptionskip}{0pt}
--->
-
-<style type="text/css">
-body,
-code.bash{
-  font-size: 8px;
-}
-pre {
-  font-size: 8px
-}
-</style>
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set( cache=TRUE )
-```
-
-\pagenumbering{gobble}
-
 # Abstract
@@
 employed for both training and evaluation using a leave-one-out
-strategy.  \textcolor{blue}{Performance measures include sensitivity,
+strategy.  \textcolor{black}{Performance measures include sensitivity,
 positive predictive value, $F_1$ score, and relative volume difference.
@@
 \clearpage
-
-# Introduction
-
-## White matter hyperintensities in TBI
-
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
-
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
-Further, volume of FLAIR lesions within the corpus
-callosum, brainstem, and thalamus in patients with severe TBI correlates
-with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
-classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
-in patients who are comatose following severe TBI
-the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].
-
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
-Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
-The development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.
-In this work, we present an automated framework for quantification of WMHs in
-multi-modal MRI using the random forest machine learning technique.}
-
-## Random forests for WMH segmentation
-
-<!--
-Machine learning and pattern recognition techniques have seen increased application
-for various medical image analysis workflows (see, for example, the annual
-Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
-Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
-Popular techniques such as support vector machines and neural networks have been applied
-successfully to clinically relevant imaging tasks such as supervised image segmentation
-(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
-Facilitating the current employment of such techniques are the number of available imaging
-data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
-SciPy [@scipy] and the R project for statistical computing [@R] and their associated
-extensions.
--->
-
-The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
-normal human brain segmentation [@yi2009]) and other computer vision applications
-(e.g., [@viola2005]).
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
-delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
-
-<!--
-Of note, these
-latter random forest approaches for brain tumor segmentation have performed well in recent
-international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
--->
-
-Random forests are conceptually straightforward [@breiman2001].
-\textcolor{blue}{The basic component of
-the random forest paradigm is the ``decision tree'' often represented by a flowchart or
-graph where internal nodes represent ``tests'', or decisions, and the edges represent the outcome of those tests.
-The final, or end, nodes represent the various classsifications produced by traversal through the decision tree.
-For the proposed application, individual
-voxels (and their corresponding feature values) are introduced at the root of a particular
-decision tree and traverse the edges and internal nodes ultimately ending up at one of
-the classification nodes according to the tests at each internal node.  A single random
-forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
-(e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly sytled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
-\textcolor{blue}{Finally, Breiman} [@breiman2001]
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
-in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
-each voxel location.}
-
-In this work, we develop a concatenated random forest framework with a
-feature image set (both spatial and intensity-based)
-for segmenting WMHs in a large TBI cohort.
-The entire framework is built on the well-known open-source
-Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
-Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.
-
-[^i1]: https://github.com/stnava/ANTs
-[^i2]: https://github.com/stnava/ANTsR
-
-# Materials and Methods
-
-## Imaging
-
-MR images utilized for this initial report were acquired from a single scanner involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (see Walker et al., this issue).
-Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue).
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
-Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
-
-\input{trainingData.tex}
-
-Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
-3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; Flip angle = 9$^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ET: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
-
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
-subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All
-multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
-The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
-\textcolor{blue}{
-which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
-
-[^1]: http://nifti.nimh.nih.gov/nifti-1
-
-## Quantitative analysis
-
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1'', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
-
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
-
-### Symmetric multi-modal templates
-
-\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}
-[@landman2011]\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
-
-![Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific asymmetry-based
-features.](Figures/MMRR.png)
-
-### Feature images for WMH segmentation
-
-\textcolor{blue}{
-Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}  For the targeted application in this work,
-tissue classification is performed at the voxelwise level.  In other words, each voxel
-within the region of interest is sent through the ensemble of decision trees and receives
-a set of classification votes from each tree thus permitting a regression or classification
-solution.  Since this procedure is performed at the voxelwise level, intensity information
-alone is insufficient for good segmentation performance due to the lack of spatial context.
-For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated lesion detection
-algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
-even though machine learning and pattern recognition techniques are extremely powerful and
-have significant potential, just as crucial to outcome is the creative construction and
-deployment of salient feature images which we detail below.
-
-\input{featureImagesTable.tex}
-
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.  In Table 2, we provide a list and basic
-categorization of the feature images used for the initial (i.e., Stage 1---more on the use
-of multiple random forest stages below) segmentation of the WMHs.
-In addition Figure 3 provides a representation of a set of feature images for a single
-subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
-with seven labels:
-
-* cerebrospinal fluid (label 1),
-* gray matter (label 2),
-* white matter (label 3),
-* deep gray matter (label 4),
-* brain stem (label 5),
-* cerebellum (label 6), and
-* white matter hyperintensities (label 7).
-
-![Representation of Stage 1 feature images for subject 01C1019.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
-application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
-The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right.  This gives a probabilistic classification of
-tissue type.  Not shown are the probability and voting images for the brain stem and
-cerebellum.](Figures/featureImages.png)
-
-As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
-acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
-the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
-are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
-N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
-range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].
-
-![Sample FLAIR acquisition image slices showing both manual and random forest
-segmentations for both stages obtained during the leave-one-out evaluation.  Manual
-segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.](Figures/sampleResults.png)
-
-The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
-  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
-CSF, gray matter, white matter, deep gray matter, brain stem, and
-cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
-first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 2).
-
-Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
-We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a Manhattan radius of one voxel given the typical size
-of individual WMHs. For each of the preprocessed
-images, we calculate the difference in intensities with the corresponding warped template
-component.  Previous success in the international brain tumor segmentation
-competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 2.  We employ the
-same strategy here.
-
-<!--
-For example, the template
-difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
-
-$$S_{FLAIR} - T_{FLAIR}\left(\phi_b^{-1}\right)$$
-
-where
-
-$$\phi_b: S  \leftrightarrow \underset{b}{\leftrightsquigarrow} T$$
-
-is the transform
-which maps from the individual subject space to the
-template space and $T_{FLAIR}$ is the FLAIR template component.
--->
-
-To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
-we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature.
-
-<!--
-For the FLAIR component, this contralateral
-difference image is calculated from
-
-$$S_{FLAIR} - S_{FLAIR}\left(\phi_b^{-1}\left(\phi_R\left(\phi_b\right)\right)\right)$$
-
-where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
-the symmetric template.
--->
-
-The segmentation probability images described above are used as feature images to
-provide a spatial context for the random forest model prediction step.  Additional spatial contextual
-feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
-deep gray matter images.  These latter images are intended to help distinguish white matter
-hyperintensities from false positives induced by the partial voluming at the gray/white
-matter interface.  A third set of images are based on the voxel location within the
-space of the template.
-\textcolor{blue}{Similar feature images were used in}
-[@Anbeek:2004aa]
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
-
-### Stacked (concatenated) random forests for improved segmentation performance
-
-In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
-concatenated supervised approach, whereby the prediction output from the first random forest
-model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
-feature images of the training data (as described previously) are used to construct the
-Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
-are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
-additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
-all three aligned preprocessed images for multivariate segmentation during the
-second stage.}
-The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
-
-### Implementation
-
-As pointed out in a recent comprehensive lesion segmentation review
-[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
-is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
-our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
-this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
-ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
-use the _randomForest_ package [@liaw2002] using the default settings with
-2000 trees per model and 500 randomly selected samples per label per image.
-Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.
-
-In addition, similar to our previous offering [^2], we plan on creating
-a self-encapsulated example to showcase the proposed methodology
-\textcolor{blue}{which will also be available on github.}[^3]  The fact that
-the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
-with the manual labelings will facilitate reproducibility on the part of the reader
-as well as any interest in extending the proposed framework to other data sets.
-
-[^2]: https://github.com/ntustison/ANTsAndArboles
-
-[^3]: https://github.com/ntustison/WatchMeHyperventilate
-
-### Evaluation protocol overview
-
-In order to evaluate the protocol described, we performed a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.
-Initial processing included the creation of all Stage 1 feature images for all subjects.
-The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
-The ``truth'' labels are the seven anatomical regions given above.
-
-The leave-one-out procedure is as follows:
-
-* Create Stage 1 feature images for all 24 subjects.
-* For each of the 24 subjects:
-    + sequester the current subject and corresponding feature images.
-    + construct the Stage 1 random forest model from the remaining 23 subjects.
-    + apply the Stage 1 random forest model to the feature images of the 23 training subjects.
-    + the previous step produces the Stage 1 voting maps for all seven labels.
-    + for each of the 23 subjects, perform a Bayesian-based segmentation with an
-      MRF spatial
-      prior using the seven voting maps as additional tissue priors.
-    + construct the Stage 2 random forest model from all the Stage 1 feature images,
-      seven voting maps, and seven posterior probability maps from the previous step.
-    + send the sequestered subject through the random forest models for both stages.
-    + compare the final results with the manually-defined white matter hyperintensity
-      regions.
-
-\textcolor{blue}{Several measures have been employed in the literature for evaluating
-automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
-For this work, due to the relatively small-size distribution} \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
-four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
-incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
-The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
-
-$$ sensitivity =  \frac{ TP }{TP + FN}, $$
-
-\textcolor{blue}{and the positive predictive value (PPV),}
-
-$$ PPV =  \frac{ TP }{TP + FP}, $$
-
-\textcolor{blue}{such that $F_1$ is given by}
-
-$$ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. $$
-
-\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
-the manual and predicted lesion volumes:}
-
-$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
-
-\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
-
-* \textcolor{blue}{Quantile 1: $[1-12)$,}
-
-* \textcolor{blue}{Quantile 2: $[12-28)$, and}
-
-* \textcolor{blue}{Quantile 3: $[28-551]$}.
-
-\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
-
-# Results
-
-## White matter hyperintensity segmentation evaluation
-
-\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
-of the previously described TBI data over the three lesion volume ranges.
-These performance measures include sensitivity,
-positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
-size ranges over which these measures are computed are meant to illustrate the variation
-in performance with lesion size.}
-
-\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
-to identify which is why the sensitivity for this range is more varied compared with the largest
-set of lesions ($> 28$ voxels).   The first three measures are based on the identification of
-entire lesions.  The relative volume difference provides a direct assessment of the accuracy of
-the volumetric estimate when comparing the manually identified lesions versus the
-automatically predicted lesions.}
-
-\textcolor{blue}{
-We averaged these measures over all lesions and all subjects which resulted in the
-following values:}
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
-
-    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
-
-    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
-
-* \textcolor{blue}{Stage 1}
-
-    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
-
-    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
-
-    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
-
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
-
-<!--
-
-In Figure 5 are the segmentation comparisons derived from manual segmentations of the same
-data.  Despite the large variability characteristic with manual labelings in related fields
-[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
-current clinical practices and the methodology proposed herein is readily adapted to refinements
-in training data.
-
-On the left of Figure 5 are the improvement in Dice values [@tustison2009],
-over all white matter hyperintensities when comparing the segmentations between the two stages
-where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
-represents the intersection between the manual/automated lesion pair.  Performing
-the second round of supervised learning improves these Dice values.  One can also note from the
-right side of Figure 4 that the total lesion load volume illustrates a few subjects that are
-severe outliers in terms of the number of false positives.  The second round helps to
-correct this issue.
-
--->
-
-![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
-section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative
-volume difference.   These quantitative assessments are given for three quantile ranges
-spanning the range of the manually-derived lesion volumes.    Overall improvement in all
-three whole lesion-based measuers is seen as the
-second Stage RF model is applied for all three quantile ranges.
-The relative volume difference corresponding to the Stage 2 results tend to predict a
-decreased predicted volume over the Stage 1 results.}](Figures/evaluationResults.png)
-
-## Ranking feature importance
-
-After performing the leave-one-out evaluation,
-we calculated the _MeanDecreaseAccuracy_ feature values for each of the 24 subjects $\times$
-2 models per subject $=48$ total models.  This measure (per feature, per model) is calculated
-during the out-of-bag phase of the random forest model construction and quantifies the decrease
-in prediction accuracy from omitting the specified feature.  In other words, this quantity
-helps determine the importance of a particular feature and, although we save such
-efforts for future work, this information provides us with guidance for future feature
-pruning and/or additions.
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 1 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile
- and illustrate the
-stability of the feature importance across the leave-one-out models.
-At this initial stage only 31 features images are used.](Figures/averageLeaveOneOutStage1.png)
-
-![Average _MeanDecreaseAccuracy_ plots generated from the creation of all 24 random
-forest models for Stage 2 during the leave-one-out evaluation.  These plots
-are useful in providing a quantitative assessment of the predictive importance of each feature.
-Features are ranked in descending order of importance.
-The horizontal error bars provide the $95^{th}$ percentile and illustrate the
-stability of the feature importance across the leave-one-out models.
-We augment the 31 feature images from the first stage by adding an additional
-7 voting maps and 7 segmentation posteriors from application of the Bayesian-based
-segmentation for a total of 45 images for the second stage.](Figures/averageLeaveOneOutStage2.png)
-
-The resulting rankings for both Stages are given in Figures 6 and 7 where the values for the
-separate stages are averaged over the entire corresponding model set.  In addition, we
-track the variance for each feature over all models to illustrate the stability of
-the chosen features during the evaluation.  This latter information is illustrated as
-horizontal errors bars providing the $95^{th}$ percentile
-Note that the reader can cross reference Table 1 for identifying corresponding feature types
-and names.
-
-# Discussion
-
-<!-- Nick could you add some detail here on how this technique compares
-to other automated or semi automated approaches for WMH detection? -->
-
-\textcolor{blue}{
-In evaluating the segmentation results, there are limited methodologies described in the
-literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
-However, we can extrapolate from similar application areas where such methodologies have a
-much longer history of development such as white matter lesion segmentation in
-multiple sclerosis )MS).  In a recent review [@Garcia-Lorenzo:2013aa], 47 total papers
-were found during a literature search and evaluated which concern automated segmentation of white
-matter lesions in MS.  The most widely-used data for algorithmic evaluation consists
-of two cohorts from two different sites used in
-the MS Lesion segmentation challenge associated with the international MICCAI 2008
-conference.  For comparison, this challenge training data set
-has a mean lesion load of 204 ($\pm$ 752) mm$^3$ per lesion (compared with
-our mean lesion load of 33.8 ($\pm$ 55.4) voxels $\times$ 1.2 mm$^3$ per voxel =
-40.56 ($\pm$ 66.48) mm$^3$ per lesion) and the resolution is
-almost twice what is used in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).
-Note that our relative volume difference numbers are comparable to the
-relative volume difference {\em between raters} of 68 percent (cf Figure 5 (d))}
-[@vanginneken:2007aa].
-
-Regarding the feature rankings, it is interesting to note some of the other top performing features for Stage 1.
-The contralateral difference FLAIR image is  highly discriminative over the set of
-evaluation random forest models (see Figure 8).  This accords with the known clinical relevance of
-FLAIR images for identifying white matter hyperintensities and the fact that such
-pathology does not typically manifest symmetrically in both hemispheres.  Interestingly, the
-posterior maps for the deep gray matter are extremely important for accurate white matter
-hyperintensity segmentation.  Perhaps the spatial specification of deep
-gray matter aids in the removal of false positives.  Inspection of the bottom of the
-plots demonstrates the lack of discriminating features associated with the
-T1 image which is also well-known in the clinical literature.
-
-![(a) FLAIR image slice illustrating WMHs which have been manually delineated.
-The region around the WMHs is enlarged (b) in the original FLAIR and the
-(c) contralateral FLAIR difference image.](Figures/FLAIRcontralaleteralWithLesionsAlternate.png)
-
-As described earlier, for Stage 2, we used the output random forest voting maps from Stage
-1 as both features themselves and as priors for input to a Bayesian-based segmentation with
-an additional MRF spatial prior.  In Figure 7, the voting maps are labeled as
-"_RFStage1VotingMaps_" where the final numeral is associated with the brain
-parenchymal labeling given previously.  Similarly, the additional RF prior segmentation
-feature probability maps are labeled as "_RFBrainSegmentationPosteriors_".
-The Stage 2 feature importance plot follows similar
-trends as that for Stage 1 with the T1 images not contributing much to the identification
-of white matter hyperintensity voxels.  The initial voting maps from Stage 1 are extremely
-important with the top 3 being the estimated locations of the 1) gray matter, 2) white matter, and
-3) white matter hyperintensities.  Since these tissue type can be conflated based on intensity
-alone it is intuitive that such features would be important.
-
-<!--
-
-* Evaluation * Proper context * We only have manual labelings from a
-single radiologist ( * TBI white matter hyperintensities are more
-difficult to segment than MS lesions * individual lesions are smaller *
-lesion load is typically smaller * not prone to enhancement * less
-distinct from surrounding tissue ? * Despite the doubts regarding the
-efficacy of a typical segmentation validation, our contribution is
-useful because * the total segmentation framework is publicly available
-within the ANTs/ANTsR framework * although only a single site is
-analyzed, the feature images are site-agnostic * we plan to apply the
-current RF models directly to the other site data * we can then build
-the training database further by enlisting other experts * Is this the
-first work exploring segmentation of white matter hyperintensities in
-TBI? * The data (including manual tracings) will be made available as
-part of the FITBIR effort.
-
--->
-
-# Conclusions
-
-The current communications describes a supervised statistical learning
-methodology for identifying WHMs within multimodal MR brain imaging.
-This effort utilized information acquired from the manual segmentation
-of WMHs from FLAIR images to help build two-stage ensembles of decision
-trees for the automated identification of these lesions. Although only a
-single expert was used to produce the manual labelings, our intent is to
-further refine the proposed paradigm by crowdsourcing with feedback from
-other experts who interact with both the data and methodology.  Also, we
-recognize that only a single site was used for evaluating the proposed
-framework.  However, we are currently processing other site data with
-the models developed for this work and the results look promising since
-the developed features are site-agnostic.
-
-As far as we know, this is the first report utilizing a novel random
-forest approach to identify WMHs in a cohort of TBI patients.  TBI WMHs
-tend to be more difficult to segment than MS lesions as the former tend
-to be smaller with an overall smaller lesion load.  Also, enhancement
-protocols with the former tend to be less successful than with the
-latter.  As mentioned previously, the work in MS lesion segmentation is
-extensive with a handful of techniques being publicly available.
-
-<!--
-Our
-framework is also available as open-source as part of well-known
-neuroimaging tools which easily allows for additions/extensions but is
-also, as far as we know, the first random forest-based technique
-available for such application.
--->
-
-Two major meta-analyses of WMHs have been published covering the periods
-prior to [@Debette:2010aa] and after 2010 [@Kloppenborg:2014aa].
-Debette & Markus [@Debette:2010aa] found that the presence of
-WMHs was related to subsequent cognitive decline, a higher risk of
-developing dementia, stroke, and of mortality. Lesion volume at baseline
-was also predictive of cognitive decline. Kloppenborg et al.
-[@Kloppenborg:2014aa] of 23 cross-sectional studies reporting MRI and
-concurrent neuropsychological results in patients with heterogeneous
-diagnoses but without previously diagnosed cognitive impairment, found
-that WMHs were associated with cognitive deficit (effect size of -0.10,
-95% CI: -0.13 to -0.08) after controlling for age.
-
-Despite the potential clinical significance of WMHs
-these lesions receive little
-attention in current clinical workflows. When reported in a standard
-neuroradiologist interpretation, they are typically handled as
-incidental findings and are assigned little clinical significance. This
-likely reflects the impracticality of performing a detailed assessment
-of number, volume, and distribution within a qualitative
-neuroradiologist interpretation as well as the lack of correlative
-information on how the presence and distribution of these lesions may
-inform a diagnosis and prognosis in the appropriate clinical setting. To
-date, automated or semi-automated tools for the detection of WMHs have
-lacked the specificity and efficiency for the mining of large-scale
-datasets to generate highly granular data on whether these lesions
-possess any true diagnostic or prognostic value in the setting of a
-specific disease process. The present communication describes a
-supervised statistical learning tool that is appropriate for the
-application to such large-scale datasets.
-
-<!--
-
-The currently described tool is just one example of how "supervised
-learning" algorithms might be applied to aid in the diagnosis of TBI and
-other disease processes through the specific identification of features
-predictive of a given disease state. It is an important demonstration of
-the potential power of these analytical approaches in the rapid but
-comprehensive mining of information from neuroimaging examinations.
-Supervised learning algorithms are presently employed across a wide
-variety of settings for the rapid identification of predictive imaging
-features [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa]. Automobile
-manufacturers utilize these types of approaches to equip self-driving
-vehicles to recognize and respond to unique external surroundings
-through the identification of visual information sufficiently similar to
-previously assimilated training data [@HadsellSBESKML09;@FarabetCNL12].
-Similarly, in the context of the neuroimaging assessments, deep learning
-approaches may allow for the rapid identification of information
-predictive of disease state in an individual patient. These approaches
-have been applied to the segmentation of macroscopically visible
-structures [@Plis:2014aa;@Suk:2015aa;@Li:2014aa;@Liu:2015aa].
-Additionally, these approaches might be applied to the interrogation of
-imaging data in the individual patient with a primary quantitative
-output metrics to include sequences such as diffusion tensor imaging
-(DTI) and its variants, functional connectivity, perfusion weighted
-imaging, and cortical thickness assessments. At present, these advanced
-neuroimaging sequences are confined to cohort-based research studies due
-to the lack of available analytical tools to assess the information in
-the setting of the individual patient [@Mayer:2014aa]. Application of
-deep learning approaches in the context of data with primary
-quantitative outputs will require large scale normative and disease
-specific databases. Building these large scale imaging libraries is
-resource intensive and requires a multi-center approach with harmonized
-scanners between sites and correlative non-imaging clinical data. Large
-scale TBI data is becoming increasingly available through activities
-such as the Chronic Effects of Neurotrauma Consortium (CENC),
-Transforming Research and Clinical Knowledge in TBI (TRACK-TBI),
-Collaborative European Neurotrauma Effectiveness Research in TBI
-(CENTER-TBI), Department of Defense Alzheimer’s Disease Neuroimaging
-Initiative (DOD-ADNI), and other data being consolidated through FITBIR.
-In concert with any available high quality normative neuroimaging data,
-deep learning algorithms may be well positioned to help transform how
-neuroimaging is interpreted for the clinical management of patients with
-this disease process.
-
--->
-
-\clearpage
-
-## Acknowledgements
-
-The authors wish to acknowledge all other members of
-the CENC Neuroimaging Steering Committee and CENC leadership (Drs. David
-X. Cifu, Ramon Diaz-Arrastia, and Rick Williams) for their support. We
-also gratefully acknowledge the assistance of Tracy Nolen, Chris Siege
-and Kevin Wilson. We would also like to thank the study participants and
-their family members. This project was jointly supported by the
-Department of Defense (W81XWH-13-2-0095), the U.S. Department of
-Veterans Affairs (I01 CX001135 and I01 RX 002174), as well as USUHS
-Grant HU 0001-08-0001.
-
-## Declaration of Interest/Disclaimer
-
-The authors report no financial
-disclosures or conflicts of interest. The views expressed here are those
-of the authors and do not necessarily reflect the official policy of
-position of the Department of the Navy, Department of Defense, nor the
-U.S. Government. This work was prepared as a part of official duties;
-Title 17 USC §105 provides that Copyright protection under this title is
-not available for any work of the U.S. Government. Title 17 USC §101
-defines a US Government work as a work prepared by a military service
-member of employee of the US Government as part of that person’s
-official duties.
-
-\clearpage
-
-<!-- \listofchanges
-
-\clearpage -->
-
-# References

### 2016-12-02 discussionX.Rmd
ENH:  Final commit.

@@
 
-\textcolor{blue}{
+\textcolor{black}{
 In evaluating the segmentation results, there are limited methodologies described in the
 literature for automatic segmentation of WMHs despite the numerous clinical studies}
-[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{blue}{exploring the connections between WMHs and TBI.
+[@Debette:2010aa;@Kloppenborg:2014aa] \textcolor{black}{exploring the connections between WMHs and TBI.
 However, we can extrapolate from similar application areas where such methodologies have a

### 2016-12-02 figureCaptions.Rmd
BUG:  Switch subject ids.
ENH:  Final commit.

@@
 
-__Figure 3:__  Representation of Stage 1 feature images for subject 01C1019.  The
+__Figure 3:__  Representation of Stage 1 feature images for subject 9.  The
 FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
@@
 spanning the range of the manually-derived lesion volumes.    Overall improvement in all
-three whole lesion-based measuers is seen as the
+three whole lesion-based measures is seen as the
 second Stage RF model is applied for all three quantile ranges.

### 2016-12-02 figureCaptions.tex
BUG:  Switch subject ids.
ENH:  Final commit.

@@
 \textbf{Figure 3:} Representation of Stage 1 feature images for subject
-01C1019. The FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-to the space of the T1 image. The three modality images are then
-preprocessed (N4 bias correction and adaptive denoising) followed by
-application of standard ANTs brain extraction and \(n\)-tissue
-segmentation protocols using the MMRR symmetric template and
-corresponding priors applied to the T1 image. The feature images are
-then generated for voxelwise input to the RF model which results in the
-voting maps illustrated on the right. This gives a probabilistic
-classification of tissue type. Not shown are the probability and voting
-images for the brain stem and cerebellum.
+9. The FLAIR, T1-, and T2-weighted images are rigidly pre-aligned to the
+space of the T1 image. The three modality images are then preprocessed
+(N4 bias correction and adaptive denoising) followed by application of
+standard ANTs brain extraction and \(n\)-tissue segmentation protocols
+using the MMRR symmetric template and corresponding priors applied to
+the T1 image. The feature images are then generated for voxelwise input
+to the RF model which results in the voting maps illustrated on the
+right. This gives a probabilistic classification of tissue type. Not
+shown are the probability and voting images for the brain stem and
+cerebellum.
 
@@
 lesion volumes. Overall improvement in all three whole lesion-based
-measuers is seen as the second Stage RF model is applied for all three
+measures is seen as the second Stage RF model is applied for all three
 quantile ranges. The relative volume difference corresponding to the

### 2016-12-02 introX.Rmd
ENH:  Final commit.

@@
 
-\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
-WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}
+\textcolor{black}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
+WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{black}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{black}{populations.}
 
-\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{.
-Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
+\textcolor{black}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{black}{.
+Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{black}{.
 Further, volume of FLAIR lesions within the corpus
@@
 classify ``outcome'' in TBI patients where ``outcome'' refers to the
-spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{blue}{. Additionally,
+spectrum of possible prognoses from death to disability to recovery} [@Moen:2014aa]\textcolor{black}{. Additionally,
 in patients who are comatose following severe TBI
@@
 
-\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
+\textcolor{black}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.
 Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical.
@@
 [@viola2005].
-\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
+\textcolor{black}{Random forest-based paradigms have been successfully employed in the
 delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
-\textcolor{blue}{for both single and multi-modal acquisition protocols.}
+\textcolor{black}{for both single and multi-modal acquisition protocols.}
 
@@
 Random forests are conceptually straightforward [@breiman2001].
-\textcolor{blue}{The basic component of
+\textcolor{black}{The basic component of
 the random forest paradigm is the ``decision tree'' often represented by a flowchart or
@@
 forest model will consist of many such trees (often refered to as an ``ensemble'').}
-\textcolor{blue}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
+\textcolor{black}{Although decision trees had been extensively studied, the success of employing collections of such weak learners for boosting machine learning performance
 (e.g., AdaBoost)} [@schapire1990;@freund1997]
-\textcolor{blue}{influenced the similarly styled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
-\textcolor{blue}{Finally, Breiman} [@breiman2001]
-\textcolor{blue}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
+\textcolor{black}{influenced the similarly styled conglomeration of decision trees into ``forests'' with randomized node optimization} [@ho1995;@amit1997].
+\textcolor{black}{Finally, Breiman} [@breiman2001]
+\textcolor{black}{improved accuracy by random sampling of training data (i.e., ``bagging'') resulting
 in the current random forest technique applied here.}
-\textcolor{blue}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
+\textcolor{black}{As voxels and their feature values are ``pushed'' through each decision tree in the forest,  votes for each label are accumulated and converted to probability values for all classification possiblilities at
 each voxel location.}

### 2016-12-02 methods.Rmd
ENH:  Final commit.

@@
 probability estimates) of the first RF model application,
-which we denote as "Stage1", is used as input (along with the original set of
+which we denote as "Stage 1", is used as input (along with the original set of
 feature images) to a refinement RF model segmentation which we denote as "Stage 2".
@@
 four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
+The first three quantities are based on the number of false positives \textcolor{blue}{($FP$), true negatives ($TN$)},
+false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.  It should be noted that the number of true negatives is not readily
 incorporated into measures of accuracy as the quantity of "true negatives" (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.

### 2016-12-02 methodsX.Rmd
ENH:  Final commit.

@@
 
-MR images utilized for this initial report were acquired from a single scanner involved in the Chronic Effects of Neurotrauma Consortium’s (CENC) observational study (see Walker et al., this issue).
-Briefly, participants were Operation Iraqi Freedom/Operation Enduring Freedom (OIF/OEF) era Service Members and Veterans between the ages of 18-60 years with prior combat exposure and deployment(s).
-The feature images \textcolor{blue}{were derived from MR acquisitions of} 26 subjects aged 39.6 $\pm$ 8.1 years (range 28--58 years).
-Within this cohort, 24 (92%) were considered positive for TBI based upon the potential concussive events (PCE) interview process described in detail in Walker et al., this issue.
-\textcolor{blue}{All lesions were isolated in the white matter of the cerebrum.
-Table 1 provides a descriptive statistical summary of the variation in lesion
-load across the selected cohort.}
+MR images utilized for this initial report were acquired from a single
+scanner involved in the Chronic Effects of Neurotrauma Consortium’s
+(CENC) observational study (see Walker et al., this issue). Briefly,
+participants were Operation Iraqi Freedom/Operation Enduring Freedom
+(OIF/OEF) era Service Members and Veterans between the ages of 18-60
+years with prior combat exposure and deployment(s). The feature images
+\textcolor{black}{were derived from MR acquisitions of} 26 subjects aged
+39.6 $\pm$ 8.1 years (range 28--58 years). Within this cohort, 24 (92%)
+were considered positive for TBI based upon the potential concussive
+events (PCE) interview process described in detail in Walker et al.,
+this issue. \textcolor{black}{All lesions were isolated in the white
+matter of the cerebrum.  Table 1 provides a descriptive statistical
+summary of the variation in lesion load across the selected cohort.}
 
@@
 
-Images were acquired on a Philips 3.0T Ingenia system with an 8-channel SENSE head coil (Philips Medical Systems, Best, Netherlands).
-3D FLAIR sequences were acquired with a turbo spin echo inversion recovery sequence with the following parameters:  repetition time (TR) = 4800 ms, echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 $\times$ 256 mm FOV.
-3D T1-weighted sequences were acquired with a fast field echo (FFE) sequence with the following parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240; flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.
-In addition, 3D T2-weighted images were acquired with a turbo spin echo sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ETL: = 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition matrix, and 256 x 256 mm FOV.
+Images were acquired on a Philips 3.0T Ingenia system with an 8-channel
+SENSE head coil (Philips Medical Systems, Best, Netherlands).  3D FLAIR
+sequences were acquired with a turbo spin echo inversion recovery
+sequence with the following parameters:  repetition time (TR) = 4800 ms,
+echo time (TE) = 325 ms, inversion time (TI) = 1650 ms; 170 sagittal
+slices with a 1.2 mm slice thickness, 256 $\times$ 256 acquisition
+matrix, and 256 $\times$ 256 mm FOV.  3D T1-weighted sequences were
+acquired with a fast field echo (FFE) sequence with the following
+parameters:  TR = 6.8 ms, TE = 3.2 ms, echo train length (ETL) = 240;
+flip angle = $9^\circ$, 170 sagittal slices with a 1.2 mm slice
+thickness, 256x240 acquisition matrix, and 256 $\times$ 256 mm FOV.  In
+addition, 3D T2-weighted images were acquired with a turbo spin echo
+sequence with the following parameters:  TR = 2500 ms, TE = 245 ms, ETL:
+= 133; 170 sagittal slices with a 1.2 mm slice thickness, 256 $\times$
+256 acquisition matrix, and 256 x 256 mm FOV.
 
-\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
-subjects.
-J.
-R.
-S.
-is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.
-All
-multi-modal MR dicom image slices were converted to the nifti file format} [^1].
-\textcolor{blue}{All nifti image volumes for each subject were rigidly aligned to the T1 image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.
-The normalized MRI volumes were then provided to
-J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
-\textcolor{blue}{
-which has multi-image overlay capabilities for visualizing all modalities in all three
-canonical views.}
+\textcolor{black}{The first author (J. R. S.) performed the manual WMH
+tracings for all 26 subjects.  J. R. S. is a radiologist certified by
+the American Board of Radiology, with a certificate of advanced
+qualification in vascular and interventional radiology, over 18 years of
+research experience in TBI, and 6 years of clinical imaging experience.
+All multi-modal MR dicom image slices were converted to the nifti file
+format} [^1]. \textcolor{black}{All nifti image volumes for each subject
+were rigidly aligned to the T1 image of that subject using the ANTs
+software} [@Avants:2014aa] \textcolor{black}{.  The normalized MRI
+volumes were then provided to J. R. S. who traced each lesion using the
+ITK-SNAP tool} [@Yushkevich:2006aa] \textcolor{black}{ which has
+multi-image overlay capabilities for visualizing all modalities in all
+three canonical views.}
 
@@
 
-\textcolor{blue}{Figure 1 provides a graphical overview of the proposed workflow.  The major components
-include offline generation of symmetric multimodal templates, the creation of
-feature images from the training data which are then employed for statistical
-prediction using a concatenated random forest framework.  This framework involves
-the use of two random forest models where the outcome (i.e., the tissue
-probability estimates) of the first RF model application,
-which we denote as ``Stage1'', is used as input (along with the original set of
-feature images) to a refinement RF model segmentation which we denote as ``Stage 2''.
-Once these offline steps
-are performed, a new (i.e., unsegmented) subject can then be processed using the proposed pipeline.}
+\textcolor{black}{Figure 1 provides a graphical overview of the proposed
+workflow.  The major components include offline generation of symmetric
+multimodal templates, the creation of feature images from the training
+data which are then employed for statistical prediction using a
+concatenated random forest framework.  This framework involves the use
+of two random forest models where the outcome (i.e., the tissue
+probability estimates) of the first RF model application, which we
+denote as ``Stage1'', is used as input (along with the original set of
+feature images) to a refinement RF model segmentation which we denote as
+``Stage 2''. Once these offline steps are performed, a new (i.e.,
+unsegmented) subject can then be processed using the proposed pipeline.}
 
-![\textcolor{blue}{Workflow illustration for the proposed pipeline.  Processing of the multi-modal
-input MRI for a single subject, using the multi-modal symmetric template, results in
-the generation of the feature images.  These feature images are used as input to the
-Stage 1 RF model producing the initial RF probability map estimates.  The Stage 1
-voting maps, the original feature images, and the Stage 2 RF model result in the
-final voting maps which includes the WMH probability estimate.  Note that the RF models
-are constructed once from a set of training data which are processed using the
-same feature-construction pipeline as the single-subject input MRI.}](Figures/wmhPipeline.png)
+![\textcolor{black}{Workflow illustration for the proposed pipeline.
+Processing of the multi-modal input MRI for a single subject, using the
+multi-modal symmetric template, results in the generation of the feature
+images.  These feature images are used as input to the Stage 1 RF model
+producing the initial RF probability map estimates.  The Stage 1 voting
+maps, the original feature images, and the Stage 2 RF model result in
+the final voting maps which includes the WMH probability estimate.  Note
+that the RF models are constructed once from a set of training data
+which are processed using the same feature-construction pipeline as the
+single-subject input MRI.}](Figures/wmhPipeline.png)
 
@@
 
-\textcolor{blue}{Following} [@Tustison:2014ab] \textcolor{blue}{and} [@Tustison:2015aa], \textcolor{blue}{optimally derived templates
-serve for both brain tissue segmentation (for the derivation of tissue prior probability maps) and generation of asymmetry feature images.
-For this work we use the multi-modal data available from the public MMRR data set}
-[@landman2011]\textcolor{blue}{.
-We used all 46 multi-modal acquisitions of that study to produce a multi-modal template according to the procedure described in} [@Avants:2010aa] \textcolor{blue}{which
-results in a mean (in terms of both shape and intensity) multivariate template representing the entire cohort.  Mid-canonical slices of the FLAIR, T1, and T2 components are illustrated in Figure 2.}
+\textcolor{black}{Following} [@Tustison:2014ab] \textcolor{black}{and}
+[@Tustison:2015aa], \textcolor{black}{optimally derived templates serve
+for both brain tissue segmentation (for the derivation of tissue prior
+probability maps) and generation of asymmetry feature images. For this
+work we use the multi-modal data available from the public MMRR data
+set} [@landman2011]\textcolor{black}{.  We used all 46 multi-modal
+acquisitions of that study to produce a multi-modal template according
+to the procedure described in} [@Avants:2010aa] \textcolor{black}{which
+results in a mean (in terms of both shape and intensity) multivariate
+template representing the entire cohort.  Mid-canonical slices of the
+FLAIR, T1, and T2 components are illustrated in Figure 2.}
 
-![Canonical views of the mutlivariate, bilaterally symmetric template constructed
-from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
-the components relevant for this work).  Template construction is detailed in
-[@Tustison:2015aa].  These images are important for specific asymmetry-based
-features.](Figures/MMRR.png)
+![Canonical views of the mutlivariate, bilaterally symmetric template
+constructed from the MMRR data set [@landman2011] (only shown are the
+FLAIR, T1, and T2 modalities--- the components relevant for this work).
+Template construction is detailed in [@Tustison:2015aa].  These images
+are important for specific asymmetry-based features.](Figures/MMRR.png)
 
@@
 
-\textcolor{blue}{
-Crucial to these supervised segmentation approaches are the creation and selection of
-``features'' as input (i.e., feature images constructed from the training data)
-in conjunction with expertly identified structures of interest
-(i.e., WMHs) for model construction.}  For the targeted application in this work,
-tissue classification is performed at the voxelwise level.  In other words, each voxel
-within the region of interest is sent through the ensemble of decision trees and receives
-a set of classification votes from each tree thus permitting a regression or classification
-solution.  Since this procedure is performed at the voxelwise level, intensity information
-alone is insufficient for good segmentation performance due to the lack of spatial context.
-For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
-periventricular caps in normal subjects which often confounds automated lesion detection
-algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
-even though machine learning and pattern recognition techniques are extremely powerful and
-have significant potential, just as crucial to outcome is the creative construction and
-deployment of salient feature images which we detail below.
+\textcolor{black}{ Crucial to these supervised segmentation approaches
+are the creation and selection of ``features'' as input (i.e., feature
+images constructed from the training data) in conjunction with expertly
+identified structures of interest (i.e., WMHs) for model construction.}
+For the targeted application in this work, tissue classification is
+performed at the voxelwise level.  In other words, each voxel within the
+region of interest is sent through the ensemble of decision trees and
+receives a set of classification votes from each tree thus permitting a
+regression or classification solution.  Since this procedure is
+performed at the voxelwise level, intensity information alone is
+insufficient for good segmentation performance due to the lack of
+spatial context. For example, as pointed out in [@Neema:2009aa], higher
+intensities can be found at the periventricular caps in normal subjects
+which often confounds automated lesion detection algorithms.  Other
+potential confounds include MR signal inhomogeneity and noise.
+Therefore, even though machine learning and pattern recognition
+techniques are extremely powerful and have significant potential, just
+as crucial to outcome is the creative construction and deployment of
+salient feature images which we detail below.
 
@@
 
-Supervised methodologies are uniquely characterized, in part, by the feature images that
-are used to identify the regions of interest.  In Table 2, we provide a list and basic
-categorization of the feature images used for the initial (i.e., Stage 1---more on the use
-of multiple random forest stages below) segmentation of the WMHs.
-In addition Figure 3 provides a representation of a set of feature images for a single
-subject analyzed in this work.  Note that in this work we categorize the brain parenchyma
-with seven labels:
+Supervised methodologies are uniquely characterized, in part, by the
+feature images that are used to identify the regions of interest.  In
+Table 2, we provide a list and basic categorization of the feature
+images used for the initial (i.e., Stage 1---more on the use of multiple
+random forest stages below) segmentation of the WMHs. In addition Figure
+3 provides a representation of a set of feature images for a single
+subject analyzed in this work.  Note that in this work we categorize the
+brain parenchyma with seven labels:
 
@@
 
-![Representation of Stage 1 feature images for subject 9.  The
-FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
-[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
-(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
-application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
-the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
-The feature images are then generated for voxelwise input to the RF model which results in
-the voting maps illustrated on the right.  This gives a probabilistic classification of
-tissue type.  Not shown are the probability and voting images for the brain stem and
+![Representation of Stage 1 feature images for subject 9.  The FLAIR,
+T1-, and T2-weighted images are rigidly pre-aligned [@Avants:2014aa] to
+the space of the FLAIR image.  The three modality images are then
+preprocessed (N4 bias correction [@Tustison:2010ac] and adaptive
+denoising [@Manjon:2010aa]) followed by application of standard ANTs
+brain extraction and $n$-tissue segmentation protocols using the MMRR
+symmetric template and corresponding priors [@Tustison:2014ab] applied
+to the T1 image. The feature images are then generated for voxelwise
+input to the RF model which results in the voting maps illustrated on
+the right.  This gives a probabilistic classification of tissue type.
+Not shown are the probability and voting images for the brain stem and
 cerebellum.](Figures/featureImages.png)
 
-As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
-acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
-the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
-are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
-N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
-range $[0,1]$.  Although we could have used an alternative intensity standardization
-algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].
+As mentioned previously, input for each subject comprises FLAIR, T1-,
+and T2-weighted acquisitions.  The T1 and T2 images are rigidly
+registered to the FLAIR image using the open-source Advanced
+Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images are
+then preprocessed using the denoising algorithm of [@Manjon:2010aa]
+followed by N4 bias correction [@Tustison:2010ac] which are then
+normalized to the intensity range $[0,1]$.  Although we could have used
+an alternative intensity standardization algorithm (e.g., [@nyul2000]),
+we found that a simple linear rescaling produced better results similar
+to previous work[@Tustison:2015aa].
 
-![Sample FLAIR acquisition image slices showing both manual and random forest
-segmentations for both stages obtained during the leave-one-out evaluation.  Manual
-segmentations were performed by one of the authors and provided the ground
-truth WMH labels for training the random forest models.](Figures/sampleResults.png)
+![Sample FLAIR acquisition image slices showing both manual and random
+forest segmentations for both stages obtained during the leave-one-out
+evaluation.  Manual segmentations were performed by one of the authors
+and provided the ground truth WMH labels for training the random forest
+models.](Figures/sampleResults.png)
 
-The T1 image is then processed via the ANTs brain
-extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
-  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
-CSF, gray matter, white matter, deep gray matter, brain stem, and
-cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
-first six tissue labels given above.  Tissue prior probability maps for segmentation
-are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 2).
+The T1 image is then processed via the ANTs brain extraction and normal
+tissue segmentation pipelines [@Tustison:2014ab]. The result is a mask
+delineating the brain parenchyma and probabilistic estimates of the CSF,
+gray matter, white matter, deep gray matter, brain stem, and cerebellum
+[@Avants:2011aa].  These provide the expertly annotated labels for the
+first six tissue labels given above.  Tissue prior probability maps for
+segmentation are from multi-model optimal symmetric shape/intensity
+templates [@Tustison:2015aa] created from the public MMRR data set
+[@landman2011] (cf Figure 2).
 
-Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
-We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
-a Manhattan radius of one voxel given the typical size
-of individual WMHs. For each of the preprocessed
-images, we calculate the difference in intensities with the corresponding warped template
-component.  Previous success in the international brain tumor segmentation
-competition [@Menze:2015aa] was based on an important set of intensity features that were created
-from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 2.  We employ the
-same strategy here.
+Feature values include the preprocessed FLAIR, T1, and T2 image voxel
+intensities. We also calculate a set of neighborhood statistics (mean,
+standard deviation, and skewness) feature images using a Manhattan
+radius of one voxel given the typical size of individual WMHs. For each
+of the preprocessed images, we calculate the difference in intensities
+with the corresponding warped template component.  Previous success in
+the international brain tumor segmentation competition [@Menze:2015aa]
+was based on an important set of intensity features that were created
+from multi-modal templates mentioned previously [@Tustison:2015aa] and
+listed in Table 2.  We employ the same strategy here.
 
-<!--
-For example, the template
-difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:
+<!-- For example, the template difference feature image for the FLAIR
+image, $S_{FLAIR}$ is calculated as:
 
@@
 
-is the transform
-which maps from the individual subject space to the
-template space and $T_{FLAIR}$ is the FLAIR template component.
--->
+is the transform which maps from the individual subject space to the
+template space and $T_{FLAIR}$ is the FLAIR template component. -->
 
-To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
-we use the symmetric templates to compute the contralateral intensity
-differences as an additional intensity feature.
+To take advantage of the gross bilateral symmetry of the normal brain
+(in terms of both shape and intensity), and the fact that WMHs do not
+generally manifest symmetrically across hemispheres, we use the
+symmetric templates to compute the contralateral intensity differences
+as an additional intensity feature.
 
-<!--
-For the FLAIR component, this contralateral
-difference image is calculated from
+<!-- For the FLAIR component, this contralateral difference image is
+calculated from
 
@@
 
-where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
-the symmetric template.
--->
+where $\phi_R$ denotes a horizontal reflection perpendicular to the
+mid-sagittal plane of the symmetric template. -->
 
-The segmentation probability images described above are used as feature images to
-provide a spatial context for the random forest model prediction step.  Additional spatial contextual
-feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
-deep gray matter images.  These latter images are intended to help distinguish white matter
-hyperintensities from false positives induced by the partial voluming at the gray/white
-matter interface.  A third set of images are based on the voxel location within the
-space of the template.
-\textcolor{blue}{Similar feature images were used in}
-[@Anbeek:2004aa]
-\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
-normalization to the standard
-coordinate system provided by the template to dramatically improve spatial specificity
-across all subjects.  To generate these images, the T1 image of each subject is
-registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
-\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
-\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}
+The segmentation probability images described above are used as feature
+images to provide a spatial context for the random forest model
+prediction step.  Additional spatial contextual feature images include
+the distance maps [@maurer2003] based on the csf, gray matter, and deep
+gray matter images.  These latter images are intended to help
+distinguish white matter hyperintensities from false positives induced
+by the partial voluming at the gray/white matter interface.  A third set
+of images are based on the voxel location within the space of the
+template. \textcolor{black}{Similar feature images were used in}
+[@Anbeek:2004aa] \textcolor{black}{although, unlike the proposed
+framework, this previous work lacks normalization to the standard
+coordinate system provided by the template to dramatically improve
+spatial specificity across all subjects.  To generate these images, the
+T1 image of each subject is registered to the T1 template component
+using a B-spline variant} [@Tustison:2013ac] \textcolor{black}{of the
+well-known ANTs Symmetric Normalization (SyN) algorithm}
+[@Avants:2011ab]. \textcolor{black}{Using the derived transforms, the
+template coordinate images are warped back to the space of the
+individual subject.}
 
@@
 
-In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
-concatenated supervised approach, whereby the prediction output from the first random forest
-model serves as partial input for a second random forest model, can significantly improve
-segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
-feature images of the training data (as described previously) are used to construct the
-Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
-"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
-are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
-tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
-additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
-\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
-all three aligned preprocessed images for multivariate segmentation during the
-second stage.}
-The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.
+In previous brain tumor segmentation work [@Tustison:2015aa], it was
+demonstrated that a concatenated supervised approach, whereby the
+prediction output from the first random forest model serves as partial
+input for a second random forest model, can significantly improve
+segmentation performance.  We do the same thing for the work described
+here where we employ two stacked random forests (or two "stages").  The
+Stage 1 feature images of the training data (as described previously)
+are used to construct the Stage 1 model.  The training data Stage 1
+features are then used to produce the voxelwise "voting maps" (i.e., the
+classification count of each decision tree for each tissue label) via
+the Stage 1 random forest model.  All the Stage 1 features plus the
+Stage 1 voting maps are used as input to the Stage 2 model.  In
+addition, we use the Stage 1 voting maps as tissue priors (i.e.,
+probabilistic estimates of the tissue spatial locations) for a second
+application of the $6$-tissue segmentation algorithm with an additional
+Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
+\textcolor{black}{In order to maximize the spatial information for the
+$n$-tissue segmentation process following the voxelwise RF
+classification of Stage 1, we use all three aligned preprocessed images
+for multivariate segmentation during the second stage.} The resulting
+seven posterior probability images constitute a third additional feature
+image set for Stage 2.
 
@@
 As pointed out in a recent comprehensive lesion segmentation review
-[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
-is quite extensive, there were only four publicly available segmentation algorithms
-at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
-our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
-this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
-ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
-use the _randomForest_ package [@liaw2002] using the default settings with
-2000 trees per model and 500 randomly selected samples per label per image.
-Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.
+[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in
+the literature is quite extensive, there were only four publicly
+available segmentation algorithms at the time of writing this article.
+In contrast to the current work, none are based on supervised learning.
+As we did for our brain tumor segmentation algorithm [@Tustison:2015aa],
+all of the code described in this work is publicly available through the
+open-source ANTs/ANTsR toolkits.  Through ANTsR (an add-on toolkit
+which, in part, bridges ANTs and the R statistical project) we use the
+_randomForest_ package [@liaw2002] using the default settings with 2000
+trees per model and 500 randomly selected samples per label per image.
+Note that we saw little variation in performance when these parameters
+were changed (i.e. up to 1000 random samples and as little as 1000
+trees) which is consistent with our previous experience.
 
@@
 a self-encapsulated example to showcase the proposed methodology
-\textcolor{blue}{which will also be available on github.}[^3]  The fact that
-the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
-with the manual labelings will facilitate reproducibility on the part of the reader
-as well as any interest in extending the proposed framework to other data sets.
+\textcolor{black}{which will also be available on github.}[^3]  The fact
+that the data will also be made available through the Federal
+Interagency Traumatic Brain Injury Research (FITBIR) repository along
+with the manual labelings will facilitate reproducibility on the part of
+the reader as well as any interest in extending the proposed framework
+to other data sets.
 
@@
 
-In order to evaluate the protocol described, we performed a leave-one-out evaluation using
-the data acquired from the 24 subjects described above.
-Initial processing included the creation of all Stage 1 feature images for all subjects.
-The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.
-The ``truth'' labels are the seven anatomical regions given above.
+In order to evaluate the protocol described, we performed a
+leave-one-out evaluation using the data acquired from the 24 subjects
+described above.  Initial processing included the creation of all Stage
+1 feature images for all subjects.  The initial brain segmentation of
+each T1 image and the manual white matter hyperintensity tracings were
+combined to provide the truth labels for the training data.  The
+``truth'' labels are the seven anatomical regions given above.
 
@@
 
-\textcolor{blue}{Several measures have been employed in the literature for evaluating
-automated white matter lesion segmentation involving such quantities as lesion load,
-voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}[@Garcia-Lorenzo:2013aa]\textcolor{blue}{.
-For this work, due to the relatively small-size distribution} \textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
-four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.
-The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.
-It should be noted that the number of true negatives is not readily
-incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments.
-The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}
+\textcolor{black}{Several measures have been employed in the literature
+for evaluating automated white matter lesion segmentation involving such
+quantities as lesion load, voxel-based overlap measures (such as the
+Dice similarity coefficient), and lesion-based
+measures.}[@Garcia-Lorenzo:2013aa]\textcolor{black}{.  For this work,
+due to the relatively small-size distribution} \textcolor{black}{of the
+lesion load in our data cohort (see Table 1), we used four lesion-based
+measures:  sensitivity, positive predictive value (i.e., precision),
+$F_1$ score, and relative volume difference.  The first three quantities
+are based on the number of false positives \textcolor{blue}{($FP$), true negatives ($TN$)},
+false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.  It
+should be noted that the number of true negatives is not readily
+incorporated into measures of accuracy as the quantity of ``true
+negatives'' (i.e., correctly identified normal brain tissue) would
+severely skew accuracy assessments. The $F_1$ score is an assessment of
+accuracy which takes into account both the sensitivity,}
 
@@
 
-\textcolor{blue}{and the positive predictive value (PPV),}
+\textcolor{black}{and the positive predictive value (PPV),}
 
@@
 
-\textcolor{blue}{such that $F_1$ is given by}
+\textcolor{black}{such that $F_1$ is given by}
 
@@
 
-\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
-the manual and predicted lesion volumes:}
+\textcolor{black}{The relative volume difference is calculated for each
+of the true positive lesions using the manual and predicted lesion
+volumes:}
 
-$$ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. $$
+$$ Relative\,\,volume\,\,difference = \frac{V_{manual} -
+V_{predicted}}{V_{manual}}. $$
 
-\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
-the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) are:}
+\textcolor{black}{To illustrate how the performance of our framework
+varies with lesion size, we calculated the above measures based on
+evenly split quantiles of the manual estimates of lesion volumes into 3
+groups.  These three size ranges (in terms of the number of voxels)
+are:}
 
-* \textcolor{blue}{Quantile 1: $[1-12)$,}
+* \textcolor{black}{Quantile 1: $[1-12)$,}
 
-* \textcolor{blue}{Quantile 2: $[12-28)$, and}
+* \textcolor{black}{Quantile 2: $[12-28)$, and}
 
-* \textcolor{blue}{Quantile 3: $[28-551]$}.
+* \textcolor{black}{Quantile 3: $[28-551]$}.
 
-\textcolor{blue}{These quantiles are used to showcase the performance (cf Figure 5).}
+\textcolor{black}{These quantiles are used to showcase the performance
+(cf Figure 5).}

### 2016-12-02 resultsX.Rmd
ENH:  Final commit.

@@
 
-\textcolor{blue}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
+\textcolor{black}{In Figure 5 we provide the segmentation evaluations derived from the leave-one-out evaluation
 of the previously described TBI data over the three lesion volume ranges.
@@
 
-\textcolor{blue}{Smaller lesions ($< 12$ voxels) are more difficult
+\textcolor{black}{Smaller lesions ($< 12$ voxels) are more difficult
 to identify which is why the sensitivity for this range is more varied compared with the largest
@@
 
-\textcolor{blue}{
+\textcolor{black}{
 We averaged these measures over all lesions and all subjects which resulted in the
@@
 
-* \textcolor{blue}{Stage 1}
+* \textcolor{black}{Stage 1}
 
-    * \textcolor{blue}{sensitivity} = 0.70 $\pm$ 0.34
+    * \textcolor{black}{sensitivity} = 0.70 $\pm$ 0.34
 
-    * \textcolor{blue}{PPV} = 0.42 $\pm$ 0.36
+    * \textcolor{black}{PPV} = 0.42 $\pm$ 0.36
 
-    * \textcolor{blue}{$F_1$} = 0.47 $\pm$ 0.36
+    * \textcolor{black}{$F_1$} = 0.47 $\pm$ 0.36
 
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 38\%
+    * \textcolor{black}{relative volume difference} = 43 $\pm$ 38\%
 
-* \textcolor{blue}{Stage 2}
+* \textcolor{black}{Stage 2}
 
-    * \textcolor{blue}{sensitivity} = 0.68 $\pm$ 0.38
+    * \textcolor{black}{sensitivity} = 0.68 $\pm$ 0.38
 
-    * \textcolor{blue}{PPV} = 0.51 $\pm$ 0.40
+    * \textcolor{black}{PPV} = 0.51 $\pm$ 0.40
 
-    * \textcolor{blue}{$F_1$} = 0.52 $\pm$ 0.36
+    * \textcolor{black}{$F_1$} = 0.52 $\pm$ 0.36
 
-    * \textcolor{blue}{relative volume difference} = 43 $\pm$ 26\%
+    * \textcolor{black}{relative volume difference} = 43 $\pm$ 26\%
 
@@
 
-![\textcolor{blue}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
+![\textcolor{black}{Evaluation measures for both Stages of the leave-one-out protocol of the described protocol in the Methods
 section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, and (d) relative

### 2016-12-02 README.R

@@ -64,7 +64,7 @@ render( stitchedFile, output_format = "pdf_document" )
 
 
 singleRmdFiles <- c(
-               "responseToReviewers.Rmd",
+               "responseToReviewers2.Rmd",
                "figureCaptions.Rmd"
    )
 

### 2017-08-06 Presentation/wmh.Rmd
ENH:  Add slides for Lisa.

@@
 
-![](./Figures/lesions.png)
+\begin{center}
+\includegraphics[width=\textwidth]{./Figures/lesions.png}
+\end{center}
+
+## Current pipeline
+
+Stone et al. _Supervised learning technique for the automated identification of white 
+matter hyperintensities in traumatic brain injury_, March, 2016.
+
+\begin{center}
+\includegraphics[width=\textwidth]{../Figures/wmhPipeline.png}
+\end{center}
+
+## Features
+
+\tiny
+\begin{center}
+\input{../featureImagesTable}
+\end{center}
+\normalsize
 
@@
 
+\begin{center}
 $label \sim_{RF} feature_{1} + \ldots + feature_{n}$
+\end{center}
 
@@
 
-![](../Figures/featureImages.png)
+\begin{center}
+\includegraphics[width=\textwidth]{../Figures/featureImages.png}
+\end{center}
 
@@
 
-![](../Figures/averageLeaveOneOut.png)
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../Figures/averageLeaveOneOut.png}
+\end{center}
 
-## Sample results
+## Sample results:  Site 1
 
-![](../Figures/sampleResults.png)
+\begin{center}
+\includegraphics[width=0.75\textwidth]{../Figures/sampleResults.png}
+\end{center}
+
+## Leave-one-out evaluation
+
+\begin{center}
+\includegraphics[width=\textwidth]{../Figures/evaluationResults.png}
+\end{center}
+
+<!--
+Evaluation measures for both Stages of the leave-one-out protocol of the described protocol 
+in the Methods section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, 
+and (d) relative volume difference.   These quantitative assessments are given for three 
+quantile ranges spanning the range of the manually-derived lesion volumes.    Overall 
+improvement in all three whole lesion-based measuers is seen as the second Stage RF model is 
+applied for all three quantile ranges.  The relative volume difference corresponding to the 
+Stage 2 results tend to predict a decreased predicted volume over the Stage 1 results.
+-->
+
+## Problems with current approach
+
+* Takes a lot of time for new data
+    1. Create initial set of feature images
+    2. Run it through Stage 1 model
+    3. Create more feature images
+    4. Run subject through Stage 2 model
+* Relatively small data set
+* Evaluation/modeling limited to a single site
+* Creativity
+    * Feature image selection
+    * Are we choosing discriminative features?
+* Imbalanced data
+* No comparison with clinical data
+
+## What about deep learning?
+
+* Potentially quicker for new data
+    * Build initial model takes time 
+    * No feature images to create for new data
+    * GPU
+* Optimization learns the features
+* Mature packages
+    * TensorFlow, CNTK, Torch, Theano, Sci-TK, Caffe, mxnet, Keras
+* Need lots of training data
+
+## Current work
+
+1. Rebuild RF models from Site 1 balanced data $\leftarrow$
+2. Employ all CENC data $\leftarrow$
+3. Apply RF approach $\leftarrow$
+4. Check clinical correlations
+5. Manually refine data from 2.
+6. Re-check clinical correlations
+7. Use data from 5. to train deep learning model
+8. Re-check clinical correlations
+
+## Imbalanced data
+
+* 7 tissue voxel labels
+    * CSF $(n = 208080, 18\%)$
+    * gray matter $(n = 457437, 39\%)$
+    * white matter $(n = 317788, 27\%)$
+    * deep gray matter $(n = 33042, 2.8\%)$
+    * brain stem $(n = 17975, 1.5\%)$
+    * cerebellum $(n = 1369656, 12\%)$
+    * WMH $(n = 901, 0.07\%)$
+* SMOTe (Synthetic Minority Over-sampling Technique)
+    * For "rare" events:  $\leq 15$\%
+    * use bootstrapping and $k$-nearest neighbor

### 2017-08-06 Presentation/wmhBuild.Rmd
ENH:  Add slides for Lisa.

@@
 ---
-title: ANTs and WMHs
-author: James Stone and Nick Tustison
+title: 'Automated segmentation of WMHs'
+subtitle:  Current status and future directions
+author: |
+  | James Stone and Nick Tustison
+institute: |  
+  | Department of Radiology and Medical Imaging
+  | University of Virginia
+date: "August 11, 2017"
 bibliography:
@@
     incremental: false
-    theme:  AnnArbor
-    includes:
-      in_header: mystyle.tex
+    theme:  Singapore
     toc: no
-  pdf_document:
-    toc: true
-    highlight: zenburn
-  html_document:
-    toc: true
-    theme: readable
-  revealjs_presentation:
-    pandoc_args: [ "--slide-level", "2" ]
-    incremental: true
-    widescreen: true
-    smaller: false
-    theme: night
-    transition: fade
-    highlight: zenburn
-    center: true
-    self_contained: false
-  ioslides_presentation:
-    css: atk.css
-    incremental: false
-    widescreen: true
-    smaller: false
-    fig_width: 5
+header-includes:
+   - \usepackage{booktabs}
 ---
 
-##  ANTs beginnings:  image registration
+## WMH segmentation
+
+\begin{center}
+\includegraphics[width=\textwidth]{./Figures/lesions.png}
+\end{center}
+
+## Current pipeline
+
+Stone et al. _Supervised learning technique for the automated identification of white 
+matter hyperintensities in traumatic brain injury_, March, 2016.
+
+\begin{center}
+\includegraphics[width=\textwidth]{../Figures/wmhPipeline.png}
+\end{center}
+
+## Features
+
+\tiny
+\begin{center}
+\input{../featureImagesTable}
+\end{center}
+\normalsize
+
+## Feature images
+
+\begin{center}
+$label \sim_{RF} feature_{1} + \ldots + feature_{n}$
+\end{center}
 
@@
 
-![](./Figures/highresdiffeos.jpg)
-
-##  ANTs tools
-
-* MRI bias correction
-* image denoising
-* image and point set registration
-* $n$-tissue segmentation
-* template building
-* multi-atlas label fusion
-* brain extraction
-* cortical thickness estimation
-* ANTsR for statistics and visualization
-
-publicly available:  https://github.com/stnava/ANTs
-
-## Competitions
-
-![](./Figures/evalhistory.png)
-
-## Competition:  BRATS 2013
-
-![](./Figures/bratsData.png)
-
-*_only team out of 20 to make code publicly available._
-
-## WMH segmentation
-
-![](./Figures/lesions.png)
-
-## Feature images
-
-$label \sim_{RF} feature_{1} + \ldots + feature_{n}$
-
-<BR><BR>
-
-![](../Figures/featureImages.png)
+\begin{center}
+\includegraphics[width=\textwidth]{../Figures/featureImages.png}
+\end{center}
 

### 2017-08-06 Presentation/README.R

@@ -4,19 +4,30 @@
 #
 library(ggplot2)
 library(shiny)
-library(ggvis)
 library(rmarkdown)
 
 srcdir<-"./"
 buildrmd='wmhBuild.Rmd'
 
+# first presentation
+# rawrmds <- c( "format.Rmd",
+#               "antsBackground.Rmd",
+#               "wmh.Rmd"
+#                )
+
+# Second presentation:  August 11, 2017
 rawrmds <- c( "format.Rmd",
-              "antsBackground.Rmd",
               "wmh.Rmd"
                )
 
+
 for ( x in 1:length(rawrmds) ) {
-  if ( x == 1 )  cmd<-paste( "cat ",rawrmds[x]," > ",buildrmd ) else cmd<-paste( "cat ",rawrmds[x]," >> ",buildrmd )
+  if ( x == 1 )  
+    {
+    cmd<-paste( "cat ",rawrmds[x]," > ",buildrmd ) 
+    } else { 
+    cmd<-paste( "cat ",rawrmds[x]," >> ",buildrmd )
+    }
   system(cmd)
 }
 render(buildrmd, clean=TRUE )

### 2017-08-08 Presentation/wmh.Rmd
ENH:  notes.

@@
 \end{center}
+
+<!--
+
+* This is an update respecting current work.
+* We published this work in March 2016.
+* Have since gained access to the larger multi-site CENC data set.
+
+-->
 
@@
 
+<!--
+
+* This is the current pipeline.
+* From training data (produced by James Stone), 
+    * we generate feature images from the multi-modal training data (T1, T2, FLAIR),
+    * we then build a 2-layer random forest model
+    * A mean shape/intensity symmetric template is used for input
+* To segment an individual subject
+    * multi-modal image input
+    * generate feature images
+    * pass through RF model
+
+-->
+
 ## Features
@@
 \normalsize
+
+<!--
+
+* For many machine learning algorithms, as in the case of 
+  random forests, one of the most creative design aspects is
+  determining which feature images one should use.
+* This table is the list of feature images.
+* It should be noted that these take an hour plus to generate.
+  This time is for both training the random forest models
+  and pushing each new subject through the finished pipeline.
+* We will contrast these two characteristics 1) feature images 
+  and 2) time as differences with future work.
+
+-->
 
@@
 
+<!--
+
+* The functional dependence is illustrated by this equation:
+    the label at each voxel (i.e., what tissue type) is dependent on the
+    the weighted combination of all the features at that voxel.
+* Below are axial examples of the set of feature images used to discriminate
+  between the different tissue types.
+
+-->
+
 ## Feature importance
@@
 
+<!--
+
+* The pipeline, as shown on slide 3, utilizes two concatenated random forest models
+  where the output of the first is used as partial input to the second.
+* When generating the RF models during training, we can get a sense of the
+  "importance" of each of the features.
+
+-->
+
 ## Sample results:  Site 1
@@
 \end{center}
+
+<!--
+
+* Here are sample results:
+   * From Site 1 only 
+   * Used a leave-one-out evaluation strategy.
+       - take the ~25 subjects generated by James Stone
+       - use 24 to create the RF models and then test on the remaining subject
+
+-->
 
@@
 <!--
-Evaluation measures for both Stages of the leave-one-out protocol of the described protocol 
-in the Methods section:  (a) sensitivity, (b) positive predictive value, (c) $F_1$ score, 
-and (d) relative volume difference.   These quantitative assessments are given for three 
-quantile ranges spanning the range of the manually-derived lesion volumes.    Overall 
-improvement in all three whole lesion-based measuers is seen as the second Stage RF model is 
-applied for all three quantile ranges.  The relative volume difference corresponding to the 
-Stage 2 results tend to predict a decreased predicted volume over the Stage 1 results.
+
+* These are the evaluation measures showing improvement for the two "stages".
+    * We used 
+        - (a) sensitivity 
+        - (b) positive predictive value
+        - (c) F_1 score
+        - (d) relative volume difference.
+* Overall improvement in all three whole lesion-based measuers is seen as the second 
+  Stage RF model is applied for all three quantile ranges. 
+* The relative volume difference corresponding to the Stage 2 results tend to predict 
+  a decreased predicted volume over the Stage 1 results.
+
 -->
@@
     4. Run subject through Stage 2 model
-* Relatively small data set
-* Evaluation/modeling limited to a single site
-* Creativity
+* We are using a relatively small training data set
+* Modeling/training is limited to a single site
+* Creative engineering is needed
     * Feature image selection
     * Are we choosing discriminative features?
-* Imbalanced data
-* No comparison with clinical data
+* We have imbalanced data (e.g., way more GM than WMH voxels)
+* No comparisons with clinical data (yet)
 
@@
 
+* Exciting new possibilities with deep learning
 * Potentially quicker for new data
-    * Build initial model takes time 
-    * No feature images to create for new data
-    * GPU
-* Optimization learns the features
+    * Building initial model takes time __but__
+    * no feature images are needed to create for new data.
+* Optimization and architecture are used to learn the features
 * Mature packages
@@
 
-## Imbalanced data
+## Quick note on imbalanced data
 

### 2017-08-08 Presentation/wmhBuild.Rmd
ENH:  notes.

@@
 
+<!--
+
+* This is an update respecting current work.
+* We published this work in March 2016.
+* Have since gained access to the larger multi-site CENC data set.
+
+-->
+
 ## Current pipeline
